### 2019-08-17 Analysis/Coherence_Calculations_New.Rmd
commit full analysis

---
title: "New_Data_Coherence"
author: "Kellen Cresswell"
date: "July 1, 2019"
output: pdf_document
---

```{r}
require(msigdbr)
require(dplyr)
require(readr)
source("./functions/GeneSample.R")
require(biomaRt)

```

```{r}
#Repeating for string

data_dir = "./data/"

#Read in new string database

string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

#Get gene set
genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

#Rename to work with our functions

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)

#Repeat for second set of proteins

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dpyr::rename(hgnc_symbol_b = hgnc_symbol)

string_filt = string %>% filter(combined_score>=500)

dim(string)
```

```{r}
#Reading in biogrid data

links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")

#Select human protein-protein interactions

links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names

protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])

protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on

links = links %>% mutate(protein1 =protein1, 
                             protein2 = protein2) %>% dplyr::select(protein1, protein2)

links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

dim(links)
```

```{r}
#Read in edges generated from Connectivity_Comparison_*.Rmd

disease_overall_string = readRDS("string_edges_new.rds")

disease_overall_string_filt = readRDS("string_filt_edges_new.rds")

disease_overall_biogrid = readRDS("biogrid_edges_new.rds")

#Calculate slopes for all diseases under each category

dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

#Running coherence for string

#Read in KEGG pathways for coherence
msig = readRDS("KEGG_String.rds")
#msig =msig %>% filter(Category == "KEGG")
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  #Filter to only include specific disease coherence
  dis = dis_slopes_string_count %>% filter(Disease == i)

  #Get closest KEGG pathways in term of gene counts
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  #Take top 10 KEGG pathways
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  #Generate random networks
  j = 1
    while(j<11) {
    #Sample gene networks equal in size to disease count
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    #Get random gene networks
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))

    #Repeat if there are no internal edges
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }
    #Calculate overall slope of random disease networks
      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      #Normalize based on random slope and the median of kegg slopes (More robust)
      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string.rds")
}

```

```{r}
msig = readRDS("KEGG_String_Filt.rds")
#msig =msig %>% filter(Category == "KEGG")
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_filt_count$Disease)) {
  print(i)
  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
    coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_filt.rds")
}

```

```{r}
msig = readRDS("KEGG_Biogrid.rds")
#msig =msig %>% filter(Category == "KEGG")
coherence_tab = bind_rows()
for (i in unique(dis_slopes_biogrid_count$Disease)) {
  print(i)
  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_biogrid.rds")
}

```

# Reactome

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  print(i)
  dis = dis_slopes_string_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_react.rds")
}

```

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_filt_count$Disease)) {
  print(i)
  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_filt_react.rds")
}

```

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_biogrid_count$Disease)) {
  print(i)
  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_biogrid_react.rds")
}

```

# CC

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  print(i)
  dis = dis_slopes_string_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_CC.rds")
}

```

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_filt_count$Disease)) {
  print(i)
  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_filt_CC.rds")
}

```

```{r}
msig = readRDS("Msigdf.rds")
msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
coherence_tab = bind_rows()
for (i in unique(dis_slopes_biogrid_count$Disease)) {
  print(i)
  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_biogrid_CC.rds")
}

```

```{r}
#Read in results generated above

#KEGG

coherence_string_filt = readRDS("coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
coherence_string = readRDS("coherence_string.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid = readRDS("coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)

#Reactome

coherence_string_filt_react = readRDS("coherence_string_filt_react.rds") %>% dplyr::select(Disease, Coherence)
coherence_string_react = readRDS("coherence_string_react.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid_react = readRDS("coherence_biogrid_react.rds") %>% dplyr::select(Disease, Coherence)

#CC

coherence_string_filt_cc = readRDS("coherence_string_filt_CC.rds") %>% dplyr::select(Disease, Coherence)
coherence_string_cc = readRDS("coherence_string_CC.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid_cc = readRDS("coherence_biogrid_CC.rds") %>% dplyr::select(Disease, Coherence)

#Full join KEGG coherence results together based on disease
coherence_full = full_join(coherence_biogrid, coherence_string, by = "Disease")

coherence_full = full_join(coherence_full, coherence_string_filt, by = "Disease")

#Give columns proper name

colnames(coherence_full) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")

#Save coherence full in this state for later manipulation
coherence_disease = coherence_full

#Read in old diseases/categories
# old_res = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(-Coherence.Biogrid, -Coherence.String.full, -Coherence.String.filt)

# #Read in old coherences
# old_coh = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(Disease = Trait, Coherence.Biogrid, Coherence.String.full, Coherence.String.filt)

#Bind old coherence to new for analysis
#test_coherence = left_join(coherence_full, old_coh)

#Replace trait with diseases for ease of combination
#colnames(old_res)[1] = "Disease"

#Join new and old results filtering NA SNPs (Non-shared traits)
#coherence_full = left_join(coherence_full, old_res) %>% filter(!is.na(Total.number.of.SNPs))

#Match original order
#coherence_full = coherence_full %>% dplyr::select(Disease, Category, Total.number.of.SNPs, Total.number.of.genes, Biogrid_Coherence, String_Coherence, String_Filt_Coherence, P.value.Biogrid, P.value.String.full, P.value.String.filt)

#Filter out forced . from read.csv
#colnames(coherence_full) = gsub("\\.", " ", #colnames(coherence_full))

#Replace _ with spaces for style
#colnames(coherence_full) = gsub("\\_", " ", #colnames(coherence_full))

#Rename to old first column name
#coherence_full = coherence_full %>% dpyr::rename(Trait = Disease)

#coherence_full[,5:7] = apply(coherence_full[,5:7], 2, function(x) ifelse(x<0, 0, ifelse (
  #x>1, 1, x
#)))

#Add reactome data

#coherence_full = left_join(coherence_full, coherence_biogrid_cc, by= c("Trait"= "Disease"))

#coherence_full = left_join(coherence_full, coherence_string_cc, by= c("Trait"= "Disease"))

#coherence_full = left_join(coherence_full, coherence_string_filt_cc, by= c("Trait"= "Disease"))

#coherence_full = left_join(coherence_full, coherence_biogrid_react, by= c("Trait"= "Disease"))

#coherence_full = left_join(coherence_full, coherence_string_react, by= c("Trait"= "Disease"))

#coherence_full = left_join(coherence_full, coherence_string_filt_react, by= c("Trait"= "Disease"))

#coherence_full = coherence_full %>% dpyr::rename(`Biogrid Coherence Reactome` = Coherence.x, `String Coherence Reactome` = Coherence.y, `String Filt Coherence Reactome` = Coherence.x.x, `Biogrid Coherence CC` = Coherence.y.y, `String Coherence CC` = Coherence.x.x.x, `String Filt Coherence CC` = Coherence.y.y.y)

#Move p-values to end
#coherence_full = coherence_full %>% dplyr::select(Trait:`String Filt Coherence`,`Biogrid Coherence Reactome`:`String Filt Coherence CC`)

#Commented out old results for final version

# coherence_full = left_join(coherence_full, old_coh, by = c("Trait" = "Disease"))
# 
# colnames(coherence_full)[14:16] = c("Biogrid Coherence Old", "String Coherence Old", "String Filt Coherence Old")
# 
# coherence_full = coherence_full[, c(1:10, 14,16,15, 11:13)]
# 
# coherence_no_big = coherence_full %>% filter(Trait != "Height") %>% filter(Trait != "Obesity-related traits")

```

```{r}
#Finish all disease plot (From save coherence disease frame)

#Joining REACTOME

coherence_disease = left_join(coherence_disease, coherence_biogrid_react, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_react, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_filt_react, by= c("Disease"= "Disease"))

#Joining CC

coherence_disease = left_join(coherence_disease, coherence_biogrid_cc, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_cc, by= c("Disease"= "Disease"))

coherence_disease = left_join(coherence_disease, coherence_string_filt_cc, by= c("Disease"= "Disease"))

coherence_disease = coherence_disease  %>% dplyr::rename(`Biogrid Coherence Reactome` = Coherence.x, `String Coherence Reactome` = Coherence.y, `String Filt Coherence Reactome` = Coherence.x.x, `Biogrid Coherence CC` = Coherence.y.y, `String Coherence CC` = Coherence.x.x.x, `String Filt Coherence CC` = Coherence.y.y.y)

#Get categories from slope frame

Cat = bind_rows(dis_slopes_string_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct(), dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct())

#Join to all diseases frame

coherence_disease =left_join(coherence_disease, Cat)

#Remove underscores from column names
colnames(coherence_disease) = gsub("\\_", " ", colnames(coherence_disease))

coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Biogrid Coherence`:`String Filt Coherence CC`) %>% dplyr::distinct()

#Adding gene counts

disease_select = read_csv("supplementary_table_diseases_selected.csv") %>% dplyr::select(Disease =`Disease name`, `Number of SNPs`, `Number of genes`)

#Join gene counts

coherence_disease = left_join(coherence_disease, disease_select)

coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Number of SNPs`, `Number of genes`, `Biogrid Coherence`:`String Filt Coherence CC`)
```

```{r}
write.csv(coherence_full %>% arrange(`Category`, `Trait`), "Coherence_New.csv", row.names = FALSE)

```

```{r}
#Random networks

kegg_string = readRDS("msigdf.rds") %>% filter(Category == "KEGG")

kegg_string_filt = readRDS("KEGG_String_Filt.rds")

kegg_biogrid = readRDS("KEGG_String.rds")

ke

#String

string_random = bind_rows()
for (i in kegg_string$Count) {   
  j = 0
  diseases = bind_rows()
    while(j<11) {
    diseases_curr = GeneSample(i, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
      string_random = bind_rows(string_random, diseases_curr)
      print(string_random)
}

saveRDS(string_random, "string_random_edges.rds")

#String Filtered

string_random_filt = bind_rows()
for (i in unique(kegg_string_filt$Count)) {   
  j = 0
    while(j<11) {
    diseases_curr = GeneSample(i, string_filt, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
      diseases_curr = data.frame(Count = i, Slope = coef(mod)[1])
      string_random_filt = bind_rows(string_random_filt, diseases_curr)
      print(string_random_filt)

    }
    }
}

#String Filtered

biogrid_random = bind_rows()
for (i in unique(kegg_biogrid$Count)) {   
  j = 0
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
      diseases_curr = data.frame(Count = i, Slope = coef(mod)[1])
      string_random = bind_rows(string_random, diseases_curr)
      print(string_random)

    }
    }
}

```

```{r}
string_random = string_random %>% group_by(Count) %>% summarise(Slope = median(Slope))
cor(string_random)

string_filt_random = string_filt_random %>% group_by(Count) %>% summarise(Slope = median(Slope))
cor(string_filt_random)
```

```{r}
msig_sum = readRDS("Msigdf.rds") %>% distinct() %>% filter(!grepl("_1", Category))

msig_slopes =msig_sum %>% group_by(Category) %>% summarise(`Mean Slope` = mean(Slope, na.rm = TRUE))

write.csv(msig_slopes, "./tables/Category_Slopes.csv", row.names = FALSE, col.names = TRUE, quote = FALSE)
```

```{r}
#Reading in data from cluster and binding

biogrid_p = bind_rows(readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/biogrid_p_psych_1k.rds") %>% as.data.frame(.))

string_p = bind_rows(readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_neuro_1k.rds") %>% as.data.frame(.))

string_filt_p = bind_rows(readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_neuro_1k.rds") %>% as.data.frame(.))

colnames(biogrid_p) = c("Disease", "Biogrid P-value", "Ave")
colnames(string_p) = c("Disease", "String P-value", "Ave")
colnames(string_filt_p) = c("Disease", "String Filt P-value", "Ave")

#Joining ignoring the average coherence column I saved

#Trait dataset 
coherence_full = left_join(coherence_full,biogrid_p %>% dplyr::select(-Ave))

coherence_full = left_join(coherence_full,string_p%>% dplyr::select(-Ave))

coherence_full = left_join(coherence_full,string_filt_p%>% dplyr::select(-Ave))

#Complete Disease dataset

coherence_disease = left_join(coherence_disease,biogrid_p %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = left_join(coherence_disease,string_p %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = left_join(coherence_disease,string_filt_p  %>% dplyr::select(-Ave), by = c("Disease" = "Disease"))

coherence_disease = coherence_disease %>% mutate_if(is.numeric, round, digits = 2)

coherence_disease = coherence_disease %>% mutate(`Biogrid P-value` = round( as.numeric(as.character(`Biogrid P-value`)),4)) %>% mutate(`String P-value` = round( as.numeric(as.character(`String P-value`)),4)) %>% mutate(`String Filt P-value` = round( as.numeric(as.character(`String Filt P-value`)),4))

coherence_disease =coherence_disease %>% arrange(-`String Coherence`)

#Rename KEGG pathways when necessary

coherence_disease = coherence_disease %>% dplyr::rename(`String Coherence KEGG` = `String Coherence`, `String Filt Coherence KEGG` = `String Filt Coherence`, `Biogrid Coherence KEGG` = `Biogrid Coherence`)

colnames(coherence_disease) = gsub("CC", "GOCC", colnames(coherence_disease))

#Writing to a tables

write_xlsx(coherence_disease, "./tables/supplementary_table_S1.xlsx")

```

```{r}
#Testing coherence and modularity relationship

#Binding disease select (Disease name) with modularity calculations

mod_string = readRDS("Modularity_String.rds")
mod_string_filt = readRDS("Modularity_String_Filt.rds")
mod_biogrid = readRDS("Modularity_Biogrid.rds")

mod_sum =data.frame(Disease = disease_select$Disease, Modularity_String = unlist(mod_string), Modularity_String_Filt = unlist(mod_string_filt), Modularity_Biogrid = unlist(mod_biogrid))

Mod_Frame = left_join(mod_sum, coherence_disease, "Disease")
Mod_Frame = Mod_Frame %>% dplyr::select(Disease, Category:`String Filt P-value`, `Modularity Biogrid` = Modularity_Biogrid, `Modularity String` = Modularity_String, `Modularity String Filt`= Modularity_String_Filt)
```

```{r}
mod_sum = ggplot(Mod_Frame, aes(x = `String Coherence KEGG`, y = `Modularity_String`)) + geom_point() + labs(x = "String Coherence KEGG", y = "Modularity String")

ggsave("./figures/Modularity_Coherence.png", mod_sum, height=5, width=7, units="in", dpi = 300)

write_xlsx(Mod_Frame, "./tables/Mod_Table.xlsx")

```

### 2019-08-17 Analysis/Connectivity_Comparison_Filtered.Rmd
commit full analysis

---
title: "Connectivity Comparison Filtered"
author: "Kellen Cresswell"
date: "December 5, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 6, fig.width = 9, digits = 4)
```

```{r}
source("./functions/perm_slopes.R")
source("./functions/GeneSample.R")
source("./functions/genes2interactions.R")
source("./functions/disease2frame.R")
require(KEGGREST)
require(KEGG.db)
require(readr)
require(dplyr)
require(biomaRt)
require(ggplot2)
require(broom)
require(stringr)
require(dplyr)
require(tidyr)
```

```{r}
#Repeating for string

data_dir = "./data/"

string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x, hgnc_symbol_b = hgnc_symbol.y)

string = string %>% filter(combined_score>=500)

dim(string)
```

```{r}
#Repeat for string

#Generating 20 random networks with 20 genes each

set.seed(123)

rand_nets = list()
for (i in 1:20) {
rand_nets[[i]] = GeneSample(5, string, directed = FALSE)
}

#Getting a dataframe of internal/external connectivity

rand_length = 1:length(rand_nets)

rand_frame_string = lapply(rand_length, function(y) {

  #Read through random list of gene networks

  x = rand_nets[[y]]

  #Place internal and external edges into data frame

  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
  colnames(sub_frame) = c("Genes", "Internal", "External")

  #Add category and "disease" name

  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
  })

#Bind list into data frame

rand_frame_string = bind_rows(rand_frame_string)
```

```{r}

#Get directory with diseases

disease_directory = "./data/Disease_Genes/"

disease_names = list.files(disease_directory)

disease_list = paste0("./data/Disease_Genes/", disease_names, "/")

#Read in manually selected diseases

disease_select = read_csv("supplementary_table_diseases_selected.csv") %>% filter(!is.na(`Disease name`))

#Change first column to make more clear

colnames(disease_select)[1] = "Disease"

#Creating a vector of file names from disease select

disease_files =  unique(paste0(disease_select$Disease, ".txt"))

#Make full path

disease_selected = paste0("./data/Disease_Genes/", disease_files)

```

```{r}
#Diseases

disease_length = 1:length(disease_selected)

#Loop through diseases and pull out internal and external edges

disease_frames_string = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = string)

  #Record ppi missing

  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

disease_mod_string_filt = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  mod = diseases2mod(disease_selected[x], links = string)

})

saveRDS(disease_mod_string_filt, "Modularity_String_Filt.rds")

#Pull out network information for overall data frame

disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)

#Convert to data frame

disease_overall_string = bind_rows(disease_protein_string)

disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")

for (i in unique(disease_overall_string$Disease)) {
 curr_dis = disease_overall_string %>% filter(Disease == i)

}

disease_overall_string = bind_rows(disease_overall_string, rand_frame_string)

#disease_overall_string = #readRDS("disease_overall_string.rds")

saveRDS(disease_overall_string, "string_filt_edges_new.rds")
```

```{r}
require(msigdbr)
dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

msig = readRDS("KEGG_String_Filt.rds")
#msig =msig %>% filter(Category == "KEGG")
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  print(i)
  dis = dis_slopes_string_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:20],]

  diseases = bind_rows()
  j = 1
    while(j<20) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string_20.rds")
}

```

```{r}

```

```{r}
#KEGG

kegg_directory = "./data/Disease_Genes/KEGG/"

kegg_names = list.files(kegg_directory)

kegg_list = paste0("./data/Disease_Genes/KEGG/", kegg_names)

kegg_length = length(kegg_list)

kegg_frames_string = lapply(1:kegg_length, function(x) {
  diseases = diseases2frame(kegg_list[x], links = string)
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

kegg_overall_string = lapply(kegg_frames_string, function(x) x$diseases %>% mutate(Category = "KEGG"))

disease_overall_string = bind_rows(disease_overall_string, kegg_overall_string) %>% distinct()
```

```{r}
#Load in normalization regression

pred_edges = readRDS("pred_edges.rds")
```

```{r}
#Remove edges with less than 5 total counts, square root internal edges and correct by predicted

disease_overall_string = disease_overall_string %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease)) %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4) %>% filter( (length(unique(Internal))>1) )

#Make names more informative

disease_overall_string =  disease_overall_string %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))

```

#Plot of all diseases

```{r}

#Calculate the slopes for diseases and categories

#Slope for categories

#Slopes for categories string
disease_overall_string = readRDS("disease_overall_string_filt.rds")

slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
    group_by(Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

#Slopes for diseases

dis_slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
    group_by(Disease, Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

#Add p-values

# dis_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Disease, Category) %>%
#     do({
#       perm = slope_perm(disease_overall_string, .$Disease[1], Norm_Reg =NULL, permutations = 1000)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)
# 
# cat_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Category) %>%
#     do({
#       perm = slope_perm(disease_overall_string, .$Category[1], Norm_Reg =NULL, permutations = 1000, Disease = FALSE)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)

#Add slopes to the overall dataset

disease_overall_string = left_join(disease_overall_string, slopes_string, "Category")
colnames(disease_overall_string)[7] = "Category_Slope"

disease_overall_string = left_join(disease_overall_string, dis_slopes_string[,c(1,3)], "Disease")
colnames(disease_overall_string)[8] = "Disease_Slope"

# disease_overall_string = left_join(disease_overall_string, dis_pvals_string[,c(1,3)], "Disease")
# colnames(disease_overall_string)[9] = "Slope_Pval"

```

```{r}
#Adding filtered results

cat_coherence = read_csv("./tables/supplementary_table_categories.csv")

gene_count_cat_string = disease_overall_string %>% group_by(Category) %>% summarise(gene_count_string_filt = n())

cat_coherence = full_join(gene_count_cat_string %>% filter(!(Category %in% c("KEGG", "Random")))  , cat_coherence, by = "Category")

string_filt_cat_p = readRDS("string_p_filt_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `String Filt Perm P-val` = perm )

cat_coherence = left_join(cat_coherence, string_filt_cat_p, by = "Category")

# cat_coherence = full_join(cat_pvals_string %>% filter(!(Category %in% c("KEGG", "Random"))), cat_coherence, by = "Category")

coherence_cat_string = slopes_string %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% filter(!(Category %in% c("KEGG", "Random"))) %>% distinct()

cat_coherence = left_join(cat_coherence, coherence_cat_string, by = "Category")

# colnames(cat_coherence)[2] = "String Filt Perm P-val"

colnames(cat_coherence)[10] = "Coherence String Filt"

#Rearrange to better order

cat_coherence = cat_coherence %>% dplyr::select(Category,  gene_count_biogrid
, gene_count_string, gene_count_string_filt, `Coherence String`, `Coherence Biogrid`, `Coherence String Filt`, `String Perm P-val`, `Biogrid Perm P-val`,  `String Filt Perm P-val`  )

cat_coherence  = cat_coherence %>% mutate(`String Filt Perm P-val` = as.numeric(as.character(`String Filt Perm P-val`))) %>% arrange(Category)

#cat_coherence[,2:10] = round(cat_coherence[,2:10], 3)
cat_coherence[,2:10] = round(cat_coherence[,2:10], 3)

write_csv(cat_coherence, "./tables/supplementary_table_categories.csv")

```

```{r}

#KEGG and random should be same color on all plots
#Order by slope

library(RColorBrewer)

#Set a custom color palette for both string and biogrid

library(randomcoloR)

#Set permanent colors for random and KEGG

col =  c("Red", "#999999", distinctColorPalette(12))
names(col) = c("Random", "KEGG", as.character(unique(disease_overall_string$Category))[!(unique(disease_overall_string$Category) %in% c("KEGG", "Random"))])

col_full = scale_color_manual(values = col)

category_plot_string = ggplot(disease_overall_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2)  + theme_light(base_size = 32) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + col_full + labs(color = "Category") +
   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank())

ggsave("./figures/category_plot_string_filtered.png", category_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of all categories (String)

```{r}
category_plot_string
```

```{r}
category_string = disease_overall_string %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)

col_sub = col[names(col) %in% category_string$Category]

cat_plot_string = ggplot(category_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + labs(color = "Category")

ggsave("./figures/category_disease_plot_string_filtered.png", cat_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of non-trait categories (String)

```{r}
cat_plot_string
```

```{r}
traits_string = disease_overall_string %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")

col_sub_string = col[names(col) %in% traits_string$Category]

trait_plot_string = ggplot(traits_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub_string) + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(), legend.text = element_text(size = 16))

ggsave("./manuscript/Fig4_category_trait_plot_string_filtered.png",trait_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of traits (String)

```{r}
trait_plot_string
```

```{r}
disease_anthro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))

col_dis_string = c(col[names(col) %in% disease_anthro_string$Disease], distinctColorPalette(length(unique(disease_anthro_string$Disease))))

names(col_dis_string)[-c(1,2)] = as.character(unique(disease_anthro_string$Disease))

anthro_plot_string = ggplot(disease_anthro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_string) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/anthro_plot_string_filtered.png",anthro_plot_string, height=5, width=7, units="in", dpi = 300)
```

```{r}
#Order and keep random and KEGG consistent from other plot

disease_psych_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease)

#Plotting psych results for string

col_dis = c(col[names(col) %in% disease_psych_string$Disease], distinctColorPalette(length(unique(disease_psych_string$Disease))))

names(col_dis)[-c(1,2)] = as.character(unique(disease_psych_string$Disease))

psych_plot_string = ggplot(disease_psych_string , aes(x = Internal, y = External, color = reorder(Disease, -Disease_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm")) + labs(color = "Disease")

ggsave("./figures/psych_plot_string_filtered.png",psych_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of psychiatric diseases (String)

```{r}
psych_plot_string
```

```{r}
#Subset to include neurological disorders

disease_neuro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))

#STRING

#Setting colors

col_neuro = c(col[names(col) %in% disease_neuro_string$Disease], distinctColorPalette(length(unique(disease_neuro_string$Disease))))

names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro_string$Disease))

#Plotting

neuro_plot_string = ggplot(disease_neuro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/neuro_plot_string_filtered.png", neuro_plot_string, height=5, width=7, units="in", dpi = 300)
```

# Plot of neuorologic diseases (String)

```{r}
neuro_plot_string
```

```{r}
#Function to create all other plots

dir.create("./figures/Category")

plot_coherence = function(Cat, dataset, protein_database, filtered = FALSE) {

  all_plots = list()
  #Subset based on category

sub_frame = dataset  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope))  %>% mutate(Disease_Slope = ifelse(Category == "KEGG", 0, Disease_Slope)) %>%
 arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == Cat | Disease == "KEGG" | Disease == "Random"))
print(sub_frame)

print(head(sub_frame))

  #Get colors

  color = c(col[names(col) %in% sub_frame$Disease], distinctColorPalette(length(unique(sub_frame$Disease))))

names(color)[-c(1,2)] = as.character(unique(sub_frame$Disease))

#Make plot

dis_plot = ggplot(sub_frame , aes(x = Internal, y = External, color =Disease, -Disease_Slope)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = color) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

all_plots[[i]] = dis_plot

if (filtered) {
ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, "_filtered", ".png"), height=5, width=7, units="in", dpi = 300) 
} else {
  ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, ".png"), height=5, width=7, units="in", dpi = 300) 
}

return(all_plots)
}

disease_overall_string = disease_overall_string %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall_string %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")

#Making plots for each category

disease_overall_string = disease_overall_string %>% arrange(as.character(Category))

cat_plots = lapply(unique(disease_overall_string$Category), function(x) {
  plot_coherence(x, dataset = disease_overall_string %>% ungroup() , protein_database = "string", filtered = TRUE)
})

saveRDS(cat_plots, "string_filt.rds")

#saveRDS(unique(disease_overall_string$Category), "disease_list.rds")
```

```{r}
#Addition of filtered p-value

orig_tab = read_csv("./tables/supplementary_table_diseases.csv")
# 
# new_tab = left_join(orig_tab, dis_pvals_string %>% ungroup() %>% dplyr::select(-Category), by = "Disease")
# colnames(new_tab)[13] = "String Filt Perm P-value"

string_p_filt = as.data.frame(readRDS("string_p_filt.rds"))
string_p_filt_psych = as.data.frame(readRDS("string_p_filt_psych.rds"))

string_over = rbind(string_p_filt, string_p_filt_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `String Filt Perm P-value` = perm ) %>% filter(!duplicated(Disease))

new_tab = left_join(orig_tab, string_over, "Disease") %>% mutate(`String Filt Perm P-value` = as.numeric(as.character(`String Filt Perm P-value`)))
```

```{r}
#Addition of coherence for string filtered

string_filt_diff = disease_overall_string %>% ungroup() %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()

new_tab = left_join(new_tab,string_filt_diff, by = "Disease")
colnames(new_tab)[14] = "String Filt Coherence"

#Place with other coherence measures

new_tab = new_tab[,c(1:10,14,11:13)]

new_tab = new_tab %>% arrange(Category, Disease)

# new_tab[,3:14] = round(new_tab[,3:14], 3)
new_tab[,3:14] = round(new_tab[,3:14], 3)

new_tab = new_tab %>% mutate(`String Filt Perm P-value` = ifelse(is.na(`String Coherence`), NA, `String Filt Perm P-value`))

write.csv(new_tab, "./tables/supplementary_table_diseases.csv", row.names = FALSE)
```

```{r}
#Category wise

disease_cats_string = unique(disease_overall_string$Category)
disease_cats_biogrid = unique(disease_overall$Category)

for (i in disease_cats) {
  gene_list = disease_overall_string
}

#Loop through diseases and pull out internal and external edges

disease_frames_string = lapply(disease_cats, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = string)

  #Record ppi missing

  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

```

```{r}

#Independent analysis of KEGG pathways

kegg_slopes = dis_slopes_string %>% filter(Category == "KEGG")

#Gsubing out KEGG prefix

kegg_slopes = kegg_slopes %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))

#Get names of KEGG pathways

xx = data.frame(path_name = names(as.list(KEGGPATHNAME2ID)), kegg_code = unlist(as.list(KEGGPATHNAME2ID)))

#Match name with code

kegg_slopes = kegg_slopes %>% mutate(kegg_code = gsub("hsa", "", Disease))

#Create table

kegg_slopes = left_join(kegg_slopes, xx, "kegg_code" )

kegg_frame = kegg_slopes %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>% mutate(Coherence = ((Slope - max(dis_slopes_string$Slope))/(min(dis_slopes_string$Slope) - max(dis_slopes_string$Slope))))

#Read in other table and merge

kegg_frame = kegg_frame %>% dplyr::select(hsa_code, String_Filt_Slope = Slope, String_Filt_Coherence = Coherence)

kegg_counts_string = bind_rows(kegg_overall_string) %>% group_by(Disease) %>% summarise(String_Filt_Count = n()) %>% rename(hsa_code  = Disease)

kegg_frame = left_join(kegg_frame, kegg_counts_string, "hsa_code")

kegg_prev = read_csv("KEGG_Summary.csv")

kegg_frame_full = left_join(kegg_frame, kegg_prev, "hsa_code")

#Rearranging

kegg_frame_full = kegg_frame_full %>% dplyr::select(hsa_code, Pathway, String_Filt_Count, String_Count, Biogrid_Count, String_Filt_Slope, String_Slope, Bio_Slope, String_Filt_Coherence, String_Coherence, Bio_Coherence) %>% arrange(-String_Filt_Coherence)

write.csv(kegg_frame_full, "./tables/KEGG_Summary.csv", row.names = FALSE)
```

```{r}

```

### 2019-08-17 Analysis/Connectivity_Comparison_Undirected.Rmd
commit full analysis

---
title: "Connectivity Comparison Undirected"
author: "Kellen Cresswell"
date: "June 30, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.height = 6, fig.width = 9, digits = 4)
```

```{r}
source("./functions/perm_slopes.R")
source("./functions/GeneSample.R")
source("./functions/genes2interactions.R")
source("./functions/disease2frame.R")
source("./functions/diseases2mod.R")

require(KEGGREST)
require(KEGG.db)
require(readr)
require(dplyr)
require(biomaRt)
require(ggplot2)
require(broom)
require(stringr)
require(dplyr)
require(tidyr)
require(igraph)
```

```{r}
#Reading in biogrid data

links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")

links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names

protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])

protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on

links = links %>% mutate(protein1 =protein1, 
                             protein2 = protein2) %>% dplyr::select(protein1, protein2)

links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

dim(links)
```

```{r}
#Repeating for string

data_dir = "./data/"

string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))

string= string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

colnames(string)[4] ="hgnc_symbol_a"

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

colnames(string)[5] ="hgnc_symbol_b"

dim(string)

```

```{r}
#Repeating for string

data_dir = "./data/"

string_old = read_table2(paste0(data_dir,"9606.protein.links.v10.5.txt.gz"))

string_old= string_old %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string_old

string_old = left_join(string_old, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

colnames(string_old)[4] ="hgnc_symbol_a"

string_old = left_join(string_old, genes, by = c("protein2" = "ensembl_peptide_id"))

colnames(string_old)[5] ="hgnc_symbol_b"

dim(string)
```

```{r}
combined_int = paste0(pmin(string_old$hgnc_symbol_a, string_old$hgnc_symbol_b), pmax(string_old$hgnc_symbol_a, string_old$hgnc_symbol_b)  ) %in%  paste0(pmin(string$hgnc_symbol_a, string$hgnc_symbol_b), pmax(string$hgnc_symbol_a, string$hgnc_symbol_b))

sum(combined_int)/length(combined_int)

string = string %>% mutate(Combined = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), "-", pmax(hgnc_symbol_a, hgnc_symbol_b))) %>% distinct()

string_old = string_old %>% mutate(Combined = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), "-", pmax(hgnc_symbol_a, hgnc_symbol_b))) %>% distinct()

string = string %>% arrange(Combined)
string_old = string_old %>% arrange(Combined)

#source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/venn.R")

#Venn2(string$Combined, string_old$Combined)

new_in_old = sum(string$Combined %in% string_old$Combined)/length(string$Combined)
old_in_new = sum(string_old$Combined %in% string$Combined)/length(string_old$Combined)
```

```{r}

#Get directory with diseases

disease_directory = "./data/Disease_Genes/"

disease_names = list.files(disease_directory)

disease_list = paste0("./data/Disease_Genes/", disease_names, "/")

#Read in manually selected diseases

disease_select = read_csv("supplementary_table_diseases_selected.csv")
eof = read_csv("supplementary_table_diseases_eof.csv")

disease_traits = disease_select %>% dplyr::select(`Disease name`, Category)

#Pull out disease and category information from select diseases

eof_counts = eof %>% dplyr::select(-Category)

#Join diseases to add categories and remove non-selected diseases

disease_select = inner_join(eof_counts, disease_traits, c("DISEASE.TRAIT" = "Disease name")) %>% distinct()

#Place category second

disease_select = disease_select[,c(1,6,2:5)]

#Summarizing

# disease_sum = disease_select %>% group_by(Category) %>% summarise(Med_SNP = median(snp_count_total), Med_Gene = median(Number.of.genes), Mean_Missing = mean(per_missing)) %>% arrange(-Mean_Missing)
# 
# disease_mann_whit = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$snp_count_total, g = .$Category, p.adjust.method = "none")))
# 
# disease_mann_whit_prop = disease_select[1:137,]  %>% mutate(Prop_Miss = 1- Number.of.genes/snp_count_total) %>% group_by(Category) %>% summarise(Med_Miss = median(Prop_Miss)) %>% arrange(-Med_Miss) # %>% do(tidy(pairwise.wilcox.test(.$Prop_Miss, g = .$Category, p.adjust.method = "none")))
# 
# disease_mann_whit_gene = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$Number.of.genes, g = .$Category, p.adjust.method = "none"))) %>% arrange(p.value)
# 
# disease_mann_whit_miss = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$per_missing, g = .$Category, p.adjust.method = "none"))) %>% arrange(p.value)

#Change first column to make more clear

colnames(disease_select)[1] = "Disease"

#Creating a vector of file names from disease select

disease_files =  unique(paste0(disease_select$Disease, ".txt"))

#Make full path

disease_selected = paste0("./data/Disease_Genes/", disease_files)

```

```{r}
#Loop through diseases and pull out internal and external edges

disease_length = 1:length(disease_selected)

disease_frames = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = links)

  #Record ppi missing

  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

#Calculate modularity

disease_mod = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  mod = diseases2mod(disease_selected[x], links = links)

})

saveRDS(disease_mod, "Modularity_Biogrid.rds")

#Pull out network information for overall data frame

disease_protein = lapply(disease_frames, function(x) x$diseases)

#Convert to data frame

disease_overall = bind_rows(disease_protein)

disease_overall = left_join(disease_overall, disease_select %>% dplyr::select(Disease, Category), "Disease")

#Add number of snps with no connection to original data frame

per_no_ppi_bio = sapply(disease_frames, function(x) x$lengths)

per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_bio = per_no_ppi_bio)

disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")

```

```{r}
#Repeat for string

#Loop through diseases and pull out internal and external edges

disease_frames_string = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  diseases = diseases2frame(disease_selected[x], links = string)

  #Record ppi missing

  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

#String

disease_mod_string = lapply(disease_length, function(x) {

  #Run function to find internal and external connectivity 
  mod = diseases2mod(disease_selected[x], links = string)

})

saveRDS(disease_mod_string, "Modularity_String.rds")

#Pull out network information for overall data frame

disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)

#Convert to data frame

disease_overall_string = bind_rows(disease_protein_string)

disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")

#Add number of snps with no connection to original data frame

per_no_ppi_string = sapply(disease_frames_string, function(x) x$lengths)

per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_string = per_no_ppi_string)

saveRDS(disease_overall_string, "string_edges_new.rds")

disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")

saveRDS(disease_overall_string, "string_edges_new.rds")

saveRDS(disease_overall, "biogrid_edges_new.rds")
```

```{r}
#Generating 20 random networks with 100 genes each

#disease_counts = disease_overall %>% group_by(Disease) %>% summarise(counts = n())

set.seed(123)

rand_nets = list()
for (i in 1:20) {
rand_nets[[i]] = GeneSample(1000, links, directed = FALSE)
}

# index = 1
# rand_nets = list()
# for (i in disease_counts$counts) {
#   rand_nets[[index]] = GeneSample(i, links, directed = FALSE)
#   index = index+1
# }

#Getting a dataframe of internal/external connectivity

rand_length = 1:length(rand_nets)

rand_frame = lapply(rand_length, function(y) {

  #Read through random list of gene networks

  x = rand_nets[[y]]

  #Place internal and external edges into data frame

  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
  colnames(sub_frame) = c("Genes", "Internal", "External")

  #Add category and "disease" name

  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
  })

#Bind list into data frame

rand_frame = bind_rows(rand_frame)

disease_overall = bind_rows(disease_overall, rand_frame)

```

```{r}
#Repeat for string

# disease_counts_string = disease_overall_string %>% group_by(Disease) %>% summarise(counts = n())

set.seed(123)

rand_nets = list()
for (i in 1:20) {
rand_nets[[i]] = GeneSample(1000, string, directed = FALSE)
}

#index = 1
# rand_nets = list()
# for (i in disease_counts_string$counts) {
#   rand_nets[[index]] = GeneSample(i, links, directed = FALSE)
#   index = index+1
# }

#Getting a dataframe of internal/external connectivity

rand_length = 1:length(rand_nets)

rand_frame_string = lapply(rand_length, function(y) {

  #Read through random list of gene networks

  x = rand_nets[[y]]

  #Place internal and external edges into data frame

  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
  colnames(sub_frame) = c("Genes", "Internal", "External")

  #Add category and "disease" name

  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
  })

#Bind list into data frame

rand_frame_string = bind_rows(rand_frame_string)

disease_overall_string = bind_rows(disease_overall_string, rand_frame_string)

```

```{r}
#Repeating for KEGG pathways

kegg_directory = "./data/Disease_Genes/KEGG/"

kegg_names = list.files(kegg_directory)

kegg_list = paste0("./data/Disease_Genes/KEGG/", kegg_names)

kegg_length = length(kegg_list)

kegg_frames = lapply(1:kegg_length, function(x) {
  diseases = diseases2frame(kegg_list[x], links = links)
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

kegg_overall = lapply(kegg_frames, function(x) x$diseases %>% mutate(Category = "KEGG"))

disease_overall = bind_rows(disease_overall, kegg_overall) %>% distinct()
```

```{r}
#Repeating for string

kegg_frames_string = lapply(1:kegg_length, function(x) {
  diseases = diseases2frame(kegg_list[x], links = string)
  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
  return(list(diseases = diseases, lengths = per_no_ppi))
  })

kegg_overall_string = lapply(kegg_frames_string, function(x) x$diseases %>% mutate(Category = "KEGG"))

disease_overall_string = bind_rows(disease_overall_string, kegg_overall_string) %>% distinct()
```

```{r}
#Load in normalization regression

pred_edges = readRDS("pred_edges.rds")
pred_edges_biogrid= readRDS("pred_edges_biogrid.rds")

```

```{r}
# #Permutation test for slope difference
# 
# source("./functions/perm_slopes.R")
# 
# set.seed(123)
# 
# #Run permutation test on each category for differences in slope
# 
# perm_cat_slope = disease_overall%>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))
# 
# perm_cat_slope_string = disease_overall_string %>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))
# 
# perm_slope = disease_overall%>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))

```

```{r}
#Normalizing by number of genes

#disease_zeros = disease_overall %>% group_by(Category) %>% summarise(Per_Zero = sum(Internal==0)/length(Internal))

#Remove edges with less than 5 total counts, square root internal edges and correct by predicted

disease_overall = disease_overall  %>% mutate(External = External-Internal) %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges_biogrid, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease)) %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4) %>% filter(length(unique(Internal))>1)

#Make names more informative

disease_overall = disease_overall %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))

#Repeat for string

disease_overall_string = disease_overall_string %>% mutate(External = External-Internal) %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease))  %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4)

disease_overall_string = disease_overall_string %>% group_by(Disease) %>% filter(length(unique(Internal))>1)

disease_overall_string =  disease_overall_string %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))

```

```{r}
disease_overall = readRDS("disease_overall_biogrid.rds")
disease_overall_string = readRDS("disease_overall_string.rds")

```

```{r}
require(msigdbr)
dis_slopes_string_count = disease_overall_string %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% filter(Internal != 0) %>% 
    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
    do({
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope) %>% filter(Count>4)

msig = readRDS("KEGG_String.rds")
#msig =msig %>% filter(Category == "KEGG")
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  print(i)
  dis = dis_slopes_string_count %>% filter(Disease == i)
  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
  msigs = msigs[1:10,]

  diseases = bind_rows()
  j = 1
    while(j<11) {
    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      mod_rand = lm(External ~ Internal-1, data = diseases)
      rand_slope = coef(mod_rand)[[1]]

      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Rand_Norm = rand_slope, Msig_Norm = median(msigs$Slope))

      coherence_tab = bind_rows(coherence, coherence_tab)
      saveRDS(coherence_tab, "coherence_string.rds")
}

```

```{r}

coherence_string_filt = readRDS("coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
coherence_string = readRDS("coherence_string.rds") %>% dplyr::select(Disease, Coherence)
coherence_biogrid = readRDS("coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)

coherence_full = full_join(coherence_string_filt, coherence_string, by = "Disease")

coherence_full = full_join(coherence_full, coherence_biogrid, by = "Disease")

coherence_full = coherence_full[,c(1,4,3,2)]

colnames(coherence_full) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")

old_res = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(-Coherence.Biogrid, -Coherence.String.full, -Coherence.String.filt)

old_coh = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(Disease = Trait, Coherence.Biogrid, Coherence.String.full, Coherence.String.filt)

test_coherence = test = left_join(coherence_full, old_coh)

colnames(old_res)[1] = "Disease"

coherence_full = left_join(coherence_full, old_res) %>% filter(!is.na(Total.number.of.SNPs))

coherence_full = coherence_full[,c(1,5, 6:7, 2:4,8:10)]

colnames(coherence_full) = gsub("\\.", " ", colnames(coherence_full))

colnames(coherence_full) = gsub("\\_", " ", colnames(coherence_full))

colnames(coherence_full)[1] = "Trait"

coherence_full[,5:7] = apply(coherence_full[,5:7], 2, function(x) ifelse(x<0, 0, ifelse (
  x>1, 1, x
)))

write.table(coherence_full, "supplementary_table_S1_New.txt", sep = "\t", row.names = FALSE)

#Read in old table I sent

old_tab = read_tsv("C:\\Users\\cresswellkg\\Downloads\\supplementary_table_S1_New.txt", col_names = TRUE)

old_tab_comb = left_join(old_tab, old_coh, "Disease")

cor(old_tab_comb$`Coherence String`, old_tab_comb$Coherence.String.filt, "complete")
```

#Plot of all diseases

```{r}

#Calculate the slopes for diseases and categories

#Slope for categories

slopes = disease_overall %>% filter(Internal != 0) %>% 
    group_by(Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

#Slope for diseases

dis_slopes = disease_overall %>% filter(Internal != 0) %>% 
    group_by(Disease, Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

#Permutation test

# dis_pvals = disease_overall %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Disease, Category) %>% 
#     do({
#       perm = slope_perm(disease_overall, .$Disease[1], database = links, Norm_Reg = NULL, permutations = 1000)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)

#Slopes for Biogrid categories

# cat_pvals = disease_overall %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Category) %>%
#     do({
#       perm = slope_perm(disease_overall, .$Category[1], Norm_Reg =NULL, database = links, permutations = 1000, Disease = FALSE)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)

#Slopes for categories string

slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
    group_by(Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

#Slopes for diseases

dis_slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
    group_by(Disease, Category) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

# #Add p-values
# 
# dis_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Disease, Category) %>% 
#     do({
#       perm = slope_perm(disease_overall_string, .$Disease[1], database = string, Norm_Reg =NULL, permutations = 1000)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)
# 
# #Category p-values
# 
# cat_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
#     group_by(Category) %>%
#     do({
#       perm = slope_perm(disease_overall_string, .$Category[1], database = Norm_Reg =NULL, permutations = 1000, Disease = FALSE)
#       data.frame(P_val= perm)
#     }) %>% arrange(P_val)

#Add slopes to the overall dataset

disease_overall = left_join(disease_overall, slopes, "Category")
colnames(disease_overall)[7] = "Category_Slope"

disease_overall = left_join(disease_overall, dis_slopes[,c(1,3)], "Disease")
colnames(disease_overall)[8] = "Disease_Slope"

# disease_overall = left_join(disease_overall, dis_pvals[,c(1,3)], "Disease")
# colnames(disease_overall)[9] = "Slope_Pval"

disease_overall_string = left_join(disease_overall_string, slopes_string, "Category")
colnames(disease_overall_string)[7] = "Category_Slope"

disease_overall_string = left_join(disease_overall_string, dis_slopes_string[,c(1,3)], "Disease")
colnames(disease_overall_string)[8] = "Disease_Slope"
# 
# disease_overall_string = left_join(disease_overall_string, dis_pvals_string[,c(1,3)], "Disease")
# colnames(disease_overall_string)[9] = "Slope_Pval"

#KEGG and random should be same color on all plots
#Order by slope

library(RColorBrewer)

#Make levels into factors

disease_overall = disease_overall %>% arrange(-Category_Slope) %>% mutate(Category = factor(Category, levels =unique(Category)))

disease_overall_string = disease_overall_string %>% arrange(-Category_Slope) %>% mutate(Category = factor(Category, levels =unique(Category)))

#Filter out KEGG categories with worse coherence than fasting glucose

disease_overall = disease_overall %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")

disease_overall_string = disease_overall_string %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall_string %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")

#Set a custom color palette for both string and biogrid

library(randomcoloR)

#Set permanent colors for random and KEGG

col =  c("Red", "#999999", distinctColorPalette(12))
names(col) = c("Random", "KEGG", as.character(unique(disease_overall_string$Category))[!(unique(disease_overall_string$Category) %in% c("KEGG", "Random"))])

col_full = scale_color_manual(values = col)

category_plot = ggplot(disease_overall, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))

category_plot_string = ggplot(disease_overall_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2)  + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + col_full

ggsave("./figures/category_plot_biogrid.png", category_plot, height=5, width=7, units="in", dpi = 300)

ggsave("./figures/category_plot_string.png", category_plot_string, height=5, width=7, units="in", dpi = 300)

```

```{r}
#Table of category coherence

#Get gene counts

gene_count_cat = disease_overall %>% group_by(Category) %>% summarise(gene_count_biogrid = n())
gene_count_cat_string = disease_overall_string %>% group_by(Category) %>% summarise(gene_count_string = n())

coherence_frame =full_join(gene_count_cat, gene_count_cat_string, "Category")

#Get coherence

coherence_cat = slopes %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% distinct()

coherence_cat_string = slopes_string %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% distinct()

coherence_frame = full_join(coherence_frame,coherence_cat, "Category")

coherence_frame = full_join(coherence_frame,coherence_cat_string, "Category")

biogrid_cat_p = readRDS("biogrid_p_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `Biogrid Perm P-value` = perm )

string_cat_p = readRDS("string_p_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `String Perm P-value` = perm )

coherence_frame = left_join(coherence_frame, biogrid_cat_p, by = "Category")
coherence_frame = full_join(coherence_frame, string_cat_p, by = "Category")

# coherence_frame = full_join(coherence_frame, cat_pvals, "Category")
# 
# coherence_frame = full_join(coherence_frame, cat_pvals_string, "Category") %>% filter(!(Category %in% c("KEGG", "Random")))

#coherence_frame = coherence_frame %>% mutate(cat_pvals = NA, cat_pvals_string = NA)

colnames(coherence_frame)[4:7] = c("Coherence Biogrid", "Coherence String",  "Biogrid Perm P-val", "String Perm P-val")

coherence_frame = coherence_frame %>% filter(!is.na(`String Perm P-val`)) %>% arrange(Category)

write.csv(coherence_frame, "./tables/supplementary_table_categories.csv", row.names = FALSE)

```

# Plot of all categories (Biogrid)

```{r}
category_plot
```

# Plot of all categories (String)

```{r}
category_plot_string
```

```{r}
category = disease_overall %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)

category_string = disease_overall_string %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)

col_sub = col[names(col) %in% category$Category]

cat_plot = ggplot(category, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))

ggsave("./figures/category_disease_plot_biogrid.png", cat_plot, height=5, width=7, units="in", dpi = 300)

cat_plot_string = ggplot(category_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))

ggsave("./figures/category_disease_plot_string.png", cat_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of non-trait categories (Biogrid)

```{r}
cat_plot
```

# Plot of non-trait categories (String)

```{r}
cat_plot_string
```

```{r}
traits = disease_overall %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")

traits_string = disease_overall_string %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")

col_sub = col[names(col) %in% traits$Category]

trait_plot = ggplot(traits, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(), legend.text = element_text(size = 16))

ggsave("./manuscript/Fig_S5B_category_trait_plot_biogrid.png",trait_plot, height=5, width=7, units="in", dpi = 300)

col_sub_string = col[names(col) %in% traits_string$Category]

trait_plot_string = ggplot(traits_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub_string)  + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(), legend.text = element_text(size = 16))

ggsave("./manuscript/Fig_S5A_category_trait_plot_string.png",trait_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of traits (Biogrid)

```{r}
trait_plot
```

# Plot of traits (String)

```{r}
trait_plot_string
```

# Plot of anthropometric measures

```{r}
#Subsetting to include anthro traits

disease_anthro = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))

#Repeat for string

disease_anthro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))

#Plotting anthropometric traits

col_dis_bio = c(col[names(col) %in% disease_anthro$Disease], distinctColorPalette(length(unique(disease_anthro$Disease))))

#Giving unique colors for each disease

names(col_dis_bio)[-c(1,2)] = as.character(unique(disease_anthro$Disease))

anthro_plot = ggplot(disease_anthro , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_bio) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/anthro_plot_biogrid.png",anthro_plot, height=5, width=7, units="in", dpi = 300)

#Repeating for string

#Giving unique colors for each disease

col_dis_string = c(col[names(col) %in% disease_anthro_string$Disease], distinctColorPalette(length(unique(disease_anthro_string$Disease))))

names(col_dis_string)[-c(1,2)] = as.character(unique(disease_anthro_string$Disease))

anthro_plot_string = ggplot(disease_anthro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_string) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/anthro_plot_string.png",anthro_plot_string, height=5, width=7, units="in", dpi = 300)

```

```{r}
#Order and keep random and KEGG consistent from other plot

disease_psych = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease) %>% filter(n()>4)

disease_psych_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease)

col_dis = c(col[names(col) %in% disease_psych$Disease], distinctColorPalette(length(unique(disease_psych$Disease))))

names(col_dis)[-c(1,2)] = as.character(unique(disease_psych$Disease))

psych_plot = ggplot(disease_psych , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

#Redo colors for string

col_dis = c(col[names(col) %in% disease_psych_string$Disease], distinctColorPalette(length(unique(disease_psych_string$Disease))))

names(col_dis)[-c(1,2)] = as.character(unique(disease_psych_string$Disease))

psych_plot_string = ggplot(disease_psych_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/psych_plot_biogrid.png",psych_plot, height=5, width=7, units="in", dpi = 300)

ggsave("./figures/psych_plot_string.png",psych_plot_string, height=5, width=7, units="in", dpi = 300)

```

# Plot of psychiatric diseases (Biogrid)

```{r}
psych_plot
```

# Plot of psychiatric diseases (String)

```{r}
psych_plot_string
```

# Plot of Neuorologic diseases

```{r}
#Subset to include neurological disorders

disease_neuro = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))

disease_neuro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))

#Setting the colors for neurologic diseases

col_neuro = c(col[names(col) %in% disease_neuro$Disease], distinctColorPalette(length(unique(disease_neuro$Disease))))

names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro$Disease))

#Biogrid

neuro_plot = ggplot(disease_neuro , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/neuro_plot_biogrid.png", neuro_plot, height=5, width=7, units="in", dpi = 300)

#Repeating for STRING

#Names for 10 diseases

col_neuro = c(col[names(col) %in% disease_neuro_string$Disease], distinctColorPalette(length(unique(disease_neuro_string$Disease))))

names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro_string$Disease))

neuro_plot_string = ggplot(disease_neuro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

ggsave("./figures/neuro_plot_string.png", neuro_plot_string, height=5, width=7, units="in", dpi = 300)
```

```{r}
#Function to create all other plots

dir.create("./figures/Category")

plot_coherence = function(Cat, dataset, protein_database, filtered = FALSE) {

  all_plots = list()
  #Subset based on category

sub_frame = dataset  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope))  %>% mutate(Disease_Slope = ifelse(Category == "KEGG", 0, Disease_Slope)) %>%
 arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == Cat | Disease == "KEGG" | Disease == "Random"))
print(sub_frame)

print(head(sub_frame))

  #Get colors

  color = c(col[names(col) %in% sub_frame$Disease], distinctColorPalette(length(unique(sub_frame$Disease))))

names(color)[-c(1,2)] = as.character(unique(sub_frame$Disease))

#Make plot

dis_plot = ggplot(sub_frame , aes(x = Internal, y = External, color =Disease, -Disease_Slope)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = color) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))

all_plots[[i]] = dis_plot

if (filtered) {
ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, "_filtered", ".png"), height=5, width=7, units="in", dpi = 300) 
} else {
  ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, ".png"), height=5, width=7, units="in", dpi = 300) 
}

return(all_plots)
}

#Making plots for each category

disease_overall = disease_overall %>% ungroup() %>% arrange(as.character(Category))

cat_plots = lapply(unique(disease_overall$Category), function(x) {
  plot_coherence(x, dataset = disease_overall %>% ungroup() , protein_database = "biogrid", filtered = FALSE)
})

saveRDS(cat_plots, "biogrid.rds")

disease_overall_string = disease_overall_string %>% ungroup() %>% arrange(as.character(Category))

cat_plots_string = lapply(unique(disease_overall_string$Category), function(x) {
  plot_coherence(x, dataset = disease_overall_string %>% ungroup() , protein_database = "string", filtered = FALSE)
})

saveRDS(cat_plots_string, "string.rds")

```
# Plot of select diseases

```{r}

# #Add all of KEGG and random
# 
# select_disease= disease_overall %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% filter(Disease %in% c("Height", "Asthma", "Breast cancer", "Coronary Heart Disease", 
# "Blood pressure", "Glaucoma", "Intelligence",
# "Type 2 diabetes", "LDL cholesterol", "Alzheimer's disease", "Schizophrenia", "KEGG", "Random")) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease)))
# 
# names(col_sub) = unique(select_disease$Disease)[-c(1,8)]
# 
# select_disease_plot = ggplot(select_disease, aes(x = Internal, y = External, color = Disease))  + geom_point(alpha = .1)+ geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(Random =  "#6F8273" , KEGG = "#999999", col_sub)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
# 
# ggsave("C:\\Users\\cresswellkg\\Documents\\disease-coherence\\manuscript\\figures\\disease_plot_inbio.png", select_disease_plot, height=5, width=7, units="in", dpi = 300)
```

```{r}
# disease_summary_biogrid = disease_overall %>% group_by(Disease) %>% summarise(Mean_Internal = mean(Internal), Median_Internal = median(Internal), Min_Internal = min(Internal), Max_Internal = max(Internal), Mean_External = mean(External), Median_External = median(External), Min_External = min(External), Max_External = max(External), Percent_Zero = sum(Internal == 0)/n(), Num_Non_Zero = sum(Internal !=0), Highest_Gene = Genes[which.max(Internal/External)])
```

```{r}
#3 Figures

#KEGG, Random and Diseases

#KEGG, Random and Traits

#Disease specifically from highest number but manually choose relevant ones too
```

```{r}
#Find number of SNPs with no connections

# disease_summary_string = disease_overall_string %>% group_by(Disease) %>% summarise(Mean_Internal = mean(Internal), Median_Internal = median(Internal), Min_Internal = min(Internal), Max_Internal = max(Internal), Mean_External = mean(External), Median_External = median(External), Min_External = min(External), Max_External = max(External), Percent_Zero = sum(Internal == 0)/n(), Num_Non_Zero = sum(Internal!=0), Highest_Gene = Genes[which.max(Internal/External)])

# disease_summary_combined = left_join(disease_summary_biogrid, disease_summary_string, "Disease")
# 
# colnames(disease_summary_combined) = gsub("\\.x", "_biogrid", colnames(disease_summary_combined))
# colnames(disease_summary_combined) = gsub("\\.y", "_string", colnames(disease_summary_combined))
# 
# disease_summary_combined = disease_summary_combined[,c(1,2,13,3,14,4,15,5,16,6,17,7,18,8,19,9,20,10,21,11,22,12,23)]
# 
# disease_summary_combined = disease_summary_combined %>% filter(!is.na(Mean_Internal_string))
```

```{r}

# #Calculating coherence
# 
dis_diff = disease_overall %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()

dis_diff_string = disease_overall_string %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()

disease_select = left_join(disease_select, dis_diff, by = "Disease")
disease_select = left_join(disease_select, dis_diff_string, by = "Disease")

colnames(disease_select)[9:10] = c("Biogrid Coherence", "String Coherence")

#Removing duplicates

row.names(disease_select) = NULL
disease_select = disease_select %>% arrange(-`String Coherence`) %>% filter(!duplicated(disease_select$Disease))

#Adding slope p-values

# disease_select = left_join(disease_select, dis_pvals %>% ungroup() %>% dplyr::select(-Category), by = "Disease") %>% ungroup()
# 
# disease_select = left_join(disease_select, dis_pvals_string %>% ungroup() %>% dplyr::select(-Category), by = "Disease") %>% ungroup()

# colnames(disease_select)[11:12] = c("Biogrid Perm P-value", "String Perm P-value")

# disease_select = disease_select %>% mutate(`Biogrid Perm P-value` = NA, `String Perm P-value` = NA)

#Read in p-values from cluster

biogrid_p = readRDS("biogrid_p.rds")
biogrid_p_neuro = readRDS("biogrid_p_neuro.rds")
biogrid_p_psych = readRDS("biogrid_p_psych.rds")

bio_over = rbind(biogrid_p, biogrid_p_neuro, biogrid_p_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `Biogrid Perm P-value` = perm ) %>% filter(!duplicated(Disease))

disease_select = full_join(disease_select, bio_over, by = "Disease")

string_p = readRDS("string_p.rds")
string_p_neuro = readRDS("string_p_neuro.rds")
string_p_psych = readRDS("string_p_psych.rds")

string_over = rbind(string_p, string_p_neuro, string_p_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `String Perm P-value` = perm ) %>% filter(!duplicated(Disease))

disease_select = full_join(disease_select, string_over, by = "Disease")

write.csv(disease_select, "./tables/supplementary_table_diseases.csv", row.names = FALSE)
```

```{r}
dir.create("./data/Disease_Genes/Categories_STRING")
dir.create("./data/Disease_Genes/Categories_Biogrid")

disease_cats_string = unique(disease_overall_string$Category)
disease_cats_biogrid = unique(disease_overall$Category)

for (i in disease_cats_string) {
  gene_list = disease_overall_string %>% filter(Category == i) %>% dplyr::select(Genes) %>% dplyr::distinct()

  write.table(gene_list$Genes, paste0("./data/Disease_Genes/Categories_STRING/", i, ".txt"), row.names = FALSE, col.names = FALSE)
}

for (i in disease_cats_biogrid) {
  gene_list = disease_overall %>% filter(Category == i) %>% dplyr::select(Genes) %>% dplyr::distinct()

  write.table(gene_list$Genes, paste0("./data/Disease_Genes/Categories_Biogrid/", i, ".txt"), row.names = FALSE, col.names = FALSE)
}

string_files = paste0("./data/Disease_Genes/Categories_STRING/", list.files("./data/Disease_Genes/Categories_STRING/"))

biogrid_files = paste0("./data/Disease_Genes/Categories_Biogrid/", list.files("./data/Disease_Genes/Categories_Biogrid/"))

```

```{r}

cat_frames_biogrid = lapply(biogrid_files, function(x) {

  #Run function to find internal and external connectivity 
  cats = diseases2frame(x, links = links)

  #Record ppi missing
  cats
  })

cat_frames_string = lapply(string_files, function(x) {

  #Run function to find internal and external connectivity 
  cats = diseases2frame(x, links = string)

  #Record ppi missing
  cats
  })

cat_frames_biogrid = bind_rows(cat_frames_biogrid)
cat_frames_string = bind_rows(cat_frames_string)

```

```{r}
cat_overall = cat_frames_biogrid  %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges_biogrid, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>%  ungroup()  %>% distinct()

cat_overall_string = cat_frames_string  %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup()  %>% distinct()

cat_slopes = cat_overall %>% filter(Internal != 0) %>% 
    group_by(Disease) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

cat_slopes_string = cat_overall_string %>% filter(Internal != 0) %>% 
    group_by(Disease) %>% 
    do({
      mod = lm(External ~ Internal-1, data = .)
      data.frame(Slope= coef(mod)[1])
    }) %>% arrange(Slope)

cat_overall = left_join(cat_overall, cat_slopes, by = "Disease")
cat_overall_string = left_join(cat_overall_string, cat_slopes, by = "Disease")

category_plot_comb = ggplot(cat_overall, aes(x = Internal, y = External, color = reorder(Disease, -Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))

category_plot_comb_string = ggplot(cat_overall_string, aes(x = Internal, y = External, color = reorder(Disease, -Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))

ggsave("./figures/category_plot_biogrid_combined.png", category_plot_comb, height=5, width=7, units="in", dpi = 300)

ggsave("./figures/category_plot_string_Combined.png", category_plot_comb_string, height=5, width=7, units="in", dpi = 300)

```

```{r}
#Manually set max and min for STRING and biogrid

min_bio = min(dis_slopes$Slope)
max_bio = max(dis_slopes$Slope)

min_string = min(dis_slopes_string$Slope)
max_string = max(dis_slopes_string$Slope)

#Independent analysis of KEGG pathways

kegg_slopes_string = dis_slopes_string %>% filter(Category == "KEGG")
kegg_slopes_bio = dis_slopes %>% filter(Category == "KEGG")

#Gsubing out KEGG prefix

kegg_slopes_string = kegg_slopes_string %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))
kegg_slopes_bio = kegg_slopes_bio %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))

#Get names of KEGG pathways

xx = data.frame(path_name = names(as.list(KEGGPATHNAME2ID)), kegg_code = unlist(as.list(KEGGPATHNAME2ID)))

#Match name with code

kegg_slopes_string = kegg_slopes_string %>% mutate(kegg_code = gsub("hsa", "", Disease))
kegg_slopes_bio = kegg_slopes_bio %>% mutate(kegg_code = gsub("hsa", "", Disease))

#Create table

kegg_slopes_string = left_join(kegg_slopes_string, xx, "kegg_code" )
kegg_slopes_bio = left_join(kegg_slopes_bio, xx, "kegg_code" )

kegg_frame_string = kegg_slopes_string %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>% mutate(Coherence = ((Slope - max_string))/(min_string - max_string))

kegg_frame_bio = kegg_slopes_bio %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>%  mutate(Coherence = ((Slope - max_bio))/(min_bio - max_bio))

#Strip biogrid to merge

kegg_frame_bio = kegg_frame_bio %>% dplyr::select(hsa_code, Bio_Slope = Slope, Bio_Coherence = Coherence)

colnames(kegg_frame_string)[3:4] = c("String_Slope", "String_Coherence")

kegg_frame = left_join(kegg_frame_string, kegg_frame_bio, "hsa_code")

#Get counts for each KEGG frame

kegg_counts_string = bind_rows(kegg_overall_string) %>% group_by(Disease) %>% summarise(String_Count = n()) %>% rename(hsa_code = Disease)

kegg_counts_biogrid = bind_rows(kegg_overall) %>% group_by(Disease) %>% summarise(Biogrid_Count = n()) %>% rename(hsa_code = Disease)

kegg_frame = left_join(kegg_frame, kegg_counts_string, "hsa_code")

kegg_frame = left_join(kegg_frame, kegg_counts_biogrid, "hsa_code")

write.csv(kegg_frame, "KEGG_Summary.csv", row.names = FALSE)
```

```{r}

```

### 2019-08-17 Analysis/GWAS_Prep.Rmd
commit full analysis

---
title: "GWAS_Prep"
output:
  html_document:
    toc: true
    # toc_float: true
    # theme: united
    theme: cerulean
    # number_sections: true
date: "`r Sys.Date()`"
author: "Mikhail Dozmorov"
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
#Read in data and packages

library(gwascat)
library(limma)
library(dplyr)
library(DO.db)
library(DBI)
library(MDmisc)
library(ontoCAT)
library(rowr)
require(readr)

data(ebicat37)
```

```{r}
#Initial summary

#Convert to data frame

ebi_dat = as.data.frame(mcols(ebicat37))

#Put comma seperated genes on different lines

ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = " - ")

#Remove extra whitespace from unmebed

ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)

#Summarise

init_sum = ebi_dat %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% summarise(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE)) - sum(unique(MAPPED_GENE =="")))

write.csv(init_sum, "supplementary_table_diseases_raw.csv", row.names = FALSE)

```

```{r }
#Converting aliases to gene names

alias2SymbolEbicat = function(ebicat37) {

  ebi_dat = as.data.frame(mcols(ebicat37))

  #Unembed comma and - seperated values

  ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
  ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = " - ")

  #Remove extra spaces

  ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)

  #Summarise the results

  #Keeping track of changes (sapply necessary to keep placeholders for unmapped genes)

  recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))

  recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))

  change = (ebi_dat$MAPPED_GENE == recomb_genes)

  sum_vec = ifelse(change == TRUE, "Same", ifelse(change == FALSE & recomb_genes == "", "No Match", "Converted"))

  ebi_dat$MAPPED_GENE = recomb_genes

  return(ebi_dat)
}

ebi_dat = alias2SymbolEbicat(ebicat37)
```

```{r}
#Pulling out useful info

main_inf = ebi_dat %>% dplyr::select(MAPPED_GENE, SNPS, CHR_ID, CHR_POS, INITIAL.SAMPLE.DESCRIPTION, MAPPED_TRAIT, MAPPED_TRAIT_URI, DISEASE.TRAIT)

#Updated EFO table

```

```{r}
#Using EFO IDs from ebicat to get a hierarchy

efo_loc = "https://raw.githubusercontent.com/EBISPOT/efo/master/efo.owl"

efo = getOntology(efo_loc)
# save(list = "efo",file =  "./data/efo.rdata")
# 
# load("./data/efo.rdata")

#Cleaning up efo data to remove urls and putting comma seperated values on new lines

main_inf = unembed(main_inf, "MAPPED_TRAIT_URI", sep = ",")
main_inf$MAPPED_TRAIT_URI = basename(main_inf$MAPPED_TRAIT_URI)

tree = c()
for (i in unique(main_inf$MAPPED_TRAIT_URI)) {

  #Get entire EFO tree

  parents = getAllTermParentsById(efo, i)

  #Pull out labels and make into vector

  parents = sapply(parents, getLabel)

  #Cbind and fill in non-matching entries with NA

  tree = rowr::cbind.fill(tree,parents, fill = NA)

}

#Remove original column

tree = tree[,-1]

colnames(tree) = unique(main_inf$MAPPED_TRAIT_URI)

#Pick out important disorder groups

imp_cats = c("Abnormality of refraction", "Abnormality of the eye", "age", "anthropometric measurement", "auditory system disease", "autoimmune disease", "body weights and measures", "bone measurement", "cancer", "carcinoma", "cardiovascular disease", "cardiovascular disease biomarker measurement", "cardiovascular measurement", "diabetes mellitus", "diabetes mellitus biomarker", "digestive system disease", "eye measurement", "glucose measurement", "head and neck disorder", "heart function measurement", "hematological measurement", "immune system disease", "infectious disease biomarker", "insulin measurement", "lipid measurement", "lipid or lipoprotein measurement", "lipoprotein measurement", "mental or behavioural disorder", "metabolic disease", "metabolite measurement", "neoplasm", "nervous system disease", "neurodegenerative disease", "pulmonary function measurement", "reproductive system disease", "respiratory disease biomarker", "skin disease", "sleep measurement", "temporal measurement")

#Output cats by pipe

#Get categories for each unique EFO ID

cat_ass = apply(tree, 2, function(x) c(na.omit(match(x, imp_cats))))

#Replace categoryless values with NA

cat_ass = lapply(cat_ass, function(x) {
  if (length(x)<1) {
    cat = NA 
  } else {
    cat = x
  }
  return(cat)
})

#Assign category names

mapped_groups = stack(cat_ass)
mapped_groups$values = imp_cats[mapped_groups$values]
colnames(mapped_groups) = c("Category", "MAPPED_TRAIT_URI")

#Add groups to original data frame

main_inf = left_join(main_inf,mapped_groups, by = "MAPPED_TRAIT_URI")

main_inf$Category = as.character(main_inf$Category)

main_inf = main_inf %>% group_by(MAPPED_TRAIT_URI) %>% mutate(Category = paste0(unique(Category), collapse = "|")) %>% distinct()

#Collapse genes and disease names for simplifcation purposes and then get unique combinations

names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()

#Loop through disease traits and print a gene list for each

dir.create('./data/Disease_Genes')

dis_dir = "./data/Disease_Genes/"

for (i in unique(names_genes$DISEASE.TRAIT)) {
genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()

genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()

genes_curr = unique(genes_curr)

write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
}

#Create map of EFO ids to terms

EFO_list = lapply(unique(main_inf$MAPPED_TRAIT_URI), function(x) {
  if (x == "GO_190252") {
    term = NA
  } else {
  term = getLabel(getTermById(efo, x))
  }
  return(term)
  })

names(EFO_list) = unique(main_inf$MAPPED_TRAIT_URI)

EFO_Frame = stack(EFO_list)

colnames(EFO_Frame) = c("EFO_Name", "MAPPED_TRAIT_URI")

main_inf = left_join(main_inf, EFO_Frame, "MAPPED_TRAIT_URI")

```

```{r}
main_sum = main_inf %>% ungroup() %>%  dplyr::select(DISEASE.TRAIT, MAPPED_GENE, Category, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% mutate(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE))) %>% dplyr::select(DISEASE.TRAIT, Category, snp_count_total, snp_count_no_gene, snp_prop_nogene, gene_count_total) %>% distinct()

write.csv(main_sum, "supplementary_table_diseases_eof.csv", row.names = FALSE)
```

```{r}
#Creating KEGG pathways

mart  = useMart("ensembl", dataset="hsapiens_gene_ensembl")
min_genes = 5

xx = as.list(KEGGPATHID2EXTID)     # All KEGG pathways
xx = xx[ grepl("hsa", names(xx)) ] # Keep hsa pathways only
xx_length = sapply(xx, length)     # Number of genes in each pathway
xx = xx[xx_length > min_genes]     # Keep pathways with minimum number of genes or higher
xx_sample = sample(1:length(xx), size = length(xx), replace = FALSE) # reshuffled indices
# Get all gene symbols for KEGG pathways
genes = getBM(attributes=c('entrezgene', 'hgnc_symbol'), filters='entrezgene', values=unique(unlist(xx)), mart=mart, uniqueRows=T)

kegg_paths = lapply(xx, function(x) left_join(data.frame(entrezgene = as.integer(x)), genes))

#Loop through kegg pathways and save genes

dir.create("./data/Disease_Genes/KEGG")

for (i in 1:length(kegg_paths)) {
  genes = kegg_paths[[i]][,2]
  write.table(genes, paste0("./data/Disease_Genes/KEGG/", names(kegg_paths)[i] ), row.names = FALSE, col.names = FALSE)

}

```

### 2019-08-17 manuscript/Analysis_Tables_New.Rmd
commit full analysis

---
title: "Secondary analysis of tables"
# author: "Author's Name"
# date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
library(readr)
library(ggplot2)
library(ggrepel)
library(gridExtra)
source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/venn.R")
library(corrplot)
col3 <- colorRampPalette(c("blue", "white", "red"))
library(openxlsx)
library(cowplot)
```

```{r}
max_genes <- 200 # Maximum number of genes for a network to be included

fileNameIn0  <- "tables/supplementary_table_S1.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S1.csv > tmp.csv
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease
```

### Coherence distribution

```{r}
mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% hist
mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% summary
```

### FigS1_corr_coherence.png

**Supplementary Figure S1.[FILE: Figures/FigS1_corr_coherence.png] Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG"))
# Remove large networks
mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

### FigS2_corr_references.png

**Supplementary Figure S2.[FILE: Figures/FigS2_corr_references.png] Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome", "Biogrid Coherence GOCC", "String Coherence GOCC", "String Filt Coherence GOCC"))
# Remove large networks
# mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

paste0("Average Pearson: ", diff_cors[upper.tri(diff_cors)] %>% mean()) # Average correlation

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS2_corr_references.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

\pagebreak

# Summary plots

```{r}
mtx_traits_selected_dotplot <- function(category = "Intelligence", mtx_traits = mtx_traits, databases = c("Biogrid", "STRING full", "STRING filtered"), reference = c("KEGG")) {
  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(reference, colnames(mtx_traits))]

  mtx_traits_selected_Biogrid <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
                                            PPI = paste0("Biogrid ", reference))

  mtx_traits_selected_STRING_full <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))], 
                                            PPI = paste0("STRING full ", reference))

  mtx_traits_selected_STRING_filt <- data.frame(Disease = rownames(mtx_traits_selected), 
                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
                                            PPI = paste0("STRING filtered ", reference))
  # Append plots from selected databases
  mtx_traits_selected_to_plot <- rbind()
  if( "Biogrid" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_Biogrid)
  }
  if( "STRING full" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_full)
  }
  if( "STRING filtered" %in% databases) {
    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_filt)
  }

  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[complete.cases(mtx_traits_selected_to_plot), ]

  Coherence_levels <- mtx_traits_selected_to_plot[mtx_traits_selected_to_plot$PPI == paste0("STRING filtered ", reference), ]
  if (length(setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) > 0) {
    Coherence_levels <- rbind(Coherence_levels, mtx_traits_selected_to_plot[(mtx_traits_selected_to_plot$Disease %in% setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) & mtx_traits_selected_to_plot$PPI != paste0("STRING filtered ", reference), , drop = FALSE])
  }

  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = Coherence_levels$Disease[order(Coherence_levels$Coherence)])
  PPI_levels <- unique(mtx_traits_selected_to_plot$PPI)
  mtx_traits_selected_to_plot$PPI <- factor(mtx_traits_selected_to_plot$PPI, levels = c(PPI_levels[grepl("filtered", PPI_levels)], PPI_levels[grepl("full", PPI_levels)], PPI_levels[grepl("Biogrid", PPI_levels)]))
  # mtx_traits_selected_to_plot$Pvalue <- -log10(mtx_traits_selected_to_plot$Pvalue)

  p <- ggplot(data=mtx_traits_selected_to_plot, aes(x=Disease, y=Coherence, colour=PPI, size = Coherence)) + # label = Pvalue, 
    geom_point() +
    facet_grid(. ~ PPI) + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ylab("Normalized coherence") +
    # geom_text_repel(size = 4, colour = "black") +
    coord_flip() +
    # theme_cowplot(12) +
    theme(legend.position="none") +
    theme(axis.text = element_text(size = 12)) + 
    ggtitle(category)
}
```

## Traits

### Fig3_summary_traits.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))

png(filename = "Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS3_summary_traits_KEGG.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "KEGG"))

png(filename = "Figures/FigS3_summary_traits_KEGG.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS4_summary_traits_GOCC.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "GOCC"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "GOCC"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "GOCC"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "GOCC"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "GOCC"))

png(filename = "Figures/FigS4_summary_traits_GOCC.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS5_summary_traits_Reactome.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "Reactome"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "Reactome"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "Reactome"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "Reactome"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "Reactome"))

png(filename = "Figures/FigS5_summary_traits_Reactome.png", height = 2500, width = 6000, res = 300)
lay1 <- rbind(c(1, 2, 3))
lay2 <- rbind(c(1, 2, 3))
g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

## Diseases

### Fig4_summary_diseases.png

```{r results='hide'}
# STRING filtered plots only
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))

png(filename = "Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS6_summary_diseases_KEGG.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "KEGG"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "KEGG"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "KEGG"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "KEGG"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "KEGG"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "KEGG"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "KEGG"))

png(filename = "Figures/FigS6_summary_diseases_KEGG.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS7_summary_diseases_GOCC.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "GOCC"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "GOCC"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "GOCC"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "GOCC"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "GOCC"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "GOCC"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "GOCC"))

png(filename = "Figures/FigS7_summary_diseases_GOCC.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### FigS8_summary_diseases_Reactome.png

```{r }
p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "Reactome"))
p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "Reactome"))
p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "Reactome"))
p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "Reactome"))
p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "Reactome"))
p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "Reactome"))
p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "Reactome"))

png(filename = "Figures/FigS8_summary_diseases_Reactome.png", height = 3000, width = 8000, res = 300)
lay1 <- rbind(c(1, 2, 3, 4))
lay2 <- rbind(c(1, 2, 3, 4))
g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

# Modularity investigation

```{r}
fileNameIn0  <- "tables/Mod_Table.xlsx"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
# iconv -f UTF-8 -t ISO-8859-1 -c Mod_Table.csv >tmp.txt
fileNameIn0  <- "tables/tmp.txt"
mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
rownames(mtx0) <- mtx0$Disease
```

```{r}
# Selected data for correlation
# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value" 
mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Modularity Biogrid", "Modularity String", "Modularity String Filt"))
# Remove large networks
mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
# Go through each pairs of columns and calculate pairwise complete observations
# pairwise_columns <- combn(ncol(mtx0_cor), 2)
# for(j in 1:ncol(pairwise_columns)) {
#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
# }
# Actual correlation
diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")

# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
# Selected correlations
png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
dev.off()
```

### 2019-08-17 manuscript/DREAM_gene_extraction.Rmd
commit full analysis

---
title: "Demo Document"
author: "Author's Name"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700, 
library(pander)
panderOptions('table.split.table', Inf)
set.seed(1)
library(dplyr)
options(stringsAsFactors = FALSE)
```

```{r libraries}
library(purrr)
library(openxlsx)
# devtools::install_github("alastairrushworth/inspectdf")
library(inspectdf)
```

# Settings

```{r settings}
# establish the location of the GWAS datasets
# data_dir <- "/home/stilianoudakisc/DREAM_module/4_gwas_datasets/gene_scores/"
data_dir <- "/Users/mdozmorov/Documents/Data/DREAM_module/4_gwas_datasets/gene_scores/"
# Cutoffs
pvalue_cutoff_max <- 5E-2 # Lower bound of p-value cutoff
pvalue_cutoff_min <- 5E-8 # Higher bound of p-value cutoff
```

# Load data

Full gene scores

```{r}
# establish complete path to read in each GWAS dataset
data_file <- file.path(data_dir,list.files(data_dir))

# establish the total number of GWAS datasets to loop through (should be 180)
data_length <- length(data_file)

# establish an empty list to fill in necessary parts from each GWAS dataset
gene_data_list <- vector(mode = "list", length = data_length)

for(i in 1:data_length){
  # print(list.files(data_dir)[i])
  data <- read.table(data_file[i], header = TRUE)

  # extract only gene name and p-value
  data <- data %>% dplyr::select(gene_id, pvalue) # %>% dplyr::arrange(pvalue)

  # remove .txt from the name of the GWAS dataset
  colnames(data) <- c("gene_id", gsub(".txt", "", basename(data_file[i])))

  # Save the current data
  gene_data_list[[i]] <- data
}

# Perform Recursive full joins on the list of datasets to merge all datasets (with complete list of genes) into one
gene_data_full <- gene_data_list %>% purrr::reduce(dplyr::full_join, by = "gene_id") # join_all(gene_data_list, by='gene_id', type='full')

# remove the ".sum.genescores" part common to all GWAS dataset names
names(gene_data_full) <- gsub(".sum.genescores.gz", "", names(gene_data_full), fixed = TRUE)
gene_data_full[1:5, 1:5]
```

# Exploring global cutoffs

```{r}
# Vectors to store p-value cutoffs and medians
pvalue_cutoffs    <- c()
num_genes_medians <- c()
# Get median number of genes depending on the p-value cutoff
for (pvalue_cutoff in seq(pvalue_cutoff_min, pvalue_cutoff_max, by = 5E-3)) {
  # Median at a given cutoff
  num_genes_median <- gene_data_full %>% dplyr::select(-gene_id) %>% sapply(., function(x) sum(x < pvalue_cutoff, na.rm = TRUE)) %>% median()
  # Save the results
  pvalue_cutoffs    <- c(pvalue_cutoffs, pvalue_cutoff)
  num_genes_medians <- c(num_genes_medians, num_genes_median)
}
# Quick visualization
plot(num_genes_medians, pvalue_cutoffs)
# View(cbind(num_genes_medians, pvalue_cutoffs))
# Select cutoff
pvalue_cutoff <- 0.001
# Number of genes for each disease using this cutoff
num_genes <- gene_data_full %>% dplyr::select(-gene_id) %>% sapply(., function(x) sum(x < pvalue_cutoff, na.rm = TRUE))
# Add column names
names(num_genes) <- gene_data_full %>% dplyr::select(-gene_id) %>% colnames() 
# Largest/smallest GWASs
num_genes %>% sort(decreasing = TRUE) %>% head
num_genes %>% sort(decreasing = FALSE) %>% head
# Quick summary statistics
summary(num_genes)
hist(num_genes)
```

# Exploring per category cutoffs

```{r}
mtx <- read.xlsx("https://www.biorxiv.org/content/biorxiv/early/2019/01/28/265553/DC1/embed/media-1.xlsx", startRow = 8)
mtx$Data.file <- gsub(".txt.gz", "", mtx$Data.file)
# Attach num_genes
# inspectdf
mtx %>% inspect_types()
mtx_cat <- mtx %>% inspect_cat()
mtx_cat
mtx %>% dplyr::select(GWAS.set, Category, Trait.type, GWAS.trait) %>% inspect_cat() %>% show_plot()
mtx_cat$levels$Category
```

```{r}
mtx_subset <- mtx %>% dplyr::filter(Category == "Anthropometric")
gene_data_subset <- gene_data_full[, colnames(gene_data_full) %in% mtx_subset$Data.file]
gene_data_subset %>% sapply(., function(x) sum(x < pvalue_cutoff, na.rm = TRUE)) %>% sort(decreasing = TRUE) %>% head
gene_data_subset %>% sapply(., function(x) sum(x < pvalue_cutoff, na.rm = TRUE)) %>% sort(decreasing = FALSE) %>% head
```

Conclusion: There is a big overlap among different GWASs targeting the same trait/disease. Different GWASs differ widely in the number of significant genes. Many diseases/traits are missing. The data does not seem useful.

### 2019-08-17 manuscript/misc/07-07-2019_Dozmorov_progress.Rmd
commit full analysis

---
# All BibTex reference files
bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/BRCA.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/education.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/statistics.bib]
# Reference template
csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
# Output template
output: 
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NIH_grant_style.docx
---

## Progress report, Dozmorov M., 07-07-2019

### Coherence of Molecular Mechanisms in Genes Identified Through GWAS In Major Human Disease and Traits

This project involves quantification of coherence of the molecular networks formed by disease-associated genes. A metric for coherence is defined as a ratio between the number of internal and external protein-protein interactions. To date, a major problem with network size affecting the coherence has been solved. The solution includes the creation of size-matched references for high and low coherences (KEGG and random networks, respectively). The project is currently in results-generating and manuscript writing stage.

### 2019-08-17 Analysis/data/get_data.sh

@@ -1,0 +1,4 @@
+wget https://stringdb-static.org/download/protein.links.v11.0/9606.protein.links.v11.0.txt.gz
+wget https://downloads.thebiogrid.org/Download/BioGRID/Release-Archive/BIOGRID-3.5.174/BIOGRID-ALL-3.5.174.mitab.zip
+
+

### 2019-08-17 Analysis/functions/GeneSample.R

@@ -1,0 +1,35 @@
+GeneSample = function(num_genes, actions, directed = TRUE) {
+  
+  require(igraph)
+  gene_list = sample(unique(actions$hgnc_symbol_a), num_genes)
+  
+  if (directed) {
+  act_sub = actions %>% 
+    filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% mutate(hgnc_acting = ifelse(a_is_acting == "t",hgnc_symbol_a, hgnc_symbol_b),
+                                                                                   hgnc_receiving = ifelse(a_is_acting == "t",
+                                                                                                          hgnc_symbol_b, hgnc_symbol_a)) %>% dplyr::select(hgnc_acting, hgnc_receiving) %>% filter(complete.cases(.))
+  gene_graph = graph_from_data_frame(act_sub, directed = directed)
+  sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+  internal_con = degree(sub_graph, mode = "in", loops = FALSE)
+  external_con = degree(gene_graph, mode = "out", v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+  }
+  
+  else {
+    link_sub = actions %>% 
+      filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% 
+      mutate(hgnc_symbol_a = ifelse(hgnc_symbol_a == "", NA, hgnc_symbol_a),
+             hgnc_symbol_b = ifelse(hgnc_symbol_b == "", NA, hgnc_symbol_b)) %>%
+      filter(complete.cases(.))  %>% dplyr::select(hgnc_symbol_a, hgnc_symbol_b) %>%
+      mutate(key = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), pmax(hgnc_symbol_a, hgnc_symbol_b), sep = "")) %>% 
+      group_by(key) %>% dplyr::slice(1) %>% ungroup() %>% dplyr::select(-key)
+    
+    gene_graph = graph_from_data_frame(link_sub, directed = directed)
+    sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+    internal_con = degree(sub_graph, loops = FALSE)
+    external_con = degree(gene_graph, v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+    
+  }
[…]

### 2019-08-17 Analysis/functions/alias2SymbolEbicat.R

@@ -1,0 +1,30 @@
+#data(ebicat37)
+#Function for converting ebicat37 mapped genes data to symbols designed specifically for the ebicatdata
+
+alias2SymbolEbicat = function(ebicat37) {
+  
+  ebi_dat = as.data.frame(mcols(ebicat37))
+  
+  #Fix spaces between - and ,
+  
+  ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
+  
+  ebi_dat$MAPPED_GENE = gsub(" -", "-", ebi_dat$MAPPED_GENE)
+  ebi_dat$MAPPED_GENE = gsub("- ", "-", ebi_dat$MAPPED_GENE)
+  
+  #Summarise the results
+  
+  #Keeping track of changes (sapply necessary to keep placeholders for unmapped genes)
+  
+  recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))
+  
+  recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))
+  
+  change = (ebi_dat$MAPPED_GENE == recomb_genes)
+  
+  sum_vec = ifelse(change == TRUE, "Same", ifelse(change == FALSE & recomb_genes == "", "No Match", "Converted"))
+  
+  ebi_dat$MAPPED_GENE = recomb_genes
+  
+  return(list(ebi_dat = ebi_dat, change = change))
+}

### 2019-08-17 Analysis/functions/diease2mod.R

@@ -1,0 +1,19 @@
+diseases2mod = function(disease_dis, links) {
+  require(igraph)
+  source('functions/genes2graph.R')
+
+  
+  #Get internal vs. external connectivity
+  
+  disease_nets = lapply(disease_dis, function(x) {
+    gene_list = read.table(x)$V1
+    genes2graph(gene_list, links, directed = FALSE)
+  })
+  
+  clusters = cluster_louvain(disease_nets[[1]])
+  
+  modularity(disease_nets[[1]], clusters$membership)
+  
+ 
+}
+

### 2019-08-17 Analysis/functions/disease2frame.R

@@ -1,0 +1,35 @@
+diseases2mod = function(disease_dis, links) {
+  
+  source('functions/genes2graph.R')
+
+  
+  #Get internal vs. external connectivity
+  
+  disease_nets = lapply(disease_dis, function(x) {
+    gene_list = read.table(x)$V1
+    genes2graph(gene_list, links, directed = FALSE)
+  })
+  
+  
+  disease_length = 1:length(disease_nets)
+  
+  #Get diseases
+  
+  disease_dis_name = gsub('.*/', '', disease_dis)
+  
+  disease_dis_name = gsub('.txt', '', disease_dis_name)
+  
+  
+  
+  disease_frame = lapply(disease_length, function(y) {
+    x = disease_nets[[y]]
+    sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
+    sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
+    colnames(sub_frame) = c("Genes", "Internal", "External")
+    sub_frame %>% mutate(Disease = disease_dis_name[y])
+  })
+  
+  disease_frame = bind_rows(disease_frame)
+  return(disease_frame)
+}
+

### 2019-08-17 Analysis/functions/disease2frame_raw.R

@@ -1,0 +1,29 @@
+diseases2frame_raw = function(disease_dis, links, disease_dis_name) {
+  require(igraph)
+  source('functions/genes2interactions.R')
+
+  print(disease_dis)
+  #Get internal vs. external connectivity
+  
+  disease_nets = lapply(disease_dis, function(x) {
+    gene_list = disease_dis
+    genes2interactions(gene_list, links, directed = FALSE)
+  })
+  
+  
+  disease_length = 1:length(disease_nets)
+  
+  #Get diseases
+  
+  disease_frame = lapply(disease_length, function(y) {
+    x = disease_nets[[y]]
+    sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
+    sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
+    colnames(sub_frame) = c("Genes", "Internal", "External")
+    sub_frame %>% mutate(Disease = disease_dis_name[y])
+  })
+
+  disease_frame = bind_rows(disease_frame) %>% filter(!is.na(Disease)) 
+  return(as.matrix(disease_frame))
+}
+

### 2019-08-17 Analysis/functions/gene2randgraph.R

@@ -1,0 +1,16 @@
+gene2randgraph = function(gene_list, actions, directed) {
+  act_sub = actions %>% 
+    filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% mutate(hgnc_acting = ifelse(a_is_acting == "t",hgnc_symbol_a, hgnc_symbol_b),
+                                                                                   hgnc_receiving = ifelse(a_is_acting == "t",
+                                                                                                           hgnc_symbol_b, hgnc_symbol_a)) %>% dplyr::select(hgnc_acting, hgnc_receiving) %>% filter(complete.cases(.))
+  gene_graph = graph_from_data_frame(act_sub, directed = directed)
+  rand_graph = sample_gnm(length(V(gene_graph)), length(E(gene_graph)), directed = directed)
+  
+  V(rand_graph)$name = V(gene_graph)$name
+  
+  sub_graph = induced_subgraph(rand_graph,vids = which(V(gene_graph)$name %in% gene_list))
+  internal_con = degree(sub_graph, mode = "in", loops = FALSE)
+  external_con = degree(rand_graph, mode = "out", v = V(rand_graph)[V(rand_graph)$name%in% gene_list], loops = FALSE)
+  return(list(internal_connectivity = internal_con,external_connectivity = external_con))
+  
+  }

### 2019-08-17 Analysis/functions/genes2graph.R

@@ -1,0 +1,33 @@
+genes2graph = function(gene_list, actions, directed = TRUE) {
+  
+  if (directed ) {
+  act_sub = actions %>% 
+    filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% mutate(hgnc_acting = ifelse(a_is_acting == "t",hgnc_symbol_a, hgnc_symbol_b),
+                                                                                   hgnc_receiving = ifelse(a_is_acting == "t",                                                                                                        hgnc_symbol_b, hgnc_symbol_a)) %>% dplyr::select(hgnc_acting, hgnc_receiving) %>% filter(complete.cases(.))
+  
+  gene_graph = graph_from_data_frame(act_sub, directed = directed)
+  sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+  internal_con = degree(sub_graph, mode = "in", loops = FALSE)
+  external_con = degree(gene_graph, mode = "out", v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+  
+  }
+  
+  else {
+    link_sub = actions %>% 
+      filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% 
+      mutate(hgnc_symbol_a = ifelse(hgnc_symbol_a == "", NA, hgnc_symbol_a),
+             hgnc_symbol_b = ifelse(hgnc_symbol_b == "", NA, hgnc_symbol_b)) %>%
+      filter(complete.cases(.))  %>% dplyr::select(hgnc_symbol_a, hgnc_symbol_b) %>%
+      mutate(key = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), pmax(hgnc_symbol_a, hgnc_symbol_b), sep = "")) %>% 
+      group_by(key) %>% dplyr::slice(1) %>% ungroup() %>% dplyr::select(-key)
+    
+    gene_graph = graph_from_data_frame(link_sub, directed = directed)
+    sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+    #internal_con = degree(sub_graph, loops = FALSE)
+    #external_con = degree(gene_graph, v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+    
+  }
+  
+  return(gene_graph)
+}
+

### 2019-08-17 Analysis/functions/genes2interactions.R

@@ -1,0 +1,33 @@
+genes2interactions = function(gene_list, actions, directed = TRUE) {
+  
+  if (directed ) {
+  act_sub = actions %>% 
+    filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% mutate(hgnc_acting = ifelse(a_is_acting == "t",hgnc_symbol_a, hgnc_symbol_b),
+                                                                                   hgnc_receiving = ifelse(a_is_acting == "t",                                                                                                        hgnc_symbol_b, hgnc_symbol_a)) %>% dplyr::select(hgnc_acting, hgnc_receiving) %>% filter(complete.cases(.))
+  
+  gene_graph = graph_from_data_frame(act_sub, directed = directed)
+  sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+  internal_con = degree(sub_graph, mode = "in", loops = FALSE)
+  external_con = degree(gene_graph, mode = "out", v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+  
+  }
+  
+  else {
+    link_sub = actions %>% 
+      filter(hgnc_symbol_a %in% gene_list | hgnc_symbol_b %in% gene_list) %>% 
+      mutate(hgnc_symbol_a = ifelse(hgnc_symbol_a == "", NA, hgnc_symbol_a),
+             hgnc_symbol_b = ifelse(hgnc_symbol_b == "", NA, hgnc_symbol_b)) %>%
+      filter(complete.cases(.))  %>% dplyr::select(hgnc_symbol_a, hgnc_symbol_b) %>%
+      mutate(key = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), pmax(hgnc_symbol_a, hgnc_symbol_b), sep = "")) %>% 
+      group_by(key) %>% dplyr::slice(1) %>% ungroup() %>% dplyr::select(-key)
+    
+    gene_graph = graph_from_data_frame(link_sub, directed = directed)
+    sub_graph = induced_subgraph(gene_graph,vids = which(V(gene_graph)$name %in% gene_list))
+    internal_con = degree(sub_graph, loops = FALSE)
+    external_con = degree(gene_graph, v = V(gene_graph)[V(gene_graph)$name%in% gene_list], loops = FALSE)
+    
+  }
+  
+  return(list(internal_connectivity = internal_con,external_connectivity = external_con))
+}
+

### 2019-08-17 Analysis/functions/perm_slopes.R

@@ -1,0 +1,67 @@
+slope_perm = function(Conn_Frame, Disease_Name, Norm_Reg = NULL, database, permutations = 5000, Disease = TRUE) {
+  
+  require(dplyr)
+  source("./functions/GeneSample.R")
+  
+  
+  if (Disease) {
+  Gene_Frame = Conn_Frame %>% filter(Disease == Disease_Name)
+  } else {
+  Gene_Frame = Conn_Frame %>% filter(Category == Disease_Name)  
+  }
+  
+  
+  if (!is.null(Norm_Reg)) {
+    norm_coef_reg = predict(Norm_Reg, newdata = data.frame(num_genes = nrow(Gene_Frame)))
+    
+    Gene_Frame = Gene_Frame %>% mutate(Internal = Internal/norm_coef_reg) 
+
+  }
+  
+  #Calculate slope for original frame
+  
+  reg = lm(External ~ Internal -1,  data = Gene_Frame)
+  slope = reg$coefficients[1]
+  
+  #Get slope from comparison
+
+  real_diff = slope
+  
+  #Perform Permutation test
+  
+  Comb_Frame = bind_rows(Gene_Frame, Rand_Frame)
+  
+  diff = c()
+  i = 1
+  
+ while(i<permutations) {
+    
+  rand_over = GeneSample(nrow(Gene_Frame), database, directed = FALSE)
+  
+  rand_over = bind_rows(rand_over$internal_connectivity, rand_over$external_connectivity)
+  rand_over = cbind.data.frame(colnames(rand_over), t(rand_over[1,]), t(rand_over[2,]))
+  colnames(rand_over) = c("Genes", "Internal", "External")
+  
+  rand_over = rand_over %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External))
+  
+  #Randomly sample
+  
+    
+  new_reg = lm(External ~ Internal-1, data = rand_over)
+  new_slope = new_reg$coefficients[1]
+  
+  if (is.na(new_slope)) {
+    next
+  } else {
+    diff[i] = new_slope
+    i = i+1
+    
+  }
+  
+  }
+  
+  p_val = (sum(diff<real_diff) + 1)/(permutations+1)
+  return(p_val)
+}
+
+

### 2019-08-18 manuscript.Rmd
Move manuscript to root

---
bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib]
csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
output:
  pdf_document: default
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/styles.doc/Arial_11_single_space_normal_margins.docx
  html_document: default
editor_options: 
  chunk_output_type: console
---

# Coherence of Molecular Mechanisms in Genes Identified Through GWAS In Major Human Disease and Traits

Dozmorov, Mikhail G.^1+^ (mikhail.dozmorov@vcuhealth.org)  
Cresswell, Kellen G.^1+^ (cresswellkg@vcu.edu)  
Reimers, Mark^2^ (reimersm@msu.edu)  
Bacanu, Silviu-Alin^3^ (silviu-alin.bacanu@vcuhealth.org)  
Craver, Carl^4^ (ccraver@wustl.edu)  
Kendler, Kenneth S.^3^ (kenneth.kendler@vcuhealth.org)

^1^ - Department of Biostatistics, Virginia Commonwealth University   
^2^ - Neuroscience Program and Dept. Biomedical Engineering Michigan State  
^3^ - Virginia Institute for Psychiatric and Behavior Genetics and the Department of Psychiatry, Virginia Commonwealth University  
^4^ - Philosophy-Neuroscience-Psychology Program, Washington University in St. Louis  
^+^ - equal contribution

# Abstract

**Background.** Complex phenotypes such as height and intelligence, are thought to be a product of the collective effects of multiple phenotype-associated genes, or, more precisely, interactions among their protein products. High/low degree of interactions is suggestive of coherent/random molecular mechanisms, respectively. Comparing the degree of interactions may help to better predict and understand the coherence of the molecular mechanisms underlying complex phenotypes. However, direct comparison of the degree of interactions is difficult due to differences in size and configuration of phenotype-associated gene networks.

**Methods.** We introduce a measure of network coherence as a ratio of internal vs. external distributions of the degree of interactions. The internal degree distribution is defined by interaction counts within a phenotype-specific gene network, while the external degree distribution counts interactions with other genes in the whole protein-protein interaction (PPI) network. We present a novel method for normalizing the internal degree distributions, making them directly comparable.

**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (with exemplary high coherence) and random networks (the lower limit of coherence). We observed a range of coherence estimates for each category of phenotypes. Metabolic traits and diseases were the most coherent, while psychiatric diseases and intelligence-related traits were the least coherent.

**Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences. Our results highlight the gaps in of our current knowledge of the genetics and molecular mechanisms of complext traits and suggest priorities for future work.

**Keywords:** GWAS, network, degree, coherence

# Introduction

Genome-wide association studies (GWAS) have significantly advanced our understanding of complex phenotypes by identifying disease- and trait-associated genetic markers and suggesting corresponding genes [@Altshuler:2008aa; @Risch:1996aa]. However, GWAS findings explain only a fraction of heritability [@Eichler:2010aa; @Maher:2008aa; @Manolio:2009aa]. Furthermore, the corresponding genes are often spread across different chromosomes with no known connection to one another, hindering understanding of the molecular mechanisms. These limitations of current generation GWAS might stem from the yet incomplete knowledge of the genetic determinants of complex phenotypes, the potential heterogeneity of complex phenotypes and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].

Many studies have shown that genetically-driven complex phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa]. This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to human phenotype, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa]. Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all molecular determinants of a certain phenotype. Consequently, methods to identify phenotype-associated networks of functionally-related genes have been developed [@Krauthammer:2004aa; @Cordell:2009aa; @Akula:2011aa].

Network properties, such as connectivity (aka degree distribution) [@Ghiassian:2015aa; @Lee:2008aa; @Barabasi:2004aa; @Feldman:2008aa], can be compared to better understand the relationships among the phenotypes [@Vidal:2011aa; @Barabasi:2011aa; @Wang:2011aa]. Networks formed by phenotype-associated genes, or, more precisely, their protein-protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa]. This intuition reflects a well-known "guilt-by-association" principle that genes (or, more precisely, their products) that form an interaction network are more likely to share similar functions [@Farber:2013aa], or co-expression patterns [@Huang:2006aa; @Michalak:2008aa]. Although this view has been criticized [@Gillis:2012aa], several studies consistently observed phenotype-associated genes to be either connected as a single network or to participate in common phenotype-specific subnetworks [@Menche:2015aa; @Lee:2011aa; @Akula:2011aa; @Ideker:2008aa; @Iossifov:2008aa]. Therefore, high connectivity among phenotype-specific genes may indicate coherent molecular mechanisms that could be targeted therapeutically. In terms of graph theory, we are asking whether a phenotype-specific network is a community - a highly connected subgraph relatively well-separated from the rest of the network [@fortunato2010community]. However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes forming networks of different sizes and configurations.

This study compares the connectivity of networks formed by disease- or trait-associated genes and gene products. We hypothesized that genes associated with (categories of) diseases share common molecular mechanisms and, hence, form connected (i.e., coherent) networks. Therefore, the level of connectivity may inform us about similarities and differences among the diseases. The terms “network connectivity,” “coherence,” or “average degree distribution” are used interchangeably as appropriate. We developed a method for normalizing the internal network connectivity that accounts for differences in network sizes formed by disease-associated genes. We compared the ratio of normalized internal vs. external network connectivity between psychiatric, neurologic, autoimmune, cancer, cardiovascular, metabolic and other diseases and traits. We used KEGG pathways [30] as a reference for a coherent network with common molecular mechanisms, and we used sets of randomly sampled genes as a reference for the absense of coherence. We found that genes associated with metabolic diseases and traits and autoimmune diseases formed coherent networks similar to KEGG pathways. On the contrary, the coherence of networks of psychiatric and neurologic disease-associated genes, as well as intelligence-associated traits, was similar to that of randomly selected genes. Cancer, anthropometric traits, eye-, and cardiovascular diseases and traits formed networks of intermediate connectivity. Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.

This study presents a novel measure of network coherence as a ratio of the normalized internal vs. external degree distributions. The internal degree distribution is defined as gene-specific interaction counts within a protein-protein interaction (PPI) network formed by phenotype-associated genes, while the external degree distribution counts interactions with other genes in the whole PPI network. We used KEGG pathways [@Kanehisa:2000aa] as a reference for a paradigmatically high-coherent network [@Lee:2003aa], while sets of randomly sampled genes were used as the reference for the absence of coherence. We hypothesized that the level of coherence may inform us about similarities and differences among phenotypes. Using two PPI databases (STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa]), we compared coherence of 133 phenotypes from GWAScatalog [@Hindorff:2009aa; @MacArthur:2017aa]. Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have the most coherence. Our method enables direct comparison of coherence measures and highlight gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.

# Results

## Phenotype-specific networks of protein-protein interactions (PPIs)

We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods). These were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).

We created networks of phenotype-associated genes using PPI information from STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa] databases (see Methods). We opted for the use of three types of PPI data. First, STRING data filtered by the interaction confidence score above 500 (middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data. This filtering step ensured that only high-confidence PPIs were selected. The advantage of using filtered data is the reliance on high-quality PPIs. The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs are removed from the analysis (see Methods). Second, unfiltered STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING full”). The advantage of using unfiltered data is that more trait-associated genes will have PPI information and, hence, can be analyzed. The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs. Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, were used. The use of different data sources (STRING and BioGrid) and filtering (full or filtered STRING data) was intended to increase the generalizability of our conclusions.

**Table 1. Summary statistics of the analyzed categories.** Average coherence estimates are shown. "NA" values indicate that a phenotype was not considered due to the low number of genes with PPIs.

| Category               | Total Phenotypes | Number of SNPs | Number of Genes | Biogrid Coherence | String Coherence | String Filtered Coherence |
|------------------------|:----------------:|:--------------:|:---------------:|:-----------------:|:----------------:|:-------------------------:|
| Disease                |                  |                |                 |                   |                  |                           |
| Metabolic disease      |         5        |       345      |       280       |        0.86       |       0.79       |           0.933           |
| Neurologic             |         8        |       382      |       343       |        0.89       |       0.511      |           0.861           |
| Eye disease            |         8        |       270      |       242       |        0.75       |       0.536      |           0.856           |
| Autoimmune             |        20        |      1415      |       1212      |       0.807       |       0.801      |           0.852           |
| Cancer                 |        18        |       713      |       610       |       0.863       |       0.643      |           0.838           |
| Cardiovascular disease |         9        |       336      |       289       |       0.705       |       0.584      |           0.832           |
| Psychiatric            |        11        |       678      |       611       |        0.83       |       0.537      |           0.817           |
| Trait                  |                  |                |                 |                   |                  |                           |
| Metabolic trait        |        10        |       790      |       646       |       0.906       |       0.776      |            0.93           |
| Eye trait              |         5        |       158      |       109       |         NA        |       0.674      |           0.903           |
| Anthropometric trait   |        19        |      2143      |       1822      |       0.793       |       0.601      |           0.841           |
| Cardiovascular trait   |        15        |       728      |       604       |       0.833       |       0.647      |           0.813           |
| Intelligence           |         5        |       236      |       198       |        0.78       |       0.312      |           0.757           |

## Degree distribution as a measure of coherence

A typical network consists of nodes connected by edges [@Barabasi:2004aa]. In genomics, nodes typically represent genes and edges correspond to some measure of interactions, e.g., gene co-expression or interactions between protein products [@Goncalves:2012aa; @Wang:2012aa]. In terms of phenotypes, an intuitive expectation is that they will be represented by coherent modules of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa]. In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the network [@fortunato2010community; @Radicchi:2004aa].

Numerous network properties have been defined to describe network structures [@Barabasi:2011aa; @Wang:2012aa]. A degree of a node, or connectivity, is one of the most fundamental characteristics defining the number of other nodes connected to a given node. A collection of degrees of all network nodes form a degree distribution. Comparing degree distributions among networks is an intuitive way to gain an understanding of network similarities and differences [@Barabasi:2004aa; @Feldman:2008aa]. We hypothesized that comparing degree distributions of networks formed by phenotype-associated gene sets may inform us about similarities and differences among phenotypes in terms of coherence of their antecedent and underlying molecular mechanisms.

## Internal vs. external degree distributions as a measure of the coherence

<!-- inspired by previous [@Menche:2015aa] and our (Supplementary Figure S3) observation that network size and configuration are irrelevant when considering interactions within the whole PPI network. -->
Given that phenotype-associated networks occur within the global network of PPIs, we developed a ratio-based metric of coherence. This metric is inspired by previous work showing that communities within a large network have high internal but low external levels of connectivity [@PhysRevE.81.046110; @Spirin:2003aa]. Thus, for each phenotype-associated PPI network, we considered the relationship between its internal and external degree distributions. The internal degree distribution is defined by considering edges within an isolated network formed by phenotype-associated genes. The external degree distribution is defined by considering edges to other genes in the whole PPI network. Intuitively, a highly coherent network is expected to have a large number of internal edges but a relatively small number of external edges. An extreme example of high coherence would be a fully internally connected network with zero interactions with external genes. Conversely, a low-coherence network is expected to have a relatively low number of internal edges compared to external edges, mirroring a random network.

To compare coherences, we measured coherence as the slope of a line through a scatterplot of internal vs. external degree distributions. That is, we visualize internal (X-axis) vs. external (Y-axis) degrees for each phenotype-specific gene on a single plot and fit a regression line through it. An example of a network with the highest coherence (full internal connectivity, zero external connections) would be represented by a horizontal line. Conversely, an example of a network with the lowest coherence would be a set of non-interacting genes (zero internal connections and full external connectivity) would be represented by a vertical line. Consequently, slopes of internal vs. external degree distributions of phenotype-associated networks would represent the phenotype-specific levels of coherence. In summary, comparing internal vs. external degree distributions represents a viable metric to quantify and compare the molecular coherence of phenotype-associated networks.

## Networks of KEGG pathways and randomly selected genes as reference for high and low coherence, respectively

The coherence of phenotype-associated gene sets should be measured with respect to the reference level of networks with paradigmatically high and low coherence. To establish a reference of high coherence, we considered networks from collections of the Molecular Signatures Database (MSigDB v6.2) [@Subramanian:2005aa]. MSigDB contains sets of genes having various types of functional relationships. We found that pathways from the Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa:2000aa] had the smallest average slope corresponding to the highest level of coherence (Supplementary Table S2). Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks of high coherence. This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathwaays (REACTOME genes) are expected to interact more frequently. In our study, KEGG networks were used as the primary reference of high coherence, in parallel with GOCC and REACTOME collections. Conversely, as a reference for low coherence, we randomly sampled genes from a pool of genes in a given PPI database. These random gene lists represent expected network coherence arising by chance. The slopes of KEGG and random networks represent reference levels for high and low coherence, respectively.

## Network size and configuration hinders direct comparison of degree distributions

Direct comparison of degree distributions of the networks formed by phenotype-associated genes is hindered by the fact they depend on network size and configuration. An intuitive example is to consider a ring and a fully connected network of size 10 and 20 nodes, respectively (Figure 1). The degree distributions of the ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).  Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have degree 9, while all nodes in the 20-node network have degree 19, Figure 1B). <!--Normalizing the degree distributions of the fully connected networks by dividing them by network size makes them identical as would be expected for networks with similar coherence. However, similar normalization of ring networks will make their degree distributions different (0.22 for the 10-node network and 0.11 for the 20-node network), leading to the illusory impression of differences in coherence between intuitively equally coherent networks. -->These observations highlight the difficulty in comparing degree distributions of phenotype-associated gene networks due to two facts: 1) network sizes differ, and 2) network configuration (fully connected, ring, or intermediate connectivity) is unknown.

+-------+
| ![](Figures/Fig1_network_distributions.png)      |
| **Figure 1. Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes. While degree distributions of the similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.   |
+-------+

To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks. We observed negative correlation of network size with the slopes of KEGG and random networks (-0.37/-0.85, respectively, Figure 2, Supplementary Table S2). These observations confirm the notion that coherence estimation using internal vs. external degree distributions should be controlled for network size.

+-------+
| ![](Figures/Size_Slope_Plot.png)      |
| **Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.   |
+-------+

## Network size-independent estimation of coherence

Given the dependence of internal and external degree distributions on network size, we designed a simple strategy to account for it. Briefly, for each phenotype-specific network of a given size, we create size-matched references of high and low coherence. Specifically, we select the ten KEGG pathways closest in size and ten random networks matched in size as references for high/low coherence for the corresponding phenotype. Consequently, we use the slopes of internal vs. external degree distributions created by these size-matched references to normalize the slope of the phenotype-specific network to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively. We found that the use of size-matched references indeed removes the dependency between coherence and network size (Supplementary Figure S1).

## Results derived from different PPI data are largely consistent

The potential drawback of this reference size-matching approach is that phenotypes with large network sizes (e.g., "Obesity-related traits", 779 genes) may not have the size-matched reference of high coherence (e.g., the largest KEGG pathways is 374 genes). Consequently, coherence estimation for such networks may be affected by the smaller size of reference of high coherence. Therefore, we use networks from GOCC and REACTOME collections that have larger maximum number of genes (Supplementary Table S2) in parallel with KEGG networks as a reference of high coherence. We found that coherence estimations using KEGG, GOCC, and REACTOME networks as references of high coherence are highly concordant (Average Pearson correlation coefficient > 0.59, Supplementary Figure S2). The use of size-matched networks from KEGG, GOCC, and REACTOME collections enables robust estimation of coherence for phenotype-specific networks of different sizes. Given the similarity of the results obtained with different PPI databases and references of high coherence, we present results obtained with STRING filtered PPI database and KEGG networks as a reference of high coherence, with other observations presented in Supplementary Table S3 and Supplementary Figures S3-S8.

## Coherence of all phenotypes significantly differs from random

To estimate the significance of coherence measures, we developed a permutation-based approach (see Methods). Notably, we observed coherence for all diseases to be highly significant (permutation p-value < 0.002, Supplementary Table S3). The distribution of coherence estimates was skewed towards 1.00 (highest coherence), with the median coherence being 0.86. "Obesity-related traits" and "Height" were on the lowest range of coherence estimates (0.42 and 0.54, respectively). "Weight", "Telomere length", and "Corneal structure" were on the highest range of coherence (1.00, Supplementary Table S3). These results indicate that coherence of phenotype-specific networks is generally high and significantly different from random.

## Metabolic and intelligence-related traits as examples of high/low coherence

We estimated coherence of 53 traits from five categories (Figure 3, Supplementary Figures S3-S5). Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence. Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose- and cholesterol-related traits (metabolic). Notably, metabolic and eye trait categories had the highest overall coherence (mean coherence 0.90 and 0.93, respectively). Several notable examples of low coherence include "Obesity-related traits", "Height", and "Longevity" (anthropometric traits), blood count-related traits (cardiovascular). Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.76). These results highlight the relatively good knowledge of the genetics of metabolic traits which corresponds to highly coherent networks. Conversely, current knowledge about the genetics of intelligence-related traits appear insufficient as reflected in the low coherence of the corresponding networks.

+-------+
| ![](Figures/Fig3_summary_traits.png)      |
| **Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).   |
+-------+

## Metabolic and neurologic diseases have high coherence

We further estimated coherence of 80 diseases from seven categories (Figure 4, Supplementary Figures S6-S8, Supplementary Table S3). Metabolic diseases had high overall coherence (average coherence 0.93), with "Obesity" and "Type 2 diabetes" being the most coherent (0.99 and 0.94, respectively). Similarly, the average coherence for neurologic diseases was 0.86, with "Alzheimer's disease" and "Parkinson's disease" being highly coherent (0.94 and 0.92, respectively). The high coherence of metabolic and neurologic disease networks highlight the relatively good knowledge of their genetics.

## Intermediate coherence of autoimmune, cancer, and eye diseases

"Rheumatoid arthritis", "Ankylosing spongylitis", "Psoriasis" were the most coherent in the "Autoimmune" category (0.94, 0.91, 0.89, respectively). On the other end of the coherence spectrum, "Multiple sclerosis" and "Allergic rinitis" were the least coherent (0.76 for both). The average coherence of autoimmune diseases was high, 0.85, with all but two autoimmune diseases having high coherence (> 0.80). Diseases in the "Cancer" category were similarly highly coherent (average coherence 0.84). "Non-small lung cancer", "Testicular germ cell tumor", "Epithelial ovarian cancer" were the most coherent (0.94, 0.91, 0.88, respectively) while "Basal cell carcinoma", "Breast cancer" "Hodgkin's lymphoma" were the least coherent (0.64, 0.78, 0.78, respectively). Overall, networks of all but "Basal cell carcinoma" cancers had high coherence (> 0.78). Diseases in the "Eye disease" category were more heterogeneous, with coherence ranging from 0.71 for "Myopia (pathological)" disease to 0.97 for "Refractive error". The average coherence of diseases in the "Eye disease" category was relatively high (0.86). These results suggest good overall knowledge of the genetics of autoimmune, cancer, and eye-related diseases and highlight limited knowledge for some diseases in these categories.

+-------+
| ![](Figures/Fig4_summary_diseases.png)      |
| **Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).   |
+-------+

## Psychiatric and cardiovascular diseases as examples of low coherence

Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.82 and 0.83, respectively). The corresponding coherence estimates were highly heterogeneous. Among cardiovascular diseases, "Atrial fibrillation", "Coronary artery disease" showed the highest level of coherence (0.98 and 0.89, respectively), while "Sudden cardiac arrest" and "Hypertension" had the lowest coherence (0.70 and 0.77, respectively). Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.93 and 0.89, respectively). Notably, "Schizophrenia" and "Major depressive disorder" showed the lowest coherence not only in the "Psychiatric disease" category (0.67 and 0.75, respectively) but among all phenotypes. Other psychiatric disorders, such as "Bipolar disorder", "Autism", "ADHD", showed intermediate coherence estimates (0.87, 0.83, 0.78, respectively). In summary, these low coherence estimates highlight the need for further research on genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms, and suggests "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies.

# Discussion

We developed a general-purpose method to measure and compare coherence estimates of the molecular mechanisms in networks of genes. Intuitively, gene networks with high internal and low external interactions would be considered highly coherent; consequently, coherence is defined as the ratio between the internal and external degree distributions (the number of interactions per gene). We normalized coherence estimates to the range of high and random coherence, exemplified by KEGG pathways and random networks. To decouple coherence from its dependence on network size, we used KEGG and random networks matched in size to phenotype-specific networks. Applied to the analysis of 133 disease- and trait-associated gene sets, we found an unexpected inverse association of phenotype complexity with coherence. Consequently, complex traits like "Height", "Intelligence", "Longevity" had overall low coherence while metabolic traits and diseases had high coherence. Diseases in "Psychiatric" category had overall low coherence, with "Schizophrenia" being least coherent. Our method allows one to quantify the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.

Our approach is similar in spirit to the well-researched community detection problem [@danon2005comparing; @fortunato2010community; @Radicchi:2004aa]. However, we approach this problem from a different angle by asking whether a given network as a whole can be considered a community within the global PPI network. This may be considered a drawback as it has been shown that many networks consist of miniature communities known as motifs [@Milo:2002aa], or graphlets [@Przulj:2007aa]. While the investigation of motif enrichment [@Milo:2002aa] is a viable approach to characterize a network, our goal was to develop a unified measure of coherence that can be compared across phenotypes.

Our method relies on the ability to link the phenotype-associated genetic variants detected in genome-wide association studies to genes thought to be affected by them. Typically, genes are linked to genetic variants by proximity (nearest gene), a strategy adopted by the NHGRI-EBI GWAS catalog [@MacArthur:2017aa]. <!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, the majority of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Khurana:2013aa]. <!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].-->The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy. The development of chromatin conformation capture technologies now allows for understanding interactions among genomic regions on a genome-wide scale [@Lieberman-Aiden:2009aa; @Zhang:2013aa], helping to prioritize disease-associated gene-variant associations [@Pan:2018aa]. Consequently, tools are being developed to consider long-distance interactions when mapping SNPs to genes (3DSNP [@Lu:2017aa], FUMA [@Watanabe:2017aa], PINES [@Bodea2016], HUGIn [@Martin:2017aa]), and studies redefining disease-associated genes using long-distance interactions started to emerge [@Fang:2019aa]. Our future goal includes considering long-distance variant-gene interactions in defining phenotype-associated gene sets and comparing their network properties.

The definition of coherence as the ratio between internal vs. external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa]. Our limited knowledge of protein-protein interactions limits the coherence estimation to phenotypes that have a sufficient number of genes with PPI annotations. Furthermore, networks formed by phenotype-associated genes should be sufficiently diverse to avoid cases in which all genes in a network have the same degree. As the different PPI databases vary significantly in the number of PPI annotations, some phenotype-associated gene sets have sufficient PPI annotations when using one but not the other database. This problem is best illustrated by the BioGrid PPI database that has considerably fewer PPIs; hence, fewer phenotypes could be analyzed. Furthermore, using smaller PPI databases like BioGrid inevitably diminishes the size of phenotype-associated networks, limiting the internal degree distributions that can be derived from them. The differences in database-specific networks are alleviated by testing the significance of coherence using the permutation test; however, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].

<!--The definition of high and random coherence also depends on the choice of PPI data. Therefore, we derived degree distributions for KEGG and random networks separately for each PPI database and implemented the max-min normalization of the coherence estimates. Intuitively, the max-min normalization converts the absolute measures of coherence (slopes of regression lines fit through the internal vs. external degree distribution plots) that are dependent on the choice of PPI databases to the percentages of the full range of coherence (from high coherence defined by KEGG networks to low coherence defined by random networks) defined using a given PPI database. The phenotype-specific coherences expressed as a percentage of the full range of coherence can be directly compared across phenotypes and databases.

Although the normalized coherence derived using different PPI databases can be compared among phenotypes, it tells nothing whether a given coherence estimate is significantly different from random. We developed a permutation test aimed to provide a measure of the stability of the coherence. Intuitively, the permutation test defines how different the coherence of a given phenotype from that of random networks of the same size. It provides additional protection against cases where coherence may appear high, yet, it may be a consequence of small network size providing unstable degree distributions. Testing the coherence estimates for significance helps against conclusions of the overall low coherence (e.g., anthropometric traits, Figure 3), when the coherence of individual traits is high (e.g., individual anthropometric traits, Figure 5B). Coupled with the permutation test, our coherence estimation provides a robust framework for quantifying the coherence estimates of phenotype-associated genes, advancing our understanding of the corresponding molecular mechanisms.-->

In addition to internal vs. external degree distributions, other network properties can be used to estimate coherence. For example, the centrality measure has been used to demonstrate that highly central hub genes are functionally essential and highly conserved [@Khurana:2013aa]. A large study of PPIs (BioPlex) compared centrality distributions among several disease categories. Cancers and immunological disease networks tended to have high centrality, while nervous system, congenital, neonatal and hereditary disease networks had low centrality [@Huttlin:2017aa]. Metabolic pathways were shown to have more duplicated copies of highly central genes making the networks tolerant to loss-of-function mutations [@Khurana:2013aa]. In summary, centrality measure parallels our observations of network coherence. Our future work will incorporate centrality and other network metrics into our framework to further refine our ability to compare coherence of phenotype-associated molecular mechanisms.

[??? Potentially. discuss: We do not account for configuration]

## Conclusion

Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes. We developed a network coherence measure that estimates the degree of internal vs. external connectivity that is robust to the difference in network sizes. Using well-curated lists of genomic variants and the associated genes, we built disease-specific networks using protein-protein interaction information from the STRING and BioGrid databases. We found a range of coherence estimates in each phenotype category, with metabolic diseases and traits being the most coherent. Psychiatric disorders were, on the other hand, the least coherent, with schizophrenia and major depressive disorder being the least coherent among all phenotypes. In summary, we provide a general-purpose method for quantifying and comparing coherence of molecular networks.

# Methods

## Data sources

Trait-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa]. "MAPPED_GENE" column was used to extract trait-associated genes. To organize traits into categories, Experimental Factor Ontology (EFO) IDs were extracted from the "MAPPED_TRAIT_URI" column. Categories were mapped to EFO IDs using the `ontoCAT` v.1.12.0 R package and the EFO database (accessed 12/04/2018). In the case of a trait being mapped into two categories, the most representative category was manually assigned. Non-canonical gene names were converted to common gene symbols using `alias2Symbol` function from the `limma` R package. Gene names which could not be mapped to gene symbols were excluded from the analysis.

Human protein-protein interactions (PPIs) were obtained from STRING database v.11.0 [@Szklarczyk:2015aa] and BioGrid database v.3.5.174 [@Chatr-Aryamontri:2017aa] (accessed 06/25/2019). The STRING data for the tax ID "9606" (human) was used, and Ensembl protein IDs were converted to gene names using BioMart. Either the full dataset (11,761,040  entries) or data filtered by "combined_score" > 500 (range 0-1000) (1,373,946 entries, the main PPI dataset) was used. The BioGrid data was filtered to include "TaxID interactor" type equal to "9606", all "Interaction Types" were included, totaling 465,660 interactions. These three PPI databases were used in parallel to evaluate disease coherence.

Trait-associated gene networks were formed by mapping trait-associated genes to PPIs. Genes without PPI annotations (e.g., C5orf4, LOC100507462), antisense transcripts (e.g., CDKN2B-AS), readthrough (e.g., NPHP3-ACAD11), non-protein-coding transcripts (e.g., LINC00478) were excluded. Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept. Traits with less than five genes were omitted. Additionally, traits with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs. external degree distributions (see below). These precautions were set to avoid situations where a small number of genes in a network can cause instability in the results. Importantly, the number of trait-associated genes having PPI annotations depends on the content of a PPI database; hence, some traits may be filtered when using one but not the other PPI database to build trait-associated networks.

## Comparing internal vs. external degree (aka connectivity) distributions as a measure of coherence

Trait-specific **internal degree distribution** was defined as the number of interactions between trait-associated genes, or, more precisely, their interacting protein products. As different traits have different numbers of associated genes forming networks of unknown configurations, the corresponding internal degree distribution was normalized to the expected interaction count for a network of that size (see Results).

Trait-specific **external degree distribution** was defined as the number of interactions between genes in a trait-associated network and all other genes in the PPI network. Being measured in the context of the entire PPI network of fixed size, the external degree distribution did not require normalization.

To compare trait-specific degree distributions, the normalized internal vs. external degree distributions were plotted on a scatterplot. The square root of the degree was taken for more informative representation. For each trait, the internal vs. external degree distribution plot was fit with a regression line through the origin to capture the slopes. Slopes were used as a measure of coherence with lower slopes being associated with higher coherence. Degree distributions of the network nodes were obtained using the `igraph` R package v. 1.0.1.

## Max-min normalization of coherence

The slopes of the internal vs. external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database. Thus, we normalized the slopes for the trait-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks). For each trait $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1]. Higher values indicate a higher level of coherence similar to that of KEGG pathways. These values represent normalized coherence and can be compared across traits and PPI databases.

## Permutation test of slopes

To test whether each slope is significantly different than random, we used a permutation test. For trait $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution. We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for trait $i$. For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs. external degree distribution regression line of random coherence. We calculate the permutation p-value as $p_i = \frac{(\sum_{j=1}^nI({\beta_{rj} > \beta_i}) + 1}{n + 1}$ [@Phipson:2010aa]. Intuitively, this approach estimates whether the slope of the trait-specific network $\beta_i$ is consistently larger than those of random networks. All analyses and visualizations were performed in the R/Bioconductor computing environment v.3.4.0 [@gentleman2006bioinformatics].

<!--For each disease, we find the absolute value of the difference between its slope and the random coherence slope. Next, we combine the genes for each disease with those of the random diseases and take two random samples of equal size to the disease gene set. We then calculate the absolute difference in slopes between the randomly sampled gene sets. This process is repeated 1000 times. Finally, we obtain a permutation p-value by taking the number of times the difference in slopes between random samples were less than the overall difference in slopes and dividing by 1000. -->

# Abbreviations

- GWAS - Genome-Wide Association Study
- KEGG - Kyoto Encyclopedia of Genes and Genomes
- PPI - Protein-Protein Interaction
- SNP - Single Nucleotide Polymorphism

# Acknowledgements

This work was supported by a Genetics and Human Agency Award from the Templeton Foundation to KSK, MR, CC, SB. The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation. The authors would like to thank Spiro Stilianoudakis for help with Figure 1.

## Competing interests

The authors declare that they have no competing interests.

# Supplementary material

**Supplementary Table S1. Summary statistics of the analyzed phenotypes.** Each worksheet corresponds to the PPI database used.

**Supplementary Table S2. Summary of networks from MSigDB categories.** Minimum, maximum, mean and median of the slopes and number of genes are shown for each collection, along with the total number of networks and the correlation between network size and slope.

**Supplementary Table S3. Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) and the reference of high coherence (KEGG, GOCC, Reactome) are specified in the corresponding columns for coherence and permutation p-value estimations.

**Supplementary Figure S1. Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

**Supplementary Figure S2. Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.

**Supplementary Figure S3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis). Missing entries indicate that for a given PPI database a phenotype-associated network could not be built.

# References

<!--
## Single networks, or networks of modules?

Our study considers a network of disease-associated genes in its entirety. However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa]. Connectivity of such modules would be different from the overall network connectivity [@Ghiassian:2015aa]. Although numerous methods exist for module detection in a network [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section Intuitively, large networks will have a higher chance to have subnetworks (modules), and vice versa. It remains to be determined whether disease-associated genes form modular networks and whether the modular structure differs between (categories of) diseases.

## Tissue specificity

The effect of disease-associated SNPs should be considered in cell-type-specific context [@Trynka:2013aa; @Dozmorov:2016aa; @Sanyal:2012aa]. The difference between neurological disease- and schizophrenia-associated SNPs was determined on the cell-type-specific level, with schizophrenia- but not neurological disease-associated SNPs being enriched in genes expressed in medium spiny neurons, pyramidal CA1, pyramidal SS and cortical interneurons [@Skene:2018aa]. While cell-type-specific gene expression profiles have been generated, protein-protein interactions remain generic and highly incomplete [@Menche:2015aa]. Even the latest efforts to compile PPIs from multiple data sources (InWeb InBioMap [@Li:2017aa]) or single-technology large-scale PPI profiling (BioPlex 2.0 [@Huttlin:2017aa]) generate collections of  non-cell-type-specific interactions. Future efforts in cell-type-specific PPI profiling are expected to boost research on the comparison of disease networks built from PPIs.

## What constitutes random networks?

The degree was found to follow a power-law [@Yook:2005aa] or a truncated power-law distribution [@Huang:2007aa].

## GWAS p-value, cutoff, sample size, harmonization issues

In GWAS, it is often assumed that local correlations between adjacent hypotheses along the genome are so strong that testing the whole genome is equivalent to performing only $10^6$ tests, regardless of the number of hypotheses tested. Under this assumption, FWER can be controlled using a fixed genome-wide significance level of $\alpha = 5*10^{-8}$, corresponding to a Bonferroni adjustment for $10^6$ tests. 
-->

### 2019-08-18 manuscript.Rmd
Modify for relative paths

@@
 ---
-bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib]
-csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
+bibliography: [refs/networks.bib,refs/brain.bib,refs/enrichment.bib,refs/3D_refs.bib]
+csl: styles.ref/genomebiology.csl
 output:
@@
   word_document:
-    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/styles.doc/Arial_11_single_space_normal_margins.docx
+    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
   html_document: default

### 2019-08-19 Analysis/GWAS_Prep.Rmd
make fully contained

@@
 
-efo_loc = "https://raw.githubusercontent.com/EBISPOT/efo/master/efo.owl"
-
-efo = getOntology(efo_loc)
-# save(list = "efo",file =  "./data/efo.rdata")
+# efo_loc = "https://raw.githubusercontent.com/EBISPOT/efo/master/efo.owl"
 # 
-# load("./data/efo.rdata")
-
-#Cleaning up efo data to remove urls and putting comma seperated values on new lines
-
-main_inf = unembed(main_inf, "MAPPED_TRAIT_URI", sep = ",")
-main_inf$MAPPED_TRAIT_URI = basename(main_inf$MAPPED_TRAIT_URI)
-
-tree = c()
-for (i in unique(main_inf$MAPPED_TRAIT_URI)) {
-
-  #Get entire EFO tree
-
-  parents = getAllTermParentsById(efo, i)
-
-  #Pull out labels and make into vector
-
-  parents = sapply(parents, getLabel)
-
-  #Cbind and fill in non-matching entries with NA
-
-  tree = rowr::cbind.fill(tree,parents, fill = NA)
-
-}
-
-#Remove original column
-
-tree = tree[,-1]
-
-colnames(tree) = unique(main_inf$MAPPED_TRAIT_URI)
-
-#Pick out important disorder groups
-
-imp_cats = c("Abnormality of refraction", "Abnormality of the eye", "age", "anthropometric measurement", "auditory system disease", "autoimmune disease", "body weights and measures", "bone measurement", "cancer", "carcinoma", "cardiovascular disease", "cardiovascular disease biomarker measurement", "cardiovascular measurement", "diabetes mellitus", "diabetes mellitus biomarker", "digestive system disease", "eye measurement", "glucose measurement", "head and neck disorder", "heart function measurement", "hematological measurement", "immune system disease", "infectious disease biomarker", "insulin measurement", "lipid measurement", "lipid or lipoprotein measurement", "lipoprotein measurement", "mental or behavioural disorder", "metabolic disease", "metabolite measurement", "neoplasm", "nervous system disease", "neurodegenerative disease", "pulmonary function measurement", "reproductive system disease", "respiratory disease biomarker", "skin disease", "sleep measurement", "temporal measurement")
-
-#Output cats by pipe
-
-#Get categories for each unique EFO ID
-
-cat_ass = apply(tree, 2, function(x) c(na.omit(match(x, imp_cats))))
-
-#Replace categoryless values with NA
-
-cat_ass = lapply(cat_ass, function(x) {
-  if (length(x)<1) {
-    cat = NA 
-  } else {
-    cat = x
-  }
-  return(cat)
-})
-
-#Assign category names
-
-mapped_groups = stack(cat_ass)
-mapped_groups$values = imp_cats[mapped_groups$values]
-colnames(mapped_groups) = c("Category", "MAPPED_TRAIT_URI")
-
-#Add groups to original data frame
-
-main_inf = left_join(main_inf,mapped_groups, by = "MAPPED_TRAIT_URI")
-
-main_inf$Category = as.character(main_inf$Category)
-
-main_inf = main_inf %>% group_by(MAPPED_TRAIT_URI) %>% mutate(Category = paste0(unique(Category), collapse = "|")) %>% distinct()
-
-#Collapse genes and disease names for simplifcation purposes and then get unique combinations
-
-names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
-
-#Loop through disease traits and print a gene list for each
-
-dir.create('./data/Disease_Genes')
-
-dis_dir = "./data/Disease_Genes/"
-
-for (i in unique(names_genes$DISEASE.TRAIT)) {
-genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
-
-genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
-
-genes_curr = unique(genes_curr)
-
-write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
-}
-
-#Create map of EFO ids to terms
-
-EFO_list = lapply(unique(main_inf$MAPPED_TRAIT_URI), function(x) {
-  if (x == "GO_190252") {
-    term = NA
-  } else {
-  term = getLabel(getTermById(efo, x))
-  }
-  return(term)
-  })
-
-names(EFO_list) = unique(main_inf$MAPPED_TRAIT_URI)
-
-EFO_Frame = stack(EFO_list)
-
-colnames(EFO_Frame) = c("EFO_Name", "MAPPED_TRAIT_URI")
-
-main_inf = left_join(main_inf, EFO_Frame, "MAPPED_TRAIT_URI")
+# 
+# #This needs direct path for some reason
+# efo = getOntology("C:\\Users\\cresswellkg\\Documents\\disease_coherence_new\\Analysis\\data\\efo.owl")
+# 
+# 
+# #Cleaning up efo data to remove urls and putting comma seperated values on new lines
+# 
+# main_inf = unembed(main_inf, "MAPPED_TRAIT_URI", sep = ",")
+# main_inf$MAPPED_TRAIT_URI = basename(main_inf$MAPPED_TRAIT_URI)
+# 
+# tree = c()
+# for (i in unique(main_inf$MAPPED_TRAIT_URI)) {
+#   
+#   #Get entire EFO tree
+#   
+#   parents = getAllTermParentsById(efo, i)
+#   
+#   #Pull out labels and make into vector
+#   
+#   parents = sapply(parents, getLabel)
+#   
+#   #Cbind and fill in non-matching entries with NA
+#   
+#   tree = rowr::cbind.fill(tree,parents, fill = NA)
+#   
+# }
+# 
+# #Remove original column
+# 
+# tree = tree[,-1]
+# 
+# colnames(tree) = unique(main_inf$MAPPED_TRAIT_URI)
+# 
+# #Pick out important disorder groups
+# 
+# imp_cats = c("Abnormality of refraction", "Abnormality of the eye", "age", "anthropometric measurement", "auditory system disease", "autoimmune disease", "body weights and measures", "bone measurement", "cancer", "carcinoma", "cardiovascular disease", "cardiovascular disease biomarker measurement", "cardiovascular measurement", "diabetes mellitus", "diabetes mellitus biomarker", "digestive system disease", "eye measurement", "glucose measurement", "head and neck disorder", "heart function measurement", "hematological measurement", "immune system disease", "infectious disease biomarker", "insulin measurement", "lipid measurement", "lipid or lipoprotein measurement", "lipoprotein measurement", "mental or behavioural disorder", "metabolic disease", "metabolite measurement", "neoplasm", "nervous system disease", "neurodegenerative disease", "pulmonary function measurement", "reproductive system disease", "respiratory disease biomarker", "skin disease", "sleep measurement", "temporal measurement")
+# 
+# #Output cats by pipe
+# 
+# #Get categories for each unique EFO ID
+# 
+# cat_ass = apply(tree, 2, function(x) c(na.omit(match(x, imp_cats))))
+# 
+# #Replace categoryless values with NA
+# 
+# cat_ass = lapply(cat_ass, function(x) {
+#   if (length(x)<1) {
+#     cat = NA 
+#   } else {
+#     cat = x
+#   }
+#   return(cat)
+# })
+# 
+# #Assign category names
+# 
+# mapped_groups = stack(cat_ass)
+# mapped_groups$values = imp_cats[mapped_groups$values]
+# colnames(mapped_groups) = c("Category", "MAPPED_TRAIT_URI")
+# 
+# #Add groups to original data frame
+# 
+# main_inf = left_join(main_inf,mapped_groups, by = "MAPPED_TRAIT_URI")
+# 
+# main_inf$Category = as.character(main_inf$Category)
+# 
+# main_inf = main_inf %>% group_by(MAPPED_TRAIT_URI) %>% mutate(Category = paste0(unique(Category), collapse = "|")) %>% distinct()
+# 
+# #Collapse genes and disease names for simplifcation purposes and then get unique combinations
+# 
+# names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
+# 
+# #Loop through disease traits and print a gene list for each
+# 
+# dir.create('./data/Disease_Genes')
+# 
+# dis_dir = "./data/Disease_Genes/"
+# 
+# 
+# for (i in unique(names_genes$DISEASE.TRAIT)) {
+# genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
+# 
+# genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
+# 
+# genes_curr = unique(genes_curr)
+# 
+# write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
+# }
+# 
+# 
+# #Create map of EFO ids to terms
+# 
+# EFO_list = lapply(unique(main_inf$MAPPED_TRAIT_URI), function(x) {
+#   if (x == "GO_190252") {
+#     term = NA
+#   } else {
+#   term = getLabel(getTermById(efo, x))
+#   }
+#   return(term)
+#   })
+# 
+# names(EFO_list) = unique(main_inf$MAPPED_TRAIT_URI)
+# 
+# EFO_Frame = stack(EFO_list)
+# 
+# colnames(EFO_Frame) = c("EFO_Name", "MAPPED_TRAIT_URI")
+# 
+# main_inf = left_join(main_inf, EFO_Frame, "MAPPED_TRAIT_URI")
 
@@
 ```{r}
-main_sum = main_inf %>% ungroup() %>%  dplyr::select(DISEASE.TRAIT, MAPPED_GENE, Category, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% mutate(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE))) %>% dplyr::select(DISEASE.TRAIT, Category, snp_count_total, snp_count_no_gene, snp_prop_nogene, gene_count_total) %>% distinct()
-
-write.csv(main_sum, "supplementary_table_diseases_eof.csv", row.names = FALSE)
+# main_sum = main_inf %>% ungroup() %>%  dplyr::select(DISEASE.TRAIT, MAPPED_GENE, Category, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% mutate(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE))) %>% dplyr::select(DISEASE.TRAIT, Category, snp_count_total, snp_count_no_gene, snp_prop_nogene, gene_count_total) %>% distinct()
+# 
+# write.csv(main_sum, "supplementary_table_diseases_eof.csv", row.names = FALSE)
 ```
-
-```{r}
-#Creating KEGG pathways
-
-mart  = useMart("ensembl", dataset="hsapiens_gene_ensembl")
-min_genes = 5
-
-xx = as.list(KEGGPATHID2EXTID)     # All KEGG pathways
-xx = xx[ grepl("hsa", names(xx)) ] # Keep hsa pathways only
-xx_length = sapply(xx, length)     # Number of genes in each pathway
-xx = xx[xx_length > min_genes]     # Keep pathways with minimum number of genes or higher
-xx_sample = sample(1:length(xx), size = length(xx), replace = FALSE) # reshuffled indices
-# Get all gene symbols for KEGG pathways
-genes = getBM(attributes=c('entrezgene', 'hgnc_symbol'), filters='entrezgene', values=unique(unlist(xx)), mart=mart, uniqueRows=T)
-
-kegg_paths = lapply(xx, function(x) left_join(data.frame(entrezgene = as.integer(x)), genes))
-
-#Loop through kegg pathways and save genes
-
-dir.create("./data/Disease_Genes/KEGG")
-
-for (i in 1:length(kegg_paths)) {
-  genes = kegg_paths[[i]][,2]
-  write.table(genes, paste0("./data/Disease_Genes/KEGG/", names(kegg_paths)[i] ), row.names = FALSE, col.names = FALSE)
-
-}
-
-```

### 2019-08-19 Analysis/data/get_data.sh
add unzip for completeness

@@ -1,4 +1,4 @@
 wget https://stringdb-static.org/download/protein.links.v11.0/9606.protein.links.v11.0.txt.gz
 wget https://downloads.thebiogrid.org/Download/BioGRID/Release-Archive/BIOGRID-3.5.174/BIOGRID-ALL-3.5.174.mitab.zip
-
+unzip BIOGRID-ALL-3.5.174.mitab.zip
 

### 2019-08-19 Analysis/functions/disease2frame.R
fix code to match old verison

@@ -1,13 +1,13 @@
-diseases2mod = function(disease_dis, links) {
+diseases2frame = function(disease_dis, links) {
   
-  source('functions/genes2graph.R')
+  source('functions/genes2interactions.R')
 
   
   #Get internal vs. external connectivity
   
   disease_nets = lapply(disease_dis, function(x) {
     gene_list = read.table(x)$V1
-    genes2graph(gene_list, links, directed = FALSE)
+    genes2interactions(gene_list, links, directed = FALSE)
   })
   
   
@@ -32,4 +32,3 @@ diseases2mod = function(disease_dis, links) {
   disease_frame = bind_rows(disease_frame)
   return(disease_frame)
 }
-

### 2019-08-19 Analysis/Connectivity_Comparison_Undirected.Rmd
fully integrate

@@
 ```{r}
-#Repeating for string
-
-data_dir = "./data/"
-
-string_old = read_table2(paste0(data_dir,"9606.protein.links.v10.5.txt.gz"))
-
-string_old= string_old %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                             protein2 = gsub("9606.", "", protein2))
-
-#Read in ensemble protein ID
-
-ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
-genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
-
-#Join gene names with string_old
-
-string_old = left_join(string_old, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
-colnames(string_old)[4] ="hgnc_symbol_a"
-
-string_old = left_join(string_old, genes, by = c("protein2" = "ensembl_peptide_id"))
-
-colnames(string_old)[5] ="hgnc_symbol_b"
-
-dim(string)
-```
-
-```{r}
-combined_int = paste0(pmin(string_old$hgnc_symbol_a, string_old$hgnc_symbol_b), pmax(string_old$hgnc_symbol_a, string_old$hgnc_symbol_b)  ) %in%  paste0(pmin(string$hgnc_symbol_a, string$hgnc_symbol_b), pmax(string$hgnc_symbol_a, string$hgnc_symbol_b))
-
-sum(combined_int)/length(combined_int)
-
-string = string %>% mutate(Combined = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), "-", pmax(hgnc_symbol_a, hgnc_symbol_b))) %>% distinct()
-
-string_old = string_old %>% mutate(Combined = paste0(pmin(hgnc_symbol_a, hgnc_symbol_b), "-", pmax(hgnc_symbol_a, hgnc_symbol_b))) %>% distinct()
-
-string = string %>% arrange(Combined)
-string_old = string_old %>% arrange(Combined)
-
-#source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/venn.R")
-
-#Venn2(string$Combined, string_old$Combined)
-
-new_in_old = sum(string$Combined %in% string_old$Combined)/length(string$Combined)
-old_in_new = sum(string_old$Combined %in% string$Combined)/length(string_old$Combined)
-```
-
-```{r}
 
@@
 
-disease_select = read_csv("supplementary_table_diseases_selected.csv")
-eof = read_csv("supplementary_table_diseases_eof.csv")
+disease_select = read_csv("./supplementary_table_diseases_selected.csv")
 
@@
 
-eof_counts = eof %>% dplyr::select(-Category)
-
-#Join diseases to add categories and remove non-selected diseases
-
-disease_select = inner_join(eof_counts, disease_traits, c("DISEASE.TRAIT" = "Disease name")) %>% distinct()
-
-#Place category second
-
-disease_select = disease_select[,c(1,6,2:5)]
-
-#Summarizing
-
-# disease_sum = disease_select %>% group_by(Category) %>% summarise(Med_SNP = median(snp_count_total), Med_Gene = median(Number.of.genes), Mean_Missing = mean(per_missing)) %>% arrange(-Mean_Missing)
-# 
-# disease_mann_whit = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$snp_count_total, g = .$Category, p.adjust.method = "none")))
-# 
-# disease_mann_whit_prop = disease_select[1:137,]  %>% mutate(Prop_Miss = 1- Number.of.genes/snp_count_total) %>% group_by(Category) %>% summarise(Med_Miss = median(Prop_Miss)) %>% arrange(-Med_Miss) # %>% do(tidy(pairwise.wilcox.test(.$Prop_Miss, g = .$Category, p.adjust.method = "none")))
-# 
-# disease_mann_whit_gene = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$Number.of.genes, g = .$Category, p.adjust.method = "none"))) %>% arrange(p.value)
-# 
-# disease_mann_whit_miss = disease_select[1:137,]  %>% do(tidy(pairwise.wilcox.test(.$per_missing, g = .$Category, p.adjust.method = "none"))) %>% arrange(p.value)
-
 #Change first column to make more clear
@@
 colnames(disease_select)[1] = "Disease"
+
+#Filtering out NA in last row
+
+disease_select = disease_select %>% filter(!is.na(Disease))
 
@@
 
-#Calculate modularity
-
-disease_mod = lapply(disease_length, function(x) {
-
-  #Run function to find internal and external connectivity 
-  mod = diseases2mod(disease_selected[x], links = links)
-
-})
-
-saveRDS(disease_mod, "Modularity_Biogrid.rds")
-
+# #Calculate modularity
+# 
+# disease_mod = lapply(disease_length, function(x) {
+# 
+#   #Run function to find internal and external connectivity 
+#   mod = diseases2mod(disease_selected[x], links = links)
+#   
+# })
+# 
+# saveRDS(disease_mod, "Modularity_Biogrid.rds")
+# 
 #Pull out network information for overall data frame
@@
 
-#Add number of snps with no connection to original data frame
+saveRDS(disease_overall, "biogrid_edges_new.rds")
 
-per_no_ppi_bio = sapply(disease_frames, function(x) x$lengths)
-
-per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_bio = per_no_ppi_bio)
-
-disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
+# #Add number of snps with no connection to original data frame
+# 
+# per_no_ppi_bio = sapply(disease_frames, function(x) x$lengths)
+# 
+# per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_bio = per_no_ppi_bio)
+# 
+# disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
 
@@
 
-disease_mod_string = lapply(disease_length, function(x) {
-
-  #Run function to find internal and external connectivity 
-  mod = diseases2mod(disease_selected[x], links = string)
-
-})
-
-saveRDS(disease_mod_string, "Modularity_String.rds")
+# disease_mod_string = lapply(disease_length, function(x) {
+# 
+#   #Run function to find internal and external connectivity 
+#   mod = diseases2mod(disease_selected[x], links = string)
+#   
+# })
+# 
+# saveRDS(disease_mod_string, "Modularity_String.rds")
 
@@
 
-per_no_ppi_string = sapply(disease_frames_string, function(x) x$lengths)
-
-per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_string = per_no_ppi_string)
+# per_no_ppi_string = sapply(disease_frames_string, function(x) x$lengths)
+# 
+# per_no_ppi_frame = data.frame(Disease = unique(disease_select$Disease), per_no_ppi_string = per_no_ppi_string)
+# 
+# saveRDS(disease_overall_string, "string_edges_new.rds")
+# 
+# disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
 
@@
 
-disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
-
-saveRDS(disease_overall_string, "string_edges_new.rds")
-
-saveRDS(disease_overall, "biogrid_edges_new.rds")
 ```
-
-```{r}
-#Generating 20 random networks with 100 genes each
-
-#disease_counts = disease_overall %>% group_by(Disease) %>% summarise(counts = n())
-
-set.seed(123)
-
-rand_nets = list()
-for (i in 1:20) {
-rand_nets[[i]] = GeneSample(1000, links, directed = FALSE)
-}
-
-# index = 1
-# rand_nets = list()
-# for (i in disease_counts$counts) {
-#   rand_nets[[index]] = GeneSample(i, links, directed = FALSE)
-#   index = index+1
-# }
-
-#Getting a dataframe of internal/external connectivity
-
-rand_length = 1:length(rand_nets)
-
-rand_frame = lapply(rand_length, function(y) {
-
-  #Read through random list of gene networks
-
-  x = rand_nets[[y]]
-
-  #Place internal and external edges into data frame
-
-  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
-  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
-  colnames(sub_frame) = c("Genes", "Internal", "External")
-
-  #Add category and "disease" name
-
-  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
-  })
-
-#Bind list into data frame
-
-rand_frame = bind_rows(rand_frame)
-
-disease_overall = bind_rows(disease_overall, rand_frame)
-
-```
-
-```{r}
-#Repeat for string
-
-# disease_counts_string = disease_overall_string %>% group_by(Disease) %>% summarise(counts = n())
-
-set.seed(123)
-
-rand_nets = list()
-for (i in 1:20) {
-rand_nets[[i]] = GeneSample(1000, string, directed = FALSE)
-}
-
-#index = 1
-# rand_nets = list()
-# for (i in disease_counts_string$counts) {
-#   rand_nets[[index]] = GeneSample(i, links, directed = FALSE)
-#   index = index+1
-# }
-
-#Getting a dataframe of internal/external connectivity
-
-rand_length = 1:length(rand_nets)
-
-rand_frame_string = lapply(rand_length, function(y) {
-
-  #Read through random list of gene networks
-
-  x = rand_nets[[y]]
-
-  #Place internal and external edges into data frame
-
-  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
-  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
-  colnames(sub_frame) = c("Genes", "Internal", "External")
-
-  #Add category and "disease" name
-
-  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
-  })
-
-#Bind list into data frame
-
-rand_frame_string = bind_rows(rand_frame_string)
-
-disease_overall_string = bind_rows(disease_overall_string, rand_frame_string)
-
-```
-
-```{r}
-#Repeating for KEGG pathways
-
-kegg_directory = "./data/Disease_Genes/KEGG/"
-
-kegg_names = list.files(kegg_directory)
-
-kegg_list = paste0("./data/Disease_Genes/KEGG/", kegg_names)
-
-kegg_length = length(kegg_list)
-
-kegg_frames = lapply(1:kegg_length, function(x) {
-  diseases = diseases2frame(kegg_list[x], links = links)
-  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
-  return(list(diseases = diseases, lengths = per_no_ppi))
-  })
-
-kegg_overall = lapply(kegg_frames, function(x) x$diseases %>% mutate(Category = "KEGG"))
-
-disease_overall = bind_rows(disease_overall, kegg_overall) %>% distinct()
-```
-
-```{r}
-#Repeating for string
-
-kegg_frames_string = lapply(1:kegg_length, function(x) {
-  diseases = diseases2frame(kegg_list[x], links = string)
-  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
-  return(list(diseases = diseases, lengths = per_no_ppi))
-  })
-
-kegg_overall_string = lapply(kegg_frames_string, function(x) x$diseases %>% mutate(Category = "KEGG"))
-
-disease_overall_string = bind_rows(disease_overall_string, kegg_overall_string) %>% distinct()
-```
-
-```{r}
-#Load in normalization regression
-
-pred_edges = readRDS("pred_edges.rds")
-pred_edges_biogrid= readRDS("pred_edges_biogrid.rds")
-
-```
-
-```{r}
-# #Permutation test for slope difference
-# 
-# source("./functions/perm_slopes.R")
-# 
-# set.seed(123)
-# 
-# #Run permutation test on each category for differences in slope
-# 
-# perm_cat_slope = disease_overall%>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))
-# 
-# perm_cat_slope_string = disease_overall_string %>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))
-# 
-# perm_slope = disease_overall%>% group_by(Category) %>% mutate(Disease = Category) %>% do(tidy(slope_perm(disease_overall_string, Disease[1], pred_edges, 1000)))
-
-```
-
-```{r}
-#Normalizing by number of genes
-
-#disease_zeros = disease_overall %>% group_by(Category) %>% summarise(Per_Zero = sum(Internal==0)/length(Internal))
-
-#Remove edges with less than 5 total counts, square root internal edges and correct by predicted
-
-disease_overall = disease_overall  %>% mutate(External = External-Internal) %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges_biogrid, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease)) %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4) %>% filter(length(unique(Internal))>1)
-
-#Make names more informative
-
-disease_overall = disease_overall %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))
-
-#Repeat for string
-
-disease_overall_string = disease_overall_string %>% mutate(External = External-Internal) %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease))  %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4)
-
-disease_overall_string = disease_overall_string %>% group_by(Disease) %>% filter(length(unique(Internal))>1)
-
-disease_overall_string =  disease_overall_string %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))
-
-```
-
-```{r}
-disease_overall = readRDS("disease_overall_biogrid.rds")
-disease_overall_string = readRDS("disease_overall_string.rds")
-
-```
-
-```{r}
-require(msigdbr)
-dis_slopes_string_count = disease_overall_string %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% filter(Internal != 0) %>% 
-    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-    do({
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
-
-msig = readRDS("KEGG_String.rds")
-#msig =msig %>% filter(Category == "KEGG")
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[1:10,]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Rand_Norm = rand_slope, Msig_Norm = median(msigs$Slope))
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string.rds")
-}
-
-```
-
-```{r}
-
-coherence_string_filt = readRDS("coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string = readRDS("coherence_string.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid = readRDS("coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)
-
-coherence_full = full_join(coherence_string_filt, coherence_string, by = "Disease")
-
-coherence_full = full_join(coherence_full, coherence_biogrid, by = "Disease")
-
-coherence_full = coherence_full[,c(1,4,3,2)]
-
-colnames(coherence_full) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")
-
-old_res = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(-Coherence.Biogrid, -Coherence.String.full, -Coherence.String.filt)
-
-old_coh = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(Disease = Trait, Coherence.Biogrid, Coherence.String.full, Coherence.String.filt)
-
-test_coherence = test = left_join(coherence_full, old_coh)
-
-colnames(old_res)[1] = "Disease"
-
-coherence_full = left_join(coherence_full, old_res) %>% filter(!is.na(Total.number.of.SNPs))
-
-coherence_full = coherence_full[,c(1,5, 6:7, 2:4,8:10)]
-
-colnames(coherence_full) = gsub("\\.", " ", colnames(coherence_full))
-
-colnames(coherence_full) = gsub("\\_", " ", colnames(coherence_full))
-
-colnames(coherence_full)[1] = "Trait"
-
-coherence_full[,5:7] = apply(coherence_full[,5:7], 2, function(x) ifelse(x<0, 0, ifelse (
-  x>1, 1, x
-)))
-
-write.table(coherence_full, "supplementary_table_S1_New.txt", sep = "\t", row.names = FALSE)
-
-#Read in old table I sent
-
-old_tab = read_tsv("C:\\Users\\cresswellkg\\Downloads\\supplementary_table_S1_New.txt", col_names = TRUE)
-
-old_tab_comb = left_join(old_tab, old_coh, "Disease")
-
-cor(old_tab_comb$`Coherence String`, old_tab_comb$Coherence.String.filt, "complete")
-```
-
-#Plot of all diseases
-
-```{r}
-
-#Calculate the slopes for diseases and categories
-
-#Slope for categories
-
-slopes = disease_overall %>% filter(Internal != 0) %>% 
-    group_by(Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-#Slope for diseases
-
-dis_slopes = disease_overall %>% filter(Internal != 0) %>% 
-    group_by(Disease, Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-#Permutation test
-
-# dis_pvals = disease_overall %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Disease, Category) %>% 
-#     do({
-#       perm = slope_perm(disease_overall, .$Disease[1], database = links, Norm_Reg = NULL, permutations = 1000)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-
-#Slopes for Biogrid categories
-
-# cat_pvals = disease_overall %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Category) %>%
-#     do({
-#       perm = slope_perm(disease_overall, .$Category[1], Norm_Reg =NULL, database = links, permutations = 1000, Disease = FALSE)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-
-#Slopes for categories string
-
-slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
-    group_by(Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-#Slopes for diseases
-
-dis_slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
-    group_by(Disease, Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-# #Add p-values
-# 
-# dis_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Disease, Category) %>% 
-#     do({
-#       perm = slope_perm(disease_overall_string, .$Disease[1], database = string, Norm_Reg =NULL, permutations = 1000)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-# 
-# #Category p-values
-# 
-# cat_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Category) %>%
-#     do({
-#       perm = slope_perm(disease_overall_string, .$Category[1], database = Norm_Reg =NULL, permutations = 1000, Disease = FALSE)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-
-#Add slopes to the overall dataset
-
-disease_overall = left_join(disease_overall, slopes, "Category")
-colnames(disease_overall)[7] = "Category_Slope"
-
-disease_overall = left_join(disease_overall, dis_slopes[,c(1,3)], "Disease")
-colnames(disease_overall)[8] = "Disease_Slope"
-
-# disease_overall = left_join(disease_overall, dis_pvals[,c(1,3)], "Disease")
-# colnames(disease_overall)[9] = "Slope_Pval"
-
-disease_overall_string = left_join(disease_overall_string, slopes_string, "Category")
-colnames(disease_overall_string)[7] = "Category_Slope"
-
-disease_overall_string = left_join(disease_overall_string, dis_slopes_string[,c(1,3)], "Disease")
-colnames(disease_overall_string)[8] = "Disease_Slope"
-# 
-# disease_overall_string = left_join(disease_overall_string, dis_pvals_string[,c(1,3)], "Disease")
-# colnames(disease_overall_string)[9] = "Slope_Pval"
-
-#KEGG and random should be same color on all plots
-#Order by slope
-
-library(RColorBrewer)
-
-#Make levels into factors
-
-disease_overall = disease_overall %>% arrange(-Category_Slope) %>% mutate(Category = factor(Category, levels =unique(Category)))
-
-disease_overall_string = disease_overall_string %>% arrange(-Category_Slope) %>% mutate(Category = factor(Category, levels =unique(Category)))
-
-#Filter out KEGG categories with worse coherence than fasting glucose
-
-disease_overall = disease_overall %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")
-
-disease_overall_string = disease_overall_string %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall_string %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")
-
-#Set a custom color palette for both string and biogrid
-
-library(randomcoloR)
-
-#Set permanent colors for random and KEGG
-
-col =  c("Red", "#999999", distinctColorPalette(12))
-names(col) = c("Random", "KEGG", as.character(unique(disease_overall_string$Category))[!(unique(disease_overall_string$Category) %in% c("KEGG", "Random"))])
-
-col_full = scale_color_manual(values = col)
-
-category_plot = ggplot(disease_overall, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))
-
-category_plot_string = ggplot(disease_overall_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2)  + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + col_full
-
-ggsave("./figures/category_plot_biogrid.png", category_plot, height=5, width=7, units="in", dpi = 300)
-
-ggsave("./figures/category_plot_string.png", category_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-```{r}
-#Table of category coherence
-
-#Get gene counts
-
-gene_count_cat = disease_overall %>% group_by(Category) %>% summarise(gene_count_biogrid = n())
-gene_count_cat_string = disease_overall_string %>% group_by(Category) %>% summarise(gene_count_string = n())
-
-coherence_frame =full_join(gene_count_cat, gene_count_cat_string, "Category")
-
-#Get coherence
-
-coherence_cat = slopes %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% distinct()
-
-coherence_cat_string = slopes_string %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% distinct()
-
-coherence_frame = full_join(coherence_frame,coherence_cat, "Category")
-
-coherence_frame = full_join(coherence_frame,coherence_cat_string, "Category")
-
-biogrid_cat_p = readRDS("biogrid_p_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `Biogrid Perm P-value` = perm )
-
-string_cat_p = readRDS("string_p_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `String Perm P-value` = perm )
-
-coherence_frame = left_join(coherence_frame, biogrid_cat_p, by = "Category")
-coherence_frame = full_join(coherence_frame, string_cat_p, by = "Category")
-
-# coherence_frame = full_join(coherence_frame, cat_pvals, "Category")
-# 
-# coherence_frame = full_join(coherence_frame, cat_pvals_string, "Category") %>% filter(!(Category %in% c("KEGG", "Random")))
-
-#coherence_frame = coherence_frame %>% mutate(cat_pvals = NA, cat_pvals_string = NA)
-
-colnames(coherence_frame)[4:7] = c("Coherence Biogrid", "Coherence String",  "Biogrid Perm P-val", "String Perm P-val")
-
-coherence_frame = coherence_frame %>% filter(!is.na(`String Perm P-val`)) %>% arrange(Category)
-
-write.csv(coherence_frame, "./tables/supplementary_table_categories.csv", row.names = FALSE)
-
-```
-
-# Plot of all categories (Biogrid)
-
-```{r}
-category_plot
-```
-
-# Plot of all categories (String)
-
-```{r}
-category_plot_string
-```
-
-```{r}
-category = disease_overall %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)
-
-category_string = disease_overall_string %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)
-
-col_sub = col[names(col) %in% category$Category]
-
-cat_plot = ggplot(category, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))
-
-ggsave("./figures/category_disease_plot_biogrid.png", cat_plot, height=5, width=7, units="in", dpi = 300)
-
-cat_plot_string = ggplot(category_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))
-
-ggsave("./figures/category_disease_plot_string.png", cat_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of non-trait categories (Biogrid)
-
-```{r}
-cat_plot
-```
-
-# Plot of non-trait categories (String)
-
-```{r}
-cat_plot_string
-```
-
-```{r}
-traits = disease_overall %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")
-
-traits_string = disease_overall_string %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")
-
-col_sub = col[names(col) %in% traits$Category]
-
-trait_plot = ggplot(traits, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
-   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_blank(),
-        panel.background = element_blank(), legend.text = element_text(size = 16))
-
-ggsave("./manuscript/Fig_S5B_category_trait_plot_biogrid.png",trait_plot, height=5, width=7, units="in", dpi = 300)
-
-col_sub_string = col[names(col) %in% traits_string$Category]
-
-trait_plot_string = ggplot(traits_string, aes(x = Internal, y = External, color = Category)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub_string)  + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
-   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_blank(),
-        panel.background = element_blank(), legend.text = element_text(size = 16))
-
-ggsave("./manuscript/Fig_S5A_category_trait_plot_string.png",trait_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of traits (Biogrid)
-
-```{r}
-trait_plot
-```
-
-# Plot of traits (String)
-
-```{r}
-trait_plot_string
-```
-
-# Plot of anthropometric measures
-
-```{r}
-#Subsetting to include anthro traits
-
-disease_anthro = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))
-
-#Repeat for string
-
-disease_anthro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))
-
-#Plotting anthropometric traits
-
-col_dis_bio = c(col[names(col) %in% disease_anthro$Disease], distinctColorPalette(length(unique(disease_anthro$Disease))))
-
-#Giving unique colors for each disease
-
-names(col_dis_bio)[-c(1,2)] = as.character(unique(disease_anthro$Disease))
-
-anthro_plot = ggplot(disease_anthro , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_bio) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/anthro_plot_biogrid.png",anthro_plot, height=5, width=7, units="in", dpi = 300)
-
-#Repeating for string
-
-#Giving unique colors for each disease
-
-col_dis_string = c(col[names(col) %in% disease_anthro_string$Disease], distinctColorPalette(length(unique(disease_anthro_string$Disease))))
-
-names(col_dis_string)[-c(1,2)] = as.character(unique(disease_anthro_string$Disease))
-
-anthro_plot_string = ggplot(disease_anthro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_string) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/anthro_plot_string.png",anthro_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-```{r}
-#Order and keep random and KEGG consistent from other plot
-
-disease_psych = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease) %>% filter(n()>4)
-
-disease_psych_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease)
-
-col_dis = c(col[names(col) %in% disease_psych$Disease], distinctColorPalette(length(unique(disease_psych$Disease))))
-
-names(col_dis)[-c(1,2)] = as.character(unique(disease_psych$Disease))
-
-psych_plot = ggplot(disease_psych , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-#Redo colors for string
-
-col_dis = c(col[names(col) %in% disease_psych_string$Disease], distinctColorPalette(length(unique(disease_psych_string$Disease))))
-
-names(col_dis)[-c(1,2)] = as.character(unique(disease_psych_string$Disease))
-
-psych_plot_string = ggplot(disease_psych_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/psych_plot_biogrid.png",psych_plot, height=5, width=7, units="in", dpi = 300)
-
-ggsave("./figures/psych_plot_string.png",psych_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of psychiatric diseases (Biogrid)
-
-```{r}
-psych_plot
-```
-
-# Plot of psychiatric diseases (String)
-
-```{r}
-psych_plot_string
-```
-
-# Plot of Neuorologic diseases
-
-```{r}
-#Subset to include neurological disorders
-
-disease_neuro = disease_overall  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))
-
-disease_neuro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))
-
-#Setting the colors for neurologic diseases
-
-col_neuro = c(col[names(col) %in% disease_neuro$Disease], distinctColorPalette(length(unique(disease_neuro$Disease))))
-
-names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro$Disease))
-
-#Biogrid
-
-neuro_plot = ggplot(disease_neuro , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/neuro_plot_biogrid.png", neuro_plot, height=5, width=7, units="in", dpi = 300)
-
-#Repeating for STRING
-
-#Names for 10 diseases
-
-col_neuro = c(col[names(col) %in% disease_neuro_string$Disease], distinctColorPalette(length(unique(disease_neuro_string$Disease))))
-
-names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro_string$Disease))
-
-neuro_plot_string = ggplot(disease_neuro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/neuro_plot_string.png", neuro_plot_string, height=5, width=7, units="in", dpi = 300)
-```
-
-```{r}
-#Function to create all other plots
-
-dir.create("./figures/Category")
-
-plot_coherence = function(Cat, dataset, protein_database, filtered = FALSE) {
-
-  all_plots = list()
-  #Subset based on category
-
-sub_frame = dataset  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope))  %>% mutate(Disease_Slope = ifelse(Category == "KEGG", 0, Disease_Slope)) %>%
- arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == Cat | Disease == "KEGG" | Disease == "Random"))
-print(sub_frame)
-
-print(head(sub_frame))
-
-  #Get colors
-
-  color = c(col[names(col) %in% sub_frame$Disease], distinctColorPalette(length(unique(sub_frame$Disease))))
-
-names(color)[-c(1,2)] = as.character(unique(sub_frame$Disease))
-
-#Make plot
-
-dis_plot = ggplot(sub_frame , aes(x = Internal, y = External, color =Disease, -Disease_Slope)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = color) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-all_plots[[i]] = dis_plot
-
-if (filtered) {
-ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, "_filtered", ".png"), height=5, width=7, units="in", dpi = 300) 
-} else {
-  ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, ".png"), height=5, width=7, units="in", dpi = 300) 
-}
-
-return(all_plots)
-}
-
-#Making plots for each category
-
-disease_overall = disease_overall %>% ungroup() %>% arrange(as.character(Category))
-
-cat_plots = lapply(unique(disease_overall$Category), function(x) {
-  plot_coherence(x, dataset = disease_overall %>% ungroup() , protein_database = "biogrid", filtered = FALSE)
-})
-
-saveRDS(cat_plots, "biogrid.rds")
-
-disease_overall_string = disease_overall_string %>% ungroup() %>% arrange(as.character(Category))
-
-cat_plots_string = lapply(unique(disease_overall_string$Category), function(x) {
-  plot_coherence(x, dataset = disease_overall_string %>% ungroup() , protein_database = "string", filtered = FALSE)
-})
-
-saveRDS(cat_plots_string, "string.rds")
-
-```
-# Plot of select diseases
-
-```{r}
-
-# #Add all of KEGG and random
-# 
-# select_disease= disease_overall %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% filter(Disease %in% c("Height", "Asthma", "Breast cancer", "Coronary Heart Disease",
-# "Blood pressure", "Glaucoma", "Intelligence",
-# "Type 2 diabetes", "LDL cholesterol", "Alzheimer's disease", "Schizophrenia", "KEGG", "Random")) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease)))
-# 
-# names(col_sub) = unique(select_disease$Disease)[-c(1,8)]
-# 
-# select_disease_plot = ggplot(select_disease, aes(x = Internal, y = External, color = Disease))  + geom_point(alpha = .1)+ geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(Random =  "#6F8273" , KEGG = "#999999", col_sub)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-# 
-# ggsave("C:\\Users\\cresswellkg\\Documents\\disease-coherence\\manuscript\\figures\\disease_plot_inbio.png", select_disease_plot, height=5, width=7, units="in", dpi = 300)
-```
-
-```{r}
-# disease_summary_biogrid = disease_overall %>% group_by(Disease) %>% summarise(Mean_Internal = mean(Internal), Median_Internal = median(Internal), Min_Internal = min(Internal), Max_Internal = max(Internal), Mean_External = mean(External), Median_External = median(External), Min_External = min(External), Max_External = max(External), Percent_Zero = sum(Internal == 0)/n(), Num_Non_Zero = sum(Internal !=0), Highest_Gene = Genes[which.max(Internal/External)])
-```
-
-```{r}
-#3 Figures
-
-#KEGG, Random and Diseases
-
-#KEGG, Random and Traits
-
-#Disease specifically from highest number but manually choose relevant ones too
-```
-
-```{r}
-#Find number of SNPs with no connections
-
-# disease_summary_string = disease_overall_string %>% group_by(Disease) %>% summarise(Mean_Internal = mean(Internal), Median_Internal = median(Internal), Min_Internal = min(Internal), Max_Internal = max(Internal), Mean_External = mean(External), Median_External = median(External), Min_External = min(External), Max_External = max(External), Percent_Zero = sum(Internal == 0)/n(), Num_Non_Zero = sum(Internal!=0), Highest_Gene = Genes[which.max(Internal/External)])
-
-# disease_summary_combined = left_join(disease_summary_biogrid, disease_summary_string, "Disease")
-# 
-# colnames(disease_summary_combined) = gsub("\\.x", "_biogrid", colnames(disease_summary_combined))
-# colnames(disease_summary_combined) = gsub("\\.y", "_string", colnames(disease_summary_combined))
-# 
-# disease_summary_combined = disease_summary_combined[,c(1,2,13,3,14,4,15,5,16,6,17,7,18,8,19,9,20,10,21,11,22,12,23)]
-# 
-# disease_summary_combined = disease_summary_combined %>% filter(!is.na(Mean_Internal_string))
-```
-
-```{r}
-
-# #Calculating coherence
-# 
-dis_diff = disease_overall %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()
-
-dis_diff_string = disease_overall_string %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()
-
-disease_select = left_join(disease_select, dis_diff, by = "Disease")
-disease_select = left_join(disease_select, dis_diff_string, by = "Disease")
-
-colnames(disease_select)[9:10] = c("Biogrid Coherence", "String Coherence")
-
-#Removing duplicates
-
-row.names(disease_select) = NULL
-disease_select = disease_select %>% arrange(-`String Coherence`) %>% filter(!duplicated(disease_select$Disease))
-
-#Adding slope p-values
-
-# disease_select = left_join(disease_select, dis_pvals %>% ungroup() %>% dplyr::select(-Category), by = "Disease") %>% ungroup()
-# 
-# disease_select = left_join(disease_select, dis_pvals_string %>% ungroup() %>% dplyr::select(-Category), by = "Disease") %>% ungroup()
-
-# colnames(disease_select)[11:12] = c("Biogrid Perm P-value", "String Perm P-value")
-
-# disease_select = disease_select %>% mutate(`Biogrid Perm P-value` = NA, `String Perm P-value` = NA)
-
-#Read in p-values from cluster
-
-biogrid_p = readRDS("biogrid_p.rds")
-biogrid_p_neuro = readRDS("biogrid_p_neuro.rds")
-biogrid_p_psych = readRDS("biogrid_p_psych.rds")
-
-bio_over = rbind(biogrid_p, biogrid_p_neuro, biogrid_p_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `Biogrid Perm P-value` = perm ) %>% filter(!duplicated(Disease))
-
-disease_select = full_join(disease_select, bio_over, by = "Disease")
-
-string_p = readRDS("string_p.rds")
-string_p_neuro = readRDS("string_p_neuro.rds")
-string_p_psych = readRDS("string_p_psych.rds")
-
-string_over = rbind(string_p, string_p_neuro, string_p_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `String Perm P-value` = perm ) %>% filter(!duplicated(Disease))
-
-disease_select = full_join(disease_select, string_over, by = "Disease")
-
-write.csv(disease_select, "./tables/supplementary_table_diseases.csv", row.names = FALSE)
-```
-
-```{r}
-dir.create("./data/Disease_Genes/Categories_STRING")
-dir.create("./data/Disease_Genes/Categories_Biogrid")
-
-disease_cats_string = unique(disease_overall_string$Category)
-disease_cats_biogrid = unique(disease_overall$Category)
-
-for (i in disease_cats_string) {
-  gene_list = disease_overall_string %>% filter(Category == i) %>% dplyr::select(Genes) %>% dplyr::distinct()
-
-  write.table(gene_list$Genes, paste0("./data/Disease_Genes/Categories_STRING/", i, ".txt"), row.names = FALSE, col.names = FALSE)
-}
-
-for (i in disease_cats_biogrid) {
-  gene_list = disease_overall %>% filter(Category == i) %>% dplyr::select(Genes) %>% dplyr::distinct()
-
-  write.table(gene_list$Genes, paste0("./data/Disease_Genes/Categories_Biogrid/", i, ".txt"), row.names = FALSE, col.names = FALSE)
-}
-
-string_files = paste0("./data/Disease_Genes/Categories_STRING/", list.files("./data/Disease_Genes/Categories_STRING/"))
-
-biogrid_files = paste0("./data/Disease_Genes/Categories_Biogrid/", list.files("./data/Disease_Genes/Categories_Biogrid/"))
-
-```
-
-```{r}
-
-cat_frames_biogrid = lapply(biogrid_files, function(x) {
-
-  #Run function to find internal and external connectivity 
-  cats = diseases2frame(x, links = links)
-
-  #Record ppi missing
-  cats
-  })
-
-cat_frames_string = lapply(string_files, function(x) {
-
-  #Run function to find internal and external connectivity 
-  cats = diseases2frame(x, links = string)
-
-  #Record ppi missing
-  cats
-  })
-
-cat_frames_biogrid = bind_rows(cat_frames_biogrid)
-cat_frames_string = bind_rows(cat_frames_string)
-
-```
-
-```{r}
-cat_overall = cat_frames_biogrid  %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges_biogrid, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>%  ungroup()  %>% distinct()
-
-cat_overall_string = cat_frames_string  %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup()  %>% distinct()
-
-cat_slopes = cat_overall %>% filter(Internal != 0) %>% 
-    group_by(Disease) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-cat_slopes_string = cat_overall_string %>% filter(Internal != 0) %>% 
-    group_by(Disease) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-cat_overall = left_join(cat_overall, cat_slopes, by = "Disease")
-cat_overall_string = left_join(cat_overall_string, cat_slopes, by = "Disease")
-
-category_plot_comb = ggplot(cat_overall, aes(x = Internal, y = External, color = reorder(Disease, -Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))
-
-category_plot_comb_string = ggplot(cat_overall_string, aes(x = Internal, y = External, color = reorder(Disease, -Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + col_full + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm"))
-
-ggsave("./figures/category_plot_biogrid_combined.png", category_plot_comb, height=5, width=7, units="in", dpi = 300)
-
-ggsave("./figures/category_plot_string_Combined.png", category_plot_comb_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-```{r}
-#Manually set max and min for STRING and biogrid
-
-min_bio = min(dis_slopes$Slope)
-max_bio = max(dis_slopes$Slope)
-
-min_string = min(dis_slopes_string$Slope)
-max_string = max(dis_slopes_string$Slope)
-
-#Independent analysis of KEGG pathways
-
-kegg_slopes_string = dis_slopes_string %>% filter(Category == "KEGG")
-kegg_slopes_bio = dis_slopes %>% filter(Category == "KEGG")
-
-#Gsubing out KEGG prefix
-
-kegg_slopes_string = kegg_slopes_string %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))
-kegg_slopes_bio = kegg_slopes_bio %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))
-
-#Get names of KEGG pathways
-
-xx = data.frame(path_name = names(as.list(KEGGPATHNAME2ID)), kegg_code = unlist(as.list(KEGGPATHNAME2ID)))
-
-#Match name with code
-
-kegg_slopes_string = kegg_slopes_string %>% mutate(kegg_code = gsub("hsa", "", Disease))
-kegg_slopes_bio = kegg_slopes_bio %>% mutate(kegg_code = gsub("hsa", "", Disease))
-
-#Create table
-
-kegg_slopes_string = left_join(kegg_slopes_string, xx, "kegg_code" )
-kegg_slopes_bio = left_join(kegg_slopes_bio, xx, "kegg_code" )
-
-kegg_frame_string = kegg_slopes_string %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>% mutate(Coherence = ((Slope - max_string))/(min_string - max_string))
-
-kegg_frame_bio = kegg_slopes_bio %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>%  mutate(Coherence = ((Slope - max_bio))/(min_bio - max_bio))
-
-#Strip biogrid to merge
-
-kegg_frame_bio = kegg_frame_bio %>% dplyr::select(hsa_code, Bio_Slope = Slope, Bio_Coherence = Coherence)
-
-colnames(kegg_frame_string)[3:4] = c("String_Slope", "String_Coherence")
-
-kegg_frame = left_join(kegg_frame_string, kegg_frame_bio, "hsa_code")
-
-#Get counts for each KEGG frame
-
-kegg_counts_string = bind_rows(kegg_overall_string) %>% group_by(Disease) %>% summarise(String_Count = n()) %>% rename(hsa_code = Disease)
-
-kegg_counts_biogrid = bind_rows(kegg_overall) %>% group_by(Disease) %>% summarise(Biogrid_Count = n()) %>% rename(hsa_code = Disease)
-
-kegg_frame = left_join(kegg_frame, kegg_counts_string, "hsa_code")
-
-kegg_frame = left_join(kegg_frame, kegg_counts_biogrid, "hsa_code")
-
-write.csv(kegg_frame, "KEGG_Summary.csv", row.names = FALSE)
-```
-
-```{r}
-
-```

### 2019-08-19 Analysis/Connectivity_Comparison_Filtered.Rmd
add string filtered update

@@
 
-dim(string)
-```
-
-```{r}
-#Repeat for string
-
-#Generating 20 random networks with 20 genes each
-
-set.seed(123)
-
-rand_nets = list()
-for (i in 1:20) {
-rand_nets[[i]] = GeneSample(5, string, directed = FALSE)
-}
-
-#Getting a dataframe of internal/external connectivity
-
-rand_length = 1:length(rand_nets)
-
-rand_frame_string = lapply(rand_length, function(y) {
-
-  #Read through random list of gene networks
-
-  x = rand_nets[[y]]
-
-  #Place internal and external edges into data frame
-
-  sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)
-  sub_frame = cbind.data.frame(colnames(sub_frame), t(sub_frame[1,]), t(sub_frame[2,]))
-  colnames(sub_frame) = c("Genes", "Internal", "External")
-
-  #Add category and "disease" name
-
-  sub_frame %>% mutate(Disease = paste0("Run ", y), Category = "Random")
-  })
-
-#Bind list into data frame
-
-rand_frame_string = bind_rows(rand_frame_string)
-```
-
-```{r}
-
-#Get directory with diseases
-
-disease_directory = "./data/Disease_Genes/"
-
-disease_names = list.files(disease_directory)
-
-disease_list = paste0("./data/Disease_Genes/", disease_names, "/")
-
-#Read in manually selected diseases
-
-disease_select = read_csv("supplementary_table_diseases_selected.csv") %>% filter(!is.na(`Disease name`))
-
-#Change first column to make more clear
-
-colnames(disease_select)[1] = "Disease"
-
-#Creating a vector of file names from disease select
-
-disease_files =  unique(paste0(disease_select$Disease, ".txt"))
-
-#Make full path
-
-disease_selected = paste0("./data/Disease_Genes/", disease_files)
-
 ```
@@
 
-disease_mod_string_filt = lapply(disease_length, function(x) {
-
-  #Run function to find internal and external connectivity 
-  mod = diseases2mod(disease_selected[x], links = string)
-
-})
-
-saveRDS(disease_mod_string_filt, "Modularity_String_Filt.rds")
+# disease_mod_string_filt = lapply(disease_length, function(x) {
+# 
+#   #Run function to find internal and external connectivity 
+#   mod = diseases2mod(disease_selected[x], links = string)
+#   
+# })
+# 
+# saveRDS(disease_mod_string_filt, "Modularity_String_Filt.rds")
 
@@
 
-for (i in unique(disease_overall_string$Disease)) {
- curr_dis = disease_overall_string %>% filter(Disease == i)
-
-}
-
-disease_overall_string = bind_rows(disease_overall_string, rand_frame_string)
-
 #disease_overall_string = #readRDS("disease_overall_string.rds")
@@
 ```
-
-```{r}
-require(msigdbr)
-dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-    do({
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
-
-msig = readRDS("KEGG_String_Filt.rds")
-#msig =msig %>% filter(Category == "KEGG")
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:20],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<20) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_20.rds")
-}
-
-```
-
-```{r}
-
-```
-
-```{r}
-#KEGG
-
-kegg_directory = "./data/Disease_Genes/KEGG/"
-
-kegg_names = list.files(kegg_directory)
-
-kegg_list = paste0("./data/Disease_Genes/KEGG/", kegg_names)
-
-kegg_length = length(kegg_list)
-
-kegg_frames_string = lapply(1:kegg_length, function(x) {
-  diseases = diseases2frame(kegg_list[x], links = string)
-  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(kegg_list[x])))
-  return(list(diseases = diseases, lengths = per_no_ppi))
-  })
-
-kegg_overall_string = lapply(kegg_frames_string, function(x) x$diseases %>% mutate(Category = "KEGG"))
-
-disease_overall_string = bind_rows(disease_overall_string, kegg_overall_string) %>% distinct()
-```
-
-```{r}
-#Load in normalization regression
-
-pred_edges = readRDS("pred_edges.rds")
-```
-
-```{r}
-#Remove edges with less than 5 total counts, square root internal edges and correct by predicted
-
-disease_overall_string = disease_overall_string %>% group_by(Disease)  %>%  mutate(Pred_Edges = predict(pred_edges, data.frame(num_genes = n())), Internal = sqrt(Internal/Pred_Edges), External = sqrt(External)) %>% ungroup() %>% mutate(Disease = ifelse(Category == "Random", "Random", Disease)) %>% filter(Internal != 0) %>% group_by(Disease) %>% filter(n() > 4) %>% filter( (length(unique(Internal))>1) )
-
-#Make names more informative
-
-disease_overall_string =  disease_overall_string %>% ungroup() %>%  mutate(Disease = gsub("Run ", "Random ", Disease), Disease = gsub("hsa", "KEGG hsa", Disease))
-
-```
-
-#Plot of all diseases
-
-```{r}
-
-#Calculate the slopes for diseases and categories
-
-#Slope for categories
-
-#Slopes for categories string
-disease_overall_string = readRDS("disease_overall_string_filt.rds")
-
-slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
-    group_by(Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-#Slopes for diseases
-
-dis_slopes_string = disease_overall_string %>% filter(Internal != 0) %>% 
-    group_by(Disease, Category) %>% 
-    do({
-      mod = lm(External ~ Internal-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope)
-
-#Add p-values
-
-# dis_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Disease, Category) %>%
-#     do({
-#       perm = slope_perm(disease_overall_string, .$Disease[1], Norm_Reg =NULL, permutations = 1000)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-# 
-# cat_pvals_string = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>%
-#     group_by(Category) %>%
-#     do({
-#       perm = slope_perm(disease_overall_string, .$Category[1], Norm_Reg =NULL, permutations = 1000, Disease = FALSE)
-#       data.frame(P_val= perm)
-#     }) %>% arrange(P_val)
-
-#Add slopes to the overall dataset
-
-disease_overall_string = left_join(disease_overall_string, slopes_string, "Category")
-colnames(disease_overall_string)[7] = "Category_Slope"
-
-disease_overall_string = left_join(disease_overall_string, dis_slopes_string[,c(1,3)], "Disease")
-colnames(disease_overall_string)[8] = "Disease_Slope"
-
-# disease_overall_string = left_join(disease_overall_string, dis_pvals_string[,c(1,3)], "Disease")
-# colnames(disease_overall_string)[9] = "Slope_Pval"
-
-```
-
-```{r}
-#Adding filtered results
-
-cat_coherence = read_csv("./tables/supplementary_table_categories.csv")
-
-gene_count_cat_string = disease_overall_string %>% group_by(Category) %>% summarise(gene_count_string_filt = n())
-
-cat_coherence = full_join(gene_count_cat_string %>% filter(!(Category %in% c("KEGG", "Random")))  , cat_coherence, by = "Category")
-
-string_filt_cat_p = readRDS("string_p_filt_cat.rds") %>% as.data.frame(.) %>% dplyr::select( Category = i, `String Filt Perm P-val` = perm )
-
-cat_coherence = left_join(cat_coherence, string_filt_cat_p, by = "Category")
-
-# cat_coherence = full_join(cat_pvals_string %>% filter(!(Category %in% c("KEGG", "Random"))), cat_coherence, by = "Category")
-
-coherence_cat_string = slopes_string %>%  ungroup() %>% mutate(Min_Max = ((Slope - max(Slope))/(min(Slope) - max(Slope)))) %>% dplyr::select(Category, Min_Max) %>% filter(!(Category %in% c("KEGG", "Random"))) %>% distinct()
-
-cat_coherence = left_join(cat_coherence, coherence_cat_string, by = "Category")
-
-# colnames(cat_coherence)[2] = "String Filt Perm P-val"
-
-colnames(cat_coherence)[10] = "Coherence String Filt"
-
-#Rearrange to better order
-
-cat_coherence = cat_coherence %>% dplyr::select(Category,  gene_count_biogrid
-, gene_count_string, gene_count_string_filt, `Coherence String`, `Coherence Biogrid`, `Coherence String Filt`, `String Perm P-val`, `Biogrid Perm P-val`,  `String Filt Perm P-val`  )
-
-cat_coherence  = cat_coherence %>% mutate(`String Filt Perm P-val` = as.numeric(as.character(`String Filt Perm P-val`))) %>% arrange(Category)
-
-#cat_coherence[,2:10] = round(cat_coherence[,2:10], 3)
-cat_coherence[,2:10] = round(cat_coherence[,2:10], 3)
-
-write_csv(cat_coherence, "./tables/supplementary_table_categories.csv")
-
-```
-
-```{r}
-
-#KEGG and random should be same color on all plots
-#Order by slope
-
-library(RColorBrewer)
-
-#Set a custom color palette for both string and biogrid
-
-library(randomcoloR)
-
-#Set permanent colors for random and KEGG
-
-col =  c("Red", "#999999", distinctColorPalette(12))
-names(col) = c("Random", "KEGG", as.character(unique(disease_overall_string$Category))[!(unique(disease_overall_string$Category) %in% c("KEGG", "Random"))])
-
-col_full = scale_color_manual(values = col)
-
-category_plot_string = ggplot(disease_overall_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2)  + theme_light(base_size = 32) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + col_full + labs(color = "Category") +
-   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_blank(),
-        panel.background = element_blank())
-
-ggsave("./figures/category_plot_string_filtered.png", category_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of all categories (String)
-
-```{r}
-category_plot_string
-```
-
-```{r}
-category_string = disease_overall_string %>% filter((!grepl("trait", Category) & !grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random") %>% arrange(-Category_Slope)
-
-col_sub = col[names(col) %in% category_string$Category]
-
-cat_plot_string = ggplot(category_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub) + theme_light(base_size = 16) + theme(legend.position=c(.21,.7), legend.key.height = unit(.2, "cm")) + labs(color = "Category")
-
-ggsave("./figures/category_disease_plot_string_filtered.png", cat_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of non-trait categories (String)
-
-```{r}
-cat_plot_string
-```
-
-```{r}
-traits_string = disease_overall_string %>% filter((grepl("trait", Category) | grepl("Intelligence", Category)) | Category == "KEGG" | Category == "Random")
-
-col_sub_string = col[names(col) %in% traits_string$Category]
-
-trait_plot_string = ggplot(traits_string, aes(x = Internal, y = External, color = reorder(Category, -Category_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + scale_color_manual(values = col_sub_string) + theme_light(base_size = 24) + theme(legend.position=c(.25,.78), legend.key.height = unit(.2, "cm")) + labs(color = "Category") +
-   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.border = element_blank(),
-        panel.background = element_blank(), legend.text = element_text(size = 16))
-
-ggsave("./manuscript/Fig4_category_trait_plot_string_filtered.png",trait_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of traits (String)
-
-```{r}
-trait_plot_string
-```
-
-```{r}
-disease_anthro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Anthropometric trait" | Disease == "KEGG" | Disease == "Random"))
-
-col_dis_string = c(col[names(col) %in% disease_anthro_string$Disease], distinctColorPalette(length(unique(disease_anthro_string$Disease))))
-
-names(col_dis_string)[-c(1,2)] = as.character(unique(disease_anthro_string$Disease))
-
-anthro_plot_string = ggplot(disease_anthro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_dis_string) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/anthro_plot_string_filtered.png",anthro_plot_string, height=5, width=7, units="in", dpi = 300)
-```
-
-```{r}
-#Order and keep random and KEGG consistent from other plot
-
-disease_psych_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Psychiatric" | Disease == "KEGG" | Disease == "Random")) %>% group_by(Disease)
-
-#Plotting psych results for string
-
-col_dis = c(col[names(col) %in% disease_psych_string$Disease], distinctColorPalette(length(unique(disease_psych_string$Disease))))
-
-names(col_dis)[-c(1,2)] = as.character(unique(disease_psych_string$Disease))
-
-psych_plot_string = ggplot(disease_psych_string , aes(x = Internal, y = External, color = reorder(Disease, -Disease_Slope))) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = c(col_dis)) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm")) + labs(color = "Disease")
-
-ggsave("./figures/psych_plot_string_filtered.png",psych_plot_string, height=5, width=7, units="in", dpi = 300)
-
-```
-
-# Plot of psychiatric diseases (String)
-
-```{r}
-psych_plot_string
-```
-
-```{r}
-#Subset to include neurological disorders
-
-disease_neuro_string = disease_overall_string  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope)) %>% arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == "Neurologic" | Disease == "KEGG" | Disease == "Random"))
-
-#STRING
-
-#Setting colors
-
-col_neuro = c(col[names(col) %in% disease_neuro_string$Disease], distinctColorPalette(length(unique(disease_neuro_string$Disease))))
-
-names(col_neuro)[-c(1,2)] = as.character(unique(disease_neuro_string$Disease))
-
-#Plotting
-
-neuro_plot_string = ggplot(disease_neuro_string , aes(x = Internal, y = External, color = Disease)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = col_neuro) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-ggsave("./figures/neuro_plot_string_filtered.png", neuro_plot_string, height=5, width=7, units="in", dpi = 300)
-```
-
-# Plot of neuorologic diseases (String)
-
-```{r}
-neuro_plot_string
-```
-
-```{r}
-#Function to create all other plots
-
-dir.create("./figures/Category")
-
-plot_coherence = function(Cat, dataset, protein_database, filtered = FALSE) {
-
-  all_plots = list()
-  #Subset based on category
-
-sub_frame = dataset  %>% mutate(Category = as.character(Category), Disease = ifelse(Category %in% c("Random", "KEGG"), Category, Disease)) %>% group_by(Disease) %>% mutate(Disease_Slope = ifelse(Category %in% c("Random", "KEGG"), Category_Slope, Disease_Slope))  %>% mutate(Disease_Slope = ifelse(Category == "KEGG", 0, Disease_Slope)) %>%
- arrange(-Disease_Slope) %>% ungroup() %>% mutate(Disease = factor(Disease, levels = unique(Disease))) %>% filter( (Category == Cat | Disease == "KEGG" | Disease == "Random"))
-print(sub_frame)
-
-print(head(sub_frame))
-
-  #Get colors
-
-  color = c(col[names(col) %in% sub_frame$Disease], distinctColorPalette(length(unique(sub_frame$Disease))))
-
-names(color)[-c(1,2)] = as.character(unique(sub_frame$Disease))
-
-#Make plot
-
-dis_plot = ggplot(sub_frame , aes(x = Internal, y = External, color =Disease, -Disease_Slope)) + geom_point(alpha = .1) + geom_smooth( method = 'lm', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2) + theme_light(base_size = 18) + theme(legend.position=c(.2,.73)) + scale_color_manual(values = color) + theme_light(base_size = 16) + theme(legend.position=c(.28,.7), legend.key.height = unit(.02, "cm"))
-
-all_plots[[i]] = dis_plot
-
-if (filtered) {
-ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, "_filtered", ".png"), height=5, width=7, units="in", dpi = 300) 
-} else {
-  ggsave(paste0("./figures/Category/", Cat, "_plot_", protein_database, ".png"), height=5, width=7, units="in", dpi = 300) 
-}
-
-return(all_plots)
-}
-
-disease_overall_string = disease_overall_string %>% mutate(Genes = ifelse( (Category == "KEGG") & (Disease_Slope > min(  (disease_overall_string %>% filter(Category != "KEGG"))$Disease_Slope)), "Incoherent", Genes )) %>% filter(Genes != "Incoherent")
-
-#Making plots for each category
-
-disease_overall_string = disease_overall_string %>% arrange(as.character(Category))
-
-cat_plots = lapply(unique(disease_overall_string$Category), function(x) {
-  plot_coherence(x, dataset = disease_overall_string %>% ungroup() , protein_database = "string", filtered = TRUE)
-})
-
-saveRDS(cat_plots, "string_filt.rds")
-
-#saveRDS(unique(disease_overall_string$Category), "disease_list.rds")
-```
-
-```{r}
-#Addition of filtered p-value
-
-orig_tab = read_csv("./tables/supplementary_table_diseases.csv")
-# 
-# new_tab = left_join(orig_tab, dis_pvals_string %>% ungroup() %>% dplyr::select(-Category), by = "Disease")
-# colnames(new_tab)[13] = "String Filt Perm P-value"
-
-string_p_filt = as.data.frame(readRDS("string_p_filt.rds"))
-string_p_filt_psych = as.data.frame(readRDS("string_p_filt_psych.rds"))
-
-string_over = rbind(string_p_filt, string_p_filt_psych) %>% as.data.frame(.) %>% dplyr::select( Disease = i, `String Filt Perm P-value` = perm ) %>% filter(!duplicated(Disease))
-
-new_tab = left_join(orig_tab, string_over, "Disease") %>% mutate(`String Filt Perm P-value` = as.numeric(as.character(`String Filt Perm P-value`)))
-```
-
-```{r}
-#Addition of coherence for string filtered
-
-string_filt_diff = disease_overall_string %>% ungroup() %>% mutate(Min_Max = ((Disease_Slope - max(Disease_Slope))/(min(Disease_Slope) - max(Disease_Slope)))) %>% dplyr::select(Disease, Min_Max) %>% distinct()
-
-new_tab = left_join(new_tab,string_filt_diff, by = "Disease")
-colnames(new_tab)[14] = "String Filt Coherence"
-
-#Place with other coherence measures
-
-new_tab = new_tab[,c(1:10,14,11:13)]
-
-new_tab = new_tab %>% arrange(Category, Disease)
-
-# new_tab[,3:14] = round(new_tab[,3:14], 3)
-new_tab[,3:14] = round(new_tab[,3:14], 3)
-
-new_tab = new_tab %>% mutate(`String Filt Perm P-value` = ifelse(is.na(`String Coherence`), NA, `String Filt Perm P-value`))
-
-write.csv(new_tab, "./tables/supplementary_table_diseases.csv", row.names = FALSE)
-```
-
-```{r}
-#Category wise
-
-disease_cats_string = unique(disease_overall_string$Category)
-disease_cats_biogrid = unique(disease_overall$Category)
-
-for (i in disease_cats) {
-  gene_list = disease_overall_string
-}
-
-#Loop through diseases and pull out internal and external edges
-
-disease_frames_string = lapply(disease_cats, function(x) {
-
-  #Run function to find internal and external connectivity 
-  diseases = diseases2frame(disease_selected[x], links = string)
-
-  #Record ppi missing
-
-  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
-  return(list(diseases = diseases, lengths = per_no_ppi))
-  })
-
-```
-
-```{r}
-
-#Independent analysis of KEGG pathways
-
-kegg_slopes = dis_slopes_string %>% filter(Category == "KEGG")
-
-#Gsubing out KEGG prefix
-
-kegg_slopes = kegg_slopes %>% ungroup() %>% mutate(Disease = gsub("KEGG ", "", Disease))
-
-#Get names of KEGG pathways
-
-xx = data.frame(path_name = names(as.list(KEGGPATHNAME2ID)), kegg_code = unlist(as.list(KEGGPATHNAME2ID)))
-
-#Match name with code
-
-kegg_slopes = kegg_slopes %>% mutate(kegg_code = gsub("hsa", "", Disease))
-
-#Create table
-
-kegg_slopes = left_join(kegg_slopes, xx, "kegg_code" )
-
-kegg_frame = kegg_slopes %>% dplyr::select(hsa_code = Disease, Pathway = path_name, Slope) %>% mutate(Coherence = ((Slope - max(dis_slopes_string$Slope))/(min(dis_slopes_string$Slope) - max(dis_slopes_string$Slope))))
-
-#Read in other table and merge
-
-kegg_frame = kegg_frame %>% dplyr::select(hsa_code, String_Filt_Slope = Slope, String_Filt_Coherence = Coherence)
-
-kegg_counts_string = bind_rows(kegg_overall_string) %>% group_by(Disease) %>% summarise(String_Filt_Count = n()) %>% rename(hsa_code  = Disease)
-
-kegg_frame = left_join(kegg_frame, kegg_counts_string, "hsa_code")
-
-kegg_prev = read_csv("KEGG_Summary.csv")
-
-kegg_frame_full = left_join(kegg_frame, kegg_prev, "hsa_code")
-
-#Rearranging
-
-kegg_frame_full = kegg_frame_full %>% dplyr::select(hsa_code, Pathway, String_Filt_Count, String_Count, Biogrid_Count, String_Filt_Slope, String_Slope, Bio_Slope, String_Filt_Coherence, String_Coherence, Bio_Coherence) %>% arrange(-String_Filt_Coherence)
-
-write.csv(kegg_frame_full, "./tables/KEGG_Summary.csv", row.names = FALSE)
-```
-
-```{r}
-
-```

### 2019-08-20 Analysis/Generate_KEGG.Rmd
finish kegg/react/cc generation

---
title: "Generate_KEGG"
author: "Kellen Cresswell"
date: "August 19, 2019"
output: pdf_document
---

```{r}
require(msigdbr)
source("./functions/disease2frame_raw.R")
```

```{r}
#Read in msig data
msig_full = msigdbr()

sig_sub = msig_full %>% filter(gs_subcat %in% c("CP:REACTOME", "CP:KEGG", "CC")) %>% 
    group_by(gs_name, gs_subcat) %>% mutate(Count = n()) %>% group_by(gs_name, gs_subcat, Count) %>%
    do({
      sub = diseases2frame_raw(.$gene_symbol, string, .$gs_name[1])
      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = sub)
      data.frame(Slope= coef(mod)[1])
    })
```

```{r}

```

### 2019-08-20 Analysis/Msigdf_Biogrid.R

@@ -1,0 +1,46 @@
+source("./functions/disease2frame.R")
+
+require(readr)
+require(dplyr)
+require(biomaRt)
+require(msigdbr)
+require(igraph)
+
+#Reading in biogrid data
+
+links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+
+links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+
+#Split based on Biogrid ID system then pull out gene names
+
+protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+
+protein1 = unlist(protein1)
+protein2 = unlist(protein2)
+
+#Place into matrix in form that disease2frame works on
+
+links = links %>% mutate(protein1 =protein1, 
+                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
+
+links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
+#Getting msigs 
+
+msig = msigdbr()
+
+msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_bio.txt", links = string)
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = msig$gs_subcat[1])
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"KEGG_Biogrid.rds")
+}

### 2019-08-20 Analysis/Msigdf_String.R

@@ -1,0 +1,48 @@
+source("./functions/disease2frame.R")
+
+require(readr)
+require(dplyr)
+require(biomaRt)
+require(msigdbr)
+require(igraph)
+
+string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
+
+genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
+
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol.x)
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol.y)
+
+#Getting msigs 
+
+msig = msigdbr()
+
+msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = msig$gs_subcat[1])
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"KEGG_String.rds")
+}

### 2019-08-20 Analysis/Msigdf_String_Filt.R

@@ -1,0 +1,51 @@
+source("./functions/disease2frame.R")
+
+require(readr)
+require(dplyr)
+require(biomaRt)
+require(msigdbr)
+require(igraph)
+
+string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
+
+genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
+
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol.x)
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol.y)
+
+string = string %>% filter(combined_score>=500)
+
+
+#Getting msigs 
+
+msig = msigdbr()
+
+msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea.txt", links = string)
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = msig$gs_subcat[1])
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"KEGG_String_Filt.rds")
+  }

### 2019-08-20 Analysis/Msigdf_Biogrid.R
small typo fix

@@ -26,6 +26,7 @@ links = links %>% mutate(protein1 =protein1,
                          protein2 = protein2) %>% dplyr::select(protein1, protein2)
 
 links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
+
 #Getting msigs 
 
 msig = msigdbr()
@@ -36,7 +37,7 @@ mod_over = bind_rows()
 for (j in unique(msig$gs_name)) {
   curr_msig = msig %>% filter(gs_name == j)
   write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-  diseases_curr = diseases2frame("temp_rea_bio.txt", links = string)
+  diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,

### 2019-08-20 Analysis/Msigdf_Biogrid.R
fix category labeling

@@ -41,7 +41,7 @@ for (j in unique(msig$gs_name)) {
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = msig$gs_subcat[1])
+                       Category = curr_msig$gs_subcat[1])
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_Biogrid.rds")
 }

### 2019-08-20 Analysis/Msigdf_String.R

@@ -42,7 +42,7 @@ for (j in unique(msig$gs_name)) {
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = msig$gs_subcat[1])
+                       Category = curr_msig$gs_subcat[1])
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_String.rds")
 }

### 2019-08-20 Analysis/Msigdf_String_Filt.R

@@ -45,7 +45,7 @@ for (j in unique(msig$gs_name)) {
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = msig$gs_subcat[1])
+                       Category = curr_msig$gs_subcat[1])
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_String_Filt.rds")
   }

### 2019-08-20 Analysis/Coherence_Calculations_New.Rmd
update coherence

@@
 msig = readRDS("KEGG_String.rds")
-#msig =msig %>% filter(Category == "KEGG")
+msig =msig %>% filter(Category == "CP:KEGG")
 coherence_tab = bind_rows()
@@
 msig = readRDS("KEGG_String_Filt.rds")
-#msig =msig %>% filter(Category == "KEGG")
+msig =msig %>% filter(Category == "CP:KEGG")
 coherence_tab = bind_rows()
@@
 msig = readRDS("KEGG_Biogrid.rds")
-#msig =msig %>% filter(Category == "KEGG")
+msig =msig %>% filter(Category == "CP:KEGG")
 coherence_tab = bind_rows()
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
-msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
+msig = readRDS("KEGG_String.rds")
+msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
 coherence_tab = bind_rows()
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
-msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
+msig = readRDS("KEGG_String_Filt.rds")
+msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
 coherence_tab = bind_rows()
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
-msig =msig %>% filter(Category == "REACTOME") %>% filter(!is.na(Slope))
+msig = readRDS("KEGG_Biogrid.rds")
+msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
 coherence_tab = bind_rows()
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
+msig = readRDS("KEGG_String.rds")
 msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
+msig = readRDS("KEGG_String_Filt.rds")
 msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
@@
 ```{r}
-msig = readRDS("Msigdf.rds")
+msig = readRDS("KEGG_Biogrid.rds")
 msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
@@
 ```{r}
-#Random networks
-
-kegg_string = readRDS("msigdf.rds") %>% filter(Category == "KEGG")
-
-kegg_string_filt = readRDS("KEGG_String_Filt.rds")
-
-kegg_biogrid = readRDS("KEGG_String.rds")
-
-ke
-
-#String
-
-string_random = bind_rows()
-for (i in kegg_string$Count) {   
-  j = 0
-  diseases = bind_rows()
-    while(j<11) {
-    diseases_curr = GeneSample(i, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
-      string_random = bind_rows(string_random, diseases_curr)
-      print(string_random)
-}
-
-saveRDS(string_random, "string_random_edges.rds")
-
-#String Filtered
-
-string_random_filt = bind_rows()
-for (i in unique(kegg_string_filt$Count)) {   
-  j = 0
-    while(j<11) {
-    diseases_curr = GeneSample(i, string_filt, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-      diseases_curr = data.frame(Count = i, Slope = coef(mod)[1])
-      string_random_filt = bind_rows(string_random_filt, diseases_curr)
-      print(string_random_filt)
-
-    }
-    }
-}
-
-#String Filtered
-
-biogrid_random = bind_rows()
-for (i in unique(kegg_biogrid$Count)) {   
-  j = 0
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-      diseases_curr = data.frame(Count = i, Slope = coef(mod)[1])
-      string_random = bind_rows(string_random, diseases_curr)
-      print(string_random)
-
-    }
-    }
-}
-
-```
-
-```{r}
-string_random = string_random %>% group_by(Count) %>% summarise(Slope = median(Slope))
-cor(string_random)
-
-string_filt_random = string_filt_random %>% group_by(Count) %>% summarise(Slope = median(Slope))
-cor(string_filt_random)
-```
-
-```{r}
-msig_sum = readRDS("Msigdf.rds") %>% distinct() %>% filter(!grepl("_1", Category))
+msig_sum = readRDS("KEGG_String.rds") %>% distinct() %>% filter(!grepl("_1", Category))
 
@@
 ```
-
-```{r}
-#Testing coherence and modularity relationship
-
-#Binding disease select (Disease name) with modularity calculations
-
-mod_string = readRDS("Modularity_String.rds")
-mod_string_filt = readRDS("Modularity_String_Filt.rds")
-mod_biogrid = readRDS("Modularity_Biogrid.rds")
-
-mod_sum =data.frame(Disease = disease_select$Disease, Modularity_String = unlist(mod_string), Modularity_String_Filt = unlist(mod_string_filt), Modularity_Biogrid = unlist(mod_biogrid))
-
-Mod_Frame = left_join(mod_sum, coherence_disease, "Disease")
-Mod_Frame = Mod_Frame %>% dplyr::select(Disease, Category:`String Filt P-value`, `Modularity Biogrid` = Modularity_Biogrid, `Modularity String` = Modularity_String, `Modularity String Filt`= Modularity_String_Filt)
-```
-
-```{r}
-mod_sum = ggplot(Mod_Frame, aes(x = `String Coherence KEGG`, y = `Modularity_String`)) + geom_point() + labs(x = "String Coherence KEGG", y = "Modularity String")
-
-ggsave("./figures/Modularity_Coherence.png", mod_sum, height=5, width=7, units="in", dpi = 300)
-
-write_xlsx(Mod_Frame, "./tables/Mod_Table.xlsx")
-
-```

### 2019-08-20 manuscript.Rmd

@@
 
-**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (with exemplary high coherence) and random networks (the lower limit of coherence).
+**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (the standard of high coherence) and random networks (the lower limit of coherence).
 We observed a range of coherence estimates for each category of phenotypes.
@@
 **Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences.
-Our results highlight the gaps in of our current knowledge of the genetics and molecular mechanisms of complext traits and suggest priorities for future work.
+Our results highlight the gaps in of our current knowledge of the genetics and molecular mechanisms of complext phenotypes and suggest priorities for future work.
 
@@
 Second, unfiltered STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING full”).
-The advantage of using unfiltered data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
+The advantage of using unfiltered data is that more phenotype-associated genes will have PPI information and, hence, can be analyzed.
 The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs.
@@
 
-Trait-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa].
-"MAPPED_GENE" column was used to extract trait-associated genes.
-To organize traits into categories, Experimental Factor Ontology (EFO) IDs were extracted from the "MAPPED_TRAIT_URI" column.
+Phenotype-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa].
+"MAPPED_GENE" column was used to extract phenotype-associated genes.
+To organize phenotypes into categories, Experimental Factor Ontology (EFO) IDs were extracted from the "MAPPED_TRAIT_URI" column.
 Categories were mapped to EFO IDs using the `ontoCAT` v.1.12.0 R package and the EFO database (accessed 12/04/2018).
-In the case of a trait being mapped into two categories, the most representative category was manually assigned.
+In the case of a phenotype being mapped into two categories, the most representative category was manually assigned.
 Non-canonical gene names were converted to common gene symbols using `alias2Symbol` function from the `limma` R package.
@@
 
-Trait-associated gene networks were formed by mapping trait-associated genes to PPIs.
+Phenotype-associated gene networks were formed by mapping phenotype-associated genes to PPIs.
 Genes without PPI annotations (e.g., C5orf4, LOC100507462), antisense transcripts (e.g., CDKN2B-AS), readthrough (e.g., NPHP3-ACAD11), non-protein-coding transcripts (e.g., LINC00478) were excluded.
 Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
-Traits with less than five genes were omitted.
-Additionally, traits with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
+Phenotypes with less than five genes were omitted.
+Additionally, phenotypes with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
 external degree distributions (see below).
 These precautions were set to avoid situations where a small number of genes in a network can cause instability in the results.
-Importantly, the number of trait-associated genes having PPI annotations depends on the content of a PPI database; hence, some traits may be filtered when using one but not the other PPI database to build trait-associated networks.
+Importantly, the number of phenotype-associated genes having PPI annotations depends on the content of a PPI database; hence, some phenotypes may be filtered when using one but not the other PPI database to build phenotype-associated networks.
 
@@
 
-Trait-specific **internal degree distribution** was defined as the number of interactions between trait-associated genes, or, more precisely, their interacting protein products.
-As different traits have different numbers of associated genes forming networks of unknown configurations, the corresponding internal degree distribution was normalized to the expected interaction count for a network of that size (see Results).
+The **internal degree distribution** was defined as the number of interactions between genes of interest, or, more precisely, their interacting protein products.
+The **external degree distribution** was defined as the number of interactions between genes of interest and all other genes in the PPI network.
 
-Trait-specific **external degree distribution** was defined as the number of interactions between genes in a trait-associated network and all other genes in the PPI network.
-Being measured in the context of the entire PPI network of fixed size, the external degree distribution did not require normalization.
-
-To compare trait-specific degree distributions, the normalized internal vs.
-external degree distributions were plotted on a scatterplot.
-The square root of the degree was taken for more informative representation.
-For each trait, the internal vs.
+To compare phenotype-specific degree distributions, the internal vs.
+external degree distributions of phenotype-specific genes were plotted on a scatterplot.
+The square root of the degree was taken for more informative visualization.
+For each phenotype, the internal vs.
 external degree distribution plot was fit with a regression line through the origin to capture the slopes.
@@
 The slopes of the internal vs.
-external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database.
-Thus, we normalized the slopes for the trait-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
-For each trait $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
+external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database and the network size.
+To compare coherences using a given PPI database, we normalized the slopes for the phenotype-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
+For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
 Higher values indicate a higher level of coherence similar to that of KEGG pathways.
-These values represent normalized coherence and can be compared across traits and PPI databases.
+These values represent normalized coherence and can be compared across phenotypes and PPI databases.
 
@@
 
-To test whether each slope is significantly different than random, we used a permutation test.
-For trait $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
-We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for trait $i$.
+To test whether phenotype-specific coherence is significantly different than random, we used a permutation test.
+For phenotype $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
+We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
 For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.
@@
 We calculate the permutation p-value as $p_i = \frac{(\sum_{j=1}^nI({\beta_{rj} > \beta_i}) + 1}{n + 1}$ [@Phipson:2010aa].
-Intuitively, this approach estimates whether the slope of the trait-specific network $\beta_i$ is consistently larger than those of random networks.
+Intuitively, this approach estimates whether the slope of the phenotype-specific network $\beta_i$ is consistently larger than those of random networks.
 All analyses and visualizations were performed in the R/Bioconductor computing environment v.3.4.0 [@gentleman2006bioinformatics].
@@
 The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation.
-The authors would like to thank Spiro Stilianoudakis for help with Figure 1.
 
@@
 **Supplementary Figure S3.
-Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+Coherence estimates of phenotypes.** Size of dots represents the level of normalized coherence (X-axis) for individual phenotypes (Y-axis).
 Missing entries indicate that for a given PPI database a phenotype-associated network could not be built.

### 2019-08-22 Analysis/Msigdf_String.R
fix issue with NAs

@@ -24,9 +24,9 @@ string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "pro
 
 string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
 
-string = string %>% rename(hgnc_symbol_a = hgnc_symbol.x)
+string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
 
-string = string %>% rename(hgnc_symbol_b = hgnc_symbol.y)
+string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
 
 #Getting msigs 
 
@@ -39,6 +39,9 @@ for (j in unique(msig$gs_name)) {
   curr_msig = msig %>% filter(gs_name == j)
   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+  if (nrow(diseases_curr) == 0) {
+    next
+  }
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,

### 2019-08-22 Analysis/Msigdf_String_Filt.R

@@ -24,9 +24,9 @@ string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "pro
 
 string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
 
-string = string %>% rename(hgnc_symbol_a = hgnc_symbol.x)
+string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
 
-string = string %>% rename(hgnc_symbol_b = hgnc_symbol.y)
+string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
 
 string = string %>% filter(combined_score>=500)
 
@@ -38,14 +38,18 @@ msig = msigdbr()
 msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
 
 mod_over = bind_rows()
-for (j in unique(msig$gs_name)) {
+for (j in rev(unique(msig$gs_name))) {
   curr_msig = msig %>% filter(gs_name == j)
   write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
   diseases_curr = diseases2frame("temp_rea.txt", links = string)
+  if (nrow(diseases_curr) == 0) {
+    next
+  }
   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
   mod_slope = coef(mod)[[1]]
   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
                        Category = curr_msig$gs_subcat[1])
+  print(mod_sum)
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_String_Filt.rds")
   }

### 2019-08-23 Analysis/Coherence_Calculations_New.Rmd
fix dplyr issue
add fully integrated coherence calculations file

@@
 
-string = string %>% dpyr::rename(hgnc_symbol_b = hgnc_symbol)
+string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)
 
@@
   #Filter to only include specific disease coherence
-  dis = dis_slopes_string_count %>% filter(Disease == i)
-
+  dis = dis_slopes_string_count %>% ungroup() %>% filter(Disease == i)
   #Get closest KEGG pathways in term of gene counts
@@
       #Normalize based on random slope and the median of kegg slopes (More robust)
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
+      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
       coherence_tab = bind_rows(coherence, coherence_tab)
@@
 
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-    coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
+      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+    coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
       coherence_tab = bind_rows(coherence, coherence_tab)
@@
 
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )      
+      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
       coherence_tab = bind_rows(coherence, coherence_tab)
@@
 
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope), Rand_Norm = rand_slope )
+      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
 
@@
 
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope)-rand_slope)
+      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
@@
 coherence_disease = coherence_full
-
-#Read in old diseases/categories
-# old_res = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(-Coherence.Biogrid, -Coherence.String.full, -Coherence.String.filt)
-
-# #Read in old coherences
-# old_coh = read.csv("./manuscript/supplementary_table_S1.csv") %>% dplyr::select(Disease = Trait, Coherence.Biogrid, Coherence.String.full, Coherence.String.filt)
-
-#Bind old coherence to new for analysis
-#test_coherence = left_join(coherence_full, old_coh)
-
-#Replace trait with diseases for ease of combination
-#colnames(old_res)[1] = "Disease"
-
-#Join new and old results filtering NA SNPs (Non-shared traits)
-#coherence_full = left_join(coherence_full, old_res) %>% filter(!is.na(Total.number.of.SNPs))
-
-#Match original order
-#coherence_full = coherence_full %>% dplyr::select(Disease, Category, Total.number.of.SNPs, Total.number.of.genes, Biogrid_Coherence, String_Coherence, String_Filt_Coherence, P.value.Biogrid, P.value.String.full, P.value.String.filt)
-
-#Filter out forced . from read.csv
-#colnames(coherence_full) = gsub("\\.", " ", #colnames(coherence_full))
-
-#Replace _ with spaces for style
-#colnames(coherence_full) = gsub("\\_", " ", #colnames(coherence_full))
-
-#Rename to old first column name
-#coherence_full = coherence_full %>% dpyr::rename(Trait = Disease)
-
-#coherence_full[,5:7] = apply(coherence_full[,5:7], 2, function(x) ifelse(x<0, 0, ifelse (
-  #x>1, 1, x
-#)))
-
-#Add reactome data
-
-#coherence_full = left_join(coherence_full, coherence_biogrid_cc, by= c("Trait"= "Disease"))
-
-#coherence_full = left_join(coherence_full, coherence_string_cc, by= c("Trait"= "Disease"))
-
-#coherence_full = left_join(coherence_full, coherence_string_filt_cc, by= c("Trait"= "Disease"))
-
-#coherence_full = left_join(coherence_full, coherence_biogrid_react, by= c("Trait"= "Disease"))
-
-#coherence_full = left_join(coherence_full, coherence_string_react, by= c("Trait"= "Disease"))
-
-#coherence_full = left_join(coherence_full, coherence_string_filt_react, by= c("Trait"= "Disease"))
-
-#coherence_full = coherence_full %>% dpyr::rename(`Biogrid Coherence Reactome` = Coherence.x, `String Coherence Reactome` = Coherence.y, `String Filt Coherence Reactome` = Coherence.x.x, `Biogrid Coherence CC` = Coherence.y.y, `String Coherence CC` = Coherence.x.x.x, `String Filt Coherence CC` = Coherence.y.y.y)
-
-#Move p-values to end
-#coherence_full = coherence_full %>% dplyr::select(Trait:`String Filt Coherence`,`Biogrid Coherence Reactome`:`String Filt Coherence CC`)
-
-#Commented out old results for final version
-
-# coherence_full = left_join(coherence_full, old_coh, by = c("Trait" = "Disease"))
-# 
-# colnames(coherence_full)[14:16] = c("Biogrid Coherence Old", "String Coherence Old", "String Filt Coherence Old")
-# 
-# coherence_full = coherence_full[, c(1:10, 14,16,15, 11:13)]
-# 
-# coherence_no_big = coherence_full %>% filter(Trait != "Height") %>% filter(Trait != "Obesity-related traits")
 
@@
 ```{r}
-write.csv(coherence_full %>% arrange(`Category`, `Trait`), "Coherence_New.csv", row.names = FALSE)
-
-```
-
-```{r}
-msig_sum = readRDS("KEGG_String.rds") %>% distinct() %>% filter(!grepl("_1", Category))
-
-msig_slopes =msig_sum %>% group_by(Category) %>% summarise(`Mean Slope` = mean(Slope, na.rm = TRUE))
-
-write.csv(msig_slopes, "./tables/Category_Slopes.csv", row.names = FALSE, col.names = TRUE, quote = FALSE)
-```
-
-```{r}
 #Reading in data from cluster and binding
@@
 
+require(writexl)
 write_xlsx(coherence_disease, "./tables/supplementary_table_S1.xlsx")

### 2019-08-25 Analysis/GWAS_Prep.Rmd
uncomment disease genes

@@
 #
-# names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
-# 
-# #Loop through disease traits and print a gene list for each
-# 
-# dir.create('./data/Disease_Genes')
-# 
-# dis_dir = "./data/Disease_Genes/"
-# 
-# 
-# for (i in unique(names_genes$DISEASE.TRAIT)) {
-# genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
-# 
-# genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
-# 
-# genes_curr = unique(genes_curr)
-# 
-# write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
-# }
-# 
+
+names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
+
+#Loop through disease traits and print a gene list for each
+
+dir.create('./data/Disease_Genes')
+
+dis_dir = "./data/Disease_Genes/"
+
+for (i in unique(names_genes$DISEASE.TRAIT)) {
+genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
+
+genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
+
+genes_curr = unique(genes_curr)
+
+write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
+}
+
 # 

### 2019-08-25 Analysis/Coherence_Calculations_New.Rmd
point to new coherence path

@@
 
-```{r}
-#Repeating for string
-
-data_dir = "./data/"
-
-#Read in new string database
-
-string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
-
-string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                             protein2 = gsub("9606.", "", protein2))
-
-#Read in ensemble protein ID
-
-ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
-#Get gene set
-genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
-
-#Join gene names with string
-
-string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
-#Rename to work with our functions
-
-string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)
-
-#Repeat for second set of proteins
-
-string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-
-string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)
-
-string_filt = string %>% filter(combined_score>=500)
-
-dim(string)
-```
-
-```{r}
-#Reading in biogrid data
-
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
-
-#Select human protein-protein interactions
-
-links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
-
-#Split based on Biogrid ID system then pull out gene names
-
-protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-
-protein1 = unlist(protein1)
-protein2 = unlist(protein2)
-
-#Place into matrix in form that disease2frame works on
-
-links = links %>% mutate(protein1 =protein1, 
-                             protein2 = protein2) %>% dplyr::select(protein1, protein2)
-
-links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
-
-dim(links)
-```
-
-```{r}
-#Read in edges generated from Connectivity_Comparison_*.Rmd
-
-disease_overall_string = readRDS("string_edges_new.rds")
-
-disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
-
-disease_overall_biogrid = readRDS("biogrid_edges_new.rds")
-
-#Calculate slopes for all diseases under each category
-
-dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-    do({
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
-
-dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-    do({
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
-
-dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-    do({
-      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-      data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
-
-#Running coherence for string
-
-#Read in KEGG pathways for coherence
-msig = readRDS("KEGG_String.rds")
-msig =msig %>% filter(Category == "CP:KEGG")
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_count$Disease)) {
-  #Filter to only include specific disease coherence
-  dis = dis_slopes_string_count %>% ungroup() %>% filter(Disease == i)
-  #Get closest KEGG pathways in term of gene counts
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  #Take top 10 KEGG pathways
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  #Generate random networks
-  j = 1
-    while(j<11) {
-    #Sample gene networks equal in size to disease count
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    #Get random gene networks
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-
-    #Repeat if there are no internal edges
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-    #Calculate overall slope of random disease networks
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      #Normalize based on random slope and the median of kegg slopes (More robust)
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_String_Filt.rds")
-msig =msig %>% filter(Category == "CP:KEGG")
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_filt_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-    coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_filt.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_Biogrid.rds")
-msig =msig %>% filter(Category == "CP:KEGG")
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_biogrid_count$Disease)) {
-  print(i)
-  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_biogrid.rds")
-}
-
-```
-
-# Reactome
-
-```{r}
-msig = readRDS("KEGG_String.rds")
-msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_react.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_String_Filt.rds")
-msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_filt_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_filt_react.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_Biogrid.rds")
-msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_biogrid_count$Disease)) {
-  print(i)
-  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_biogrid_react.rds")
-}
-
-```
-
-# CC
-
-```{r}
-msig = readRDS("KEGG_String.rds")
-msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_CC.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_String_Filt.rds")
-msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_string_filt_count$Disease)) {
-  print(i)
-  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_string_filt_CC.rds")
-}
-
-```
-
-```{r}
-msig = readRDS("KEGG_Biogrid.rds")
-msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-coherence_tab = bind_rows()
-for (i in unique(dis_slopes_biogrid_count$Disease)) {
-  print(i)
-  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
-  diseases = bind_rows()
-  j = 1
-    while(j<11) {
-    diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-    if (any(diseases_curr$Internal !=0)) {
-      j = j+1
-      diseases = bind_rows(diseases, diseases_curr)
-
-    }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-      coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-
-      coherence_tab = bind_rows(coherence, coherence_tab)
-      saveRDS(coherence_tab, "coherence_biogrid_CC.rds")
-}
-
-```
+# ```{r}
+# #Repeating for string
+# 
+# data_dir = "./data/"
+# 
+# #Read in new string database
+# 
+# string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
+# 
+# string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+#                              protein2 = gsub("9606.", "", protein2))
+# 
+# #Read in ensemble protein ID
+# 
+# ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
+# 
+# #Get gene set
+# genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
+# 
+# 
+# #Join gene names with string
+# 
+# string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+# 
+# #Rename to work with our functions
+# 
+# string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)
+# 
+# #Repeat for second set of proteins
+# 
+# string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+# 
+# string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)
+# 
+# string_filt = string %>% filter(combined_score>=500)
+# 
+# dim(string)
+# ```
+# 
+# ```{r}
+# #Reading in biogrid data
+# 
+# links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+# 
+# #Select human protein-protein interactions
+# 
+# links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+# 
+# #Split based on Biogrid ID system then pull out gene names
+# 
+# protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+# protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+# 
+# protein1 = unlist(protein1)
+# protein2 = unlist(protein2)
+# 
+# #Place into matrix in form that disease2frame works on
+# 
+# links = links %>% mutate(protein1 =protein1, 
+#                              protein2 = protein2) %>% dplyr::select(protein1, protein2)
+# 
+# links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
+# 
+# 
+# dim(links)
+# ```
+# 
+# ```{r}
+# #Read in edges generated from Connectivity_Comparison_*.Rmd
+# 
+# disease_overall_string = readRDS("string_edges_new.rds")
+# 
+# disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
+# 
+# disease_overall_biogrid = readRDS("biogrid_edges_new.rds")
+# 
+# #Calculate slopes for all diseases under each category
+# 
+# dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
+#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+#     do({
+#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+#       data.frame(Slope= coef(mod)[1])
+#     }) %>% arrange(Slope) %>% filter(Count>4)
+# 
+# dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
+#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+#     do({
+#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+#       data.frame(Slope= coef(mod)[1])
+#     }) %>% arrange(Slope) %>% filter(Count>4)
+# 
+# dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
+#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+#     do({
+#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+#       data.frame(Slope= coef(mod)[1])
+#     }) %>% arrange(Slope) %>% filter(Count>4)
+# 
+# 
+# #Running coherence for string
+# 
+# #Read in KEGG pathways for coherence
+# msig = readRDS("KEGG_String.rds")
+# msig =msig %>% filter(Category == "CP:KEGG")
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_count$Disease)) {
+#   #Filter to only include specific disease coherence
+#   dis = dis_slopes_string_count %>% ungroup() %>% filter(Disease == i)
+#   #Get closest KEGG pathways in term of gene counts
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   #Take top 10 KEGG pathways
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   #Generate random networks
+#   j = 1
+#     while(j<11) {
+#     #Sample gene networks equal in size to disease count
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     #Get random gene networks
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     
+#     #Repeat if there are no internal edges
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#     #Calculate overall slope of random disease networks
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       #Normalize based on random slope and the median of kegg slopes (More robust)
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string.rds")
+# }
+#   
+# 
+# 
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_String_Filt.rds")
+# msig =msig %>% filter(Category == "CP:KEGG")
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_filt_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#     coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string_filt.rds")
+# }
+#   
+#   
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_Biogrid.rds")
+# msig =msig %>% filter(Category == "CP:KEGG")
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_biogrid_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_biogrid.rds")
+# }
+#   
+#   
+# ```
+# 
+# # Reactome
+# 
+# ```{r}
+# msig = readRDS("KEGG_String.rds")
+# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string_react.rds")
+# }
+#   
+# 
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_String_Filt.rds")
+# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_filt_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string_filt_react.rds")
+# }
+#   
+#   
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_Biogrid.rds")
+# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_biogrid_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_biogrid_react.rds")
+# }
+#   
+#   
+# ```
+# 
+# # CC
+# 
+# ```{r}
+# msig = readRDS("KEGG_String.rds")
+# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string_CC.rds")
+# }
+#   
+# 
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_String_Filt.rds")
+# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_filt_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_string_filt_CC.rds")
+# }
+#   
+#   
+# ```
+# 
+# ```{r}
+# msig = readRDS("KEGG_Biogrid.rds")
+# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_biogrid_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+#   
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+#       
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
+#       
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "coherence_biogrid_CC.rds")
+# }
+#   
+#   
+# ```
 
@@
 
-coherence_string_filt = readRDS("coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string = readRDS("coherence_string.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid = readRDS("coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string_filt = readRDS("./data/Coherence_Results/coherence_string_filt.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string = readRDS("./data/Coherence_Results/coherence_string.rds") %>% dplyr::select(Disease, Coherence)
+coherence_biogrid = readRDS("./data/Coherence_Results/coherence_biogrid.rds") %>% dplyr::select(Disease, Coherence)
 
@@
 
-coherence_string_filt_react = readRDS("coherence_string_filt_react.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string_react = readRDS("coherence_string_react.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid_react = readRDS("coherence_biogrid_react.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string_filt_react = readRDS("./data/Coherence_Results/coherence_string_filt_react.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string_react = readRDS("./data/Coherence_Results/coherence_string_react.rds") %>% dplyr::select(Disease, Coherence)
+coherence_biogrid_react = readRDS("./data/Coherence_Results/coherence_biogrid_react.rds") %>% dplyr::select(Disease, Coherence)
 
@@
 
-coherence_string_filt_cc = readRDS("coherence_string_filt_CC.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string_cc = readRDS("coherence_string_CC.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid_cc = readRDS("coherence_biogrid_CC.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string_filt_cc = readRDS("./data/Coherence_Results/coherence_string_filt_CC.rds") %>% dplyr::select(Disease, Coherence)
+coherence_string_cc = readRDS("./data/Coherence_Results/coherence_string_CC.rds") %>% dplyr::select(Disease, Coherence)
+coherence_biogrid_cc = readRDS("./data/Coherence_Results/coherence_biogrid_CC.rds") %>% dplyr::select(Disease, Coherence)
 

### 2019-08-25 Analysis/Coherence_Calculations_New.Rmd
various fixes

@@
 
-# ```{r}
+```{r}
 # #Repeating for string
@@
 # dim(string)
-# ```
 # 
-# ```{r}
+```
+
+```{r}
 # #Reading in biogrid data
@@
 # dim(links)
-# ```
 # 
-# ```{r}
+```
+
+```{r}
 # #Read in edges generated from Connectivity_Comparison_*.Rmd
@@
 #   
-# ```
 # 
-# ```{r}
+```
+
+```{r}
 # msig = readRDS("KEGG_Biogrid.rds")
@@
 #   
-# ```
 #
-# # Reactome
-# 
-# ```{r}
+```
+
+# Reactome
+
+```{r}
 # msig = readRDS("KEGG_String.rds")
@@
 # 
-# ```
 #
-# ```{r}
+```
+
+```{r}
 # msig = readRDS("KEGG_String_Filt.rds")
@@
 #   
-# ```
 # 
-# ```{r}
+```
+```{r}
 # msig = readRDS("KEGG_Biogrid.rds")
@@
 #   
-# ```
+# 
+```
 
@@
 
-disease_select = read_csv("supplementary_table_diseases_selected.csv") %>% dplyr::select(Disease =`Disease name`, `Number of SNPs`, `Number of genes`)
+disease_select = read_csv("./data/supplementary_table_diseases_selected.csv") %>% dplyr::select(Disease =`Disease name`, `Number of SNPs`, `Number of genes`)
 
@@
 
-require(writexl)
-write_xlsx(coherence_disease, "./tables/supplementary_table_S1.xlsx")
+dir.create("../manuscript/tables")
+write.csv(coherence_disease, ".././manuscript/tables/supplementary_table_S1.csv", 
+          row.names = FALSE)
 

### 2019-08-25 Analysis/Connectivity_Comparison_Filtered.Rmd
fix igraph issue

@@
 source("./functions/disease2frame.R")
-require(KEGGREST)
-require(KEGG.db)
 require(readr)
@@
 require(biomaRt)
-require(ggplot2)
 require(broom)
@@
 require(tidyr)
+require(igraph)
 ```
@@
 ```{r}
+
+#Get directory with diseases
+
+disease_directory = "./data/Disease_Genes/"
+
+disease_names = list.files(disease_directory)
+
+disease_list = paste0("./data/Disease_Genes/", disease_names, "/")
+
+#Read in manually selected diseases
+
+disease_select = read_csv("./data/supplementary_table_diseases_selected.csv")
+
+disease_traits = disease_select %>% dplyr::select(`Disease name`, Category)
+
+#Pull out disease and category information from select diseases
+
+#Change first column to make more clear
+
+colnames(disease_select)[1] = "Disease"
+
+#Filtering out NA in last row
+
+disease_select = disease_select %>% filter(!is.na(Disease))
+
+#Creating a vector of file names from disease select
+
+disease_files =  unique(paste0(disease_select$Disease, ".txt"))
+
+#Make full path
+
+disease_selected = paste0("./data/Disease_Genes/", disease_files)
+
+```
+```
+
+```{r}
 #Diseases
@@
 
-# disease_mod_string_filt = lapply(disease_length, function(x) {
-# 
-#   #Run function to find internal and external connectivity 
-#   mod = diseases2mod(disease_selected[x], links = string)
-#   
-# })
-# 
-# saveRDS(disease_mod_string_filt, "Modularity_String_Filt.rds")
-
 #Pull out network information for overall data frame

### 2019-08-25 Analysis/Coherence_Calculations_New.Rmd
uncomment slope count code

@@
 ```{r}
-# #Read in edges generated from Connectivity_Comparison_*.Rmd
-# 
-# disease_overall_string = readRDS("string_edges_new.rds")
-# 
-# disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
-# 
-# disease_overall_biogrid = readRDS("biogrid_edges_new.rds")
-# 
-# #Calculate slopes for all diseases under each category
-# 
-# dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-#     do({
-#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-#       data.frame(Slope= coef(mod)[1])
-#     }) %>% arrange(Slope) %>% filter(Count>4)
-# 
-# dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-#     do({
-#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-#       data.frame(Slope= coef(mod)[1])
-#     }) %>% arrange(Slope) %>% filter(Count>4)
-# 
-# dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>% 
-#     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-#     do({
-#       mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-#       data.frame(Slope= coef(mod)[1])
-#     }) %>% arrange(Slope) %>% filter(Count>4)
-# 
-# 
+#Read in edges generated from Connectivity_Comparison_*.Rmd
+
+disease_overall_string = readRDS("string_edges_new.rds")
+
+disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
+
+disease_overall_biogrid = readRDS("biogrid_edges_new.rds")
+
+#Calculate slopes for all diseases under each category
+
+dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+    do({
+      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+      data.frame(Slope= coef(mod)[1])
+    }) %>% arrange(Slope) %>% filter(Count>4)
+
+dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+    do({
+      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+      data.frame(Slope= coef(mod)[1])
+    }) %>% arrange(Slope) %>% filter(Count>4)
+
+dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+    do({
+      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+      data.frame(Slope= coef(mod)[1])
+    }) %>% arrange(Slope) %>% filter(Count>4)
+
 # #Running coherence for string

### 2019-08-25 Analysis/data/get_data.sh
gzip BIOGRID and cleanup

@@ -1,4 +1,5 @@
 wget https://stringdb-static.org/download/protein.links.v11.0/9606.protein.links.v11.0.txt.gz
 wget https://downloads.thebiogrid.org/Download/BioGRID/Release-Archive/BIOGRID-3.5.174/BIOGRID-ALL-3.5.174.mitab.zip
 unzip BIOGRID-ALL-3.5.174.mitab.zip
-
+gzip BIOGRID-ALL-3.5.174.mitab.txt
+rm BIOGRID-ALL-3.5.174.mitab.zip

### 2019-08-25 Analysis/Connectivity_Comparison.Rmd
Unify unfiltered and filtered connectivity comparison

--- 2019-08-19 Analysis/Connectivity_Comparison_Undirected.Rmd
@@
 
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
 
@@
 
-disease_select = read_csv("./supplementary_table_diseases_selected.csv")
+disease_select = read_csv("./data/supplementary_table_diseases_selected.csv")
 
@@
 ```
+
+```{r}
+string = string %>% filter(combined_score>=500)
+```
+
+```{r}
+#Diseases
+
+disease_length = 1:length(disease_selected)
+
+#Loop through diseases and pull out internal and external edges
+
+disease_frames_string = lapply(disease_length, function(x) {
+
+  #Run function to find internal and external connectivity 
+  diseases = diseases2frame(disease_selected[x], links = string)
+
+  #Record ppi missing
+
+  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
+  return(list(diseases = diseases, lengths = per_no_ppi))
+  })
+
+#Pull out network information for overall data frame
+
+disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)
+
+#Convert to data frame
+
+disease_overall_string = bind_rows(disease_protein_string)
+
+disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")
+
+#disease_overall_string = #readRDS("disease_overall_string.rds")
+
+saveRDS(disease_overall_string, "string_filt_edges_new.rds")
+```

### 2019-08-25 Analysis/Coherence_Calculations_New.Rmd
change string to links for biogrid

@@
 # dim(links)
-# 
+
 ```
@@
 #     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
 #     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
+#       print(coherence_tab)
 #       saveRDS(coherence_tab, "coherence_biogrid.rds")
 # }
-#   
-#   
-#
+
 ```
@@
 ```
+
 ```{r}
@@
 # 
-# ```
-# 
-# # CC
-# 
-# ```{r}
+```
+
+# CC
+
+```{r}
 # msig = readRDS("KEGG_String.rds")
@@
 #   
-# ```
-# 
-# ```{r}
+```
+
+```{r}
 # msig = readRDS("KEGG_Biogrid.rds")
@@
 # }
-#   
-#   
-# 
 ```

### 2019-08-26 Analysis/Coherence_Calculations_New.Rmd
string->string_filt

@@
 #     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
+#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
 #     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
@@
 # }
-#   
-#   
-# 
+
 ```
@@
 #
-#   
+
 ```

### 2019-08-26 Analysis/Category_Tables.Rmd
add code for Table 1 and S1

---
title: "Category_Tables"
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

```{r}
require(dplyr)
```

```{r}
#Creating the category table

Disease_Sum = read.csv("../manuscript/tables/supplementary_table_S1.csv")

#Summarizinig Categories

Disease_String_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Coherence.KEGG) %>% filter(!is.na(String.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence)

Disease_String_Filt_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Filt.Coherence.KEGG) %>% filter(!is.na(String.Filt.Coherence.KEGG))   %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Filt.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Filt.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Filt.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Filt.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence)

Disease_Biogrid_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, Biogrid.Coherence.KEGG) %>% filter(!is.na(Biogrid.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(Biogrid.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(Biogrid.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(Biogrid.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(Biogrid.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence) %>% filter(Total_Networks != 0)

#Formatting

#String

Disease_String_Sum = Disease_String_Sum %>% rename(`Minimum Coherence` = Min_Coherence, `Maximum Coherence` = Max_Coherence, `Mean Coherence` = Mean_Coherence, `Median Coherence` =  Median_Coherence, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Total Networks` = Total_Networks)

Disease_String_Sum = Disease_String_Sum %>% mutate(`Minimum Coherence` = format(`Minimum Coherence`, digits = 3)) %>% mutate(`Maximum Coherence` = format(`Maximum Coherence`, digits = 3)) %>% mutate(`Mean Coherence` = format(`Mean Coherence`, digits = 3)) %>% mutate(`Median Coherence` = format(`Median Coherence`, digits = 3)) %>% mutate(`Mean # Genes` = format(`Mean # Genes`, digits = 4)) %>% mutate(`Mean # SNPs` = format(`Mean # SNPs`, digits = 4))

#String Filtered

Disease_String_Filt_Sum = Disease_String_Filt_Sum %>% rename(`Minimum Coherence` = Min_Coherence, `Maximum Coherence` = Max_Coherence, `Mean Coherence` = Mean_Coherence, `Median Coherence` =  Median_Coherence, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Total Networks` = Total_Networks)

Disease_String_Filt_Sum = Disease_String_Filt_Sum %>% mutate(`Minimum Coherence` = format(`Minimum Coherence`, digits = 3)) %>% mutate(`Maximum Coherence` = format(`Maximum Coherence`, digits = 3)) %>% mutate(`Mean Coherence` = format(`Mean Coherence`, digits = 3)) %>% mutate(`Median Coherence` = format(`Median Coherence`, digits = 3)) %>% mutate(`Mean # Genes` = format(`Mean # Genes`, digits = 4)) %>% mutate(`Mean # SNPs` = format(`Mean # SNPs`, digits = 4))

#Biogrid

Disease_Biogrid_Sum = Disease_Biogrid_Sum %>% rename(`Minimum Coherence` = Min_Coherence, `Maximum Coherence` = Max_Coherence, `Mean Coherence` = Mean_Coherence, `Median Coherence` =  Median_Coherence, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Minimum # SNPs` = Min_SNPs, `Maximum # SNPs` = Max_SNPs, `Mean # SNPs` = Mean_SNPs, `Median # SNPs` = Median_SNPs, `Total Networks` = Total_Networks)

Disease_Biogrid_Sum = Disease_Biogrid_Sum %>% mutate(`Minimum Coherence` = format(`Minimum Coherence`, digits = 3)) %>% mutate(`Maximum Coherence` = format(`Maximum Coherence`, digits = 3)) %>% mutate(`Mean Coherence` = format(`Mean Coherence`, digits = 3)) %>% mutate(`Median Coherence` = format(`Median Coherence`, digits = 3)) %>% mutate(`Mean # Genes` = format(`Mean # Genes`, digits = 4)) %>% mutate(`Mean # SNPs` = format(`Mean # SNPs`, digits = 4))

#Writing to workbook

row.names(Disease_String_Filt_Sum) = NULL
row.names(Disease_String_Sum) = NULL
row.names(Disease_Biogrid_Sum) = NULL

require(xlsx)
wb = createWorkbook()
sheet  = createSheet(wb, sheetName="STRING")
addDataFrame(Disease_String_Sum, sheet)
sheet  = createSheet(wb, sheetName="STRING Filtered")
addDataFrame(Disease_String_Filt_Sum, sheet)
sheet  = createSheet(wb, sheetName="Biogrid")
addDataFrame(Disease_Biogrid_Sum, sheet)

saveWorkbook(wb, "../manuscript/tables/Supplementary_Table_S1.xlsx")

```

```{r}
#In text version

Cat_Within = Disease_Sum %>% group_by(Category) %>% summarise(`Number of SNPs`= sum(`Number.of.SNPs`), `Number of Genes` = sum(Number.of.genes), `Total Diseases` = n(),  `Mean Biogrid Coherence` =mean(Biogrid.Coherence.KEGG, na.rm = TRUE), `Mean String Coherence` = mean(String.Coherence.KEGG, na.rm = TRUE), `Mean String Filtered Coherence` = mean(String.Filt.Coherence.KEGG, na.rm = TRUE)) %>% arrange(Category)

write.csv(Cat_Within, "../manuscript/tables/Table_1.csv", row.names = FALSE)
```

### 2019-08-26 Analysis/Category_Tables.Rmd
Convert tables to match text

@@
 
-Disease_Sum = read.csv("../manuscript/tables/supplementary_table_S1.csv")
+Disease_Sum = read.csv("../manuscript/tables/supplementary_table_S3.csv")
 
@@
 
-saveWorkbook(wb, "../manuscript/tables/Supplementary_Table_S1.xlsx")
+saveWorkbook(wb, "../manuscript/tables/supplementary_table_S1.xlsx")
 
@@
 
-write.csv(Cat_Within, "../manuscript/tables/Table_1.csv", row.names = FALSE)
+write.csv(Cat_Within, "../manuscript/tables/table_1.csv", row.names = FALSE)
 ```

### 2019-08-26 manuscript/Analysis_Tables_New.Rmd
Convert tables to match text

@@
 
-fileNameIn0  <- "tables/supplementary_table_S1.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
-# iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S1.csv > tmp.csv
+fileNameIn0  <- "tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
+# iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S3.csv > tmp.csv
 mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()

### 2019-08-26 Analysis/Msigdf_Biogrid.R
add code for all msig cats for tab 2

@@ -45,3 +45,21 @@ for (j in unique(msig$gs_name)) {
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_Biogrid.rds")
 }
+
+#Redoing with every category at once
+
+msig = msigdbr()
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = curr_msig$gs_subcat[1])
+  print(mod_sum)
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"All_Msig_Biogrid.rds")
+}

### 2019-08-26 Analysis/Msigdf_String.R

@@ -49,3 +49,24 @@ for (j in unique(msig$gs_name)) {
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_String.rds")
 }
+
+#Redoing with every category at once
+
+msig = msigdbr()
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+  if (nrow(diseases_curr) == 0) {
+    next
+  }
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = curr_msig$gs_subcat[1])
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"All_Msig_String.rds")
+}
+

### 2019-08-26 Analysis/Msigdf_String_Filt.R

@@ -53,3 +53,24 @@ for (j in rev(unique(msig$gs_name))) {
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"KEGG_String_Filt.rds")
   }
+
+#Redoing with every category at once
+
+msig = msigdbr()
+
+mod_over = bind_rows()
+for (j in rev(unique(msig$gs_name))) {
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea.txt", links = string)
+  if (nrow(diseases_curr) == 0) {
+    next
+  }
+  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+  mod_slope = coef(mod)[[1]]
+  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+                       Category = curr_msig$gs_subcat[1])
+  print(mod_sum)
+  mod_over = bind_rows(mod_over, mod_sum)
+  saveRDS(mod_over,"All_Msig_String_Filt.rds")
+}

### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
Convert tables to match text
convert back to csv
add in random and full string

@@
 # 
-# 
 # ```
-# 
+
+#STRING All Msig
+
+# ```{r}
+# for (j in unique(msig$gs_name)) {
+#   curr_msig = msig %>% filter(gs_name == j)
+#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+#   if (nrow(diseases_curr) == 0) {
+#     next
+#   }
+#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+#   mod_slope = coef(mod)[[1]]
+#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+#                        Category = curr_msig$gs_subcat[1])
+#   mod_over = bind_rows(mod_over, mod_sum)
+#   saveRDS(mod_over,"All_Msig_String.rds")
+# }
+# ```
+
 # ```{r}
@@
 ```{r}
+#Random networks
+
+# #Read in KEGG pathways for coherence
+# msig = readRDS("KEGG_String.rds")
+# msig = msig %>% filter(Category == "CP:KEGG")
+# string_random = bind_rows()
+# for (i in msig$Count) {   
+#   j = 0
+#   diseases = bind_rows()
+#     while(j<11) {
+#     diseases_curr = GeneSample(i, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
+#       string_random = bind_rows(string_random, diseases_curr)
+#       print(string_random)
+#       saveRDS(string_random, "string_random_edges.rds")
+# }
+
+```
+
+```{r}
 #Read in results generated above
@@
 dir.create("../manuscript/tables")
-write.csv(coherence_disease, ".././manuscript/tables/supplementary_table_S1.csv", 
+write.csv(coherence_disease, ".././manuscript/tables/supplementary_table_S3.csv", 
           row.names = FALSE)

### 2019-08-27 Analysis/GWAS_Prep.Rmd
Delete unnecessary table
Make release quality

--- 2019-08-19 Analysis/GWAS_Prep.Rmd
@@
 ```{r}
-#Read in data and packages
-
-library(gwascat)
-library(limma)
-library(dplyr)
-library(DO.db)
-library(DBI)
-library(MDmisc)
-library(ontoCAT)
-library(rowr)
-require(readr)
-
-data(ebicat37)
+# Read in data and packages
+# install.packages("Biocmanager")
+library(gwascat) # BiocManager::install("gwascat")
+library(limma)   # BiocManager::install("limma")
+library(dplyr)   # install.packages("dplyr")
+library(DO.db)   # BiocManager::install("DO.db")
+library(DBI)     # BiocManager::install("DBI")
+source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/unembed.R") # library(MDmisc)  
+# install.packages("rJava")
+# Download https://www.bioconductor.org/packages/2.10/bioc/src/contrib/ontoCAT_1.8.0.tar.gz
+# Install "ontoCAT" manually from the file
+library(ontoCAT) # BiocManager::install("ontoCAT")
+library(rowr)    # install.packages("rowr")
+library(readr)   # install.packages("readr")
 ```
@@
 #Initial summary
+data(ebicat37)
 
 #Convert to data frame
-
 ebi_dat = as.data.frame(mcols(ebicat37))
@@
 #Put comma seperated genes on different lines
-
 ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
@@
 #Remove extra whitespace from unmebed
-
 ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)
@@
 #Summarise
-
 init_sum = ebi_dat %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% summarise(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE)) - sum(unique(MAPPED_GENE =="")))
-
-write.csv(init_sum, "supplementary_table_diseases_raw.csv", row.names = FALSE)
-
 ```
@@
 #Converting aliases to gene names
-
 alias2SymbolEbicat = function(ebicat37) {
@@
   #Unembed comma and - seperated values
-
   ebi_dat = unembed(ebi_dat, "MAPPED_GENE", sep = ",")
@@
   #Remove extra spaces
-
   ebi_dat$MAPPED_GENE = gsub(" ", "", ebi_dat$MAPPED_GENE)
@@
   #Summarise the results
-
   #Keeping track of changes (sapply necessary to keep placeholders for unmapped genes)
-
   recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))
-
   recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))
-
   change = (ebi_dat$MAPPED_GENE == recomb_genes)
-
   sum_vec = ifelse(change == TRUE, "Same", ifelse(change == FALSE & recomb_genes == "", "No Match", "Converted"))
-
   ebi_dat$MAPPED_GENE = recomb_genes
-
   return(ebi_dat)
@@
 ebi_dat = alias2SymbolEbicat(ebicat37)
-```
 
-```{r}
 #Pulling out useful info
-
 main_inf = ebi_dat %>% dplyr::select(MAPPED_GENE, SNPS, CHR_ID, CHR_POS, INITIAL.SAMPLE.DESCRIPTION, MAPPED_TRAIT, MAPPED_TRAIT_URI, DISEASE.TRAIT)
-
-#Updated EFO table
 
@@
 # 
-# #Collapse genes and disease names for simplifcation purposes and then get unique combinations
-# 
-# names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
-# 
-# #Loop through disease traits and print a gene list for each
-# 
-# dir.create('./data/Disease_Genes')
-# 
-# dis_dir = "./data/Disease_Genes/"
-# 
-# 
-# for (i in unique(names_genes$DISEASE.TRAIT)) {
-# genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
-# 
-# genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
-# 
-# genes_curr = unique(genes_curr)
-# 
-# write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
-# }
-# 
-# 
+```
+
+```{r}
+#Collapse genes and disease names for simplifcation purposes and then get unique combinations
+names_genes = main_inf %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% distinct()
+
+#Loop through disease traits and print a gene list for each
+dir.create('./data/Disease_Genes')
+dis_dir = "./data/Disease_Genes/"
+
+for (i in unique(names_genes$DISEASE.TRAIT)) {
+  genes_curr = names_genes %>% filter(DISEASE.TRAIT == i & MAPPED_GENE != "") %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE) %>% mutate(DISEASE.TRAIT = gsub("\\/", "-", DISEASE.TRAIT), DISEASE.TRAIT = gsub("\\*", "-", DISEASE.TRAIT), File_Path =  DISEASE.TRAIT[1]) %>% ungroup() %>% dplyr::select(DISEASE.TRAIT, MAPPED_GENE, File_Path) %>% distinct()
+  genes_disease = genes_curr %>% ungroup() %>% dplyr::select(MAPPED_GENE) %>% distinct()
+  genes_curr = unique(genes_curr)
+  write.table(genes_disease, paste0(dis_dir, substr(genes_curr$File_Path[1],1,100), ".txt"), row.names = FALSE, col.names = FALSE)
+}
+```
+
+```{r}
 # #Create map of EFO ids to terms

### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
fix paths in commented sections

--- 2019-08-26 Analysis/Coherence_Calculations_New.Rmd
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string.rds")
 # }
@@
 # 
-# 
 # ```
-# 
+
+#STRING All Msig
+
+# ```{r}
+# for (j in unique(msig$gs_name)) {
+#   curr_msig = msig %>% filter(gs_name == j)
+#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+#   if (nrow(diseases_curr) == 0) {
+#     next
+#   }
+#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+#   mod_slope = coef(mod)[[1]]
+#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+#                        Category = curr_msig$gs_subcat[1])
+#   mod_over = bind_rows(mod_over, mod_sum)
+#   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
+# }
+# ```
+
 # ```{r}
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string_filt.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
 # }
@@
 #       print(coherence_tab)
-#       saveRDS(coherence_tab, "coherence_biogrid.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string_react.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_react.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string_filt_react.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_react.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_biogrid_react.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_react.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string_CC.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_CC.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_string_filt_CC.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_CC.rds")
 # }
@@
 #       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "coherence_biogrid_CC.rds")
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_CC.rds")
 # }
+```
+
+```{r}
+#Random networks
+
+# #Read in KEGG pathways for coherence
+# msig = readRDS("KEGG_String.rds")
+# msig = msig %>% filter(Category == "CP:KEGG")
+# string_random = bind_rows()
+# for (i in msig$Count) {   
+#   j = 0
+#   diseases = bind_rows()
+#     while(j<11) {
+#     diseases_curr = GeneSample(i, string, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+#   
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
+#       string_random = bind_rows(string_random, diseases_curr)
+#       print(string_random)
+#       saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
+# }
+
 ```

### 2019-08-27 Analysis/Category_Tables.Rmd
add table 2

@@
 
-#Summarizinig Categories
+#Summarizing Categories
 
-Disease_String_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Coherence.KEGG) %>% filter(!is.na(String.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence)
+Disease_String_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Coherence.KEGG) %>% filter(!is.na(String.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
-Disease_String_Filt_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Filt.Coherence.KEGG) %>% filter(!is.na(String.Filt.Coherence.KEGG))   %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Filt.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Filt.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Filt.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Filt.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence)
+Disease_String_Filt_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Filt.Coherence.KEGG) %>% filter(!is.na(String.Filt.Coherence.KEGG))   %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Filt.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Filt.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Filt.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Filt.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
-Disease_Biogrid_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, Biogrid.Coherence.KEGG) %>% filter(!is.na(Biogrid.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(Biogrid.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(Biogrid.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(Biogrid.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(Biogrid.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes), Max_Genes = max(Number.of.genes), Mean_Genes = mean(Number.of.genes), Median_Genes = median(Number.of.genes) , Min_SNPs = min(Number.of.SNPs), Max_SNPs = max(Number.of.SNPs), Mean_SNPs = mean(Number.of.SNPs), Median_SNPs = median(Number.of.SNPs), Total_Networks = n()) %>% arrange(-Median_Coherence) %>% filter(Total_Networks != 0)
+Disease_Biogrid_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, Biogrid.Coherence.KEGG) %>% filter(!is.na(Biogrid.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(Biogrid.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(Biogrid.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(Biogrid.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(Biogrid.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
@@
 ```
+
+```{r}
+msig = readRDS("./data/Coherence_Results/All_Msig_String.rds")
+
+#Correlation betweeen categories and size
+
+msig_cor = msig %>% group_by(Category) %>% summarise(Slope_Cor = cor(Count, Slope, "complete")) %>% arrange(Slope_Cor)
+
+colnames(msig_cor)[2] = "Size Correlation"
+
+```
+
+```{r}
+options(scipen = 2)
+
+msig_sum = msig %>% group_by(Category) %>% summarise(Min_Slope = min(Slope, na.rm = TRUE), Max_Slope = max(Slope, na.rm = TRUE), Mean_Slope = mean(Slope, na.rm = TRUE), Median_Slope = median(Slope, na.rm = TRUE), Min_Genes = min(Count), Max_Genes = max(Count), Mean_Genes = mean(Count), Median_Genes = median(Count), Total_Sets = n()) %>% arrange(Mean_Slope)
+
+msig_sum = left_join(msig_sum, msig_cor)
+
+#Adding in random slope
+
+rand_string = readRDS("./data/Coherence_Results/string_random_edges.rds") %>% mutate(Category = "Random") %>% dplyr::select(Category, Count, Slope)
+
+rand_sum = rand_string %>% group_by(Category) %>% summarise(Min_Slope = min(Slope, na.rm = TRUE), Max_Slope = max(Slope, na.rm = TRUE), Mean_Slope = mean(Slope, na.rm = TRUE), Median_Slope = median(Slope, na.rm = TRUE), Min_Genes = min(Count), Max_Genes = max(Count), Mean_Genes = mean(Count), Median_Genes = median(Count), Total_Sets = n(),) %>% arrange(Mean_Slope)
+
+rand_sum = rand_sum %>% mutate(`Size Correlation` = cor(rand_string$Count, rand_string$Slope))
+
+msig_sum = bind_rows(msig_sum, rand_sum)
+
+msig_sum = msig_sum %>% rename(`Minimum Slope` = Min_Slope, `Maximum Slope` = Max_Slope, `Mean Slope` = Mean_Slope, `Median Slope` =  Median_Slope, `Minimum # Genes` = Min_Genes, `Maximum # Genes` = Max_Genes, `Mean # Genes` = Mean_Genes, `Median # Genes` = Median_Genes, `Total Networks` = Total_Sets)
+
+msig_sum = msig_sum %>% mutate(`Minimum Slope` = format(`Minimum Slope`, digits = 3)) %>% mutate(`Maximum Slope` = format(`Maximum Slope`, digits = 3)) %>% mutate(`Mean Slope` = format(`Mean Slope`, digits = 3)) %>% mutate(`Median Slope` = format(`Median Slope`, digits = 3)) %>% mutate(`Mean # Genes` = format(`Mean # Genes`, digits = 4)) %>% mutate(`Size Correlation` = format(`Size Correlation`, digits = 2))
+
+require(writexl)
+
+write_xlsx(msig_sum, "../manuscript/tables/supplementary_table_S2.xlsx")
+
+```

### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
wip to make release ready

--- 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
@@
 ```{r}
-require(msigdbr)
-require(dplyr)
-require(readr)
+library(msigdbr) # install.packages("msigdbr")
+library(dplyr)   # install.packages("dplyr")
+library(readr)   # install.packages("readr")
 source("./functions/GeneSample.R")
-require(biomaRt)
-
+library(biomaRt) # BiocManager::install("biomaRt")
 ```
@@
 # dim(links)
-
 ```
@@
 #Read in edges generated from Connectivity_Comparison_*.Rmd
-
-disease_overall_string = readRDS("string_edges_new.rds")
-
-disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
-
-disease_overall_biogrid = readRDS("biogrid_edges_new.rds")
+disease_overall_string      <- readRDS("string_edges_new.rds")
+disease_overall_string_filt <- readRDS("string_filt_edges_new.rds")
+disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")
 
 #Calculate slopes for all diseases under each category
-
 dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%

### 2019-08-27 Analysis/Category_Tables.Rmd

@@
 ```
+
+```{r}
+#Creating plot of slopes from tables
+
+#Read in counts
+
+kegg_string  = msig %>% filter(Category == "CP:KEGG") %>% dplyr::select(Count, Slope)
+
+Slope_Sum = bind_rows(rand_string %>% mutate(Method = "Random"), kegg_string %>% dplyr::select(Count, Slope) %>% mutate(Method = "KEGG"))
+```
+
+```{r}
+require(ggplot2)
+
+#Plotting
+
+slope_plot = ggplot(Slope_Sum, aes(x = Count, y = Slope, color = Method)) + geom_point(size = 2) + theme_bw() + labs(y = "Slope", x = "Network Size") + geom_smooth( method = 'loess', se = FALSE, formula=y~x-1, fullrange = TRUE, size = 2)  + theme_light(base_size = 18) +
+   theme(axis.line = element_line(colour = "black"), panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_blank(),
+        panel.background = element_blank(), legend.position = c(.75, .75))
+
+ggsave("../Figures/Size_Slope_Plot.png", slope_plot, height=5, width=7, units="in", dpi = 600)
+```

### 2019-08-27 Analysis/Category_Tables.Rmd
make everything writexl

@@
 
-require(xlsx)
-wb = createWorkbook()
-sheet  = createSheet(wb, sheetName="STRING")
-addDataFrame(Disease_String_Sum, sheet)
-sheet  = createSheet(wb, sheetName="STRING Filtered")
-addDataFrame(Disease_String_Filt_Sum, sheet)
-sheet  = createSheet(wb, sheetName="Biogrid")
-addDataFrame(Disease_Biogrid_Sum, sheet)
+require(writexl)
 
-saveWorkbook(wb, "../manuscript/tables/supplementary_table_S1.xlsx")
+wb = list("STRING" = Disease_String_Sum %>% arrange(Category), "STRING Filtered" = Disease_String_Filt_Sum  %>% arrange(Category), "Biogrid" = Disease_Biogrid_Sum  %>% arrange(Category))
+
+write_xlsx(wb, "../manuscript/tables/supplementary_table_S1.xlsx")
 
@@
 
-require(writexl)
-
-write_xlsx(msig_sum, "../manuscript/tables/supplementary_table_S2.xlsx")
+write_xlsx(msig_sum,"../manuscript/tables/supplementary_table_S2.xlsx")
 

### 2019-08-27 Analysis/Connectivity_Comparison.Rmd

--- 2019-08-19 Analysis/Connectivity_Comparison_Undirected.Rmd
@@
 author: "Kellen Cresswell"
-date: "June 30, 2018"
+date: "`r Sys.Date()`"
 output: pdf_document
@@
 
-require(KEGGREST)
-require(KEGG.db)
-require(readr)
-require(dplyr)
-require(biomaRt)
-require(ggplot2)
-require(broom)
-require(stringr)
-require(dplyr)
-require(tidyr)
-require(igraph)
+# install.packages("png")
+# Download https://bioconductor.org/packages/release/bioc/src/contrib/KEGGREST_1.24.0.tar.gz
+# Manually install "KEGGREST" from the file
+library(KEGGREST) # BiocManager::install("KEGGREST")
+library(KEGG.db)  # BiocManager::install("KEGG.db")
+library(biomaRt)  # BiocManager::install("biomaRt")
+library(readr)    # install.packages("readr")
+library(dplyr)    # install.packages("dplyr")
+library(ggplot2)  # install.packages("ggplot2")
+library(broom)    # install.packages("broom")
+library(stringr)  # install.packages("stringr")
+library(tidyr)    # install.packages("tidyr")
+library(igraph)   # install.packages("igraph")
 ```
@@
 #Reading in biogrid data
-
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
-
+links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
 links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
@@
 #Split based on Biogrid ID system then pull out gene names
-
 protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
 protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-
 protein1 = unlist(protein1)
@@
 #Place into matrix in form that disease2frame works on
-
-links = links %>% mutate(protein1 =protein1, 
-                             protein2 = protein2) %>% dplyr::select(protein1, protein2)
-
+links = links %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
 links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
@@
 #Repeating for string
-
 data_dir = "./data/"
-
 string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
-
-string= string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                             protein2 = gsub("9606.", "", protein2))
+string= string %>% mutate(protein1 = gsub("9606.", "", protein1), protein2 = gsub("9606.", "", protein2))
 
 #Read in ensemble protein ID
-
 ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
 genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
@@
 #Join gene names with string
-
 string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
 colnames(string)[4] ="hgnc_symbol_a"
-
 string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-
 colnames(string)[5] ="hgnc_symbol_b"
-
 dim(string)
-
 ```
@@
 ```{r}
-
 #Get directory with diseases
-
 disease_directory = "./data/Disease_Genes/"
-
 disease_names = list.files(disease_directory)
-
 disease_list = paste0("./data/Disease_Genes/", disease_names, "/")
@@
 #Read in manually selected diseases
-
-disease_select = read_csv("./supplementary_table_diseases_selected.csv")
-
+disease_select = read_csv("./data/supplementary_table_diseases_selected.csv")
 disease_traits = disease_select %>% dplyr::select(`Disease name`, Category)
@@
 #Change first column to make more clear
-
 colnames(disease_select)[1] = "Disease"
@@
 #Filtering out NA in last row
-
 disease_select = disease_select %>% filter(!is.na(Disease))
@@
 #Creating a vector of file names from disease select
-
 disease_files =  unique(paste0(disease_select$Disease, ".txt"))
@@
 #Make full path
-
 disease_selected = paste0("./data/Disease_Genes/", disease_files)
-
 ```
@@
 #Loop through diseases and pull out internal and external edges
-
 disease_length = 1:length(disease_selected)
-
 disease_frames = lapply(disease_length, function(x) {
@@
   #Record ppi missing
-
   per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
@@
 # saveRDS(disease_mod, "Modularity_Biogrid.rds")
-# 
+
 #Pull out network information for overall data frame
-
 disease_protein = lapply(disease_frames, function(x) x$diseases)
@@
 #Convert to data frame
-
 disease_overall = bind_rows(disease_protein)
-
 disease_overall = left_join(disease_overall, disease_select %>% dplyr::select(Disease, Category), "Disease")
-
 saveRDS(disease_overall, "biogrid_edges_new.rds")
@@
 # disease_select = left_join(disease_select, per_no_ppi_frame, "Disease")
-
 ```
@@
 #Loop through diseases and pull out internal and external edges
-
 disease_frames_string = lapply(disease_length, function(x) {
@@
 #Pull out network information for overall data frame
-
 disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)
@@
 #Convert to data frame
-
 disease_overall_string = bind_rows(disease_protein_string)
-
 disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")
@@
 saveRDS(disease_overall_string, "string_edges_new.rds")
+```
 
+```{r}
+#Repeat for string filtered
+string = string %>% filter(combined_score>=500)
+
+#Diseases
+disease_length = 1:length(disease_selected)
+
+#Loop through diseases and pull out internal and external edges
+disease_frames_string = lapply(disease_length, function(x) {
+
+  #Run function to find internal and external connectivity 
+  diseases = diseases2frame(disease_selected[x], links = string)
+
+  #Record ppi missing
+  per_no_ppi = 1- (nrow(diseases)/nrow(read.table(disease_selected[x])))
+  return(list(diseases = diseases, lengths = per_no_ppi))
+})
+
+#Pull out network information for overall data frame
+disease_protein_string = lapply(disease_frames_string, function(x) x$diseases)
+
+#Convert to data frame
+disease_overall_string = bind_rows(disease_protein_string)
+disease_overall_string = left_join(disease_overall_string, disease_select %>% dplyr::select(Disease, Category), "Disease")
+
+#disease_overall_string = #readRDS("disease_overall_string.rds")
+saveRDS(disease_overall_string, "string_filt_edges_new.rds")
 ```

### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
change order and add comments

@@
 ```{r}
-#Reading in data from cluster and binding
 
+#Adding p-values from permutation test
+#Biogrid
 biogrid_p = bind_rows(readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/biogrid_p_psych_1k.rds") %>% as.data.frame(.))
 
+#String
 string_p = bind_rows(readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_neuro_1k.rds") %>% as.data.frame(.))
 
+#String Filtered
 string_filt_p = bind_rows(readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_neuro_1k.rds") %>% as.data.frame(.))
 
+#Give appropriate name
 colnames(biogrid_p) = c("Disease", "Biogrid P-value", "Ave")
@@
 
-#Complete Disease dataset
+#Complete Disease dataset by adding p-values
 
@@
 
+#Rounding data for table
 coherence_disease = coherence_disease %>% mutate_if(is.numeric, round, digits = 2)
@@
 
-coherence_disease =coherence_disease %>% arrange(-`String Coherence`)
+#Arranging by disease and category
+coherence_disease =coherence_disease %>% arrange(`Category`, `Disease`)
 
@@
 
+#Rename CC to GOCC to better reflect source
 colnames(coherence_disease) = gsub("CC", "GOCC", colnames(coherence_disease))

### 2019-08-27 Analysis/Node_Plots.R
repush node plots

@@ -1,0 +1,51 @@
+remotes::install_github("schochastics/graphlayouts")
+require(graphlayouts)
+require(igraph)
+require(ggraph)
+
+#Creating ring networks
+#10 vertices
+ring1 = make_ring(10)
+#20 vertices
+ring2 = make_ring(20)
+
+#Plotting each relationship
+
+#To change line color change the "colour" parameter in geom_edge_link
+#To change edge color change value in scale_fill_manual
+ring1_plot = ggraph(ring1) +
+  geom_edge_link(width=.5,colour="red")+
+  geom_node_point(shape=21,size=3, aes(fill = "green")) +
+  theme_graph() + 
+  scale_fill_manual(values = "forestgreen") + 
+  guides(fill = FALSE)
+
+ring2_plot = ggraph(ring2) +
+  geom_edge_link(width=.5,colour="red")+
+  geom_node_point(shape=21,size=3, aes(fill = "green")) +
+  theme_graph() + 
+  scale_fill_manual(values = "forestgreen") + 
+  guides(fill = FALSE)
+
+#Full graphs
+
+full1_plot = ggraph(make_full_graph(10)) +
+  geom_edge_link(width=.5,colour="red")+
+  geom_node_point(shape=21,size=3, aes(fill = "green")) +
+  theme_graph() +
+  scale_fill_manual(values = "forestgreen") + 
+  guides(fill = FALSE)
+
+full2_plot = ggraph(make_full_graph(20)) +
+  geom_edge_link(width=.5,colour="yellow3")+
+  geom_node_point(shape=21,size=3, aes(fill = "green")) +
+  theme_graph() + 
+  scale_fill_manual(values = "forestgreen") + 
+  guides(fill = FALSE)
+
+#Saving plots
+ggsave(ring1_plot, file = "ring1.png", width = 2.5, height = 2.5, units = "in")
+ggsave(ring2_plot, file = "ring2.png", width = 2.5, height = 2.5, units = "in")
+ggsave(full1_plot, file = "full1.png", width = 2.5, height = 2.5, units = "in")
+ggsave(full2_plot, file = "full2.png", width = 2.5, height = 2.5, units = "in")
+

### 2019-08-27 Analysis/Node_Plots.R
Polish plots

@@ -1,4 +1,4 @@
-remotes::install_github("schochastics/graphlayouts")
+# remotes::install_github("schochastics/graphlayouts")
 require(graphlayouts)
 require(igraph)
 require(ggraph)
@@ -10,37 +10,37 @@ ring1 = make_ring(10)
 ring2 = make_ring(20)
 
 #Plotting each relationship
-
+set.seed(1)
 #To change line color change the "colour" parameter in geom_edge_link
 #To change edge color change value in scale_fill_manual
-ring1_plot = ggraph(ring1) +
-  geom_edge_link(width=.5,colour="red")+
+ring1_plot = ggraph(ring1, layout = "stress") +
+  geom_edge_link(width=.5,colour="dodgerblue", alpha = 0.5)+
   geom_node_point(shape=21,size=3, aes(fill = "green")) +
   theme_graph() + 
-  scale_fill_manual(values = "forestgreen") + 
+  scale_fill_manual(values = "darkorange") + 
   guides(fill = FALSE)
 
-ring2_plot = ggraph(ring2) +
-  geom_edge_link(width=.5,colour="red")+
+ring2_plot = ggraph(ring2, layout = "stress") +
+  geom_edge_link(width=.5,colour="dodgerblue", alpha = 0.5)+
   geom_node_point(shape=21,size=3, aes(fill = "green")) +
   theme_graph() + 
-  scale_fill_manual(values = "forestgreen") + 
+  scale_fill_manual(values = "darkorange") + 
   guides(fill = FALSE)
 
 #Full graphs
 
-full1_plot = ggraph(make_full_graph(10)) +
-  geom_edge_link(width=.5,colour="red")+
-  geom_node_point(shape=21,size=3, aes(fill = "green")) +
-  theme_graph() +
-  scale_fill_manual(values = "forestgreen") + 
-  guides(fill = FALSE)
-
-full2_plot = ggraph(make_full_graph(20)) +
-  geom_edge_link(width=.5,colour="yellow3")+
+full1_plot = ggraph(make_full_graph(10), layout = "stress") +
+  geom_edge_link(width=.5,colour="dodgerblue", alpha = 0.5)+
   geom_node_point(shape=21,size=3, aes(fill = "green")) +
   theme_graph() + 
-  scale_fill_manual(values = "forestgreen") + 
+  scale_fill_manual(values = "darkorange") + 
+  guides(fill = FALSE)
+
[…]

### 2019-08-27 Analysis/Coherence_Calculations_New.Rmd
move all coherence to .R files

@@
 library(biomaRt) # BiocManager::install("biomaRt")
-```
-
-```{r}
-# #Repeating for string
-# 
-# data_dir = "./data/"
-# 
-# #Read in new string database
-# 
-# string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))
-# 
-# string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-#                              protein2 = gsub("9606.", "", protein2))
-# 
-# #Read in ensemble protein ID
-# 
-# ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-# 
-# #Get gene set
-# genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
-# 
-# 
-# #Join gene names with string
-# 
-# string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-# 
-# #Rename to work with our functions
-# 
-# string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)
-# 
-# #Repeat for second set of proteins
-# 
-# string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-# 
-# string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)
-# 
-# string_filt = string %>% filter(combined_score>=500)
-# 
-# dim(string)
-# 
-```
-
-```{r}
-# #Reading in biogrid data
-# 
-# links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
-# 
-# #Select human protein-protein interactions
-# 
-# links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
-# 
-# #Split based on Biogrid ID system then pull out gene names
-# 
-# protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-# protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-# 
-# protein1 = unlist(protein1)
-# protein2 = unlist(protein2)
-# 
-# #Place into matrix in form that disease2frame works on
-# 
-# links = links %>% mutate(protein1 =protein1,
-#                              protein2 = protein2) %>% dplyr::select(protein1, protein2)
-# 
-# links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
-# 
-# 
-# dim(links)
 ```
@@
     }) %>% arrange(Slope) %>% filter(Count>4)
-
-# #Running coherence for string
-# 
-# #Read in KEGG pathways for coherence
-# msig = readRDS("KEGG_String.rds")
-# msig =msig %>% filter(Category == "CP:KEGG")
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_count$Disease)) {
-#   #Filter to only include specific disease coherence
-#   dis = dis_slopes_string_count %>% ungroup() %>% filter(Disease == i)
-#   #Get closest KEGG pathways in term of gene counts
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   #Take top 10 KEGG pathways
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-#   
-#   diseases = bind_rows()
-#   #Generate random networks
-#   j = 1
-#     while(j<11) {
-#     #Sample gene networks equal in size to disease count
-#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-#     #Get random gene networks
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     
-#     #Repeat if there are no internal edges
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-#     #Calculate overall slope of random disease networks
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-#       
-#       #Normalize based on random slope and the median of kegg slopes (More robust)
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )      
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string.rds")
-# }
-#   
-# 
-# ```
-
-#STRING All Msig
-
-# ```{r}
-# for (j in unique(msig$gs_name)) {
-#   curr_msig = msig %>% filter(gs_name == j)
-#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
-#   if (nrow(diseases_curr) == 0) {
-#     next
-#   }
-#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-#   mod_slope = coef(mod)[[1]]
-#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-#                        Category = curr_msig$gs_subcat[1])
-#   mod_over = bind_rows(mod_over, mod_sum)
-#   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
-# }
-# ```
-
-# ```{r}
-# msig = readRDS("KEGG_String_Filt.rds")
-# msig =msig %>% filter(Category == "CP:KEGG")
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_filt_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-# 
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-# 
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-# 
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#     coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
-# }
-
-```
-
-```{r}
-# msig = readRDS("KEGG_Biogrid.rds")
-# msig =msig %>% filter(Category == "CP:KEGG")
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_biogrid_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-# 
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-# 
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-# 
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       print(coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
-# }
-
-```
-
-# Reactome
-
-```{r}
-# msig = readRDS("KEGG_String.rds")
-# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_string_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-#   
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-#   
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-#       
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-#       
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_react.rds")
-# }
-#   
-# 
-#
-```
-
-```{r}
-# msig = readRDS("KEGG_String_Filt.rds")
-# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_filt_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-#   
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-#   
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-#       
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-#       
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_react.rds")
-# }
-#   
-#   
-# 
-```
-
-```{r}
-# msig = readRDS("KEGG_Biogrid.rds")
-# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_biogrid_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-# 
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-# 
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-# 
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-# 
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_react.rds")
-# }
-# 
-# 
-```
-
-# CC
-
-```{r}
-# msig = readRDS("KEGG_String.rds")
-# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_string_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-#   
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-#   
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-#       
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
-#       
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_CC.rds")
-# }
-#   
-# 
-# ```
-# 
-# ```{r}
-# msig = readRDS("KEGG_String_Filt.rds")
-# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_string_filt_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-# 
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-# 
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-# 
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-# 
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt_CC.rds")
-# }
-#
-
-```
-
-```{r}
-# msig = readRDS("KEGG_Biogrid.rds")
-# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
-# coherence_tab = bind_rows()
-# for (i in unique(dis_slopes_biogrid_count$Disease)) {
-#   print(i)
-#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
-#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
-#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-# 
-#   diseases = bind_rows()
-#   j = 1
-#     while(j<11) {
-#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-# 
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       rand_slope = coef(mod_rand)[[1]]
-# 
-#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence)
-# 
-#       coherence_tab = bind_rows(coherence, coherence_tab)
-#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid_CC.rds")
-# }
-```
-
-```{r}
-#Random networks
-
-# #Read in KEGG pathways for coherence
-# msig = readRDS("KEGG_String.rds")
-# msig = msig %>% filter(Category == "CP:KEGG")
-# string_random = bind_rows()
-# for (i in msig$Count) {   
-#   j = 0
-#   diseases = bind_rows()
-#     while(j<11) {
-#     diseases_curr = GeneSample(i, string, directed = FALSE)
-#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
-#     if (any(diseases_curr$Internal !=0)) {
-#       j = j+1
-#       diseases = bind_rows(diseases, diseases_curr)
-# 
-#     }
-#     }
-#   
-#       mod_rand = lm(External ~ Internal-1, data = diseases)
-#       diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
-#       string_random = bind_rows(string_random, diseases_curr)
-#       print(string_random)
-#       saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
-# }
 

### 2019-08-27 Analysis/Msigdf_Biogrid.R

@@ -63,3 +63,104 @@ for (j in unique(msig$gs_name)) {
   mod_over = bind_rows(mod_over, mod_sum)
   saveRDS(mod_over,"All_Msig_Biogrid.rds")
 }
+
+# #Calculating normalized coherence
+
+# #KEGG
+# msig = readRDS("KEGG_Biogrid.rds")
+# msig =msig %>% filter(Category == "CP:KEGG")
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_biogrid_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+# 
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+# 
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+# 
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       print(coherence_tab)
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
+# }
+
+# #GOCC
+# msig = readRDS("KEGG_Biogrid.rds")
+# msig =msig %>% filter(Category == "CC") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_biogrid_count$Disease)) {
+#   print(i)
[…]

### 2019-08-27 Analysis/Msigdf_String.R

@@ -52,21 +52,170 @@ for (j in unique(msig$gs_name)) {
 
 #Redoing with every category at once
 
-msig = msigdbr()
+# msig = msigdbr()
+# 
+# msig = msig %>% filter(gs_subcat != "")
+# mod_over = bind_rows()
+# for (j in unique(msig$gs_name)) {
+#   curr_msig = msig %>% filter(gs_name == j)
+#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
+#   if (nrow(diseases_curr) == 0) {
+#     next
+#   }
+#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+#   mod_slope = coef(mod)[[1]]
+#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+#                        Category = curr_msig$gs_subcat[1])
+#   mod_over = bind_rows(mod_over, mod_sum)
+#   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
+# }
 
-mod_over = bind_rows()
-for (j in unique(msig$gs_name)) {
-  curr_msig = msig %>% filter(gs_name == j)
-  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
-  if (nrow(diseases_curr) == 0) {
-    next
-  }
-  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-  mod_slope = coef(mod)[[1]]
-  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = curr_msig$gs_subcat[1])
-  mod_over = bind_rows(mod_over, mod_sum)
-  saveRDS(mod_over,"All_Msig_String.rds")
-}
+#Calculating normalized coherence
+
+disease_overall_string  <- readRDS("string_edges_new.rds")
+
+#Reading in slopes for diseases
+
+#Calculate slopes for all diseases under each category
+dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
[…]

### 2019-08-27 Analysis/Msigdf_String_Filt.R

@@ -54,23 +54,104 @@ for (j in rev(unique(msig$gs_name))) {
   saveRDS(mod_over,"KEGG_String_Filt.rds")
   }
 
-#Redoing with every category at once
+# #Normalized coherence
+# #KEGG
+# msig = readRDS("KEGG_String_Filt.rds")
+# msig =msig %>% filter(Category == "CP:KEGG")
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_filt_count$Disease)) {
+#   print(i)
+#   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
+#   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+#   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
+# 
+#   diseases = bind_rows()
+#   j = 1
+#     while(j<11) {
+#     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
+#     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
+#     if (any(diseases_curr$Internal !=0)) {
+#       j = j+1
+#       diseases = bind_rows(diseases, diseases_curr)
+# 
+#     }
+#     }
+# 
+#       mod_rand = lm(External ~ Internal-1, data = diseases)
+#       rand_slope = coef(mod_rand)[[1]]
+# 
+#       coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
+#     coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
+#       coherence_tab = bind_rows(coherence, coherence_tab)
+#       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
+# }
 
-msig = msigdbr()
+# #Reactome
+# msig = readRDS("KEGG_String_Filt.rds")
+# msig =msig %>% filter(Category == "CP:REACTOME") %>% filter(!is.na(Slope))
+# coherence_tab = bind_rows()
+# for (i in unique(dis_slopes_string_filt_count$Disease)) {
+#   print(i)
[…]

### 2019-08-28 Analysis/Coherence_Calculations_New.Rmd
remove all unneeded packages and clean

@@
 ```{r}
-library(msigdbr) # install.packages("msigdbr")
 library(dplyr)   # install.packages("dplyr")
 library(readr)   # install.packages("readr")
-source("./functions/GeneSample.R")
-library(biomaRt) # BiocManager::install("biomaRt")
 ```
@@
 #Calculate slopes for all diseases under each category
+
+#STRING
 dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
@@
 
+#STRING Filtered
 dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
@@
 
+#BIOGRID
 dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
@@
 ```{r}
-#Read in results generated above
+#Read in normalized coherence (Generated from Msigdf_*.R)
 

### 2019-08-28 Analysis/Connectivity_Comparison.Rmd
remove all unneeded packages and clean

@@
 ```{r}
-source("./functions/perm_slopes.R")
-source("./functions/GeneSample.R")
-source("./functions/genes2interactions.R")
 source("./functions/disease2frame.R")
-source("./functions/diseases2mod.R")
 
@@
 # Manually install "KEGGREST" from the file
-library(KEGGREST) # BiocManager::install("KEGGREST")
-library(KEGG.db)  # BiocManager::install("KEGG.db")
+# library(KEGGREST) # BiocManager::install("KEGGREST")
+# library(KEGG.db)  # BiocManager::install("KEGG.db")
 library(biomaRt)  # BiocManager::install("biomaRt")
@@
 library(dplyr)    # install.packages("dplyr")
-library(ggplot2)  # install.packages("ggplot2")
-library(broom)    # install.packages("broom")
 library(stringr)  # install.packages("stringr")

### 2019-08-28 Analysis/GWAS_Prep.Rmd
remove all unneeded packages and clean

@@
 library(gwascat) # BiocManager::install("gwascat")
-library(limma)   # BiocManager::install("limma")
 library(dplyr)   # install.packages("dplyr")
-library(DO.db)   # BiocManager::install("DO.db")
-library(DBI)     # BiocManager::install("DBI")
 source("https://raw.githubusercontent.com/mdozmorov/MDmisc/master/R/unembed.R") # library(MDmisc)  
 # install.packages("rJava")
-# Download https://www.bioconductor.org/packages/2.10/bioc/src/contrib/ontoCAT_1.8.0.tar.gz
-# Install "ontoCAT" manually from the file
-library(ontoCAT) # BiocManager::install("ontoCAT")
 library(rowr)    # install.packages("rowr")
@@
 ```{r}
-#Using EFO IDs from ebicat to get a hierarchy
-
-# efo_loc = "https://raw.githubusercontent.com/EBISPOT/efo/master/efo.owl"
-# 
-# 
-# #This needs direct path for some reason
-# efo = getOntology("C:\\Users\\cresswellkg\\Documents\\disease_coherence_new\\Analysis\\data\\efo.owl")
-# 
-# 
-# #Cleaning up efo data to remove urls and putting comma seperated values on new lines
-# 
-# main_inf = unembed(main_inf, "MAPPED_TRAIT_URI", sep = ",")
-# main_inf$MAPPED_TRAIT_URI = basename(main_inf$MAPPED_TRAIT_URI)
-# 
-# tree = c()
-# for (i in unique(main_inf$MAPPED_TRAIT_URI)) {
-#   
-#   #Get entire EFO tree
-#   
-#   parents = getAllTermParentsById(efo, i)
-#   
-#   #Pull out labels and make into vector
-#   
-#   parents = sapply(parents, getLabel)
-#   
-#   #Cbind and fill in non-matching entries with NA
-#   
-#   tree = rowr::cbind.fill(tree,parents, fill = NA)
-#   
-# }
-# 
-# #Remove original column
-# 
-# tree = tree[,-1]
-# 
-# colnames(tree) = unique(main_inf$MAPPED_TRAIT_URI)
-# 
-# #Pick out important disorder groups
-# 
-# imp_cats = c("Abnormality of refraction", "Abnormality of the eye", "age", "anthropometric measurement", "auditory system disease", "autoimmune disease", "body weights and measures", "bone measurement", "cancer", "carcinoma", "cardiovascular disease", "cardiovascular disease biomarker measurement", "cardiovascular measurement", "diabetes mellitus", "diabetes mellitus biomarker", "digestive system disease", "eye measurement", "glucose measurement", "head and neck disorder", "heart function measurement", "hematological measurement", "immune system disease", "infectious disease biomarker", "insulin measurement", "lipid measurement", "lipid or lipoprotein measurement", "lipoprotein measurement", "mental or behavioural disorder", "metabolic disease", "metabolite measurement", "neoplasm", "nervous system disease", "neurodegenerative disease", "pulmonary function measurement", "reproductive system disease", "respiratory disease biomarker", "skin disease", "sleep measurement", "temporal measurement")
-# 
-# #Output cats by pipe
-# 
-# #Get categories for each unique EFO ID
-# 
-# cat_ass = apply(tree, 2, function(x) c(na.omit(match(x, imp_cats))))
-# 
-# #Replace categoryless values with NA
-# 
-# cat_ass = lapply(cat_ass, function(x) {
-#   if (length(x)<1) {
-#     cat = NA 
-#   } else {
-#     cat = x
-#   }
-#   return(cat)
-# })
-# 
-# #Assign category names
-# 
-# mapped_groups = stack(cat_ass)
-# mapped_groups$values = imp_cats[mapped_groups$values]
-# colnames(mapped_groups) = c("Category", "MAPPED_TRAIT_URI")
-# 
-# #Add groups to original data frame
-# 
-# main_inf = left_join(main_inf,mapped_groups, by = "MAPPED_TRAIT_URI")
-# 
-# main_inf$Category = as.character(main_inf$Category)
-# 
-# main_inf = main_inf %>% group_by(MAPPED_TRAIT_URI) %>% mutate(Category = paste0(unique(Category), collapse = "|")) %>% distinct()
-# 
-```
-
-```{r}
 #Collapse genes and disease names for simplifcation purposes and then get unique combinations
@@
 ```{r}
-# #Create map of EFO ids to terms
-# 
-# EFO_list = lapply(unique(main_inf$MAPPED_TRAIT_URI), function(x) {
-#   if (x == "GO_190252") {
-#     term = NA
-#   } else {
-#   term = getLabel(getTermById(efo, x))
-#   }
-#   return(term)
-#   })
-# 
-# names(EFO_list) = unique(main_inf$MAPPED_TRAIT_URI)
-# 
-# EFO_Frame = stack(EFO_list)
-# 
-# colnames(EFO_Frame) = c("EFO_Name", "MAPPED_TRAIT_URI")
-# 
-# main_inf = left_join(main_inf, EFO_Frame, "MAPPED_TRAIT_URI")
-
-```
-
-```{r}
 # main_sum = main_inf %>% ungroup() %>%  dplyr::select(DISEASE.TRAIT, MAPPED_GENE, Category, SNPS)  %>% distinct() %>% group_by(DISEASE.TRAIT) %>% mutate(snp_count_total = length(unique(SNPS)), snp_count_no_gene = sum(MAPPED_GENE == ""), snp_prop_nogene = snp_count_no_gene/snp_count_total,gene_count_total = length(unique(MAPPED_GENE))) %>% dplyr::select(DISEASE.TRAIT, Category, snp_count_total, snp_count_no_gene, snp_prop_nogene, gene_count_total) %>% distinct()

### 2019-08-28 Analysis/Msigdf_Biogrid.R
fix packages

@@ -1,10 +1,10 @@
 source("./functions/disease2frame.R")
 
-require(readr)
-require(dplyr)
-require(biomaRt)
-require(msigdbr)
-require(igraph)
+library(readr) #install.packages("readr")
+library(dplyr) #install.packages("dplyr")
+library(biomaRt) #BiocManager::install("biomaRt")
+library(msigdbr) #BiocManager::install("msigdbr")
+library(igraph) #install.packages("igraph")
 
 #Reading in biogrid data
 

### 2019-08-28 Analysis/Msigdf_String_Filt.R

@@ -1,10 +1,10 @@
 source("./functions/disease2frame.R")
 
-require(readr)
-require(dplyr)
-require(biomaRt)
-require(msigdbr)
-require(igraph)
+library(readr) #install.packages("readr")
+library(dplyr) #install.packages("dplyr")
+library(biomaRt) #BiocManager::install("biomaRt")
+library(msigdbr) #BiocManager::install("msigdbr")
+library(igraph) #install.packages("igraph")
 
 string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
 

### 2019-08-28 Analysis/Node_Plots.R

@@ -1,7 +1,6 @@
-# remotes::install_github("schochastics/graphlayouts")
-require(graphlayouts)
-require(igraph)
-require(ggraph)
+library(graphlayouts) # remotes::install_github("schochastics/graphlayouts")
+require(igraph) #install.packages("igraph")
+require(ggraph) #install.packages("ggraph")
 
 #Creating ring networks
 #10 vertices

### 2019-08-28 Analysis/functions/alias2SymbolEbicat.R
comment functions

@@ -1,8 +1,10 @@
-#data(ebicat37)
 #Function for converting ebicat37 mapped genes data to symbols designed specifically for the ebicatdata
+#You must first run data(ebicat37) from the gwascat package
+
 
 alias2SymbolEbicat = function(ebicat37) {
   
+  #Convert ebicat37 to a data frame
   ebi_dat = as.data.frame(mcols(ebicat37))
   
   #Fix spaces between - and ,
@@ -18,12 +20,16 @@ alias2SymbolEbicat = function(ebicat37) {
   
   recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))
   
+  #Find where there are multiple genes mapped to a symbol and combine them with a space delimeter
   recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))
   
+  #Identify location of changed genes
   change = (ebi_dat$MAPPED_GENE == recomb_genes)
   
+  #Create a vector containing all changed values
   sum_vec = ifelse(change == TRUE, "Same", ifelse(change == FALSE & recomb_genes == "", "No Match", "Converted"))
   
+  #Replace old genes with the new mapped genes
   ebi_dat$MAPPED_GENE = recomb_genes
   
   return(list(ebi_dat = ebi_dat, change = change))

### 2019-08-28 Analysis/functions/disease2frame.R

@@ -1,3 +1,5 @@
+#disease_dis is the path to gene sets (text file containing a gene on each line)
+#Links protein protein intraction network
 diseases2frame = function(disease_dis, links) {
   
   source('functions/genes2interactions.R')
@@ -13,14 +15,13 @@ diseases2frame = function(disease_dis, links) {
   
   disease_length = 1:length(disease_nets)
   
-  #Get diseases
+  #Get diseases by removing file extension
   
   disease_dis_name = gsub('.*/', '', disease_dis)
   
   disease_dis_name = gsub('.txt', '', disease_dis_name)
   
-  
-  
+  #Place into data frame
   disease_frame = lapply(disease_length, function(y) {
     x = disease_nets[[y]]
     sub_frame = bind_rows(x$internal_connectivity, x$external_connectivity)

### 2019-08-29 Analysis/Msigdf_Biogrid.R
fx msigdf files

@@ -29,138 +29,131 @@ links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein
 
 #Getting msigs 
 
-msig = msigdbr()
+# msig = msigdbr()
+# 
+# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+# 
+# mod_over = bind_rows()
+# for (j in unique(msig$gs_name)) {
+#   curr_msig = msig %>% filter(gs_name == j)
+#   write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+#   diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
+#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+#   mod_slope = coef(mod)[[1]]
+#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+#                        Category = curr_msig$gs_subcat[1])
+#   mod_over = bind_rows(mod_over, mod_sum)
+#   saveRDS(mod_over,"KEGG_Biogrid.rds")
+# }
 
-msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")
 
-mod_over = bind_rows()
-for (j in unique(msig$gs_name)) {
-  curr_msig = msig %>% filter(gs_name == j)
-  write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-  diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
-  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-  mod_slope = coef(mod)[[1]]
-  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = curr_msig$gs_subcat[1])
-  mod_over = bind_rows(mod_over, mod_sum)
-  saveRDS(mod_over,"KEGG_Biogrid.rds")
+#BIOGRID edge counts
+
+dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
[…]

### 2019-08-29 Analysis/Msigdf_String.R

@@ -30,31 +30,10 @@ string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
 
 #Getting msigs 
 
-msig = msigdbr()
-
-msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-
-mod_over = bind_rows()
-for (j in unique(msig$gs_name)) {
-  curr_msig = msig %>% filter(gs_name == j)
-  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
-  if (nrow(diseases_curr) == 0) {
-    next
-  }
-  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-  mod_slope = coef(mod)[[1]]
-  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-                       Category = curr_msig$gs_subcat[1])
-  mod_over = bind_rows(mod_over, mod_sum)
-  saveRDS(mod_over,"KEGG_String.rds")
-}
-
-#Redoing with every category at once
-
 # msig = msigdbr()
 # 
-# msig = msig %>% filter(gs_subcat != "")
+# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+# 
 # mod_over = bind_rows()
 # for (j in unique(msig$gs_name)) {
 #   curr_msig = msig %>% filter(gs_name == j)
@@ -68,8 +47,29 @@ for (j in unique(msig$gs_name)) {
 #   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
 #                        Category = curr_msig$gs_subcat[1])
 #   mod_over = bind_rows(mod_over, mod_sum)
-#   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
+#   saveRDS(mod_over,"KEGG_String.rds")
 # }
+
+#Redoing with every category at once
+
+msig = msigdbr()
+
+msig = msig %>% filter(gs_subcat != "")
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  curr_msig = msig %>% filter(gs_name == j)
[…]

### 2019-08-29 Analysis/Msigdf_String_Filt.R

@@ -33,125 +33,125 @@ string = string %>% filter(combined_score>=500)
 
 #Getting msigs 
 
-msig = msigdbr()
+# msig = msigdbr()
+# 
+# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+# 
+# mod_over = bind_rows()
+# for (j in rev(unique(msig$gs_name))) {
+#   curr_msig = msig %>% filter(gs_name == j)
+#   write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+#   diseases_curr = diseases2frame("temp_rea.txt", links = string)
+#   if (nrow(diseases_curr) == 0) {
+#     next
+#   }
+#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
+#   mod_slope = coef(mod)[[1]]
+#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
+#                        Category = curr_msig$gs_subcat[1])
+#   print(mod_sum)
+#   mod_over = bind_rows(mod_over, mod_sum)
+#   saveRDS(mod_over,"KEGG_String_Filt.rds")
+#   }
 
-msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+#Normalized coherence
+#KEGG
+msig = readRDS("KEGG_String_Filt.rds")
+msig =msig %>% filter(Category == "CP:KEGG")
+coherence_tab = bind_rows()
+for (i in unique(dis_slopes_string_filt_count$Disease)) {
+  print(i)
+  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
+  msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
+  msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
 
-mod_over = bind_rows()
-for (j in rev(unique(msig$gs_name))) {
-  curr_msig = msig %>% filter(gs_name == j)
-  write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-  diseases_curr = diseases2frame("temp_rea.txt", links = string)
[…]

### 2019-08-29 Analysis/GWAS_Prep.Rmd
change alias2symbol

@@
   #Keeping track of changes (sapply necessary to keep placeholders for unmapped genes)
-  recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2Symbol(x, species = "Hs"))
+  recomb_genes = sapply(ebi_dat$MAPPED_GENE, function(x) alias2SymbolEbicat(x, species = "Hs"))
   recomb_genes = unlist(lapply(recomb_genes, function(x) paste0(x,collapse=' ')))

### 2019-08-29 Analysis/Msigdf_Biogrid.R
change to .gz

@@ -8,7 +8,7 @@ library(igraph) #install.packages("igraph")
 
 #Reading in biogrid data
 
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
 
 links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
 

### 2019-08-29 Analysis/Msigdf_String_Filt.R
bring in stirng edges filt

@@ -28,8 +28,17 @@ string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
 
 string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
 
-string = string %>% filter(combined_score>=500)
+string_filt = string %>% filter(combined_score>=500)
 
+disease_overall_string_filt <- readRDS("string_filt_edges_new.rds")
+
+#STRING Filtered
+dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+  do({
+    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+    data.frame(Slope= coef(mod)[1])
+  }) %>% arrange(Slope) %>% filter(Count>4)
 
 #Getting msigs 
 
@@ -64,26 +73,26 @@ for (i in unique(dis_slopes_string_filt_count$Disease)) {
   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
   msigs = msig %>% mutate(Num_Genes = abs(Count-dis$Count)) %>% arrange(Num_Genes)
   msigs = msigs[msigs$Num_Genes %in% unique(msigs$Num_Genes)[1:10],]
-
+  
   diseases = bind_rows()
   j = 1
-    while(j<11) {
+  while(j<11) {
     diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
     if (any(diseases_curr$Internal !=0)) {
       j = j+1
       diseases = bind_rows(diseases, diseases_curr)
-
+      
     }
-    }
-
-      mod_rand = lm(External ~ Internal-1, data = diseases)
-      rand_slope = coef(mod_rand)[[1]]
-
-      coherence = (dis$Slope - rand_slope)/(median(msigs$Slope, na.rm = TRUE)-rand_slope)
-    coherence = data.frame(Disease = dis$Disease, Coherence = coherence, Msig_Norm = median(msigs$Slope, na.rm = TRUE), Rand_Norm = rand_slope )
[…]

### 2019-08-29 Analysis/Msigdf_String_Filt.R
fix comment

@@ -50,7 +50,7 @@ dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal !
 # for (j in rev(unique(msig$gs_name))) {
 #   curr_msig = msig %>% filter(gs_name == j)
 #   write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-#   diseases_curr = diseases2frame("temp_rea.txt", links = string)
+#   diseases_curr = diseases2frame("temp_rea.txt", links = string_filt)
 #   if (nrow(diseases_curr) == 0) {
 #     next
 #   }

### 2019-08-30 Analysis/functions/genes2interactions.R
update comments

@@ -1,3 +1,4 @@
+#Function that takes a geneset and converts it to 
 genes2interactions = function(gene_list, actions, directed = TRUE) {
   
   if (directed ) {

### 2019-09-04 Analysis/Category_Tables.Rmd
remove all unneeded packages and clean
Move Figures folder to manuscript
Save tables in capital Tables

@@
 ```{r}
-require(dplyr)
+library(dplyr) #install.packages("dplyr")
 ```
@@
 
-write_xlsx(wb, "../manuscript/tables/supplementary_table_S1.xlsx")
+write_xlsx(wb, "../manuscript/Tables/supplementary_table_S1.xlsx")
 
@@
 
-write.csv(Cat_Within, "../manuscript/tables/table_1.csv", row.names = FALSE)
+write.csv(Cat_Within, "../manuscript/Tables/table_1.csv", row.names = FALSE)
 ```
@@
 
-write_xlsx(msig_sum,"../manuscript/tables/supplementary_table_S2.xlsx")
+write_xlsx(msig_sum,"../manuscript/Tables/supplementary_table_S2.xlsx")
 
@@
 
-ggsave("../Figures/Size_Slope_Plot.png", slope_plot, height=5, width=7, units="in", dpi = 600)
+ggsave("../manuscript/Figures/Size_Slope_Plot.png", slope_plot, height=5, width=7, units="in", dpi = 600)
 ```

### 2019-09-04 Analysis/Coherence_Calculations_New.Rmd
fix warnings
Save tables in capital Tables

@@
 library(readr)   # install.packages("readr")
+options(stringsAsFactors = FALSE)
 ```
@@
 dir.create("../manuscript/tables")
-write.csv(coherence_disease, ".././manuscript/tables/supplementary_table_S3.csv", 
+write.csv(coherence_disease, ".././manuscript/Tables/supplementary_table_S3.csv", 
           row.names = FALSE)

### 2019-09-04 Analysis/Analysis_Tables_New.Rmd
Adjust paths

--- 2019-08-26 manuscript/Analysis_Tables_New.Rmd
@@
 
-fileNameIn0  <- "tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
+fileNameIn0  <- "../manuscript/Tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
 # iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S3.csv > tmp.csv
@@
 # Selected correlations
-png(filename = "Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
+png(filename = "../manuscript/Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
 # corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
@@
 # Selected correlations
-png(filename = "Figures/FigS2_corr_references.png", height = 2000, width = 2000, res = 300)
+png(filename = "../manuscript/Figures/FigS2_corr_references.png", height = 2000, width = 2000, res = 300)
 # corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
@@
 
-png(filename = "Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
+png(filename = "../manuscript/Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 
-png(filename = "Figures/FigS3_summary_traits_KEGG.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS3_summary_traits_KEGG.png", height = 2500, width = 6000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 
-png(filename = "Figures/FigS4_summary_traits_GOCC.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS4_summary_traits_GOCC.png", height = 2500, width = 6000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 
-png(filename = "Figures/FigS5_summary_traits_Reactome.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS5_summary_traits_Reactome.png", height = 2500, width = 6000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 
-png(filename = "Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
+png(filename = "../manuscript/Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 
-png(filename = "Figures/FigS6_summary_diseases_KEGG.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS6_summary_diseases_KEGG.png", height = 3000, width = 8000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 
-png(filename = "Figures/FigS7_summary_diseases_GOCC.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS7_summary_diseases_GOCC.png", height = 3000, width = 8000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 
-png(filename = "Figures/FigS8_summary_diseases_Reactome.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS8_summary_diseases_Reactome.png", height = 3000, width = 8000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 
-```{r}
+```{r eval=FALSE}
 fileNameIn0  <- "tables/Mod_Table.xlsx"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
@@
 rownames(mtx0) <- mtx0$Disease
-```
 
-```{r}
 # Selected data for correlation

### 2019-09-07 manuscript/manuscript.Rmd
Adjust Fig2 name
grammarly edits

--- 2019-08-20 manuscript.Rmd
@@
 ---
-bibliography: [refs/networks.bib,refs/brain.bib,refs/enrichment.bib,refs/3D_refs.bib]
-csl: styles.ref/genomebiology.csl
+title: 'TADCompare: an R package for differential analysis and characterization of Topologically Associated Domains'
+fontsize: 12pt
+mainfont: Times New Roman
+csl: styles.ref/genome-research.csl
+header-includes:
+   - \usepackage{float}
+   - \usepackage{setspace}\doublespacing
 output:
-  pdf_document: default
   word_document:
     reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
-  html_document: default
-editor_options: 
-  chunk_output_type: console
+  pdf_document:
+    fig_caption: no
+  html_document: null
+bibliography: [3D_refs.bib,tad_references.bib,difftad_references.bib]
 ---
 
-# Coherence of Molecular Mechanisms in Genes Identified Through GWAS In Major Human Disease and Traits
+```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
+# Set up the environment
+library(knitr)
+opts_chunk$set(cache.path='cache/', fig.path='figures/', cache=F, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700,
+library(pander)
+panderOptions('table.split.table', Inf)
+set.seed(1)
+library(dplyr)
+options(stringsAsFactors = FALSE)
+```
 
-Dozmorov, Mikhail G.^1+^ (mikhail.dozmorov@vcuhealth.org)  
-Cresswell, Kellen G.^1+^ (cresswellkg@vcu.edu)  
-Reimers, Mark^2^ (reimersm@msu.edu)  
-Bacanu, Silviu-Alin^3^ (silviu-alin.bacanu@vcuhealth.org)  
-Craver, Carl^4^ (ccraver@wustl.edu)  
-Kendler, Kenneth S.^3^ (kenneth.kendler@vcuhealth.org)
+Kellen G. Cresswell^1,2^ (cresswellkg@vcu.edu), Mikhail G. Dozmorov^1,2\*^ (mikhail.dozmorov@vcuhealth.org)
 
-^1^ - Department of Biostatistics, Virginia Commonwealth University   
-^2^ - Neuroscience Program and Dept. Biomedical Engineering Michigan State  
-^3^ - Virginia Institute for Psychiatric and Behavior Genetics and the Department of Psychiatry, Virginia Commonwealth University  
-^4^ - Philosophy-Neuroscience-Psychology Program, Washington University in St. Louis  
-^+^ - equal contribution
+^1^ Dept. of Biostatistics, Virginia Commonwealth University, Richmond, VA, 23298, USA  
+^2^ Equal contribution  
+^\*^ To whom correspondence should be addressed: Virginia Commonwealth University, Richmond, VA, 23298, 804-827-2055, mikhail.dozmorov@vcuhealth.org
+
+**Keywords:** Hi-C, chromosome conformation capture, topologically associated domains, TADs, differential analysis, TADCompare
+
+**Running Title:** TADCompare: an R package for differential TAD detection
 
@@
 
-**Background.** Complex phenotypes such as height and intelligence, are thought to be a product of the collective effects of multiple phenotype-associated genes, or, more precisely, interactions among their protein products.
-High/low degree of interactions is suggestive of coherent/random molecular mechanisms, respectively.
-Comparing the degree of interactions may help to better predict and understand the coherence of the molecular mechanisms underlying complex phenotypes.
-However, direct comparison of the degree of interactions is difficult due to differences in size and configuration of phenotype-associated gene networks.
+Recent research using chromatin conformation capture technologies, such as Hi-C, has demonstrated the importance of topologically associated domains (TADs).
+TADs are considered to be relatively stable three-dimensional (3D) genomic structures orchestrating gene expression regulation; yet, many dynamically reorganize during development or disease, and exhibit cell- and condition-specific differences.
+Quantification of the dynamic behavior of TADs will help to better understand genome regulation.
+Methods for comparing TADs between cells and conditions are highly limited.
+We developed `TADCompare`, a method for differential analysis of TAD boundaries between Hi-C datasets.
+`TADCompare` is based on a spectral clustering-derived measure called the eigenvector gap, which enables a loci-by-loci comparison of TAD boundary differences between datasets.
+Using this measure, we introduce methods for identifying differential and consensus TAD boundaries, and tracking TAD boundary changes over time.
+We further propose a novel framework for the systematic classification of TAD boundary changes.
+Colocalization- and gene enrichment analysis of different types of TAD boundary changes demonstrated distinct biological functionality associated with them.
+`TADCompare` is available on https://github.com/dozmorovlab/TADCompare.
 
-**Methods.** We introduce a measure of network coherence as a ratio of internal vs.
-external distributions of the degree of interactions.
-The internal degree distribution is defined by interaction counts within a phenotype-specific gene network, while the external degree distribution counts interactions with other genes in the whole protein-protein interaction (PPI) network.
-We present a novel method for normalizing the internal degree distributions, making them directly comparable.
-
-**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (the standard of high coherence) and random networks (the lower limit of coherence).
-We observed a range of coherence estimates for each category of phenotypes.
-Metabolic traits and diseases were the most coherent, while psychiatric diseases and intelligence-related traits were the least coherent.
-
-**Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences.
-Our results highlight the gaps in of our current knowledge of the genetics and molecular mechanisms of complext phenotypes and suggest priorities for future work.
-
-**Keywords:** GWAS, network, degree, coherence
+### Keywords
 
@@
 
-Genome-wide association studies (GWAS) have significantly advanced our understanding of complex phenotypes by identifying disease- and trait-associated genetic markers and suggesting corresponding genes [@Altshuler:2008aa; @Risch:1996aa].
-However, GWAS findings explain only a fraction of heritability [@Eichler:2010aa; @Maher:2008aa; @Manolio:2009aa].
-Furthermore, the corresponding genes are often spread across different chromosomes with no known connection to one another, hindering understanding of the molecular mechanisms.
-These limitations of current generation GWAS might stem from the yet incomplete knowledge of the genetic determinants of complex phenotypes, the potential heterogeneity of complex phenotypes and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].
+Recent research indisputably proves importance of the 3D genome organization in regulating gene expression and other genomic processes [@Schoenfelder:2010ab; @Steensel:2011aa; @Franke:2016aa; @Symmons:2014aa; @Sexton:2015aa; @Li:2012aa; @Papantonis:2013aa; @Laat:2003aa; @Mora:2016aa; @Mifsud:2015aa; @Shavit:2014aa; @Osborne:2004aa; @Schoenfelder:2010aa; @Tanizawa:2010aa].
+The 3D genomic structures consists of chromosome territories [@Cremer:2010aa], A/B compartments corresponding to active/repressed chromatin [@Rao:2014aa; @Lieberman-Aiden:2009aa], topologically associated domains (TADs) [@Dixon:2012aa; @Jackson:1998aa; @Ma:1998aa; @Nora:2012aa; @Sexton:2012aa; @Bonev:2017aa; @Dekker:2002aa], smaller sub-TADs [@Phillips-Cremins:2013aa; @Rao:2014aa] and chromatin loops [@Rao:2014aa; @Dowen:2014aa; @Ji:2016aa; @Denker:2016aa].
+These structures help to regulate global gene expression [@Schoenfelder:2010ab; @Steensel:2011aa; @Franke:2016aa; @Symmons:2014aa; @Sexton:2015aa; @Li:2012aa; @Papantonis:2013aa; @Laat:2003aa; @Mora:2016aa; @Mifsud:2015aa; @Shavit:2014aa; @Osborne:2004aa; @Schoenfelder:2010aa; @Tanizawa:2010aa].
+Consequently, coordinated changes in the 3D structures [@Yaffe:2011aa; @Symmons:2014aa; @Dai:2012aa] determine cell type-specific gene expression and identity [@Dowen:2014aa; @Ji:2016aa; @Phillips-Cremins:2013aa; @Rao:2014aa; @Vietri-Rudan:2015aa; @Jin:2013aa; @Schoenfelder:2010aa; @Dekker:2013aa], guide recombination [@Jhunjhunwala:2009aa], X chromosome inactivation [@Nora:2012aa; @Crane:2015aa].
+Many 3D structures are largely invariant between different cell types, and even conserved between mammalian species [@Dixon:2012aa; @Nora:2012aa; @Pope:2014aa; @Naumova:2013aa; @Rao:2014aa; @Vietri-Rudan:2015aa], indicating their high biological importance during genome evolution.
 
-Many studies have shown that genetically-driven complex phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa].
-This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to human phenotype, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
-Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all molecular determinants of a certain phenotype.
-Consequently, methods to identify phenotype-associated networks of functionally-related genes have been developed [@Krauthammer:2004aa; @Cordell:2009aa; @Akula:2011aa].
+Despite the high level of conservation, recent research uncovered the dynamic nature of the 3D genomic structures, and this plasticity accompanies various biological functions and phenomena [@Yu:2017aa].
+In Drosophila, exposure to heat-shock caused local changes in certain TAD boundaries resulting in TAD merging [@Li:2015ac].
+Another recent study showed that during motor neuron (MN) differentiation in mammals, TAD, and sub-TAD boundaries in the Hox cluster are not rigid and their plasticity is linked to changes in gene expression during differentiation [@Narendra:2016aa].
+The global organization of the 3D genomic structure is found in mitosis [@Nagano:2017aa], <!--zygotic development [@Hug:2017aa; @Flyamer:2017aa], -->earliest stages of mammalian lineage development [@Dixon:2015aa; @Du:2017aa; @Ke:2017aa; @Bonev:2017aa], and somatic cell reprogramming of pluripotent stem cells [@Zhang:2018aa; @Novo:2018aa].
+Fusion of TADs [@Flavahan:2016aa; @Fudenberg:2016aa; @Sanborn:2015aa; @Dowen:2014aa; @Guo:2015aa; @Tang:2015aa; @Nora:2012aa], creation or destruction of sub-TADs within existing TAD boundaries [@Taberlay:2016aa; @Lupianez:2016aa], and/or switching TAD states between active and inactive conformations [@Dixon:2012aa; @Lieberman-Aiden:2009aa] has been associated with a variety of phenotypes [@Krijger:2016aa; @Misteli:2010aa; @Spielmann:2018aa], ranging from limb malformation [@Lupianez:2016aa], congenital disorders [@Ibn-Salem:2014aa], to cancer [@Mitelman:2000aa; @Valton:2016aa; @Rickman:2012aa; @Hnisz:2016aa; @Groschel:2014aa; @Corces:2016aa; @Lupianez:2016aa; @Krijger:2016aa; @Flavahan:2016aa; @Barutcu:2015aa].
+These observations highlight the importance of studying changes in TADs as a means to understand genomic regulation.
+However, methods for identifying changes in TAD boundaries remain underdeveloped.
 
-Network properties, such as connectivity (aka degree distribution) [@Ghiassian:2015aa; @Lee:2008aa; @Barabasi:2004aa; @Feldman:2008aa], can be compared to better understand the relationships among the phenotypes [@Vidal:2011aa; @Barabasi:2011aa; @Wang:2011aa].
-Networks formed by phenotype-associated genes, or, more precisely, their protein-protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa].
-This intuition reflects a well-known "guilt-by-association" principle that genes (or, more precisely, their products) that form an interaction network are more likely to share similar functions [@Farber:2013aa], or co-expression patterns [@Huang:2006aa; @Michalak:2008aa].
-Although this view has been criticized [@Gillis:2012aa], several studies consistently observed phenotype-associated genes to be either connected as a single network or to participate in common phenotype-specific subnetworks [@Menche:2015aa; @Lee:2011aa; @Akula:2011aa; @Ideker:2008aa; @Iossifov:2008aa].
-Therefore, high connectivity among phenotype-specific genes may indicate coherent molecular mechanisms that could be targeted therapeutically.
-In terms of graph theory, we are asking whether a phenotype-specific network is a community - a highly connected subgraph relatively well-separated from the rest of the network [@fortunato2010community].
-However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes forming networks of different sizes and configurations.
+To our knowledge, there are only three methods that can be adapted for detecting changes in TAD boundaries: `localtadsim` [@Sauerwald:2018aa2], `HiCDB` [@Chen:2018aa] and `DiffTAD` [@Zaborowski:2016aa].
+Of the three methods, none provide an intuitive, easy to use way of calling differential TADs.
+Both `localtadsim` and `DiffTAD` are two-step procedures requiring separately defined TADs and comparing them using a command-line utility.
+`HiCDB` has a built-in TAD caller but does not allow for comparisons of chromosome-specific contact matrices.
+All three methods require highly specific data types and file names to be able to run.
+The lack of usability is compounded with issues such as a lack of upkeep, slow runtimes, and lack of statistical rigor (Supplementary methods).
 
-This study compares the connectivity of networks formed by disease- or trait-associated genes and gene products.
-We hypothesized that genes associated with (categories of) diseases share common molecular mechanisms and, hence, form connected (i.e., coherent) networks.
-Therefore, the level of connectivity may inform us about similarities and differences among the diseases.
-The terms “network connectivity,” “coherence,” or “average degree distribution” are used interchangeably as appropriate.
-We developed a method for normalizing the internal network connectivity that accounts for differences in network sizes formed by disease-associated genes.
-We compared the ratio of normalized internal vs.
-external network connectivity between psychiatric, neurologic, autoimmune, cancer, cardiovascular, metabolic and other diseases and traits.
-We used KEGG pathways [30] as a reference for a coherent network with common molecular mechanisms, and we used sets of randomly sampled genes as a reference for the absense of coherence.
-We found that genes associated with metabolic diseases and traits and autoimmune diseases formed coherent networks similar to KEGG pathways.
-On the contrary, the coherence of networks of psychiatric and neurologic disease-associated genes, as well as intelligence-associated traits, was similar to that of randomly selected genes.
-Cancer, anthropometric traits, eye-, and cardiovascular diseases and traits formed networks of intermediate connectivity.
-Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.
+As costs of Hi-C data continue to drop, several studies started to investigate the dynamics of 3D changes over time.
+The most notable applications include cell differentiation studies [@Bonev:2017aa], embryonic development [@Du:2017aa;@Hug:2017aa;@Ke:2017aa], cancer progression [@Zhou:2019aa].
+Typically, TAD boundary changes over time are quantified by overlap [@Hug:2017aa;@Du:2017aa] and classified into distinct patterns [@Zhou:2019aa].
+However, general-purpose methods for systematic analysis of TAD boundary changes over time do not exist.
+Furthermore, the number of replicates for a given experiment continues to rise, requiring methods for defining and comparing TAD boundaries across replicates of Hi-C data.
+<!--Previous research has shown inconsistencies between both technological and biological replicates [@Dixon:2012aa;@Rao:2014aa;@Sauerwald:2018aa2].
+As a result, researchers must take this variation into account and make full use of the data.
+-->Traditionally, two approaches have been developed to identify TAD boundaries across replicates.
+The first is to call TADs on individual replicates and aggregate them (e.g., `Arrowhead` [@Rao:2014aa]).
+The second approach involves combining all replicates into a consensus contact matrix <!--using utilities such as `Juicer` [@Durand:2016aa] -->and then calling TADs (`Arboretum` [@Fotuhi-Siahpirani:2016aa]).
+[???
+TBD - are the tools mentioned to the corresponding methods?] However, it is focused on identifying TADs across cell-lines and species as opposed to across replicates.
+Methods for detecting consensus TADs across Hi-C datasets remain underdeveloped.
 
-This study presents a novel measure of network coherence as a ratio of the normalized internal vs.
-external degree distributions.
-The internal degree distribution is defined as gene-specific interaction counts within a protein-protein interaction (PPI) network formed by phenotype-associated genes, while the external degree distribution counts interactions with other genes in the whole PPI network.
-We used KEGG pathways [@Kanehisa:2000aa] as a reference for a paradigmatically high-coherent network [@Lee:2003aa], while sets of randomly sampled genes were used as the reference for the absence of coherence.
-We hypothesized that the level of coherence may inform us about similarities and differences among phenotypes.
-Using two PPI databases (STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa]), we compared coherence of 133 phenotypes from GWAScatalog [@Hindorff:2009aa; @MacArthur:2017aa].
-Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have the most coherence.
-Our method enables direct comparison of coherence measures and highlight gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.
+We developed `TADCompare`, an R package aimed at providing a fast, accurate, user-friendly and well-documented approach to differential TAD analysis.
+We introduce a method based on the boundary score statistic [@Cresswell:2019aa] and use it to identify five types of TAD boundary changes.
+The method is extended to allow for calling consensus TAD boundaries and comparing them between groups of Hi-C replicates.
+We further demonstrate how the TAD boundary score statistic may be used to analyze TAD dynamics over the time course.
+For both differential TAD boundary detection and time course analysis, we provide novel terminology for the classification of TAD boundary changes.
+We demonstrated the robustness of `TADCompare` using simulated data with pre-defined TADs [@Forcato:2017aa] and its ability to reveal distinct biological roles of different TAD boundary changes.
+In summary, `TADCompare` provides an all-in-one pipeline from TAD calling to differential boundary detection, including time course, that supports replicated Hi-C data analysis.
+The output is formatted in a commonly used BED format that allows for flexible downstream analyses and visualization.
+The `TADCompare` R package is freely available on GitHub (https://github.com/dozmorovlab/TADCompare).
+
+# Methods
+
+### Representation of Hi-C data as a graph
+
+For a given Hi-C experiment, Hi-C data is represented by a chromosome-specific contact matrix $C$ of non-overlapping regions (aka bins) of size $r$ (resolution of the data).
+Each entry $C_{ij}$ corresponds to the number of contacts between region $i$ and region $j$.
+Previous work has shown that this contact matrix is essentially an analog of the adjacency matrix found in graph theory and Hi-C data can be thought of as a naturally occurring graph where edges are contacts and vertices are genomic regions [@Boulos:2013aa; @wang2017network; @Wang:2013:TPC:2506583.2506633;@Cresswell:2019aa].
+The graph representation of Hi-C data is the foundation of our method and allows us to use a graph-clustering based approach to identify and analyze TADs.
+
+### Calculating the graph spectrum
+
+The first step of our method is to calculate the graph spectrum, defined as the eigenvectors of the Laplacian of an adjacency matrix.
+Using the interpretation of the contact matrix as a naturally occurring adjacency matrix, we calculate the Laplacian directly from the contact data.
+Briefly, the graph spectrum for a given contact matrix is calculated as follows:
+
+1. Calculate the normalized Laplacian $\bar{L}$:
+
+$$
+\bar{L} = D^{-\frac{1}{2}}CD^{-\frac{1}{2}}
+$$
+
+where $D = diag(\textbf{1}^{T}C)$, where $\textbf{1}$ is a column vector of size $C$ where each entry is 1. $D$ can be thought of as a vector containing the sum of the degrees for a given node.
+
+2. Perform an eigendecompositon of the Laplacian:
+
+$$
+\bar{L}v = \lambda v
+$$
+
+In practice, we calculate the first two eigenvectors with the largest absolute values of eigenvalues and organize them into a matrix $\bar{V}$ with dimensions $i \times 2$, where $i$ is the number of regions in the contact matrix.
+$\bar{V}$ is referred to as the graph spectrum of the contact matrix.
+
+### Eigenvector gap as a measure of pattern change
+
+We can think of each row of the matrix $\bar{V}$ as a quantification of the pattern of contacts in each region of the contact matrix.
+Previous work [@Cresswell:2019aa] has demonstrated that by taking the Euclidean distance between row $V_{i.}$ and its neighboring row $V_{(i+1).}$, one can measure the similarity in the pattern of contacts between region $i$ and region $i+1$ of the chromosome, termed "eigenvector gap".
+A TAD boundary manifests itself as a sudden break in the pattern of contacts.
+This pattern is reflected in the eigenvector gap by a spike in gap size followed by and preceded by smaller gaps (Figure 1).
+The eigenvector gap quantifies the degree of this break, acting as a proxy for TAD boundary likelihood.
+To calculate the eigenvector gaps, we perform the following procedure:
+
+1. Normalize columns of $\bar{V}$ to sum to 1:
+
+$$
+\begin{aligned}
+\hat{V_{ij}} = \frac{\bar{V_{ij}}}{ \Vert{\bar{V_{.j}\Vert}}}
+\end{aligned}
+$$
+
+where the subscript $.j$ corresponds to column $j$.
+
+2. Normalize $\hat{V}$ and project onto a unit circle:
+
+$$
+\tilde{Z} = Diag(diag^{-\frac{1}{2}}(\hat{V_{i.}}\hat{V_{i.}}^{T}))\hat{V_{i.}}
+$$
+<!--$\tilde{Z}$ is the projection of $\bar{V}$ onto a unit circle such that all values are bounded by [-1,1].-->
+
+3. Calculate the distance between neighboring regions (rows $i$ and $i-1$ of $\tilde{Z}$) and store in a vector $D_{i}$:
+
+$$
+D_{i}=\sqrt{(\tilde{Z}_{i1}-\tilde{Z}_{(i-1)1})^2 + (\tilde{Z}_{i2}-\tilde{Z}_{(i-1)2})^2}
+$$
+
+We refer to $D$ as the vector where each entry $D_i$ is referred to as an eigenvector gap.
+Formally, an eigenvector gap is the Euclidean distance between each successive row of the first two eigenvectors.
+In practical terms, the eigenvector gap for a given locus is a measure of how likely that locus is a TAD boundary.
+
+To maintain the association of each entry of the vector with its corresponding matrix region, a placeholder is used in the first entry of the vector.
+This is necessary because we cannot calculate an eigenvector gap for the first entry of the contact matrix due to a lack of a left-bound neighbor.
+In mathematical terms, this means that for a matrix of size $n$ the total number of eigenvector gaps is $n-1$.
+
+<!--
+Graph spectrum calculation scales poorly for large matrices, having a quadratic complexity [@Wu:2018aa].
+By using a windowed approach, the complexity becomes linear, relying mainly on the window size to determine the speed of the algorithm.
+The window starts at the beginning of the contact matrix.
+After calculating the eigenvector gaps within the window, the window moves forward $w-1$ bins across the diagonal where $w$ is the size of the window.
+We move $w-1$ spaces because for a given window it is impossible to calculate the eigenvector gap of the last bin.
+Moving $w-1$ allows us to include that last bin in the next window and calculate this value.
+We continue this process until the end of the contact matrix is reached.
+-->
+
+### Converting eigenvector gaps to boundary scores
+
+We showed that the distribution of eigenvector gaps can be approximated by a log-normal distribution (Supplementary Figure 1, Supplementary Material).
+The log-normality allows us to convert the eigenvector gap values into boundary scores:
+
+$$
+B_{i} = \frac{ (ln(D_i) - \mu)} {\sigma^{2}}
+$$
+
+where $ln(D) \sim N(\mu, \sigma^{2})$ where $\mu$ and $\sigma^{2}$ are the mean and variance of the distribution of the natural log of the eigenvector gaps, respectively, and $B$ is a vector of boundary scores with a N(0,1) distribution.
+In practice, this value is simply the Z-score for the natural log of eigenvector gaps.
+
+### Sliding window eigenvector gap calculation
+
+Frequency of interactions decays following power law as the distance between the interacting regions in a linear genome increases [@Lajoie:2015aa].
+This decay leads to noisy and non-informative interactions farther off-diagonal of the contact matrix.
+To alleviate the effect of noisy distant interactions, we perform spectral decomposition within a fixed-size window that moves along the diagonal of the matrix.
+For instance, a window size of 15 bins (default setting, Supplementary Figure 2, Supplementary Material) means that only values within 15 bins of the diagonal will be used to calculate the eigenvector gap.
+The sliding window approach improves the performance of eigenvector gap calculation [@Cresswell:2019aa].
+Additionally, it provides for faster calculations, operating on many small matrices instead of one large matrix.
+
+<!-- ### Measuring contact matrix similarity using boundary scores -->
+
+<!-- Oftentimes, it is of interest to measure the similarity of entire contact matrices.
+This is useful for comparing technical or biological replicates and assessing the consistency of Hi-C sequencing.
+Calculating similarity is complicated by the need to capture the topological similarity of contact matrices and certain issues such as distance decay and noise.
+Previous methods have attempted to quantify this similarity, such as Quasar-Rep [@Sauria:2015aa], Hi-C-Spector[@Yardimci:2017aa], Hi-CRep [@Yang:2017aa] and GenomeDISCO [@Ursu:2018aa].
+Both the Quasar-Rep and approach takes two contact matrices, transforms them by stratifying based on distance, and then takes a Pearson correlation.
+GenomeDISCO and Hi-C-Spector are graph-based methods, with GenomeDISCO measuring similarity based on the distance between graph-diffusion smoothed contact matrices and Hi-C-Spector measuring the distance between eigenvectors of the Laplacians of each contact matrix.
+We propose a method based on cross-correlation of boundaries to calculate similarity.
+-->
+
+<!-- Given two contact matrices $P$ and $R$ with boundary scores of $B_P$ and $B_R$ having means of $\mu_{P}$ and $\mu_{R}$ respectively, we can calculate matrix similarity using a modified form of cross-correlation that takes into account small shifts in TAD boundaries.
+For this example, we first define the basic cross-correlation of two sets of boundary scores with lag $l$: -->
+
+<!-- $$ -->
+<!-- S = \frac{ \sum[(B_{P}(i)-\mu_{P})*(B_{R}(i-l)-\mu_{R})] -->
+
+<!-- }{\sqrt{\sum(B_{P}(i)-\mu_{P})^{2}}\sqrt{\sum(B_{R}(i-l)-\mu_{R})^{2}}} -->
+<!-- $$ -->
+
+<!-- where $i$ is the corresponding vector entry and $l$ is the lag.
+Using this framework, we incorporate the fact that shifts of less than 5 are biologically insignificant by replacing all instances of $B_{R}(i-l)$ with $B_{min}(i) = min(|(B_{P}(i)-\mu_{P})-(B_{R}(i-l)-\mu_{R})|)$ for l = 1,...,5.
+The resulting formula is the following: -->
+
+<!-- $$ -->
+<!-- S_{adj} = \frac{ \sum_{i}[(B_{P}(i)-\mu_{P})*(B_{min}(i)-\mu_{R})] -->
+
+<!-- }{\sqrt{\sum_{i}(B_{P}(i)-\mu_{P})^{2}}\sqrt{\sum_{i}(B_{min}(i)-\mu_{R})^{2}}} -->
+<!-- $$ -->
+
+<!-- $S_{adj}$ gives us a measure of similarity between contact matrices that gives equal weighting to lagged and non-lagged TAD boundaries, finding the region of maximum similarity.
+This score effectively provides a measure of similarity between contact matrices purely in terms of TAD similarity.
+The adjustment allows us to treat shifted boundaries the same as non-shifted boundaries and more accurately capture contact matrix similarity.
+-->
+
+### Differential analysis using boundary scores
+
+To define the differences between two contact matrices, $P$ and $R$, we compare their eigenvector gaps $D_P$ and $D_R$, respectively.
+Given that $ln(D_P) \sim N(\mu_{P}, \sigma_{P}^{2})$ and $ln(D_R) \sim N(\mu_{R}, \sigma_{R}^{2})$, it follows that $ln(D_P) - ln(D_R) \sim N(\mu_{P}-\mu_{R}, \sigma_{P}^{2}+\sigma_{R}^{2})$.
+These results allow us to calculate a vector of differential boundary scores:
+
+$$
+DB_{i} = \frac{ (ln(D_{Pi}) - ln(D_{Ri})) - (\mu_{P}-\mu_{R}) } {\sigma_{P}^{2}+\sigma_{R}^{2}}
+$$
+
+or more simply,
+
+$$
+DB_{i} = \frac{\sigma_{P}^{2}B_{P}-\sigma_{R}^{2}B_{R}}{\sigma_{P}^{2}+\sigma_{R}^{2}}
+$$
+
+where $B_{P}$ and $B_{R}$ are the boundary scores for the $P$ and $R$ matrices, respectively.
+This score can be thought of as the difference in TAD boundary likelihood for a given locus in two data sets.
+Due to the aforementioned normality of the difference in log eigenvector gaps, $DB_i$ can be thought of as a simple z-score where $DB \sim N(0,1)$.
+
+### Time course boundary changes
+
+Boundary scores provide a convenient method for modeling the change of TAD boundaries over time.
+For a given TAD boundary, or, any region of the genome, we can monitor the trajectory of the boundary score.
+Over time, we can define boundary score changes based on their deviation from a baseline level (typically, the first time point).
+It is expected that these scores will be relatively constant over time except in regions where a boundary appears or disappears.
+The trend across time points can be recorded, and the pattern of change classified accordingly.
+Our implementation of time course boundary analysis allows for the usage of multiple replicates for a given time point.
+Briefly, at each region of the genome, the consensus boundary score is calculated, defined as the median of consensus scores across all replicates, and is then used to identify boundaries.
+
+<!--The overall trend can be calculated by taking the full set of TAD boundaries across every time point and calculating the average boundary score for each individual point.
+If this score is increasing, then it suggests the appearance of TAD boundaries over time.
+Conversely, if it is decreasing, then there is a disappearance of TAD boundaries.
+We can also incorporate the consensus boundary score framework into boundary score calculations over time.
+Given a set of replicates at multiple different time points, we can call consensus boundary scores at for each time points and aggregate them.
+We can also quantify the change in TAD strength for the entire contact matrix over time by modeling the trend of mean boundary scores across each of the time point.
+A positive trend indicates an increase in TAD strength or abundance and a decrease indicates a drop.-->
+
+<!--
+### Simulated data for differential TAD boundary detection
+
+To simulate differential TAD boundaries, we used 25 simulated contact matrices with manually annotated TADs downloaded from Hi-CToolsCompare [@Forcato:2017aa].
+These matrices were generated at five levels of noise (4%, 8%, 12%, 16%, and 20%) with five simulated matrices per noise level.
+Noise is generated by randomly selecting entries in the contact matrix and adding two contacts.
+The percentage value corresponds to the percentage of matrix entries with constant noise added.
+
+Within each noise level, we take each possible combination of contact matrices and calculate differential TADs between them.
+It is expected that an accurate differential boundary detection method will find all manually annotated TADs which differ between each combination of simulated matrices.
+For each combination of simulated matrices, we define true differential TAD boundaries as the set of non-shared boundaries between the matrices.
+To assess the ability of our method to discriminate between true differential TADs and non-differential TADs, we use the Matthews correlation coefficient.
+[???
+TBD, rationale for MCC]
+-->
+<!-- ### Primary and replicate experimental data for differential TAD boundary detection -->
+
+<!-- Primary and replicate data from [@Rao:2014aa] was used to test the percentage of differential TAD boundaries between primary and replicate data.
+TAD boundaries were called using the `SpectralTAD` TAD caller [@Cresswell:2019aa].
+The raw percentage of shared boundaries at the first level of the hierarchy were calculated between the primary and replicate.
+To test whether these boundaries were truly differential, we then ran `TADCompare` and calculated the percentage of non-shared boundaries which were also detected by TADCompare.
+-->
+
+<!-- We hypothesized that non-shared boundaries between primary and replicates are largely the result of differences in the TAD hierarchy.
+If true, it is expected that TAD boundaries which are not shared between primary and replicate and are not detected by `TADCompare` will actually appear somewhere in both the primary and replicate's hierarchy of TADs.
+[???
+Shall we forgo hierarchy?] -->
+<!--
+### Handling of gaps
+
+Gaps refer to loci with either complete absence of loci or levels of sparsity which make analysis impossible.
+For our analysis, we defined gaps as any region with less than 80\% of its contacts greater than zero.
+Since we are not particularly interested in long-range contacts, this percentage is calculated based on regions within our sliding window.
+These gaps can introduce instability in the algorithm and lack important information.
+To counter this, we simply remove these regions before the analysis.
+This is done for both contact matrices such that, if one contact matrix contains a gap at a given locus and the other does not, we remove it from both.
+This allows us to still make a one-to-one comparison of regions.
+-->
+
+### Data sources
+
+All simulated data were downloaded from `HiCToolsCompare` repository [@Forcato:2017aa].
+In total, we used 25 simulated matrices with varying levels of noise.
+For sparsity and downsampling analysis matrices were manually created based on matrices from `HiCToolsCompare` matrices with the minimum noise level (see [@Cresswell:2019aa] for methods description).
+Data for comparisons across cell lines, replicates, and tissues were taken from [@Schmitt:2016aa], generated at 40kb resolution (Supplementary Table 1).
+Time course data was taken from [@Rao:2017aa], HCT-116 human colon cancer cell-line at four time points after auxin-treatment withdrawal (20, 40, 60, 180 minutes).
+Contact matrices were generated at 25kb, 50kb, and 100kb using the `straw` tool from `Juicer` [@Durand:2016aa].
+Chromatin state data were taken from chromHMM [@Ernst:2010aa].
+Histone modifications and transcription factor binding sites were downloaded from the Encyclopedia of DNA Elements (ENCODE) [@Davis:2018aa] (Supplementary Table 2).
+
+### Gene enrichment testing
+
+All gene enrichment testing was performed using the `rGREAT` (Version 2.0) R package.
+Briefly, we detect genes within 5kb upstream and 1kb downstream of each type of TAD boundary changes, similar to work of others [@Chen:2018aa].
+For each Gene Ontology (GO) and pathway (MSigDb [???
+Describe version, add ref]), a hypergeometric test is then performed to determine over-representation of TAD boundary-associated genes.
+For all figures, we report results for GO Biological Processes.
+Results for GO Molecular Function, GO Cellular Component, MSigDB and PANTHER pathways are reported in tables.
+
+### Colocalization enrichment testing
+
+A permutation test was used to quantify the enrichment of colocalization of TAD boundaries of interest with genomic annotations.
+Briefly, we flank each type of TAD boundary changes (differential or time course) by 50kb on each side and calculate the mean number of genomic annotations across those regions (observed enrichment).
+Next, we generate two sets of bins, one the size of the TAD boundaries which we are testing (considering the flanking) and another the size of all other bins.
+The difference in the mean number of genomic annotations colocalized with TAD boundaries of interest was calculated for each set (expected enrichment).
+We repeat this procedure 10,000 times.
+We calculate the permutation p-value by taking the number of times the expected enrichment was greater than the observed enrichment, and dividing by 10000.
+$\alpha = 0.05$ was set to assess statistical significance.
 
@@
 
-## Phenotype-specific networks of protein-protein interactions (PPIs)
+### A modified spectral clustering approach is better suited for TAD boundary detection than other approaches
 
-We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods).
-These were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).
+Our previous work on TAD detection using spectral clustering, implemented as a `SpectralTAD` R package [@Cresswell:2019aa], introduced the concept of the boundary score statistic, adapted here for differential boundary detection.
+Briefly, the boundary score is calculated for each bin by sliding a window across the diagonal of the contact matrix, calculating the eigenvectors of the Laplacian matrix, finding the distance between consecutive eigenvectors (eigenvector gap) and converting them into Z-scores (boundary score, see Methods).
+The boundary score is a continuous measure of TAD boundary likelihood.
 
-We created networks of phenotype-associated genes using PPI information from STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa] databases (see Methods).
-We opted for the use of three types of PPI data.
-First, STRING data filtered by the interaction confidence score above 500 (middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data.
-This filtering step ensured that only high-confidence PPIs were selected.
-The advantage of using filtered data is the reliance on high-quality PPIs.
-The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs are removed from the analysis (see Methods).
-Second, unfiltered STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING full”).
-The advantage of using unfiltered data is that more phenotype-associated genes will have PPI information and, hence, can be analyzed.
-The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs.
-Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, were used.
-The use of different data sources (STRING and BioGrid) and filtering (full or filtered STRING data) was intended to increase the generalizability of our conclusions.
+In contrast to other metrics for boundary identification that rely on finding inflection points of monotonic functions, such as directionality index [@Dixon:2012aa], insulation score [@Crane:2015aa], RobusTAD score [@Dali:2017aa] (Supplementary Material), our boundary score weakly oscillates within TADs and spikes at the TAD boundary (Figure 1).
+This unique behavior enables easy distinction between TAD boundaries and non-TAD boundaries.
+An additional advantage of the boundary score is that its magnitude is directly interpretable as a "boundary strength".
+This is in contrast to other methods which are only interpretable relative to neighboring points.
+We can use this interpretability for parametric modeling of TAD boundary behavior.
+Our previous work has shown that the boundary score is robust to noise, sparsity, and changes in sequencing depth of Hi-C data [@Cresswell:2019aa].
+Thus, the boundary score is ideal when finding differences in TAD boundaries.
 
-**Table 1. Summary statistics of the analyzed categories.** Average coherence estimates are shown. "NA" values indicate that a phenotype was not considered due to the low number of genes with PPIs.
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_1}
+\end{figure}
+**Figure 1.
+Boundary score distinguishes TAD boundaries better than monotonic metrics.** TAD boundary scores calculated with four methods: directionality index, insulation score, RobusTAD and TADCompare boundary scores are shown.
+X-axis - Distance from TAD boundary, measured in bins (40kb each), Y-axis - Score (signed log10 values centered at zero).
+Results from five simulated contact matrices, 40kb resolution, with manually annotated TAD boundaries [@Forcato:2017aa] are shown.
 
-| Category               | Total Phenotypes | Number of SNPs | Number of Genes | Biogrid Coherence | String Coherence | String Filtered Coherence |
-|------------------------|:----------------:|:--------------:|:---------------:|:-----------------:|:----------------:|:-------------------------:|
-| Disease                |                  |                |                 |                   |                  |                           |
-| Metabolic disease      |         5        |       345      |       280       |        0.86       |       0.79       |           0.933           |
-| Neurologic             |         8        |       382      |       343       |        0.89       |       0.511      |           0.861           |
-| Eye disease            |         8        |       270      |       242       |        0.75       |       0.536      |           0.856           |
-| Autoimmune             |        20        |      1415      |       1212      |       0.807       |       0.801      |           0.852           |
-| Cancer                 |        18        |       713      |       610       |       0.863       |       0.643      |           0.838           |
-| Cardiovascular disease |         9        |       336      |       289       |       0.705       |       0.584      |           0.832           |
-| Psychiatric            |        11        |       678      |       611       |        0.83       |       0.537      |           0.817           |
-| Trait                  |                  |                |                 |                   |                  |                           |
-| Metabolic trait        |        10        |       790      |       646       |       0.906       |       0.776      |            0.93           |
-| Eye trait              |         5        |       158      |       109       |         NA        |       0.674      |           0.903           |
-| Anthropometric trait   |        19        |      2143      |       1822      |       0.793       |       0.601      |           0.841           |
-| Cardiovascular trait   |        15        |       728      |       604       |       0.833       |       0.647      |           0.813           |
-| Intelligence           |         5        |       236      |       198       |        0.78       |       0.312      |           0.757           |
+### Differential boundary scores translate to five types of TAD boundary changes
 
-## Degree distribution as a measure of coherence
+Differential boundary score is a measure of the difference between TAD boundaries between two samples.
+This score follows a standard normal centered at 0 (see Methods, Supplementary Figure 1).
+Differential TADs are detected by finding regions with the differential boundary score is greater than 2 (Supplementary Figure 2), which intuitively corresponds to differences with a p-value smaller than 0.05.
 
-A typical network consists of nodes connected by edges [@Barabasi:2004aa].
-In genomics, nodes typically represent genes and edges correspond to some measure of interactions, e.g., gene co-expression or interactions between protein products [@Goncalves:2012aa; @Wang:2012aa].
-In terms of phenotypes, an intuitive expectation is that they will be represented by coherent modules of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
-In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the network [@fortunato2010community; @Radicchi:2004aa].
+We divide TAD boundary changes into five categories (complex, split, merge, shifted, strength change; Figure 2, Supplementary Figure 3).
+A similar strategy was used in Ke et al.
+[@Ke:2017aa].
+A TAD can be **split** between the datasets meaning it exists as a continuous TAD in one and is split into two or more TADs in another.
+In practice, this situation requires two shared TAD boundaries and a differential TAD between them.
+**Merging** is the opposite of splitting and arises when a TAD boundary surrounded by two non-differential TAD boundaries disappears in one of the contact matrices.
+Classification of TAD boundary change as merged and split depends on the reference contact matrix being compared to.
+Finally, a TAD can be split in a **complex** way meaning they are neither split or merged but instead taking on an entirely new structure.
+<!-- If we are comparing two contact matrices and flip the reference matrix, all splits become merges and vice-versa.--> Merged and split boundaries represent the structural change (splitting or merging) of the same TAD as opposed to complex boundaries which we consider to be part of a completely different TAD.
+The "complex", "merge", and "split" TAD boundaries are considered to be the most disruptive changes in the 3D structure of the genome.
 
-Numerous network properties have been defined to describe network structures [@Barabasi:2011aa; @Wang:2012aa].
-A degree of a node, or connectivity, is one of the most fundamental characteristics defining the number of other nodes connected to a given node.
-A collection of degrees of all network nodes form a degree distribution.
-Comparing degree distributions among networks is an intuitive way to gain an understanding of network similarities and differences [@Barabasi:2004aa; @Feldman:2008aa].
-We hypothesized that comparing degree distributions of networks formed by phenotype-associated gene sets may inform us about similarities and differences among phenotypes in terms of coherence of their antecedent and underlying molecular mechanisms.
+A **shifted** TAD boundary is defined as the non-overlapping boundary that lies within five bins (or another user-defined threshold) of a boundary in the contact matrix in which it is being compared to.
+A **strength change** occurs when a TAD boundary is present in both contact matrices, but its differential boundary score magnitude is greater than the differential threshold of 2.
+The other cases are considered to be non-differential TAD boundaries.
+This framework allows us to systematically compare and classify TAD boundary changes.
 
-## Internal vs. external degree distributions as a measure of the coherence
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_2}
+\end{figure}
+**Figure 2.
+Five types of TAD boundary changes.** Complex, split, and merge TAD boundary changes are considered as the major differences, while shifted and strength changes are considered as the minor differences.
 
-<!-- inspired by previous [@Menche:2015aa] and our (Supplementary Figure S3) observation that network size and configuration are irrelevant when considering interactions within the whole PPI network.
+<!--### Selecting cutoffs for TAD boundary and differential boundary detection
+[??? We already referred to supplementary Figure S2. This paragraph is a mess, we don't have panels C and D on that figure]
+Once we have an established boundary score, TAD detection, and differential TAD detection, become a matter of choosing a cutoff that maximizes the number of true TAD boundaries detected while controlling for false detection of TAD boundaries.
+To this end, we tested a range of cutoffs on data with simulated noise and sparsity.
+The viability of each cutoff was calculated using the Youden index which is simply specificity+sensitivity-1.
+We find that for differential TADs, a cutoff of either 1.5 or 2 maximizes our ability to balance true and false positives (Supplementary Figure 2A & 2B).
+Accordingly, we choose 2 as our cutoff in this manuscript and by default in the `TADCompare` package.
+While a cutoff of 1.5 is viable, 2 allows us to be slightly more conservative in choosing differential TADs.
+In the case of normal TAD boundary detection, a cutoff of 2.5 results in the best performance at realistic levels of noise and sparsity (Supplementary Figure 2C & 2D).
 -->
-Given that phenotype-associated networks occur within the global network of PPIs, we developed a ratio-based metric of coherence.
-This metric is inspired by previous work showing that communities within a large network have high internal but low external levels of connectivity [@PhysRevE.81.046110; @Spirin:2003aa].
-Thus, for each phenotype-associated PPI network, we considered the relationship between its internal and external degree distributions.
-The internal degree distribution is defined by considering edges within an isolated network formed by phenotype-associated genes.
-The external degree distribution is defined by considering edges to other genes in the whole PPI network.
-Intuitively, a highly coherent network is expected to have a large number of internal edges but a relatively small number of external edges.
-An extreme example of high coherence would be a fully internally connected network with zero interactions with external genes.
-Conversely, a low-coherence network is expected to have a relatively low number of internal edges compared to external edges, mirroring a random network.
 
-To compare coherences, we measured coherence as the slope of a line through a scatterplot of internal vs.
-external degree distributions.
-That is, we visualize internal (X-axis) vs.
-external (Y-axis) degrees for each phenotype-specific gene on a single plot and fit a regression line through it.
-An example of a network with the highest coherence (full internal connectivity, zero external connections) would be represented by a horizontal line.
-Conversely, an example of a network with the lowest coherence would be a set of non-interacting genes (zero internal connections and full external connectivity) would be represented by a vertical line.
-Consequently, slopes of internal vs.
-external degree distributions of phenotype-associated networks would represent the phenotype-specific levels of coherence.
-In summary, comparing internal vs.
-external degree distributions represents a viable metric to quantify and compare the molecular coherence of phenotype-associated networks.
+### TAD boundaries are highly consistent in both technical and biological replicates
 
-## Networks of KEGG pathways and randomly selected genes as reference for high and low coherence, respectively
+Previous studies have shown that the overlap between TAD boundaries in replicate data ranges from around 60\% to 70\% [@Dixon:2012aa; @Rao:2014aa; @Sauerwald:2018aa2].
+Additionally, technical replicates have been shown to have a slightly higher proportion of shared TAD boundaries (~65%) than biological replicates (~60%) [@Sauerwald:2018aa2].
+We have tested and confirmed these observations by showing that significantly more TADs were non-differential in technical replicates than in biological replicates (73% vs.
+65.7%).
+Similarly, 9.3%/8.1% of boundaries showed significant strength change, while 7.8%/6.1% were shifted in the biological/technical replicates, respectively.
+A similar trend was observed for complex and merge-split boundaries.
+In summary, only 17.2%/12.8% of TAD boundaries were differential in biological/technical replicates, respectively (Figure 3A), confirming the higher stability of TAD structures in technical replicates.
+<!--These results indicate that biological replicates are generally less stable than technical replicates as expected.
+Using our method, we find that the percentage of TADs which are truly differential is much lower than previous methods estimated.
+In total, 83% of TADs are preserved (Shifted, Strength Change, Non-Differential) in biological replicates and 87% in technical replicates.
+Our estimate differs from previous works which showed much less consistency among TAD boundaries.
+Additionally, we find that using our method, the percentage of shared TADs is consistent among different chromosomes.
+Among technical replicates, the percentage of preserved TADs range from 79.6% to 95.5% with a mean of 86.6% (SD = 0.07).
+For biological replicates the numbers ranged from 72.8% to 87.2% with a mean of 82.3% (SD = 0.07).
+These discrepancies suggest other approaches such as TAD calling and then using Jaccard indices to compare TADs (Dixon, Rao) or calling TADs and using an information-based metric (Sauerwald) provide an over-estimate of differentiality.
+Likely reasoning for these issues is a combination of the error caused by variation in TAD callers and a failure to account for small shifts in TAD boundaries which are likely due to noise.
+-->
 
-The coherence of phenotype-associated gene sets should be measured with respect to the reference level of networks with paradigmatically high and low coherence.
-To establish a reference of high coherence, we considered networks from collections of the Molecular Signatures Database (MSigDB v6.2) [@Subramanian:2005aa].
-MSigDB contains sets of genes having various types of functional relationships.
-We found that pathways from the Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa:2000aa] had the smallest average slope corresponding to the highest level of coherence (Supplementary Table S2).
-Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks of high coherence.
-This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathwaays (REACTOME genes) are expected to interact more frequently.
-In our study, KEGG networks were used as the primary reference of high coherence, in parallel with GOCC and REACTOME collections.
-Conversely, as a reference for low coherence, we randomly sampled genes from a pool of genes in a given PPI database.
-These random gene lists represent expected network coherence arising by chance.
-The slopes of KEGG and random networks represent reference levels for high and low coherence, respectively.
+### TAD boundaries are more similar within cells than tissues
 
-## Network size and configuration hinders direct comparison of degree distributions
+Previous research showed that TADs are largely invariant across cell lines and to a lesser extent tissue types [@Pope:2014aa; @Rao:2014aa; @Schmitt:2016aa].
+However, the types of TAD boundary changes remained undefined.
+We compared Hi-C matrices of 7 different cell-lines and 18 different tissue types [@Schmitt:2016aa] (Supplementary Table 3).
+In total, the average percentage of differential TAD boundaries was significantly less in cell lines (22.5%) than tissue samples (39.7%, Figure 3B).
+As expected, these percentages were higher than those for biological (17.2%) and technical replicates (12.8%).
+These results suggest that the variability of TAD boundaries mirrors the homogeneity of data types (technical replicates, biological replicates, cell lines, and tissues, in that order).
+<!--Consequently, clustering cell type- and tissue-specific Hi-C datasets using Pearson correlation of their genome-wide boundary score showed replicates from cell types clustering with each other, in contrast to tissue-specific replicates (Supplementary Figure).
+[???
+More the current Figure 6 to supplemental] The mean Pearson correlation of boundary scores for cells was 0.89 and 0.52 for tissues.
+The larger variability in tissue-specific Hi-C datasets likely comes from the fact that tissue samples contain mixtures of cells while cell lines provide homogeneous samples.
+-->
 
-Direct comparison of degree distributions of the networks formed by phenotype-associated genes is hindered by the fact they depend on network size and configuration.
-An intuitive example is to consider a ring and a fully connected network of size 10 and 20 nodes, respectively (Figure 1).
-The degree distributions of the ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).
-Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have degree 9, while all nodes in the 20-node network have degree 19, Figure 1B).
-<!--Normalizing the degree distributions of the fully connected networks by dividing them by network size makes them identical as would be expected for networks with similar coherence.
-However, similar normalization of ring networks will make their degree distributions different (0.22 for the 10-node network and 0.11 for the 20-node network), leading to the illusory impression of differences in coherence between intuitively equally coherent networks.
--->These observations highlight the difficulty in comparing degree distributions of phenotype-associated gene networks due to two facts: 1) network sizes differ, and 2) network configuration (fully connected, ring, or intermediate connectivity) is unknown.
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_3}
+\end{figure}
+**Figure 3.
+Biological replicates and cell lines have more differential TADs than technical replicates and tissues, respectively.** Differential TADs were calculated between Hi-C datasets of biological and technical replicates (Panel A, HCT-116 cell line, 50kb resolution, chr 1-22 [@Rao:2017aa]) and between cell lines and tissues (Panel B, various cell lines, 40kb resolution, chr1-22, [@Schmitt:2016aa]).
+Types of TAD boundary changes were recorded, and the proportions of TAD differences for each type were summarized across chromosomes.
+[??? Check figure!!! We first present replicates, then cell lines. Currently, the panels are swapped]
 
-+-------+
-| ![](Figures/Fig1_network_distributions.png)      |
-| **Figure 1.
-Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
-While degree distributions of the similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
-|
-+-------+
+<!--
+Overall, we observed a higher level of non-differential TAD boundaries [???
+Changed from "conservation" - is this change correct?] in cell lines (Mean = 70.5%, SD = 0.08%) than in tissues (Mean = 54.6%, SD = 0.04%).
 
-To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
-We observed negative correlation of network size with the slopes of KEGG and random networks (-0.37/-0.85, respectively, Figure 2, Supplementary Table S2).
-These observations confirm the notion that coherence estimation using internal vs.
-external degree distributions should be controlled for network size.
+To better understand the similarity of the TAD structures across cell and tissue types, we clustered cells and tissues by Pearson correlation of their genome-wide boundary score profiles (Figure 6).
+The mean correlation for cells was 0.89 and 0.52 for tissues.
+While replicates of cell line samples were clustered with each other, this was not the case with tissues.
+To examine the mechanisms behind poor clustering, we found differential boundaries and calculated the percentage of each type.
+Replicates which failed to cluster tended to have a high level of shifted TAD boundaries.
+For instance, 21.2% TAD boundaries were shifted between replicates of lung samples, while 22.0% were shifted between replicates of pancreas samples.
+This was higher than the overall mean percentage of shifted boundaries (14.7%).
+All comparisons are summarized in Supplementary Table 2.
+These results suggest that a large percentage of inconsistency between certain tissue-based replicates may be due to small shifts in boundaries.
+The implications of these findings are two-fold.
+One is that methods which cannot identify shifted TAD boundaries will overstate the degree of difference between tissues.
+This is because they will incorrectly consider shifts as differential boundaries without considering the distance of boundary movement.
+The second is that there is a need for methods which can take a set of replicates and correct for the high number of shifted boundaries among replicates.
+One approach for achieving this is our proposed consensus TAD boundary calling method.
+With enough replicates it is expected that the consensus TAD boundary for a given region will converge to the true boundary, making shifts have less of an effect.
+-->
 
-+-------+
-| ![](Figures/Size_Slope_Plot.png)      |
-| **Figure 2.
-Network size is inversely associated with coherence.** Slopes of internal vs.
-external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
-|
-+-------+
+### Each type of differential TAD boundaries is associated with different levels of epigenomic enrichment
 
-## Network size-independent estimation of coherence
+To understand the biological relevance of types of TAD boundary changes, we identified differences between the GM12878 and IMR90 cell lines (chr1-22, 40kb resolution, [@Schmitt:2016aa]) and categorized them according to the type of change.
+For each change type, we assessed the number of overlapping peaks and calculated the enrichment of four genome annotation marks known to co-locate with TAD boundaries - CTCF, RAD21, insulators, and heterochromatin states.
 
-Given the dependence of internal and external degree distributions on network size, we designed a simple strategy to account for it.
-Briefly, for each phenotype-specific network of a given size, we create size-matched references of high and low coherence.
-Specifically, we select the ten KEGG pathways closest in size and ten random networks matched in size as references for high/low coherence for the corresponding phenotype.
-Consequently, we use the slopes of internal vs.
-external degree distributions created by these size-matched references to normalize the slope of the phenotype-specific network to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed removes the dependency between coherence and network size (Supplementary Figure S1).
+We found that non-differential boundaries had a higher average number of overlapping peaks for all four marks, followed by "strength change" TAD boundaries (Figure 4A).
+Similarly, enrichment of non-differential TAD boundaries was the most significant (Figure 4B).
+Notably, the number of peaks for each mark was highly variable on “strength change” TAD boundaries (Figure 4A), suggesting their biological relevance is less certain.
+Similarly, "shifted" TAD boundaries had the lowest average number of peaks, suggesting that they may be detected due to noise and, consequently, be less biologically significant.
+In contrast, "complex" and "merge-split" TAD boundaries had a moderate number of overlapping peaks and were moderately enriched in them (Figure 4).
+These results highlight the varied biological relevance of different types of TAD boundary changes and suggest "complex" and "merge-split" changes are biologically important alterations of TAD structure.
 
-## Results derived from different PPI data are largely consistent
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_4}
+\end{figure}
+**Figure 4.
+Non-differential TAD boundaries are more enriched for selected genome annotation marks than other types of differential TAD boundaries.** Differential TAD boundariess were called between GM12878 and IMR90 cell lines and categorized based on differential boundary type.
+(A) Number of peaks at TAD boundaries and (B) permutation p-values (-log10) are shown.
+Data from [@Schmitt:2016aa], 40kb resolution, chr 1-22.
 
-The potential drawback of this reference size-matching approach is that phenotypes with large network sizes (e.g., "Obesity-related traits", 779 genes) may not have the size-matched reference of high coherence (e.g., the largest KEGG pathways is 374 genes).
-Consequently, coherence estimation for such networks may be affected by the smaller size of reference of high coherence.
-Therefore, we use networks from GOCC and REACTOME collections that have larger maximum number of genes (Supplementary Table S2) in parallel with KEGG networks as a reference of high coherence.
-We found that coherence estimations using KEGG, GOCC, and REACTOME networks as references of high coherence are highly concordant (Average Pearson correlation coefficient > 0.59, Supplementary Figure S2).
-The use of size-matched networks from KEGG, GOCC, and REACTOME collections enables robust estimation of coherence for phenotype-specific networks of different sizes.
-Given the similarity of the results obtained with different PPI databases and references of high coherence, we present results obtained with STRING filtered PPI database and KEGG networks as a reference of high coherence, with other observations presented in Supplementary Table S3 and Supplementary Figures S3-S8.
+### Each type of differential TAD boundaries is associated with distinct biological functionality
 
-## Coherence of all phenotypes significantly differs from random
+To test the biological significance of different types of TAD boundary changes, we compared mesenchymal stem cells (MSC) against neural progenitor cells (NPC) [@Schmitt:2016aa].
+Altogether, we found that the vast majority of boundaries are either complex (38.6\%) or non-differential (32.6\%).
+Shifted (17.5\%), merge-split (7.7\%) and strength change (3.5\%) were less common (Figure 5A).
+We investigated the enrichment of genes in the proximity of each type of differential TAD boundary in biological processes and other gene ontology- and pathway types using GREAT [@McLean:2010aa] (see Methods).
+As NPCs are more advanced on differentiation path than MSCs, we expected that TAD boundaries changed between them would be associated with genes responsible for neural development-related processes.
+Indeed, genes around "merge" and "complex" TAD boundary changes, as well as the "non-differential" TAD boundaries were enriched in a variety of developmental processes (e.g., "cellular developmental process", etc.), including neural-specific ("nervous system development", Figure 5A).
+Notably, "split" TAD boundary changes were not enriched in these processes, indicating the importance of directionality of TAD boundary changes.
+Genes around "merge" and "non-differential", but not "complex", TAD boundaries were enriched in differentiation-related processes (e.g., "positive regulation of cell differentiation"), while "forebrain radial glial cell differentiation" and "neural tube development" processes were exclusively enriched in genes around "merged" TAD boundaries (Figure 5A).
+In this case, "merge" indicates boundaries enriched in the NPC cell-line causing a separation of TADs in MSC and "split" indicates a split in NPC caused by a boundary enriched in MSC.
+As expected, genes around "noisy" TAD boundary changes ("shifted" and "strength change") lacked enrichment in any biological processes (Figure 5A, Supplementary Table 4).
+These results emphasize the importance of classifying TAD boundary changes into distinct patterns which tend to be associated with distinct biological functionality.
 
-To estimate the significance of coherence measures, we developed a permutation-based approach (see Methods).
-Notably, we observed coherence for all diseases to be highly significant (permutation p-value < 0.002, Supplementary Table S3).
-The distribution of coherence estimates was skewed towards 1.00 (highest coherence), with the median coherence being 0.86.
-"Obesity-related traits" and "Height" were on the lowest range of coherence estimates (0.42 and 0.54, respectively).
-"Weight", "Telomere length", and "Corneal structure" were on the highest range of coherence (1.00, Supplementary Table S3).
-These results indicate that coherence of phenotype-specific networks is generally high and significantly different from random.
+To further test whether different types of TAD boundary changes reflect biology of an experimental system, we used  post-auxin treatment time course experiment from Rao et al.
+2017 study (HCT-116 cell line, 40kb resolution, 20, 40, 60, and 180 minutes following auxin withdrawal, 4 replicates at each time point) [@Rao:2017aa].
+Auxin treatment eliminates CTCF binding genome-wide; consequently, the majority of TAD boundaries should be absent and gradually re-appear following auxin withdrawal.
+To identify biological processes associated with re-appearing of TAD boundaries, we compared first and last time points (20 and 180 minutes) following auxin withdrawal.
+As TAD boundaries were reported to be enriched in housekeeping genes [@Jin:2013aa], we expected genes around appearing TAD boundaries to be enriched in general cellular processes.
+Indeed, the vast majority of differential TADs were complex (41.4\%) and non-differential (34.7\%) (Supplementary Figure 4).
+We found that only genes around "non-differential" and "complex" TAD boundary changes showed some level of enrichment (Supplementary Figure 4, Supplementary Table 5).
+[???
+Some tabs are empty?] As expected, "metabolic processes" and various developmental and housekeeping processes were specifically enriched in genes around complex TAD boundary changes, while cyclic AMP synthesis and metabolic processes were enriched in genes around "non-differential" TADs.
+From these results, we show that `TADCompare` can correctly classify non-essential TAD boundary changes ("shifted", "strength change") and detect distinct TAD boundary changes associated with shared and unique biological processes.
 
-## Metabolic and intelligence-related traits as examples of high/low coherence
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_5}
+\end{figure}
+**Figure 5.
+Gene enrichment analysis of differential TADs.** Differential TADs were called between A) neural progenitor cell (NPC) and mesenchymal stem cells (MSC) (chr1-22, 50kb resolution, [@Rao:2017aa]), and B) between the first and last time point [???
+Adjust, that is wrong] of auxin-treated cells from the HCT-116 cell-line (chr1-22, 40kb resolution, [@Schmitt:2016aa].
+For each type of TAD boundary change, -log10-transformed enrichment p-values (`rGREAT`, see Methods) are shown as heatmaps.
+The top 30 gene ontology biological processes, in terms of average enrichment, are shown.
 
-We estimated coherence of 53 traits from five categories (Figure 3, Supplementary Figures S3-S5).
-Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
-Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose- and cholesterol-related traits (metabolic).
-Notably, metabolic and eye trait categories had the highest overall coherence (mean coherence 0.90 and 0.93, respectively).
-Several notable examples of low coherence include "Obesity-related traits", "Height", and "Longevity" (anthropometric traits), blood count-related traits (cardiovascular).
-Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.76).
-These results highlight the relatively good knowledge of the genetics of metabolic traits which corresponds to highly coherent networks.
-Conversely, current knowledge about the genetics of intelligence-related traits appear insufficient as reflected in the low coherence of the corresponding networks.
+### Time course analysis framework
 
-+-------+
-| ![](Figures/Fig3_summary_traits.png)      |
-| **Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).   |
-+-------+
+Time course analysis of TADs refers to the analysis of TAD boundary dynamics over time.
+The quantitative nature of boundary score allows us to monitor its changes at TAD boundaries across any number of time points.
+We recommend taking a union of TAD boundaries detected at each time point and monitor boundary score changes for each TAD boundary.
+Monitoring TAD boundary scores across time points provides an opportunity to quantify patterns of TAD boundary changes.
 
-## Metabolic and neurologic diseases have high coherence
+<!-- Using boundary scores, we can also go region-by-region and identify time points which contain TADs.
+For each region, we use the pattern of change over time to classify them into four temporal TAD boundary categories (early appearing, late appearing, highly common and dynamic).
+These categories give us a simple summarization of the evolution of each TAD boundary over time.
+An additional function of `TADCompare` is the ability to plot TAD boundaries over time and visualize the overall trend.
+This gives users the ability to visually assess their data and determine the effect of time.
+We also give users the ability to view heatmaps giving a region-by-region visualization of changes with the type of temporal boundary type labeled.
+All temporal TAD functions in `TADCompare` are fully integrated with consensus boundary calling meaning if we have multiple replicates at a given time point it is possible to use consensus boundary scores instead of traditional boundary scores.
+-->
 
-We further estimated coherence of 80 diseases from seven categories (Figure 4, Supplementary Figures S6-S8, Supplementary Table S3).
-Metabolic diseases had high overall coherence (average coherence 0.93), with "Obesity" and "Type 2 diabetes" being the most coherent (0.99 and 0.94, respectively).
-Similarly, the average coherence for neurologic diseases was 0.86, with "Alzheimer's disease" and "Parkinson's disease" being highly coherent (0.94 and 0.92, respectively).
-The high coherence of metabolic and neurologic disease networks highlight the relatively good knowledge of their genetics.
+Using the boundary score cutoff of 3 for TAD boundary definition, we define six patterns of temporal TAD boundary changes (adapted from [@Zhou:2019aa], Table 2, Figure 6).
+_Highly common_ TAD boundaries refer to boundaries present across all time points or in three out of four time points.
+_Early appearing_ TAD boundaries switch from non-boundary to boundary at second time points and stay as boundaries for the rest of the time points.
+Conversely, _early disappearing_ TAD boundaries switch from boundary to non-boundary at the second time point and stay as non-boundaries.
+_Late appearing_ TAD boundaries switch from non-boundaries to boundaries at the last or the second to last time point.
+Conversely, _late disappearing_ TADs switch from boundaries to non-boundaries at the last of the second to last time point.
+Finally, _dynamic_ TAD boundaries are those which have inconsistent boundary status and do not follow any of the aforementioned patterns (Figure 7).
+These six patterns of temporal changes can be easily adapted for a larger number of time points.
 
-## Intermediate coherence of autoimmune, cancer, and eye diseases
+**Table 1.
+Six patterns of temporal TAD boundary changes.** Each column corresponds to a point in time.
+"1" refers to the presence of a TAD and "0" refers to the absence of a TAD at considered time point.
+Total column shows percentage of occurrences in CTCF degradation-recovery time course, HCT-116 cell line, chr1-22 [@Rao:2017aa].
 
-"Rheumatoid arthritis", "Ankylosing spongylitis", "Psoriasis" were the most coherent in the "Autoimmune" category (0.94, 0.91, 0.89, respectively).
-On the other end of the coherence spectrum, "Multiple sclerosis" and "Allergic rinitis" were the least coherent (0.76 for both).
-The average coherence of autoimmune diseases was high, 0.85, with all but two autoimmune diseases having high coherence (> 0.80).
-Diseases in the "Cancer" category were similarly highly coherent (average coherence 0.84).
-"Non-small lung cancer", "Testicular germ cell tumor", "Epithelial ovarian cancer" were the most coherent (0.94, 0.91, 0.88, respectively) while "Basal cell carcinoma", "Breast cancer" "Hodgkin's lymphoma" were the least coherent (0.64, 0.78, 0.78, respectively).
-Overall, networks of all but "Basal cell carcinoma" cancers had high coherence (> 0.78).
-Diseases in the "Eye disease" category were more heterogeneous, with coherence ranging from 0.71 for "Myopia (pathological)" disease to 0.97 for "Refractive error".
-The average coherence of diseases in the "Eye disease" category was relatively high (0.86).
-These results suggest good overall knowledge of the genetics of autoimmune, cancer, and eye-related diseases and highlight limited knowledge for some diseases in these categories.
+| Temporal TAD Type  | Time Point 1 | Time Point 2 | Time Point 3 | Time Point 4 | Total (% Occurrence) |
+|--------------------|:------------:|:------------:|:------------:|:------------:|:--------------------:|
+| Highly Common      | 1            | 1            | 1            | 1            | 326 (17.35%)         |
+|                    | 1            | 0            | 1            | 1            |                      |
+| Early Appearing    | 0            | 1            | 1            | 1            | 184 (9.79%)          |
+| Early Disappearing | 1            | 0            | 0            | 0            | 133 (7.08%)          |      
+| Late Appearing     | 0            | 0            | 1            | 1            | 1047 (55.72%)        |
+|                    | 0            | 0            | 0            | 1            |                      |
+| Late Disappearing  | 1            | 1            | 0            | 0            | 79 (4.20%)           |
+|                    | 1            | 1            | 1            | 0            |                      |
+| Dynamic            | 1            | 0            | 1            | 0            | 110 (5.86%)          |
+|                    | 1            | 0            | 0            | 1            |                      |
 
-+-------+
-| ![](Figures/Fig4_summary_diseases.png)      |
-| **Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).   |
-+-------+
+\begin{figure}[H]
+\begin{center}
+\includegraphics[width=6in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_6}
+\end{center}
+\end{figure}
+**Figure 6.
+Six patterns of TAD boundary score changes across time course.** Average trajectories for each pattern of boundary score change are shown.
+The red horizontal line indicates the cutoff for TAD boundary detection.
+HCT-116 cell line, 40kb resolution, chr 1-22.
 
-## Psychiatric and cardiovascular diseases as examples of low coherence
+<!-- An alternative approach to classifying boundaries is simply clustering the trend of boundary scores.
+First, we place each boundary score in a matrix where each column is a time point of each row is a TAD boundary.
+Since we are specifically interested in the trend profile, we z-normalize the boundary scores on a boundary-by-boundary basis.
+This normalization transforms each row into a summary of the trend over time for the corresponding boundary score.
+The matrix is then transformed into a Euclidean distance matrix, and hierarchical clustering with Ward's minimum variance criterion is used to assign clusters.
+Users can then choose their desired number of clusters and choose their own labels.
+-->
 
-Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.82 and 0.83, respectively).
-The corresponding coherence estimates were highly heterogeneous.
-Among cardiovascular diseases, "Atrial fibrillation", "Coronary artery disease" showed the highest level of coherence (0.98 and 0.89, respectively), while "Sudden cardiac arrest" and "Hypertension" had the lowest coherence (0.70 and 0.77, respectively).
-Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.93 and 0.89, respectively).
-Notably, "Schizophrenia" and "Major depressive disorder" showed the lowest coherence not only in the "Psychiatric disease" category (0.67 and 0.75, respectively) but among all phenotypes.
-Other psychiatric disorders, such as "Bipolar disorder", "Autism", "ADHD", showed intermediate coherence estimates (0.87, 0.83, 0.78, respectively).
-In summary, these low coherence estimates highlight the need for further research on genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms, and suggests "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies.
+### Temporal TAD boundary types are associated with different levels of epigenomic enrichment
+
+To evaluate the biological relevance of temporal patterns of TAD boundaries, we used post-auxin treatment time course experiment introduced above.
+Briefly, HCT-116 cells were treated with auxin to eliminate TAD boundaries, and Hi-C measures were obtained at 20, 40, 60, and 180 minutes following auxin withdrawal and subsequent TAD boundary reappearance [@Rao:2017aa].
+Accordingly, we expected to detect some number of highly common TAD boundaries (already existing at 20 minutes) and boundaries appearing at different stages of post-auxin withdrawal (early/late appearing).
+Conversely, dynamic and early/late disappearing TAD boundaries should be rare and may potentially constitute noise in TAD boundary detection.
+
+Boundary scores were calculated for auxin-treated cells 20, 40, 60, and 180 minutes after withdrawal.
+Taking the union of TAD boundaries (boundaries detected at one or more time points), we calculated temporal patterns for each boundary.
+We found that the vast majority of boundaries were late appearing TADs (55.7%) (Table 2, Figure 5B).
+Early appearing TADs (9.8%) and highly common TADs (17.3%) made up most of the other TADs present at the end of the time course.
+~20% of TAD boundaries were highly common, i.e., resistant to auxin treatment, a number similar to previous works [@Nora:2017aa].
+Meanwhile, 5.9% of TAD boundaries were dynamic, 7.1% were early disappearing, and 4.2% were late disappearing, highlighting potential errors in TAD boundary detection.
+In summary, some TAD boundaries can be detected at 20 minutes post-auxin treatment and remain present through all time points; however, the timing of TAD boundary restoration varies by TAD.
+
+To test whether TAD boundaries associated with different temporal patterns have different functional roles, we investigated their overlap with and enrichment in the common marks of TAD boundaries (CTCF, RAD21, insulators, heterochromatin, Figure 7A).
+For highly common, early- and late appearing TAD boundaries, we observed more overlaps with CTCF and RAD21 sites, insulator and heterochromatin states (Supplementary Table 7).
+Similarly, these types of TAD boundaries were highly enriched in the aforementioned genomic annotations (Figure 7B).
+Conversely, dynamic, early and late disappearing TAD boundaries showed less overlap with CTCF, RAD21, insulator and heterochromatin marks, and were less enriched in them.
+These observations suggest that disappearing and dynamic TAD boundaries are likely detected due to noise in the data, while TADs appearing after auxin treatment expectedly represent the biologically relevant signal.
+
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_7}
+\end{figure}
+**Figure 7.
+Common and appearing TAD boundaries show stronger enrichment in known epigenomic marks.** The number of peaks at TAD boundaries (A), and permutation p-values (B) within 50kb of boundaries in each temporal classification are shown.
+Hi-C data from [@Rao:2017aa], 50kb resolution, HCT-116 cell-line, chr 1-22.
+
+### Temporal TAD boundary types are associated with distinct biological functionality
+
+We further investigated these results using gene enrichment analysis (Supplementary Table 8) for temporal boundary types.
+We found that, with a few exceptions, all significant GO Biological pathways were enriched in late appearing TADs or early appearing TADs (Figure 5B, Supplementary Table 8), which make up the majority of TAD boundaries (Table 2, Figure 5B).
+Both early and late appearing TAD boundaries were enriched in metabolism-related processes, such as "cellular metabolic process", "oxidation-reduction process".
+Late appearing TADs, on the other hand, were enriched in "cellular component organization", "protein complex biogenesis" and the like processes (Figure 5B).
+These results are expected as cells may be activating metabolic and biogenesis pathways to recover after the destruction of TAD boundaries by auxin.
+These results confirm that `TADCompare` can accurately classify biologically relevant temporal TAD boundary changes and discern them from noisy changes.
+
+### Consensus boundary score for defining robust TAD boundaries across multiple Hi-C datasets
+
+The sizeable proportion of noisy "shifted" and "strength change" TAD boundary changes across Hi-C datasets (Figure 3) highlights the need to identify TAD boundaries that are robustly detected.
+The consensus boundary score, defined as the median of boundary scores across replicates, addresses this challenge.
+Intuitively, higher consensus boundary scores correspond to TAD boundaries supported by evidence from multiple replicates (Table 1).
+This is in contrast to a union of TAD boundaries, where TAD boundaries detected in at least one Hi-C dataset are pooled together.
+Consensus boundary scores allow us to filter out boundaries with insufficient support from multiple replicates, thus "denoise" the detected TAD boundaries.
+Given the fact that boundary scores are log-normally distributed (Supplementary Figure 1, Supplementary Methods), the consensus boundary scores will also be asymptotically normal.
+The consensus boundary score can be used as a proxy for the normal boundary score for the analysis of replicated Hi-C datasets.
+Consequently, the consensus boundary scores may be compared to define TAD boundary changes between groups of replicated Hi-C datasets.
+
+**Table 2.
+Consensus (aka median) boundary score is supported by high boundary scores from multiple replicates.** Examples of boundary scores across five regions in three replicates, and the corresponding consensus boundary score.
+Both union and consensus TAD boundaries are calculated using a cutoff of 3.
+
+|| Boundary Score 1 | Boundary Score 2 | Boundary Score 3 | Consensus Boundary Score | Union TAD Boundary?  | Consensus TAD Boundary? |
+|---------|------------------|------------------|------------------|--------------------------|----------------------|-------------------------|
+||         1        |         2        |         1        |              1           |             No       |                 No      |
+||         3        |         2        |         1        |              2           |             Yes      |                 No      |
+||         5        |         5        |         4        |              5           |             Yes      |                 Yes      |
+||         3        |         3        |         3        |              3           |             Yes      |                 Yes     |
+||         6        |         0        |         0        |              0           |             Yes      |                 No      |
+
+### Consensus TAD boundaries are supported by strong biological evidence
+
+To investigate the biological relevance of TAD boundaries defined using consensus boundary score, we defined consensus TAD boundaries across 7 cell-lines (17 matrices total) [@Schmitt:2016aa].
+These boundaries represent cell type-invariant TAD boundaries supported by evidence from multiple datasets.
+Regions of the genome were separated into three categories based on the level of their consensus boundary score (<2, 2-4 and >4).
+In total, there were 65,336 non-gap [???
+What's this?] bins (40kb resolution).
+Expectedly, the majority (62,791 bins, 96.1% of all regions) were in the <2 category, 2,032 (3.1%) regions were in the 2-4 category, and 513 (0.8%) regions were in the >4 category.
+We assessed the number of overlapping peaks and the enrichment of CTCF, RAD21, insulators, and heterochromatin states in different categories of consensus TAD boundaries.
+Expectedly, we observed increasing average number of peaks overlapping TAD boundaries selected at more stringent consensus boundary score thresholds (Figure 8, Supplementary Table 6).
+Similarly, TAD boundaries with higher consensus boundary scores have stronger enrichment in genome annotations.
+These results suggest that TAD boundaries with higher consensus boundary scores (i.e., supported by evidence from multiple Hi-C datasets) are more biologically relevant.
+
+<!-- Based on these results, we chose a default consensus boundary score cutoff of 3.
+This cutoff is used for all analyses in this work and is the default in the `TADCompare` package.
+In practice, the cutoff of 3 is not necessarily the only choice.
+For instance, a cutoff of 2 would include weaker strength TAD boundaries at the risk of more false positives and a cutoff of 4 would ensure only strong boundaries at risk of more false negatives.
+Due to the normality (Supplementary methods) of consensus boundary scores, we can interpret these cutoffs in terms of standard deviations from the mean boundary score.
+For instance, a boundary score of 3 is 3 standard deviations greater than the mean boundary score.
+-->
+
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_8}
+\end{figure}
+**Figure 8.
+TAD boundaries defined at higher consensus boundary score thresholds show stronger overlap with and enrichment in known epigenomic marks.** TAD boundaries were classified based on the range of their consensus boundary score.
+Enrichment of genomic factors known to occur near TAD boundaries was calculated.
+(A) The number of TAD peaks within 40kb of TAD boundaries with the corresponding consensus score range and (B) the permutation p-values for each score range are shown.
+Data from seven cell lines, chr1-22, 40kb resolution, [@Schmitt:2016aa].
+
+### The union of TAD boundaries is supported by weaker biological evidence than consensus boundaries
+
+The union of TAD boundaries called in individual Hi-C datasets represents an alternative method of defining TAD boundaries across multiple datasets (Table 1).
+The union method may be useful for analysis of time course data, where TAD boundaries are expected to change across individual datasets.
+We hypothesized that the union method would select for the less biologically relevant set of TAD boundaries because many may be detected due to noise in Hi-C data.
+<!--
+One advantage of the consensus boundary score is its ability to filter out weak TAD boundaries that only occur in a minority of replicates and are likely artifacts of the sequencing process.
+Traditionally, researchers have called TAD boundaries on individual contact matrices, taken a union of all boundaries called and then ran analyses.
+Contrary to what one would expect, this approach may actually reduce the quality of TAD boundary data.
+By adding in new replicates, we can potentially introduce noisy boundaries which are unique to each specific replicate and not biologically significant.
+In fact, we find a decline in quality of TAD boundaries as more replicates are added due to these noisy boundaries (Supplementary Figure 3).
+Consensus differs from the union method in that it automatically filters out TAD boundaries which only occur in a small portion of replicates while union calls all TADs that appear in at least one replicate (Supplementary Table 1).
+-->
+
+To evaluate the biological relevance of TAD boundaries called using both methods, we call consensus and union TADs on a set of replicates (four cell lines, 40kb resolution, 3 replicates each, data from [@Schmitt:2016aa]).
+Consensus scores were calculated separately for each cell line among the 3 replicates.
+Expectedly, the consensus method filtered out 38% of TAD boundaries (4906 vs.
+3059, Supplementary Figure 5), suggesting that many TAD boundaries are detected in single datasets.
+We found that TAD boundaries called using consensus boundary score overlapped significantly more with CTCF sites (P = 0.0006) and RAD21 (P = 0.0002) than those called using the union method (Figure 9A).
+While the enrichment results were similar for consensus- and union-defined TAD boundaries, consensus TAD boundaries were more significantly enriched in "heterochromatin" (Figure 9B).
+Together with previous observations (Figure 6), these results strengthen our conclusion that consensus boundary scores are more effective in removing "noisy" TAD boundaries that otherwise would be captured using the union method.
+
+\begin{figure}[H]
+\center
+\includegraphics[width=7in,keepaspectratio]{Finished_Figures/Final_Figures//Figure_9}
+\end{figure}
+**Figure 9.
+Consensus TAD boundaries show stronger overlap with and enrichment in known epigenomic marks than union of TAD boundaries.** (A) Number of peaks at TAD boundaries and (B) permutation p-values (-log10) are shown.
+Data from [@Schmitt:2016aa], four cell lines, 40kb resolution, chr 1-22.
+
+<!--
+[??? That is too much. don't report it]
+An alternative, much more conservative, the approach is only to use TAD boundaries called in all or the majority of replicates or cell lines.
+This approach gives a much smaller set of TADs but with a reduction in noisy one-off-boundaries.
+To test the effectiveness of consensus boundary calling versus this method, we called TADs across 17 cell lines and 20 tissue samples and compared enrichment of genomic features at boundaries.
+TAD boundaries were separated based on the number of occurrences and binned into 3 categories (1-5, 6-10, and 11-20 occurrences).
+Among cells, we find a non-significant increase in enrichment of CTCF, RAD21, and heterochromatin and a significant increase in insulators (P = 0.0460) between consensus and the highest quality 11-20 group.
+For tissues we find a significant increase in CTCF, RAD21, heterochromatin and insulators (P = <0.0001, P = <0.0001, P = 0.0002, P = 0.0006) between consensus and the 11-20 group.
+In general, we find that TADs occurring in a large number of tissue samples or cell lines have higher enrichment than more rare TAD boundaries suggesting an even stronger need for filtering out weak boundaries.
+Consensus boundary score also allows us to get a much more refined set of TADs in cells (1158 vs.
+1268) and tissues (119 vs.
+557) when compared with taking the majority of occurrences.
+These results show consensus boundary score is a powerful method for summarizing TADs across many different sources and provides an improvement over naive methods such as using a union of TADs based on number of occurrences (Supplementary Figure 4).
+-->
+
+### Runtime performance of TADcompare
+
+When ran on data from [@Rao:2014aa], without parallelization, both consensus TAD calling and differential TAD detection were exceptionally fast.
+In total, for the entire genome, differential TAD detection took ~6 seconds on 100kb data, ~9 seconds on 50kb data, ~17 seconds on 25kb data and ~312 seconds on 10kb data.
+In the case of consensus TAD calling, `TADCompare` took ~17 seconds to run on 50kb data for 4 matrices, ~32 seconds for 8 matrices and ~45 seconds for 12 matrices.
+On 10kb data, it took ~611 seconds to run for 4 matrices, ~1152 seconds for 8 matrices and ~1680 seconds for 12 matrices.
+For a full summary of runtimes across all resolutions, see Supplementary Figure 6.
 
@@
 
-We developed a general-purpose method to measure and compare coherence estimates of the molecular mechanisms in networks of genes.
-Intuitively, gene networks with high internal and low external interactions would be considered highly coherent; consequently, coherence is defined as the ratio between the internal and external degree distributions (the number of interactions per gene).
-We normalized coherence estimates to the range of high and random coherence, exemplified by KEGG pathways and random networks.
-To decouple coherence from its dependence on network size, we used KEGG and random networks matched in size to phenotype-specific networks.
-Applied to the analysis of 133 disease- and trait-associated gene sets, we found an unexpected inverse association of phenotype complexity with coherence.
-Consequently, complex traits like "Height", "Intelligence", "Longevity" had overall low coherence while metabolic traits and diseases had high coherence.
-Diseases in "Psychiatric" category had overall low coherence, with "Schizophrenia" being least coherent.
-Our method allows one to quantify the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.
+The initial development of Hi-C technologies focused on investigating individual genomes.
+While several key properties have been discovered (chromosome territories, A/B compartments, TADs, chromatin loops), the next steps include investigating changes in the 3D structure across multiple conditions.
+We [@Stansfield:2018aa;@Stansfield:2019aa] and others [@Djekidel:2018aa;@Lun:2015aa] started to develop methods for comparative analysis of the 3D structures.
+However, to our knowledge, no methods are available for differential analysis of TAD boundaries.
+In this work, we introduce a method for differential TAD boundary analysis, including time course, that supports replicated Hi-C data.
+The method is based on a novel boundary score metric that provides a continuous measure of TAD boundary likelihood [@Cresswell:2019aa].
+We introduce unique terminology for classifying differential and temporal TAD boundary changes.
+We show that our approach is robust and effective at identifying distinct biology associated with different types of TAD boundary changes.
+Our method is implemented in the `TADCompare` R package available on Bioconductor, filling a vital gap in intuitive R-based software for TAD detection and comparison.
 
-Our approach is similar in spirit to the well-researched community detection problem [@danon2005comparing; @fortunato2010community; @Radicchi:2004aa].
-However, we approach this problem from a different angle by asking whether a given network as a whole can be considered a community within the global PPI network.
-This may be considered a drawback as it has been shown that many networks consist of miniature communities known as motifs [@Milo:2002aa], or graphlets [@Przulj:2007aa].
-While the investigation of motif enrichment [@Milo:2002aa] is a viable approach to characterize a network, our goal was to develop a unified measure of coherence that can be compared across phenotypes.
+The boundary score concept developed in our work addresses three main problems: differential TAD boundary detection, time course analysis of TAD boundary changes, and consensus TAD boundary calling.
+Yet, it has a wider scope of applications.
+Future work will expand the utility of boundary score by developing a similarity/reproducibility score to measure the agreement between (multiple) Hi-C matrices, in the same vein as HiCRep [@Yang:2017aa],<!-- which uses spectral methods and other similarity score metrics such as--> Selfish [@Ardakany:2019aa], GenomeDISCO [@Ursu:2017aa], HiC-Spector [@Yan:2017aa], QuASAR-Rep [@Sauria:2015aa]<!-- which use different approaches-->.
+Furthermore, for differential TAD boundary detection, our method is still limited to the comparison of two profiles of (consensus) TAD boundary scores.
+This approach will eventually be expanded to include comparisons of many contact matrices, similar to the concept of comparing groups of multiple replicates in RNA-seq data.
+Finally, there is still room for expansion of time course boundary analysis.
+The continuous nature of boundary score allows for adopting time course analysis methods developed for gene expression studies [@Bar-Joseph:2012aa].
+More flexible classification of temporal trends may be considered, such as 24 temporal patterns proposed by Zhou et al.
+2019 [@Zhou:2019aa], or fuzzy clustering techniques that do not require a pattern to belong to a specific cluster [@Abu-Jamous:2018aa].
+In summary, our work enables further development of various aspects of 3D genome analysis.
 
-Our method relies on the ability to link the phenotype-associated genetic variants detected in genome-wide association studies to genes thought to be affected by them.
-Typically, genes are linked to genetic variants by proximity (nearest gene), a strategy adopted by the NHGRI-EBI GWAS catalog [@MacArthur:2017aa].
-<!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, the majority of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Khurana:2013aa].
-<!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].-->The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy.
-The development of chromatin conformation capture technologies now allows for understanding interactions among genomic regions on a genome-wide scale [@Lieberman-Aiden:2009aa; @Zhang:2013aa], helping to prioritize disease-associated gene-variant associations [@Pan:2018aa].
-Consequently, tools are being developed to consider long-distance interactions when mapping SNPs to genes (3DSNP [@Lu:2017aa], FUMA [@Watanabe:2017aa], PINES [@Bodea2016], HUGIn [@Martin:2017aa]), and studies redefining disease-associated genes using long-distance interactions started to emerge [@Fang:2019aa].
-Our future goal includes considering long-distance variant-gene interactions in defining phenotype-associated gene sets and comparing their network properties.
+One difficulty in our work is how to accurately quantify the biological relevance of TAD boundaries (differential, time-varying, and consensus) that we detect.
+There is no natural gold standard for TAD boundaries, but there are known genomic features that form the building blocks of TADs (CTCF, RAD21).
+In practice, we can use enrichment near boundaries of these as a proxy for "true boundaries".
+To test whether enrichment is different than random (non-boundaries), we use a permutation test and present these p-values.
+However, we can not compare these values between groups because the variance of random samples is affected by sample sizes.
+As a result, all p-values presented in the manuscript must be considered only within the context of the boundary type itself and not compared between boundary types or resolutions.
 
-The definition of coherence as the ratio between internal vs.
-external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa].
-Our limited knowledge of protein-protein interactions limits the coherence estimation to phenotypes that have a sufficient number of genes with PPI annotations.
-Furthermore, networks formed by phenotype-associated genes should be sufficiently diverse to avoid cases in which all genes in a network have the same degree.
-As the different PPI databases vary significantly in the number of PPI annotations, some phenotype-associated gene sets have sufficient PPI annotations when using one but not the other database.
-This problem is best illustrated by the BioGrid PPI database that has considerably fewer PPIs; hence, fewer phenotypes could be analyzed.
-Furthermore, using smaller PPI databases like BioGrid inevitably diminishes the size of phenotype-associated networks, limiting the internal degree distributions that can be derived from them.
-The differences in database-specific networks are alleviated by testing the significance of coherence using the permutation test; however, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].
+Our results in this manuscript demonstrate the ability of `TADCompare` to provide accurate, biologically relevant results.
+The methods implemented span differential, time-course, and consensus analysis.
+To date, `TADCompare` is the only actively maintained and publicly available tool to provide any of this functionality.
+We intend for `TADCompare` to be a one-stop tool for comparison of HiC datasets, providing simple, easy-to-interpret, results promptly.
+As a one-of-a-kind tool, `TADCompare` will increase the ability of researchers to extract important biological insights from the structure of TAD boundaries.
 
-<!--The definition of high and random coherence also depends on the choice of PPI data.
-Therefore, we derived degree distributions for KEGG and random networks separately for each PPI database and implemented the max-min normalization of the coherence estimates.
-Intuitively, the max-min normalization converts the absolute measures of coherence (slopes of regression lines fit through the internal vs.
-external degree distribution plots) that are dependent on the choice of PPI databases to the percentages of the full range of coherence (from high coherence defined by KEGG networks to low coherence defined by random networks) defined using a given PPI database.
-The phenotype-specific coherences expressed as a percentage of the full range of coherence can be directly compared across phenotypes and databases.
+# References
 
-Although the normalized coherence derived using different PPI databases can be compared among phenotypes, it tells nothing whether a given coherence estimate is significantly different from random.
-We developed a permutation test aimed to provide a measure of the stability of the coherence.
-Intuitively, the permutation test defines how different the coherence of a given phenotype from that of random networks of the same size.
-It provides additional protection against cases where coherence may appear high, yet, it may be a consequence of small network size providing unstable degree distributions.
-Testing the coherence estimates for significance helps against conclusions of the overall low coherence (e.g., anthropometric traits, Figure 3), when the coherence of individual traits is high (e.g., individual anthropometric traits, Figure 5B).
-Coupled with the permutation test, our coherence estimation provides a robust framework for quantifying the coherence estimates of phenotype-associated genes, advancing our understanding of the corresponding molecular mechanisms.-->
+<div id="refs"></div>
 
-In addition to internal vs.
-external degree distributions, other network properties can be used to estimate coherence.
-For example, the centrality measure has been used to demonstrate that highly central hub genes are functionally essential and highly conserved [@Khurana:2013aa].
-A large study of PPIs (BioPlex) compared centrality distributions among several disease categories.
-Cancers and immunological disease networks tended to have high centrality, while nervous system, congenital, neonatal and hereditary disease networks had low centrality [@Huttlin:2017aa].
-Metabolic pathways were shown to have more duplicated copies of highly central genes making the networks tolerant to loss-of-function mutations [@Khurana:2013aa].
-In summary, centrality measure parallels our observations of network coherence.
-Our future work will incorporate centrality and other network metrics into our framework to further refine our ability to compare coherence of phenotype-associated molecular mechanisms.
+\newpage
 
-[??? Potentially. discuss: We do not account for configuration]
+# Supplementary Legends
 
-## Conclusion
+**Supplementary Figure 1.
+Log-normal distribution of eigenvector gaps converted to boundary Z-scores.** Eigenvector gaps were calculated for contact matrices across three resolutions (10kb, 25kb, and 50kb, Hi-C data from [@Rao:2014aa], GM12878 cell line, chr 1-22).
+Density plots are shown for the (A) Natural log of the eigenvector gaps and (B) Boundary scores derived from the same data, separated by resolution.
+Regions of non-TADs are highlighted by a yellow bar, moderate strength TADs (2 < boundary score cutoff < 3) are highlighted by a red bar, and strong TADs (cutoff > 3) are shown using a green bar.
+We see a slightly right-skewed distribution due to the filtering of gaps for plotting purposes.
 
-Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
-We developed a network coherence measure that estimates the degree of internal vs.
-external connectivity that is robust to the difference in network sizes.
-Using well-curated lists of genomic variants and the associated genes, we built disease-specific networks using protein-protein interaction information from the STRING and BioGrid databases.
-We found a range of coherence estimates in each phenotype category, with metabolic diseases and traits being the most coherent.
-Psychiatric disorders were, on the other hand, the least coherent, with schizophrenia and major depressive disorder being the least coherent among all phenotypes.
-In summary, we provide a general-purpose method for quantifying and comparing coherence of molecular networks.
+**Supplementary Figure 2.
+Window size of 15 units of Hi-C data resolution and TAD boundary score cutoff of 2 yields consistent TAD boundary detection.** Differential TAD boundaries were compared between two simulated data sets with window size sizes ranging from 10 to 25 and boundary score cutoff ranging from 1.5 to 4.
+Youden index (balanced sensitivity and specificity metric) was calculated for each combination and plotted to show agreement with ground-truth annotations.
+Results are shown for noise-injected matrices (A) and sparsity-injected matrices (B).
 
-# Methods
+**Supplementary Figure 3.
+Visualization of different types of boundary score patterns.** Patterns of raw boundary scores are shown for 5 different types of differential boundaries (Merge, split, complex, shifted, and strength change).
+The red horizontal line corresponds to the minimum cutoff for a TAD boundary (40kb resolution, human neural progenitor cell line, data from [@Schmitt:2016aa]).
+Data from chromosome 22 with the most representative examples chosen.
 
-## Data sources
+**Supplementary Figure 4.
+Heatmap of gene ontology enrichment at the first and last time point in auxin-treated data.** Differential boundary identification was performed on auxin-treated data at the time of application (first time point) and complete withdrawal (last time point) (HCT-116 cell line, chr1-22, 40kb resolution, [@Schmitt:2016aa]).
+A barplot of the proportion of each boundary type (A) and FDR-adjusted hypergeometric p-values (B) obtained from gene ontology enrichment analysis using `rGREAT` (See methods) are shown.
+The top 30 pathways, in terms of average enrichment, are shown and clustered using Ward clustering.
 
-Phenotype-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa].
-"MAPPED_GENE" column was used to extract phenotype-associated genes.
-To organize phenotypes into categories, Experimental Factor Ontology (EFO) IDs were extracted from the "MAPPED_TRAIT_URI" column.
-Categories were mapped to EFO IDs using the `ontoCAT` v.1.12.0 R package and the EFO database (accessed 12/04/2018).
-In the case of a phenotype being mapped into two categories, the most representative category was manually assigned.
-Non-canonical gene names were converted to common gene symbols using `alias2Symbol` function from the `limma` R package.
-Gene names which could not be mapped to gene symbols were excluded from the analysis.
+**Supplementary Figure 5.
+Venn diagram of union and consensus TAD counts.** Consensus and union TADs were called across four different cell lines (hesc, mesynchymal, npc, trophectoderm) and the number of union and consensus TADs were recorded.
+The Venn diagram shows the complete overlap of consensus TADs within union TADs.
+(40kb resolution, data from [@Schmitt:2016aa])
 
-Human protein-protein interactions (PPIs) were obtained from STRING database v.11.0 [@Szklarczyk:2015aa] and BioGrid database v.3.5.174 [@Chatr-Aryamontri:2017aa] (accessed 06/25/2019).
-The STRING data for the tax ID "9606" (human) was used, and Ensembl protein IDs were converted to gene names using BioMart.
-Either the full dataset (11,761,040  entries) or data filtered by "combined_score" > 500 (range 0-1000) (1,373,946 entries, the main PPI dataset) was used.
-The BioGrid data was filtered to include "TaxID interactor" type equal to "9606", all "Interaction Types" were included, totaling 465,660 interactions.
-These three PPI databases were used in parallel to evaluate disease coherence.
+**Supplementary Figure 6.
+Runtime of TADCompare.** Plot containing the runtime of two-way comparison (A) and consensus TADs called on 4, 8, 12, and 16 replicates (B).
+Each point represents the runtime for a specific chromosome.
+X-axis - Chromosome, Y-axis - Runtime in seconds.
+Hi-C data from [@Rao:2014aa], chr 1-22, 10kb, 25kb, 50kb, and 100kb resolution.
 
-Phenotype-associated gene networks were formed by mapping phenotype-associated genes to PPIs.
-Genes without PPI annotations (e.g., C5orf4, LOC100507462), antisense transcripts (e.g., CDKN2B-AS), readthrough (e.g., NPHP3-ACAD11), non-protein-coding transcripts (e.g., LINC00478) were excluded.
-Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
-Phenotypes with less than five genes were omitted.
-Additionally, phenotypes with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
-external degree distributions (see below).
-These precautions were set to avoid situations where a small number of genes in a network can cause instability in the results.
-Importantly, the number of phenotype-associated genes having PPI annotations depends on the content of a PPI database; hence, some phenotypes may be filtered when using one but not the other PPI database to build phenotype-associated networks.
+**Supplementary Table 1.
+Contact matrix data sources.** The source of all contact matrices, experimental and simulated, used in this paper are provided.
+Experimental data are separated based on the study and cell line.
 
-## Comparing internal vs. external degree (aka connectivity) distributions as a measure of coherence
+**Supplementary Table 2. Genomic annotation data sources.** The sources, with download links, for all genomic annotation used in this paper are included.
 
-The **internal degree distribution** was defined as the number of interactions between genes of interest, or, more precisely, their interacting protein products.
-The **external degree distribution** was defined as the number of interactions between genes of interest and all other genes in the PPI network.
+**Supplementary Table 3.
+Summary of differential boundary types across tissues and cell-lines.** The percentage of each type of differential boundary for all tissue-tissue and cell line-cell line comparisons is reported.
+Results are aggregated over all chromosomes.
+Hi-C data from Schmitt et al.
+[@Schmitt:2016aa], 40kb resolution, chr 1-22.
 
-To compare phenotype-specific degree distributions, the internal vs.
-external degree distributions of phenotype-specific genes were plotted on a scatterplot.
-The square root of the degree was taken for more informative visualization.
-For each phenotype, the internal vs.
-external degree distribution plot was fit with a regression line through the origin to capture the slopes.
-Slopes were used as a measure of coherence with lower slopes being associated with higher coherence.
-Degree distributions of the network nodes were obtained using the `igraph` R package v.
-1.0.1.
+**Supplementary Table 4.
+Gene ontology enrichment for differential boundary types.** Differential boundaries were identified between the neural progenitor cells (NPC) and mesenchymal stem cells (MSC) [@Schmitt:2016aa].
+Pathway analysis was performed using `rGREAT` (Methods), and results are separated by ontology.
+Boundaries with an FDR adjusted p-value of <0.3 are shown.
+40kb resolution, chr1-22.
 
-## Max-min normalization of coherence
+**Supplementary Table 5.
+Gene ontology enrichment between the first and last time point in auxin-treated data .** Differential boundaries were identified between the first and last time point of auxin-treated data [@Rao:2017aa].
+Pathway analysis was performed using `rGREAT` (Methods), and results are separated by ontology.
+Boundaries with an FDR adjusted p-value of <0.3 are shown.
+50kb resolution, chr1-22.
 
-The slopes of the internal vs.
-external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database and the network size.
-To compare coherences using a given PPI database, we normalized the slopes for the phenotype-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
-For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
-Higher values indicate a higher level of coherence similar to that of KEGG pathways.
-These values represent normalized coherence and can be compared across phenotypes and PPI databases.
+**Supplementary Table 6.
+Enrichment across different temporal boundary types.** Temporal boundary types were identified across four time points in auxin-treated data [@Rao:2017aa].
+Results are shown for four types of temporal TAD (Early Appearing, Late Appearing, Highly Common, Dynamic).
+Permutation p-values, along with enrichment or depletion designations, are reported.
+HCT-116 cell line, 40kb resolution, chr 1-22.
 
-## Permutation test of slopes
+**Supplementary Table 7.
+Gene ontology enrichment for different temporal boundary types.** Temporal boundary types were identified across four time points in auxin-treated data [@Rao:2017aa].
+For each temporal boundary type, pathway analysis was performed using `rGREAT` (Methods), and results are separated by ontology.
+Boundaries with an FDR adjusted p-value of <0.3 are shown.
+HCT-116 cell line, 50kb resolution, chr1-22.
 
-To test whether phenotype-specific coherence is significantly different than random, we used a permutation test.
-For phenotype $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
-We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
-For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.
-external degree distribution regression line of random coherence.
-We calculate the permutation p-value as $p_i = \frac{(\sum_{j=1}^nI({\beta_{rj} > \beta_i}) + 1}{n + 1}$ [@Phipson:2010aa].
-Intuitively, this approach estimates whether the slope of the phenotype-specific network $\beta_i$ is consistently larger than those of random networks.
-All analyses and visualizations were performed in the R/Bioconductor computing environment v.3.4.0 [@gentleman2006bioinformatics].
-
-<!--For each disease, we find the absolute value of the difference between its slope and the random coherence slope.
-Next, we combine the genes for each disease with those of the random diseases and take two random samples of equal size to the disease gene set.
-We then calculate the absolute difference in slopes between the randomly sampled gene sets.
-This process is repeated 1000 times.
-Finally, we obtain a permutation p-value by taking the number of times the difference in slopes between random samples were less than the overall difference in slopes and dividing by 1000.
--->
-
-# Abbreviations
-
-- GWAS - Genome-Wide Association Study
-- KEGG - Kyoto Encyclopedia of Genes and Genomes
-- PPI - Protein-Protein Interaction
-- SNP - Single Nucleotide Polymorphism
+**Supplementary Table 8.
+Enrichment across different consensus scores.** Consensus scores were called across 17 contact matrices representing 7 different cell lines.
+Results were dichotomized into three groups (<2, 2-4, >4) based on consensus boundary scores.
+Permutation p-values, along with enrichment or depletion designations, are reported.
+Hi-C data from Schmitt et al.
+[@Schmitt:2016aa], 40kb resolution, chr 1-22.
 
@@
 
-This work was supported by a Genetics and Human Agency Award from the Templeton Foundation to KSK, MR, CC, SB.
-The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation.
+This work is supported in part by the PhRMA Foundation Research Informatics Award. The authors would like to acknowledge Dr. Katarzyna Tyc for helpful feedback.
 
-## Competing interests
+_Conflict of Interest._ None.
 
-The authors declare that they have no competing interests.
+# Author Contributions Statement
 
-# Supplementary material
-
-**Supplementary Table S1. Summary statistics of the analyzed phenotypes.** Each worksheet corresponds to the PPI database used.
-
-**Supplementary Table S2.
-Summary of networks from MSigDB categories.** Minimum, maximum, mean and median of the slopes and number of genes are shown for each collection, along with the total number of networks and the correlation between network size and slope.
-
-**Supplementary Table S3.
-Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) and the reference of high coherence (KEGG, GOCC, Reactome) are specified in the corresponding columns for coherence and permutation p-value estimations.
-
-**Supplementary Figure S1. Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.
-
-**Supplementary Figure S2.
-Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.
-
-**Supplementary Figure S3.
-Coherence estimates of phenotypes.** Size of dots represents the level of normalized coherence (X-axis) for individual phenotypes (Y-axis).
-Missing entries indicate that for a given PPI database a phenotype-associated network could not be built.
+MGD and KGC conceived the project, KGC implemented `TADcompare` and wrote the analysis scripts. MGD and KGC wrote the manuscript.
 
 # References
-
-<!--
-## Single networks, or networks of modules?
-
-Our study considers a network of disease-associated genes in its entirety.
-However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa].
-Connectivity of such modules would be different from the overall network connectivity [@Ghiassian:2015aa].
-Although numerous methods exist for module detection in a network [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section Intuitively, large networks will have a higher chance to have subnetworks (modules), and vice versa.
-It remains to be determined whether disease-associated genes form modular networks and whether the modular structure differs between (categories of) diseases.
-
-## Tissue specificity
-
-The effect of disease-associated SNPs should be considered in cell-type-specific context [@Trynka:2013aa; @Dozmorov:2016aa; @Sanyal:2012aa].
-The difference between neurological disease- and schizophrenia-associated SNPs was determined on the cell-type-specific level, with schizophrenia- but not neurological disease-associated SNPs being enriched in genes expressed in medium spiny neurons, pyramidal CA1, pyramidal SS and cortical interneurons [@Skene:2018aa].
-While cell-type-specific gene expression profiles have been generated, protein-protein interactions remain generic and highly incomplete [@Menche:2015aa].
-Even the latest efforts to compile PPIs from multiple data sources (InWeb InBioMap [@Li:2017aa]) or single-technology large-scale PPI profiling (BioPlex 2.0 [@Huttlin:2017aa]) generate collections of  non-cell-type-specific interactions.
-Future efforts in cell-type-specific PPI profiling are expected to boost research on the comparison of disease networks built from PPIs.
-
-## What constitutes random networks?
-
-The degree was found to follow a power-law [@Yook:2005aa] or a truncated power-law distribution [@Huang:2007aa].
-
-## GWAS p-value, cutoff, sample size, harmonization issues
-
-In GWAS, it is often assumed that local correlations between adjacent hypotheses along the genome are so strong that testing the whole genome is equivalent to performing only $10^6$ tests, regardless of the number of hypotheses tested.
-Under this assumption, FWER can be controlled using a fixed genome-wide significance level of $\alpha = 5*10^{-8}$, corresponding to a Bonferroni adjustment for $10^6$ tests.
--->

### 2019-09-08 manuscript/manuscript.Rmd
Replace manuscript, adjust for latest paths

--- 2019-08-18 manuscript.Rmd
@@
 ---
-bibliography: [refs/networks.bib,refs/brain.bib,refs/enrichment.bib,refs/3D_refs.bib]
-csl: styles.ref/genomebiology.csl
+bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib]
+csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
 output:
+  word_document:
+    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/styles.doc/Arial_11_single_space_normal_margins.docx
   pdf_document: default
-  word_document:
-    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
   html_document: default
@@
 
+<!--Tables/table_1.csv-->
 **Table 1. Summary statistics of the analyzed categories.** Average coherence estimates are shown. "NA" values indicate that a phenotype was not considered due to the low number of genes with PPIs.
 
-| Category               | Total Phenotypes | Number of SNPs | Number of Genes | Biogrid Coherence | String Coherence | String Filtered Coherence |
-|------------------------|:----------------:|:--------------:|:---------------:|:-----------------:|:----------------:|:-------------------------:|
+| Category               | Total Phenotypes | Number of SNPs | Number of Genes | Biogrid | String | String Filtered |
+|------------------------|:----------------:|:--------------:|:---------------:|:-------:|:------:|:---------------:|
 | Disease                |                  |                |                 |         |        |                 |
-| Metabolic disease      |         5        |       345      |       280       |        0.86       |       0.79       |           0.933           |
-| Neurologic             |         8        |       382      |       343       |        0.89       |       0.511      |           0.861           |
-| Eye disease            |         8        |       270      |       242       |        0.75       |       0.536      |           0.856           |
-| Autoimmune             |        20        |      1415      |       1212      |       0.807       |       0.801      |           0.852           |
-| Cancer                 |        18        |       713      |       610       |       0.863       |       0.643      |           0.838           |
-| Cardiovascular disease |         9        |       336      |       289       |       0.705       |       0.584      |           0.832           |
-| Psychiatric            |        11        |       678      |       611       |        0.83       |       0.537      |           0.817           |
+| Metabolic disease      |         5        |       345      |       280       |  0.515  |  0.776 |      0.845      |
+| Autoimmune             |        20        |      1415      |       1212      |  0.470  |  0.801 |      0.702      |
+| Eye disease            |         8        |       270      |       242       |  0.300  |  0.509 |      0.654      |
+| Neurologic             |         8        |       382      |       343       |  0.710  |  0.498 |      0.637      |
+| Cancer                 |        18        |       713      |       610       |  0.548  |  0.658 |      0.604      |
+| Cardiovascular disease |         9        |       336      |       289       |  0.130  |  0.577 |      0.602      |
+| Psychiatric            |        11        |       678      |       611       |  0.560  |  0.528 |      0.583      |
 | Trait                  |                  |                |                 |         |        |                 |
-| Metabolic trait        |        10        |       790      |       646       |       0.906       |       0.776      |            0.93           |
-| Eye trait              |         5        |       158      |       109       |         NA        |       0.674      |           0.903           |
-| Anthropometric trait   |        19        |      2143      |       1822      |       0.793       |       0.601      |           0.841           |
-| Cardiovascular trait   |        15        |       728      |       604       |       0.833       |       0.647      |           0.813           |
-| Intelligence           |         5        |       236      |       198       |        0.78       |       0.312      |           0.757           |
+| Metabolic trait        |        10        |       790      |       646       |  0.766  |  0.779 |      0.842      |
+| Eye trait              |         5        |       158      |       109       |    NA   |  0.678 |      0.698      |
+| Anthropometric trait   |        19        |      2143      |       1822      |  0.600  |  0.624 |      0.678      |
+| Cardiovascular trait   |        15        |       728      |       604       |  0.463  |  0.655 |      0.537      |
+| Intelligence           |         5        |       236      |       198       |  0.310  |  0.334 |      0.373      |
 
@@
 +-------+
-| ![](Figures/Size_Slope_Plot.png)      |
+| ![](Figures/Fig2_Size_Slope_Plot.png)      |
 | **Figure 2.

### 2019-09-10 Analysis/Msigdf_Biogrid.R
combine and then slope

@@ -30,22 +30,37 @@ links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein
 
 #Getting msigs 
 
-# msig = msigdbr()
-# 
-# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-# 
-# mod_over = bind_rows()
-# for (j in unique(msig$gs_name)) {
-#   curr_msig = msig %>% filter(gs_name == j)
-#   write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-#   diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
-#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-#   mod_slope = coef(mod)[[1]]
-#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-#                        Category = curr_msig$gs_subcat[1])
-#   mod_over = bind_rows(mod_over, mod_sum)
-#   saveRDS(mod_over,"KEGG_Biogrid.rds")
-# }
+msig = msigdbr()
+
+msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+
+mod_over = bind_rows()
+for (j in unique(msig$gs_name)) {
+  print(j)
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
+  diseases_curr = diseases_curr %>% mutate(Disease = j)
+  mod_over = bind_rows(mod_over,diseases_curr)
+}
+
+#Applying categories
+
+msig_cats  = msig %>% 
+  dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
+  distinct() #Get categories in order
+
+#Joining categories
+
+mod_over = left_join(mod_over, msig_cats)
+
+#Getting gene counts
+
+mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
+
+#Saving
+
+saveRDS(mod_over, "KEGG_Biogrid.rds")
 
 disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")
 
[…]

### 2019-09-10 Analysis/Msigdf_String.R

@@ -29,29 +29,21 @@ string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
 
 string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
 
-#Getting msigs 
-
-# msig = msigdbr()
-# 
-# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-# 
-# mod_over = bind_rows()
-# for (j in unique(msig$gs_name)) {
-#   curr_msig = msig %>% filter(gs_name == j)
-#   write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-#   diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
-#   if (nrow(diseases_curr) == 0) {
-#     next
-#   }
-#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-#   mod_slope = coef(mod)[[1]]
-#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-#                        Category = curr_msig$gs_subcat[1])
-#   mod_over = bind_rows(mod_over, mod_sum)
-#   saveRDS(mod_over,"KEGG_String.rds")
-# }
-
 #Redoing with every category at once
+
+#Calculating normalized coherence
+
+disease_overall_string  <- readRDS("string_edges_new.rds")
+
+#Reading in slopes for diseases
+
+#Calculate slopes for all diseases under each category
+dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+  do({
+    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+    data.frame(Slope= coef(mod)[1])
+  }) %>% arrange(Slope) %>% filter(Count>4)
 
 msig = msigdbr()
 
@@ -72,126 +64,6 @@ for (j in unique(msig$gs_name)) {
   saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
 }
 
-#Calculating normalized coherence
-
-disease_overall_string  <- readRDS("string_edges_new.rds")
-
-#Reading in slopes for diseases
-
[…]

### 2019-09-10 Analysis/Msigdf_String_Filt.R

@@ -43,41 +43,65 @@ dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal !
 
 #Getting msigs 
 
-# msig = msigdbr()
-# 
-# msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-# 
-# mod_over = bind_rows()
-# for (j in rev(unique(msig$gs_name))) {
-#   curr_msig = msig %>% filter(gs_name == j)
-#   write.table(curr_msig$gene_symbol, file = "temp_rea.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
-#   diseases_curr = diseases2frame("temp_rea.txt", links = string_filt)
-#   if (nrow(diseases_curr) == 0) {
-#     next
-#   }
-#   mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
-#   mod_slope = coef(mod)[[1]]
-#   mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
-#                        Category = curr_msig$gs_subcat[1])
-#   print(mod_sum)
-#   mod_over = bind_rows(mod_over, mod_sum)
-#   saveRDS(mod_over,"KEGG_String_Filt.rds")
-#   }
+msig = msigdbr()
 
-#Normalized coherence
-#KEGG
+msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
+
+mod_over = list()
+index = 1
+for (j in unique(msig$gs_name)) {
+  print(j)
+  curr_msig = msig %>% filter(gs_name == j)
+  write.table(curr_msig$gene_symbol, file = "temp_rea_string_filt.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+  diseases_curr = diseases2frame("temp_rea_string_filt.txt", links = string_filt)
+  diseases_curr = diseases_curr %>% mutate(Disease = j)
+  mod_over[[index]] = diseases_curr
+  index = index+1
+}
+
+mod_over = bind_rows(mod_over)
+
+#Applying categories
+
+msig_cats  = msig %>% 
+  dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
+  distinct() #Get categories in order
+
+#Joining categories
+
+mod_over = left_join(mod_over, msig_cats)
+
+#Getting gene counts
+
[…]

### 2019-09-10 Analysis/Msigdf_String.R
fix saved file path

@@ -177,5 +177,5 @@ for (i in unique(dis_slopes_string_count$Disease)) {
                          Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
   coherence_tab = bind_rows(coherence, coherence_tab)
   print(coherence_tab)
-  saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
+  saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string.rds")
 }

### 2019-09-10 Analysis/Msigdf_Biogrid.R
make quicker

@@ -34,15 +34,18 @@ msig = msigdbr()
 
 msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
 
-mod_over = bind_rows()
+mod_over = list()
+index = index+1
 for (j in unique(msig$gs_name)) {
   print(j)
   curr_msig = msig %>% filter(gs_name == j)
   write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
   diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
   diseases_curr = diseases_curr %>% mutate(Disease = j)
-  mod_over = bind_rows(mod_over,diseases_curr)
+  mod_over[[index]] = diseases_curr
 }
+
+mod_over = bind_rows(mod_over)
 
 #Applying categories
 

### 2019-09-11 Analysis/Msigdf_Biogrid.R
update to new approach with variables

@@ -6,6 +6,14 @@ library(dplyr) #install.packages("dplyr")
 library(biomaRt) #BiocManager::install("biomaRt")
 library(msigdbr) #BiocManager::install("msigdbr")
 library(igraph) #install.packages("igraph")
+library(tidyr) #install.packages("tidyr")
+library(broom) #install.packages("broom")
+
+#Option for number of msig pathways to take from each category
+num_msig = 10
+#Option for number of random networks to generate (Will generate num_rand-1 networks)
+num_rand = 11
+
 
 #Reading in biogrid data
 
@@ -35,7 +43,7 @@ msig = msigdbr()
 msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
 
 mod_over = list()
-index = index+1
+index = 1
 for (j in unique(msig$gs_name)) {
   print(j)
   curr_msig = msig %>% filter(gs_name == j)
@@ -43,6 +51,7 @@ for (j in unique(msig$gs_name)) {
   diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
   diseases_curr = diseases_curr %>% mutate(Disease = j)
   mod_over[[index]] = diseases_curr
+  index =  index+1
 }
 
 mod_over = bind_rows(mod_over)
@@ -82,7 +91,9 @@ dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>%
 #Get msig pathway genes for KEGG, Reactome and GOCC
 msig = readRDS("KEGG_Biogrid.rds")
 #Get pathway specific counts
-Count_Sum = msig %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
+Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
+  group_by(Disease) %>%
+  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
 #Splitting into list by categories
 Count_Sum = split(Count_Sum, Count_Sum$Category)
 coherence_tab = bind_rows()
@@ -91,27 +102,39 @@ for (i in unique(dis_slopes_biogrid_count$Disease)) {
   dis = dis_slopes_biogrid_count %>% filter(Disease == i)
   #Get 30 closest msigdf networks taking 10 from each database and bind
[…]

### 2019-09-11 Analysis/Msigdf_String.R
updated version of string

@@ -7,6 +7,11 @@ library(biomaRt) #BiocManager::install("biomaRt")
 library(msigdbr) #BiocManager::install("msigdbr")
 library(igraph) #install.packages("igraph")
 
+#Option for number of msig pathways to take from each category
+num_msig = 10
+#Option for number of random networks to generate (Will generate num_rand-1 networks)
+num_rand = 11
+
 string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
 
 string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
@@ -137,11 +142,15 @@ mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
 
 saveRDS(mod_over, "KEGG_String.rds")
 
+#Calculating normalized coherence
+
 # KEGG
 #Get msig pathway genes for KEGG, Reactome and GOCC
 msig = readRDS("KEGG_String.rds")
 #Get pathway specific counts
-Count_Sum = msig %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
+Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
+  group_by(Disease) %>%
+  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
 #Splitting into list by categories
 Count_Sum = split(Count_Sum, Count_Sum$Category)
 coherence_tab = bind_rows()
@@ -150,27 +159,39 @@ for (i in unique(dis_slopes_string_count$Disease)) {
   dis = dis_slopes_string_count %>% filter(Disease == i)
   #Get 30 closest msigdf networks taking 10 from each database and bind
   close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
-                                  arrange(Diff) %>% head(.,10)))
+                                  arrange(Diff) %>% head(.,num_msig)))
   #Subsetting msig to only include close pathways
   msig_sub = msig %>% filter(Disease %in% close_nets$Disease)
-  #Square rooting because its not already square rooted
-  mod_msig = lm(I(sqrt(External)) ~ Internal-1, data = msig_sub)
-  #Pull out coefficient
-  mod_slope = coef(mod_msig)[[1]]
-  diseases = bind_rows()
+  #Taking slopes for each category
+  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
[…]

### 2019-09-11 Analysis/Msigdf_String_Filt.R
new method for string filt

@@ -7,6 +7,12 @@ library(biomaRt) #BiocManager::install("biomaRt")
 library(msigdbr) #BiocManager::install("msigdbr")
 library(igraph) #install.packages("igraph")
 
+#Option for number of msig pathways to take from each category
+num_msig = 10
+#Option for number of random networks to generate (Will generate num_rand-1 networks)
+num_rand = 11
+
+
 string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
 
 string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
@@ -79,11 +85,15 @@ mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
 
 saveRDS(mod_over, "KEGG_String_Filt.rds")
 
+#Calculating normalized coherence
+
 # KEGG
 #Get msig pathway genes for KEGG, Reactome and GOCC
 msig = readRDS("KEGG_String_Filt.rds")
 #Get pathway specific counts
-Count_Sum = msig %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
+Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
+  group_by(Disease) %>%
+  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
 #Splitting into list by categories
 Count_Sum = split(Count_Sum, Count_Sum$Category)
 coherence_tab = bind_rows()
@@ -92,27 +102,39 @@ for (i in unique(dis_slopes_string_filt_count$Disease)) {
   dis = dis_slopes_string_filt_count %>% filter(Disease == i)
   #Get 30 closest msigdf networks taking 10 from each database and bind
   close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
-                                  arrange(Diff) %>% head(.,10)))
+                                  arrange(Diff) %>% head(.,num_msig)))
   #Subsetting msig to only include close pathways
   msig_sub = msig %>% filter(Disease %in% close_nets$Disease)
-  #Square rooting because its not already square rooted
-  mod_msig = lm(I(sqrt(External)) ~ Internal-1, data = msig_sub)
-  #Pull out coefficient
-  mod_slope = coef(mod_msig)[[1]]
-  diseases = bind_rows()
+  #Taking slopes for each category
[…]

### 2019-09-11 Analysis/Msigdf_Biogrid.R
Save results to a file, use 1000 random networks

@@ -12,7 +12,10 @@ library(broom) #install.packages("broom")
 #Option for number of msig pathways to take from each category
 num_msig = 10
 #Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 11
+num_rand = 1001
+# Construct file name to save the results
+run <- 1 # Placeholder for the run ID
+fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
 
 
 #Reading in biogrid data
@@ -144,5 +147,5 @@ for (i in unique(dis_slopes_biogrid_count$Disease)) {
       saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
 }
 
-
+write_csv(coherence_tab, fileNameOut)
 

### 2019-09-13 Analysis/Msigdf_String.Rmd
convert to rmd and comment

---
  title: "Msigdf String"
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

```{r}
source("./functions/disease2frame.R")
source("./functions/GeneSample.R")

library(readr) #install.packages("readr")
library(dplyr) #install.packages("dplyr")
library(biomaRt) #BiocManager::install("biomaRt")
library(msigdbr) #BiocManager::install("msigdbr")
library(igraph) #install.packages("igraph")
```

```{r}
#Option for number of msig pathways to take from each category
num_msig = 10
#Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 11
```

```{r}
string = read_table2("./data/9606.protein.links.v11.0.txt.gz")

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                           protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)

```

# Preparing disease-specific data

```{r}

#Reading in disease-specific edges
disease_overall_string  <- readRDS("string_edges_new.rds")

#Get counts and slopes from edge data frame
dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)

```

#Getting msigdf specific coherence

```{r}
#This code is for the table of msigdf coherences

#Read in msig object
msig = msigdbr()

#Get all non-empty msigdf categories

msig = msig %>% filter(gs_subcat != "")
mod_over = bind_rows()
for (j in unique(msig$gs_name)) {
  #Subset to current pathway name
  curr_msig = msig %>% filter(gs_name == j) 
  #Write to data frame that is formatted for disease2frame
  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE) 
  #Run disease 2 frame and get internal and external edges
  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
  #Skip situations where there are no genes in ppi network
  if (nrow(diseases_curr) == 0) {
    next
  }
  #Calculate slope
  mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = diseases_curr)
  #Pull out slope
  mod_slope = coef(mod)[[1]]
  #Place slopes into data frames with counts
  mod_sum = data.frame(Disease = j, Count =nrow(diseases_curr), Slope = mod_slope,
                       Category = curr_msig$gs_subcat[1])
  #Bind together slopes and counts
  mod_over = bind_rows(mod_over, mod_sum)
  #Save
  saveRDS(mod_over,"./data/Coherence_Results/All_Msig_String.rds")
}

```

# Calculating random slopes

```{r}
# Random networks

#Read in KEGG pathways for reference
msig = readRDS("KEGG_String.rds")
#Pull outt KEGG pathways
msig = msig %>% filter(Category == "CP:KEGG")
string_random = bind_rows()
for (i in msig$Count) {
  #Start looping through KEGG network sizes
  #Initialize the number of non-empty networks
  j = 0
  diseases = bind_rows()
    while(j<11) {
    #Sample genes
    diseases_curr = GeneSample(i, string, directed = FALSE)
    #Put into frame with edges
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = sqrt(diseases_curr$internal_connectivity), External = sqrt(diseases_curr$external_connectivity))
    #Make sure at least  one gene has non-zero internal edges. If not retry
    if (any(diseases_curr$Internal !=0)) {
      j = j+1
      #Bind connnectivities together
      diseases = bind_rows(diseases, diseases_curr)

    }
    }

      #Get slope
      mod_rand = lm(External ~ Internal-1, data = diseases)
      #Place number of genes and slope into data frame
      diseases_curr = data.frame(Count = i, Slope = coef(mod_rand)[1])
      #Bind together
      string_random = bind_rows(string_random, diseases_curr)
      #Save
      saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
}

```

```{r}
#Getting msigs

msig = msigdbr()

#Reread msig and subset to include GOCC, KEGG and REACTOME
msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )

#Create empty list
mod_over = list()
#Set up indexing
index = 1
for (j in unique(msig$gs_name)) {
  #Loop through msig names
  curr_msig = msig %>% filter(gs_name == j)
  #Create table in form disease2frame takes
  write.table(curr_msig$gene_symbol, file = "temp_rea_string.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
  #Get internal and external edges
  diseases_curr = diseases2frame("temp_rea_string.txt", links = string)
  #Add results to data frame and assign disease name
  diseases_curr = diseases_curr %>% mutate(Disease = j)
  #Save results to list
  mod_over[[index]] = diseases_curr
  index = index+1
}

mod_over = bind_rows(mod_over)

#Applying categories

msig_cats  = msig %>% 
  dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
  distinct() #Get categories in order

#Joining categories

mod_over = left_join(mod_over, msig_cats)

#Getting gene counts

mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())

#Saving

saveRDS(mod_over, "KEGG_String.rds")

#Calculating normalized coherence

# KEGG
#Get msig pathway genes for KEGG, Reactome and GOCC
msig = readRDS("KEGG_String.rds")
#Get pathway specific counts
Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
#Splitting into list by categories
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_count$Disease)) {
  #Get disease specific genes
  dis = dis_slopes_string_count %>% filter(Disease == i)
  #Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
                                  arrange(Diff) %>% head(.,num_msig)))
  #Subsetting msig to only include close pathways
  msig_sub = msig %>% filter(Disease %in% close_nets$Disease)
  #Taking slopes for each of the 3 msig categories
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  #Take median
  mod_slope = median(mod_msig$x)
  #Creating 10 random networks
  j = 1
  slope_list = list()
  while(j<num_rand) {
    #Empty data frame for disease networks
    diseases = bind_rows()
    #Create random networks until there are the same non-zero edges as the disease
    while(nrow(diseases)<dis$Count) {
      #Generate random diseases network
      diseases_curr = GeneSample(dis$Count, string, directed = FALSE)
      #Save internal and external edges
      diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
      #Filter out zeros
      diseases_curr = diseases_curr %>% filter(Internal !=0)
      #Bind together non-zero internal/external edges
      diseases = bind_rows(diseases, diseases_curr)
    }
    #Save slope of current network
    slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
    #Iterate
    j = j+1

  }
  .
  #Get median of aggregated slopes
  rand_slope = median(unlist(slope_list))

  #Get normalized coherence
  coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
  #save into data frame
  coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                         Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
  coherence_tab = bind_rows(coherence, coherence_tab)
  saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string.rds")
}

### 2019-09-13 Analysis/Msigdf_String_Filt.Rmd
some clean up

---
title: "Msigdf String" Filt
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

# Reading in packages
```{r}
source("./functions/disease2frame.R")
source("./functions/GeneSample.R")

library(readr) #install.packages("readr")
library(dplyr) #install.packages("dplyr")
library(biomaRt) #BiocManager::install("biomaRt")
library(msigdbr) #BiocManager::install("msigdbr")
library(igraph) #install.packages("igraph")
```

#Setting options
```{r}
#Option for number of msig pathways to take from each category
num_msig = 10
#Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 11
```

# Reading in data
```{r}

string = read_table2("./data/9606.protein.links.v11.0.txt.gz")

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                           protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)

string_filt = string %>% filter(combined_score>=500)

disease_overall_string_filt <- readRDS("string_filt_edges_new.rds")

#STRING Filtered
dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)

```

#Getting msig pathway coherence

```{r}

#Read in msig pathways
msig = msigdbr()

#Subset to only include GOCC, KEGG and REACTOME

msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )

#Create empty list
mod_over = list()
index = 1
#Loop over each ontology
for (j in unique(msig$gs_name)) {
  #Subset to only inc\lude specific onotlogies
  curr_msig = msig %>% filter(gs_name == j)
  #Write to table in disease2frame friendly format
  write.table(curr_msig$gene_symbol, file = "temp_rea_string_filt.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
  #Get internal/external connectivity
  diseases_curr = diseases2frame("temp_rea_string_filt.txt", links = string_filt)
  #Add column with ontology name
  diseases_curr = diseases_curr %>% mutate(Disease = j)
  #Add to list
  mod_over[[index]] = diseases_curr
  index = index+1
}

#Collapse list
mod_over = bind_rows(mod_over)

#Applying categories

msig_cats  = msig %>% 
  dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
  distinct() #Get categories in order

#Joining categories

mod_over = left_join(mod_over, msig_cats)

#Getting gene counts

mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())

#Saving

saveRDS(mod_over, "KEGG_String_Filt.rds")

```

```{r}
#Calculating normalized coherence

# KEGG
#Read in KEGG pathway data
msig = readRDS("KEGG_String_Filt.rds")
#Get pathway specific counts but without internal edges
Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
#Splitting into list by categories
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
for (i in unique(dis_slopes_string_filt_count$Disease)) {
  #Get disease specific genes
  dis = dis_slopes_string_filt_count %>% filter(Disease == i)
  #Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
                                  arrange(Diff) %>% head(.,num_msig)))
  #Subsetting msig to only include closest pathways
  msig_sub = msig %>% filter(Disease %in% close_nets$Disease)
  #Taking slopes for each category
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  #Take median
  mod_slope = median(mod_msig$x)
  #Creating 10 random networks
  j = 1
  slope_list = list()
  while(j<num_rand) {
    #Empty data frame for disease networks
    diseases = bind_rows()
    #Generate random networks until the number of genes matches the diseases
    while(nrow(diseases)<dis$Count) {
      #Generate random diseases network
      diseases_curr = GeneSample(dis$Count, string_filt, directed = FALSE)
      #Save internal and external edges
      diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
      #Filter out zeros
      diseases_curr = diseases_curr %>% filter(Internal !=0)
      #Bind together non-zero internal/external edges
      diseases = bind_rows(diseases, diseases_curr)
    }
    #Save slope of current network
    slope_list[[j]] = coef(lm(I(sqrt(Exter.nal)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
    #Iterate
    j = j+1

  }

  #Unlist and get median of combined slopes
  rand_slope = median(unlist(slope_list))

  #Get normalized coherence
  coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
  #Place into data frame with disease name, coherence, msig normalization value and random normalizetion value
  coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                         Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
  coherence_tab = bind_rows(coherence, coherence_tab)
  print(coherence_tab)
  saveRDS(coherence_tab, "./data/Coherence_Results/coherence_string_filt.rds")
}

```

### 2019-09-13 Analysis/Msigdf_Biogrid.Rmd
make msig .rmd with comments

---
title: "Msigdf Biogrid"
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

# Read in functions
```{r}  
source("./functions/disease2frame.R")
source("./functions/GeneSample.R")

library(readr) #install.packages("readr")
library(dplyr) #install.packages("dplyr")
library(biomaRt) #BiocManager::install("biomaRt")
library(msigdbr) #BiocManager::install("msigdbr")
library(igraph) #install.packages("igraph")
library(tidyr) #install.packages("tidyr")
library(broom) #install.packages("broom")

```

# Setting options

```{r}
#Option for number of msig pathways to take from each category
num_msig = 10
#Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 1001
# Construct file name to save the results
run <- 1 # Placeholder for the run ID
fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
```

```{r}

#Reading in biogrid data

links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")

links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names

protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])

protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on

links = links %>% mutate(protein1 =protein1, 
                         protein2 = protein2) %>% dplyr::select(protein1, protein2)

links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

#Getting msigs

msig = msigdbr()

#Subset to include msig
msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )

#Create empty list
mod_over = list()
#Set index for list entries
index = 1
#Loop through each pathway
for (j in unique(msig$gs_name)) {
  #Subset to only include current pathway
  curr_msig = msig %>% filter(gs_name == j)
  #Write to table in format for disease2frame
  write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
  #Get internal/external connectivity
  diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
  #Get disease name
  diseases_curr = diseases_curr %>% mutate(Disease = j)
  mod_over[[index]] = diseases_curr
  index =  index+1
}

#Bind together list

mod_over = bind_rows(mod_over)

#Applying categories

msig_cats  = msig %>% 
  dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
  distinct() #Get categories in order

#Joining categories

mod_over = left_join(mod_over, msig_cats)

#Getting gene counts

mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())

#Saving

saveRDS(mod_over, "KEGG_Biogrid.rds")

#Get edges for diseases from Biogrid (Connectivity_Comparison.Rmd)
disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")

#Get Biogrid slopes and edge counts

dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)

#Calculating normalized coherence

# KEGG
#Get msig pathway genes for KEGG, Reactome and GOCC
msig = readRDS("KEGG_Biogrid.rds")
#Get pathway specific counts
Count_Sum = msig %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
#Splitting into list by categories
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
for (i in unique(dis_slopes_biogrid_count$Disease)) {
  #Get disease specific genes
  dis = dis_slopes_biogrid_count %>% filter(Disease == i)
  #Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
    arrange(Diff) %>% head(.,num_msig)))
  #Subsetting msig to only include close pathways
  msig_sub = msig %>% filter(Disease %in% close_nets$Disease)
  #Taking slopes for each category
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  #Take median
  mod_slope = median(mod_msig$x)
  #Creating 10 random networks
  j = 1
  slope_list = list()
    while(j<num_rand) {
    #Empty data frame for disease networks
      diseases = bind_rows()
      #Generate random networks until there as many internal edges as current diseases
      while(nrow(diseases)<dis$Count) {
    #Generate random diseases network
    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
    #Save internal and external edges
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
    #Filter out zeros
    diseases_curr = diseases_curr %>% filter(Internal !=0)
    #Bind together non-zero internal/external edges
    diseases = bind_rows(diseases, diseases_curr)
      }
      #Save slope of current network
      slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
      #Iterate
      j = j+1

    }

      #Get median of all slopes
      rand_slope = median(unlist(slope_list))

      #Calculate coherence
      coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
      #Place into data frame with disease, coherence, msig normalization value and random slope normalization value
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                             Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
      coherence_tab = bind_rows(coherence, coherence_tab)
      print(coherence_tab)
      saveRDS(coherence_tab, "./data/Coherence_Results/coherence_biogrid.rds")
}

write_csv(coherence_tab, fileNameOut)

### 2019-09-13 Analysis/Msigdf_Biogrid.R
remove old .R files

@@ -1,151 +1,0 @@
-source("./functions/disease2frame.R")
-source("./functions/GeneSample.R")
-
-library(readr) #install.packages("readr")
-library(dplyr) #install.packages("dplyr")
-library(biomaRt) #BiocManager::install("biomaRt")
-library(msigdbr) #BiocManager::install("msigdbr")
-library(igraph) #install.packages("igraph")
-library(tidyr) #install.packages("tidyr")
-library(broom) #install.packages("broom")
-
-#Option for number of msig pathways to take from each category
-num_msig = 10
-#Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 1001
-# Construct file name to save the results
-run <- 1 # Placeholder for the run ID
-fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
-
-
-#Reading in biogrid data
-
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
-
-links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
-
-#Split based on Biogrid ID system then pull out gene names
-
-protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-
-protein1 = unlist(protein1)
-protein2 = unlist(protein2)
-
-#Place into matrix in form that disease2frame works on
-
-links = links %>% mutate(protein1 =protein1, 
-                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
-
-links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
-
-#Getting msigs 
-
-msig = msigdbr()
-
-msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-
-mod_over = list()
-index = 1
-for (j in unique(msig$gs_name)) {
-  print(j)
-  curr_msig = msig %>% filter(gs_name == j)
[…]

### 2019-09-13 Analysis/Msigdf_String.R

@@ -1,205 +1,0 @@
-source("./functions/disease2frame.R")
-source("./functions/GeneSample.R")
-
-library(readr) #install.packages("readr")
-library(dplyr) #install.packages("dplyr")
-library(biomaRt) #BiocManager::install("biomaRt")
-library(msigdbr) #BiocManager::install("msigdbr")
-library(igraph) #install.packages("igraph")
-
-#Option for number of msig pathways to take from each category
-num_msig = 10
-#Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 11
-
-string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
-
-string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                           protein2 = gsub("9606.", "", protein2))
-
-#Read in ensemble protein ID
-
-ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
-genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
-
-
-#Join gene names with string
-
-string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
-string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-
-string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
-
-string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
-
-#Redoing with every category at once
-
-#Calculating normalized coherence
-
-disease_overall_string  <- readRDS("string_edges_new.rds")
-
-#Reading in slopes for diseases
-
-#Calculate slopes for all diseases under each category
-dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
-  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-  do({
-    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-    data.frame(Slope= coef(mod)[1])
-  }) %>% arrange(Slope) %>% filter(Count>4)
-
-msig = msigdbr()
-
-msig = msig %>% filter(gs_subcat != "")
[…]

### 2019-09-13 Analysis/Msigdf_String_Filt.R

@@ -1,146 +1,0 @@
-source("./functions/disease2frame.R")
-source("./functions/GeneSample.R")
-
-library(readr) #install.packages("readr")
-library(dplyr) #install.packages("dplyr")
-library(biomaRt) #BiocManager::install("biomaRt")
-library(msigdbr) #BiocManager::install("msigdbr")
-library(igraph) #install.packages("igraph")
-
-#Option for number of msig pathways to take from each category
-num_msig = 10
-#Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 11
-
-
-string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
-
-string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                           protein2 = gsub("9606.", "", protein2))
-
-#Read in ensemble protein ID
-
-ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
-genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
-
-
-#Join gene names with string
-
-string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
-string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-
-string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
-
-string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
-
-string_filt = string %>% filter(combined_score>=500)
-
-disease_overall_string_filt <- readRDS("string_filt_edges_new.rds")
-
-#STRING Filtered
-dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
-  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
-  do({
-    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
-    data.frame(Slope= coef(mod)[1])
-  }) %>% arrange(Slope) %>% filter(Count>4)
-
-#Getting msigs 
-
-msig = msigdbr()
-
-msig = msig %>% filter(gs_subcat %in% c("CC", "CP:KEGG", "CP:REACTOME") )
-
[…]

### 2019-09-13 Analysis/Msigdf_all.R

@@ -1,3 +1,0 @@
-source("Msigdf_Biogrid.R")
-source("Msigdf_String.R")
-source("Msigdf_String_Filt.R")

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
put everything in functions

---
title: "Msigdf Biogrid"
author: "Kellen Cresswell"
date: "July 19, 2019"
output: pdf_document
---

# Read in functions

```{r}  
source("./functions/disease2frame.R")
source("./functions/GeneSample.R")
source("./functions/get_msig_slopes.R")

library(readr) #install.packages("readr")
library(dplyr) #install.packages("dplyr")
library(biomaRt) #BiocManager::install("biomaRt")
library(msigdbr) #BiocManager::install("msigdbr")
library(igraph) #install.packages("igraph")
library(tidyr) #install.packages("tidyr")
library(broom) #install.packages("broom")

```

# Setting options

```{r}
#Option for number of msig pathways to take from each category
num_msig = 10
#Option for number of random networks to generate (Will generate num_rand-1 networks)
num_rand = 1001
# Construct file name to save the results
run <- 1 # Placeholder for the run ID
fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_", num_msig, "_rand_", num_rand, "_", run, ".csv"))
```

# Read in biogrid data

```{r}
#Reading in biogrid data

links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")

links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")

#Split based on Biogrid ID system then pull out gene names

protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])

protein1 = unlist(protein1)
protein2 = unlist(protein2)

#Place into matrix in form that disease2frame works on

links = links %>% mutate(protein1 =protein1, 
                         protein2 = protein2) %>% dplyr::select(protein1, protein2)

links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)

```

# Read in string data and create the string filtered frame

```{r}
string = read_table2("./data/9606.protein.links.v11.0.txt.gz")

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                           protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)

string_filt = string %>% filter(combined_score>=500)
```

```{r}

get_msig_slopes = function(sub_cats, database) {
  require(msigdbr)
  msig = msigdbr()

  #Subset to include msig
  msig = msig %>% filter(gs_subcat %in% sub_cats)

  #Create empty list
  mod_over = list()
  #Set index for list entries
  index = 1
  #Loop through each pathway
  for (j in unique(msig$gs_name)) {
    #Subset to only include current pathway
    curr_msig = msig %>% filter(gs_name == j)
    #Write to table in format for disease2frame
    write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
    #Get internal/external connectivity
    diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
    #Get disease name
    diseases_curr = diseases_curr %>% mutate(Disease = j)
    mod_over[[index]] = diseases_curr
    index =  index+1
  } 
  #Bind together list

  mod_over = bind_rows(mod_over)

  #Applying categories

  msig_cats  = msig %>% 
    dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
    distinct() #Get categories in order

  #Joining categories

  mod_over = left_join(mod_over, msig_cats)

  #Getting gene counts

  mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
  return(mod_over)
}

```

```{r}
#Setting the three categories we want

subcats = c("CC", "CP:KEGG", "CP:REACTOME")

#Getting biogrid msig slopes

mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links)

#Getting string msig slopes

mod_over_string= get_msig_slopes(subcats = subcats, database = string)

#Getting string filtered msig slopes

mod_over_string_filt = get_msig_slopes(subcats = subcats, database = string_filt)

#Saving the files

saveRDS(mod_over_biogrid, "Biogrid_Msig_Slopes.rds")
saveRDS(mod_over_string, "String_Msig_Slopes.rds")
saveRDS(mod_over_string_filt, "String_Filt_Msig_Slopes.rds")
```

#Define function for calculating slopes and counts from internal/external edges
```{r}
slopes_from_edges = function(edges) {

  edges %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
  group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
  do({
    mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
    data.frame(Slope= coef(mod)[1])
  }) %>% arrange(Slope) %>% filter(Count>4)

}
```

# Get edges for diseases from Biogrid (Connectivity_Comparison.Rmd)

```{r}
#Biogrid

disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")

#Get Biogrid slopes and edge counts

dis_slopes_biogrid_count = slopes_from_edges(disease_overall_biogrid)

#String

disease_overall_string     <- readRDS("string_edges_new.rds")

#Get string slopes and edge counts

dis_slopes_string_count = slopes_from_edges(disease_overall_string)

#String Filtered

disease_overall_string_filt     <- readRDS("string_filt_edges_new.rds")

#Get string filtered slopes and edge counts

dis_slopes_string_filt_count = slopes_from_edges(disease_overall_string_filt)

```

# Define function for calculating normalized coherence and counts ignoring zero internal edges

```{r}
calc_norm_coherence = function(msig_frame, dis_frame, num_msig = 10, num_rand = 10) {
  Count_Sum = msig_frame %>% dplyr::filter(Internal>0) %>% 
  group_by(Disease) %>%
  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
#Splitting into list by categories
Count_Sum = split(Count_Sum, Count_Sum$Category)
coherence_tab = bind_rows()
for (i in unique(dis_frame$Disease)) {
  #Get disease specific genes
  dis = dis_frame %>% filter(Disease == i)
  #Get 30 closest msigdf networks taking 10 from each database and bind
  close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
    arrange(Diff) %>% head(.,num_msig)))
  #Subsetting msig to only include close pathways
  msig_sub = msig_frame %>% filter(Disease %in% close_nets$Disease)
  #Taking slopes for each category
  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
  ))
  #Take median
  mod_slope = median(mod_msig$x)
  #Creating 10 random networks
  j = 1
  slope_list = list()
    while(j<num_rand) {
    #Empty data frame for disease networks
      diseases = bind_rows()
      #Generate random networks until there as many internal edges as current diseases
      while(nrow(diseases)<dis$Count) {
    #Generate random diseases network
    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
    #Save internal and external edges
    diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
    #Filter out zeros
    diseases_curr = diseases_curr %>% filter(Internal !=0)
    #Bind together non-zero internal/external edges
    diseases = bind_rows(diseases, diseases_curr)
      }
      #Save slope of current network
      slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
      #Iterate
      j = j+1

    }

      #Get median of all slopes
      rand_slope = median(unlist(slope_list))

      #Calculate coherence
      coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
      #Place into data frame with disease, coherence, msig normalization value and random slope normalization value
      coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
                             Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
      coherence_tab = bind_rows(coherence, coherence_tab)
      return(coherence_tab)
}
}
```

# Calculating normalized coherence

```{r}
#Get read in msig pathway genes for KEGG, Reactome and GOCC
msig_frame_biogrid = readRDS("Biogrid_Msig_Slopes.rds")
msig_frame_string = readRDS("String_Msig_Slopes.rds")
msig_frame_string_filt = readRDS("String_Filt_Msig_Slopes.rds")

#Calculating normalized coherence
norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string = calc_norm_coherence(msig_frame_string, dis_slopes_string_count, num_msig = num_msig, num_rand = num_rand)
norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_count,num_msig = num_msig, num_rand = num_rand)

#Saving to .rds files
saveRDS(norm_coh_biogrid, "./data/Coherence_Results/coherence_biogrid.rds")
saveRDS(norm_coh_string, "./data/Coherence_Results/coherence_string.rds")
saveRDS(norm_coh_string_filt, "./data/Coherence_Results/coherence_string_filt.rds")
```

# Write general function for saving runs to .csv

```{r}
save_norm_coh = function(coherence_tab, num_msig, num_rand, run, database) {
  require(dplyr)
  fileNameOut <- file.path("./data/Coherence_Results/", paste0("coherence_msig_",database,"_",num_msig, "_rand_", num_rand, "_", run, ".csv"))
  write_csv(coherence_tab, fileNameOut)
}

```

# Save the runs in .csv form

```{r}
#Saving biogrid results to csv
save_norm_coh(norm_coh_biogrid, num_msig = num_msig, num_rand = num_rand, run = run, database = "Biogrid")

#Saving string results to csv
save_norm_coh(norm_coh_string, num_msig = num_msig, num_rand = num_rand, run = run, database = "String")

#Saving string filtered results to csv
save_norm_coh(norm_coh_string_filt, num_msig = num_msig, num_rand = num_rand, run = run, database = "String_Filtered")
```

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
??? to fix

@@
 ---
-title: "Msigdf Biogrid"
+title: "Msigdf Biogrid" [??? Is it the right title?]
 author: "Kellen Cresswell"
-date: "July 19, 2019"
+date: "July 19, 2019" [??? Is it the right date?]
 output: pdf_document
@@
 ```{r}  
+[??? What those functions do?]
 source("./functions/disease2frame.R") 
@@
 
-library(readr) #install.packages("readr")
-library(dplyr) #install.packages("dplyr")
-library(biomaRt) #BiocManager::install("biomaRt")
-library(msigdbr) #BiocManager::install("msigdbr")
-library(igraph) #install.packages("igraph")
-library(tidyr) #install.packages("tidyr")
-library(broom) #install.packages("broom")
-
+library(readr)   # install.packages("readr")
+library(dplyr)   # install.packages("dplyr")
+library(biomaRt) # BiocManager::install("biomaRt")
+library(msigdbr) # BiocManager::install("msigdbr")
+library(igraph)  # install.packages("igraph")
+library(tidyr)   # install.packages("tidyr")
+library(broom)   # install.packages("broom")
 ```
@@
 ```{r}
-#Option for number of msig pathways to take from each category
+# Option for number of msig pathways to take from each category
 num_msig = 10
-#Option for number of random networks to generate (Will generate num_rand-1 networks)
+# Option for number of random networks to generate (Will generate num_rand-1 networks)
 num_rand = 1001
@@
 
-# Read in biogrid data
+# Read in Biogrid data
 
 ```{r}
-#Reading in biogrid data
-
+# Reading in Biogrid data
 links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
-
 links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
@@
 #Place into matrix in form that disease2frame works on
-
-links = links %>% mutate(protein1 =protein1, 
-                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
-
+links = links %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
 links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
-
 ```
 
-# Read in string data and create the string filtered frame
+# Read in String data and create the String filtered dataset
 
 ```{r}
+# Reading in Biogrid data
 string = read_table2("./data/9606.protein.links.v11.0.txt.gz")
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), protein2 = gsub("9606.", "", protein2))
 
-string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
-                           protein2 = gsub("9606.", "", protein2))
-
-#Read in ensemble protein ID
-
+# Read in ensemble protein ID, convert them to gene symbols
 ensembl = useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")
-
 genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)
 
-#Join gene names with string
-
+# Join gene names with string
 string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
-
 string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
-
 string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol.x)
-
 string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol.y)
-
 string_filt = string %>% filter(combined_score>=500)
@@
 
+[??? Why do you read all data at once?]
+
 ```{r}
+[??? What this function does? Write roxygen help]
 
@@
 
-  #Subset to include msig
+  # Subset to include msig
   msig = msig %>% filter(gs_subcat %in% sub_cats)
 
-  #Create empty list
+  # Create empty list
   mod_over = list()
-  #Set index for list entries
+  # Set index for list entries
   index = 1
-  #Loop through each pathway
+  # Loop through each pathway
   for (j in unique(msig$gs_name)) {
@@
   }
-  #Bind together list
 
+  # Bind together list
   mod_over = bind_rows(mod_over)
 
-  #Applying categories
-
+  # Applying categories
   msig_cats  = msig %>% 
@@
 
-  #Joining categories
-
+  # Joining categories
   mod_over = left_join(mod_over, msig_cats)
@@
   #Getting gene counts
-
   mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
@@
 }
-
 ```
 
+[??? What the following chunk does?]
+
 ```{r}
-#Setting the three categories we want
-
+# Setting the three categories we want
 subcats = c("CC", "CP:KEGG", "CP:REACTOME")
 
-#Getting biogrid msig slopes
+# Getting biogrid msig slopes [??? Do not understand, elaborate in text outside the chunk]
+mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links) [??? Could you refactor "links" variable]
 
-mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links)
-
-#Getting string msig slopes
-
+# Getting string msig slopes
 mod_over_string= get_msig_slopes(subcats = subcats, database = string)
 
-#Getting string filtered msig slopes
-
+# Getting string filtered msig slopes
 mod_over_string_filt = get_msig_slopes(subcats = subcats, database = string_filt)
 
-#Saving the files
-
+# Saving the files
 saveRDS(mod_over_biogrid, "Biogrid_Msig_Slopes.rds")
@@
 
-#Define function for calculating slopes and counts from internal/external edges
+# Define function for calculating slopes and counts from internal/external edges
+
 ```{r}
+[??? What this function does? Write roxygen help]
 slopes_from_edges = function(edges) {
-
+  [??? What is "KEGG" category?]
   edges %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
@@
   }) %>% arrange(Slope) %>% filter(Count>4)
-
+  [??? Do not understand, elaborate in text outside the chunk]
 }
@@
 ```{r}
-#Biogrid
-
+# Biogrid
 disease_overall_biogrid     <- readRDS("biogrid_edges_new.rds")
-
-#Get Biogrid slopes and edge counts
-
+# Get Biogrid slopes and edge counts
 dis_slopes_biogrid_count = slopes_from_edges(disease_overall_biogrid)
 
-#String
-
+# String
 disease_overall_string     <- readRDS("string_edges_new.rds")
-
-#Get string slopes and edge counts
-
+# Get string slopes and edge counts
 dis_slopes_string_count = slopes_from_edges(disease_overall_string)
 
-#String Filtered
-
+# String Filtered
 disease_overall_string_filt     <- readRDS("string_filt_edges_new.rds")
-
-#Get string filtered slopes and edge counts
-
+# Get string filtered slopes and edge counts
 dis_slopes_string_filt_count = slopes_from_edges(disease_overall_string_filt)
-
 ```
@@
 ```{r}
+[??? What this function does? Write roxygen help]
+[??? Fix indentations]
 calc_norm_coherence = function(msig_frame, dis_frame, num_msig = 10, num_rand = 10) {
@@
   group_by(Disease) %>%
-  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct()
-#Splitting into list by categories
+  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct() [??? What is that?]
+# Splitting into list by categories [??? Fix indentation]
 Count_Sum = split(Count_Sum, Count_Sum$Category)
 coherence_tab = bind_rows()
+[??? What this loop does?]
 for (i in unique(dis_frame$Disease)) {
-  #Get disease specific genes
+  # Get disease specific genes
   dis = dis_frame %>% filter(Disease == i)
-  #Get 30 closest msigdf networks taking 10 from each database and bind
+  # Get 30 closest msigdf networks taking 10 from each database and bind
   close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
-    arrange(Diff) %>% head(.,num_msig)))
-  #Subsetting msig to only include close pathways
+    arrange(Diff) %>% head(.,num_msig))) [??? Where is 30? Should be a setting to tweak. Describe better what 30 means]
+  # Subsetting msig to only include close pathways
   msig_sub = msig_frame %>% filter(Disease %in% close_nets$Disease)
-  #Taking slopes for each category
+  # Taking slopes for each category
   mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
@@
   ))
-  #Take median
+  # Take median
   mod_slope = median(mod_msig$x)
-  #Creating 10 random networks
+  # Creating num_rand random networks
   j = 1
   slope_list = list()
-    while(j<num_rand) {
-    #Empty data frame for disease networks
+    while(j < num_rand) {
+    # Empty data frame for disease networks
       diseases = bind_rows()
-      #Generate random networks until there as many internal edges as current diseases
+      # Generate random networks until there as many internal edges as current diseases
       while(nrow(diseases)<dis$Count) {
-    #Generate random diseases network
-    diseases_curr = GeneSample(dis$Count, links, directed = FALSE)
-    #Save internal and external edges
+    # Generate random diseases network [??? Fix indentation]
+    diseases_curr = GeneSample(dis$Count, links, directed = FALSE) [??? What this function does?]
+    # Save internal and external edges
     diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
-    #Filter out zeros
-    diseases_curr = diseases_curr %>% filter(Internal !=0)
-    #Bind together non-zero internal/external edges
+    # Filter out zeros
+    diseases_curr = diseases_curr %>% filter(Internal !=0) [??? What happens if, after filtering zeros, the network will be removed? Detail]
+    # Bind together non-zero internal/external edges
     diseases = bind_rows(diseases, diseases_curr)
       }
-      #Save slope of current network
+      # Save slope of current network
       slope_list[[j]] = coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]]
-      #Iterate
+      # Iterate
       j = j+1
@@
 
-      #Get median of all slopes
+      # Get median of all slopes
       rand_slope = median(unlist(slope_list))
 
-      #Calculate coherence
+      # Calculate coherence
       coherence = (dis$Slope - rand_slope)/(mod_slope-rand_slope)
-      #Place into data frame with disease, coherence, msig normalization value and random slope normalization value
+
+      # Place into data frame with disease, coherence, msig normalization value and random slope normalization value
       coherence = data.frame(Disease = dis$Disease, Coherence = coherence, 
@@
 ```{r}
-#Get read in msig pathway genes for KEGG, Reactome and GOCC
+# Get read in msig pathway genes for KEGG, Reactome and GOCC
+[??? Why read them if they are created above? Rename variables above. It looks like you do not need to save/read those objects at all]
 msig_frame_biogrid = readRDS("Biogrid_Msig_Slopes.rds")
@@
 
-#Calculating normalized coherence
+# Calculating normalized coherence 
 norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, num_msig = num_msig, num_rand = num_rand)
@@
 }
-
 ```

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
More bugs ???

@@
 source("./functions/GeneSample.R")
-source("./functions/get_msig_slopes.R")
+source("./functions/get_msig_slopes.R") [??? This does not exist!]
 
@@
 
-get_msig_slopes = function(sub_cats, database) {
+get_msig_slopes = function(sub_cats, database) { [??? Calls to this function use "subcat", not "sub_cat" argument - correct]
   require(msigdbr)
@@
 
+[??? After hacking around, normalized coherence objects still contain data for ONE disease, not many. What is going on?]
+
 # Write general function for saving runs to .csv

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
move msig slope calculations into get_msig_slopes and other fixes

--- 2019-09-14 Analysis/Normalized_Coherence.Rmd
@@
 ---
-title: "Msigdf Biogrid" [??? Is it the right title?]
+title: "Normalized Coherence"
 author: "Kellen Cresswell"
-date: "July 19, 2019" [??? Is it the right date?]
+date: "September 13, 2019"
 output: pdf_document
@@
 ```{r}  
-[??? What those functions do?]
+#Read in function for converting genes into data frame of internal/external edges
 source("./functions/disease2frame.R") 
+#Read in function for randomly sampling gene networks
 source("./functions/GeneSample.R")
-source("./functions/get_msig_slopes.R")
 
@@
 # Reading in Biogrid data
-links = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
-links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+biogrid = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+biogrid = biogrid %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
 
 #Split based on Biogrid ID system then pull out gene names
-protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
-protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein1 = lapply(strsplit(biogrid$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(biogrid$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
 protein1 = unlist(protein1)
@@
 #Place into matrix in form that disease2frame works on
-links = links %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
-links = links %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
+biogrid = biogrid %>% mutate(protein1 =protein1, protein2 = protein2) %>% dplyr::select(protein1, protein2)
+biogrid = biogrid %>% dplyr::select(hgnc_symbol_a = protein1,hgnc_symbol_b = protein2)
 ```
@@
     #Get internal/external connectivity
-    diseases_curr = diseases2frame("temp_rea_bio.txt", links = links)
+    diseases_curr = diseases2frame("temp_rea_bio.txt", links = biogrid)
     #Get disease name
@@
   #Getting gene counts
-  mod_over = mod_over %>% group_by(Disease) %>% mutate(Num_Genes = n())
+  mod_over = mod_over %>% group_by(Disease) %>% filter(Internal != 0) %>% mutate(Num_Genes = n())
+
+  #Getting slopes from each diseases
+  mod_over = mod_over %>% group_by(Disease, Category, Num_Genes) %>% do(tidy(coef(lm(sqrt(External)~ sqrt(Internal)-1, data = .))[[1]]))
+  #Rename x (Default do tidy result) to Slope
+  mod_over = mod_over %>% rename(Slope = x)
   return(mod_over)
@@
 
-[??? What the following chunk does?]
+# Getting slope between internal/external edges from GOCC, KEGG and REACTOME pathways
 
@@
 # Getting biogrid msig slopes [??? Do not understand, elaborate in text outside the chunk]
-mod_over_biogrid = get_msig_slopes(subcats = subcats, database = links) [??? Could you refactor "links" variable]
+msig_frame_biogrid = get_msig_slopes(sub_cats = subcats, database = biogrid)
 
 # Getting string msig slopes
-mod_over_string= get_msig_slopes(subcats = subcats, database = string)
+msig_frame_string = get_msig_slopes(sub_cats = subcats, database = string)
 
 # Getting string filtered msig slopes
-mod_over_string_filt = get_msig_slopes(subcats = subcats, database = string_filt)
+msig_frame_string_filt = get_msig_slopes(sub_cats = subcats, database = string_filt)
 
 # Saving the files
-saveRDS(mod_over_biogrid, "Biogrid_Msig_Slopes.rds")
-saveRDS(mod_over_string, "String_Msig_Slopes.rds")
-saveRDS(mod_over_string_filt, "String_Filt_Msig_Slopes.rds")
+saveRDS(msig_frame_biogrid, "Biogrid_Msig_Slopes.rds")
+saveRDS(msig_frame_string, "String_Msig_Slopes.rds")
+saveRDS(msig_frame_string_filt, "String_Filt_Msig_Slopes.rds")
 ```
@@
 slopes_from_edges = function(edges) {
-  [??? What is "KEGG" category?]
-  edges %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+  edges %>% filter(Internal != 0)  %>%
   group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
@@
   }) %>% arrange(Slope) %>% filter(Count>4)
-  [??? Do not understand, elaborate in text outside the chunk]
 }
 ```
+
+The above function takes internal and external edges from each gene for each disease and calculates the total number of genes and the slope of internal vs.
+external edges.
+Genes with zero internal edges are filtered and diseases with less than 5 non-zero internal edges are filtered.
 
@@
 [??? What this function does? Write roxygen help]
-[??? Fix indentations]
-calc_norm_coherence = function(msig_frame, dis_frame, num_msig = 10, num_rand = 10) {
-  Count_Sum = msig_frame %>% dplyr::filter(Internal>0) %>% 
-  group_by(Disease) %>%
-  mutate(Num_Genes = n()) %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct() [??? What is that?]
-# Splitting into list by categories [??? Fix indentation]
-Count_Sum = split(Count_Sum, Count_Sum$Category)
+calc_norm_coherence = function(msig_frame, dis_frame, database, num_msig = 10, num_rand = 10) {
+  # Splitting into list by categories 
+  Count_Sum = split(msig_frame %>% dplyr::select(Disease,Category, Num_Genes) %>% distinct() , msig_frame$Category)
   coherence_tab = bind_rows()
-[??? What this loop does?]
+  #Loop through each disease name and get msig normalization and random normalization factors
   for (i in unique(dis_frame$Disease)) {
@@
     dis = dis_frame %>% filter(Disease == i)
-  # Get 30 closest msigdf networks taking 10 from each database and bind
+    # Get msigdf pathways with cloest number of genes, taking num_msig from each category
     close_nets = bind_rows(lapply(Count_Sum,  function(x) x %>% mutate(Diff = abs(Num_Genes-dis$Count)) %>%
-    arrange(Diff) %>% head(.,num_msig))) [??? Where is 30? Should be a setting to tweak. Describe better what 30 means]
+      arrange(Diff) %>% head(.,num_msig)))
     # Subsetting msig to only include close pathways
     msig_sub = msig_frame %>% filter(Disease %in% close_nets$Disease)
-  # Taking slopes for each category
-  mod_msig = msig_sub %>% group_by(Disease) %>% do(tidy(
-    coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = .))[[1]]
-  ))
     # Take median
-  mod_slope = median(mod_msig$x)
+    mod_slope = median(msig_sub$Slope)
     # Creating num_rand random networks
@@
       while(nrow(diseases)<dis$Count) {
-    # Generate random diseases network [??? Fix indentation]
-    diseases_curr = GeneSample(dis$Count, links, directed = FALSE) [??? What this function does?]
+        # Generate random diseases network from PPI database[??? What this function does?]
+        diseases_curr = GeneSample(dis$Count, database, directed = FALSE) 
         # Save internal and external edges
@@
         # Filter out zeros
-    diseases_curr = diseases_curr %>% filter(Internal !=0) [??? What happens if, after filtering zeros, the network will be removed? Detail]
-    # Bind together non-zero internal/external edges
+        diseases_curr = diseases_curr %>% filter(Internal !=0) 
+        # Bind together non-zero internal/external edges (If all zero then diseases_curr is null and nothing happens)
         diseases = bind_rows(diseases, diseases_curr)
@@
       coherence_tab = bind_rows(coherence, coherence_tab)
+      print(coherence_tab)
       return(coherence_tab)
@@
 ```{r}
-# Get read in msig pathway genes for KEGG, Reactome and GOCC
-[??? Why read them if they are created above? Rename variables above. It looks like you do not need to save/read those objects at all]
-msig_frame_biogrid = readRDS("Biogrid_Msig_Slopes.rds")
-msig_frame_string = readRDS("String_Msig_Slopes.rds")
-msig_frame_string_filt = readRDS("String_Filt_Msig_Slopes.rds")
-
 # Calculating normalized coherence 
-norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, num_msig = num_msig, num_rand = num_rand)
-norm_coh_string = calc_norm_coherence(msig_frame_string, dis_slopes_string_count, num_msig = num_msig, num_rand = num_rand)
-norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_count,num_msig = num_msig, num_rand = num_rand)
+norm_coh_biogrid = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_count, database = biogrid, num_msig = num_msig, num_rand = num_rand)
+norm_coh_string = calc_norm_coherence(msig_frame_string, dis_slopes_string_count, database = string, num_msig = num_msig, num_rand = num_rand)
+norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_count, database = string_filt, num_msig = num_msig, num_rand = num_rand)
 

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
fix issue where calc_norm_coherence only prints 1 row

@@
       coherence_tab = bind_rows(coherence, coherence_tab)
-      print(coherence_tab)
+  }
   return(coherence_tab)
-}
+
 }

### 2019-09-14 Analysis/Normalized_Coherence.Rmd
Tweaks to make it run

@@
 author: "Kellen Cresswell"
-date: "September 13, 2019"
+date: "`r Sys.Date()`"
 output: pdf_document
 ---
+
+```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
+# Set up the environment
+library(knitr)
+opts_chunk$set(cache.path='cache/', fig.path='img/', cache=T, tidy=T, fig.keep='high', echo=F, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6) #out.width=700,
+library(pander)
+panderOptions('table.split.table', Inf)
+set.seed(1)
+library(dplyr)
+options(stringsAsFactors = FALSE)
+```
 
@@
 # Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 1001
+num_rand = 11
 # Construct file name to save the results
@@
 # Reading in Biogrid data
-biogrid = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
+biogrid = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
 biogrid = biogrid %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
@@
 
-[??? Why do you read all data at once?]
-
 ```{r}
-[??? What this function does? Write roxygen help]
+# [??? What this function does? Write roxygen help]
 
@@
 ```{r}
-[??? What this function does? Write roxygen help]
+# [??? What this function does? Write roxygen help]
 slopes_from_edges = function(edges) {
@@
 ```{r}
-[??? What this function does? Write roxygen help]
+# [??? What this function does? Write roxygen help]
 calc_norm_coherence = function(msig_frame, dis_frame, database, num_msig = 10, num_rand = 10) {

### 2019-09-15 Analysis/Coherence_Calculations_New.Rmd
update to match new output

@@
 #STRING
-dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+dis_slopes_string_count = disease_overall_string %>% filter(Internal != 0) %>%
     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
@@
 #STRING Filtered
-dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+dis_slopes_string_filt_count = disease_overall_string_filt %>% filter(Internal != 0) %>%
     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
@@
 #BIOGRID
-dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0) %>% filter(Category != "Random") %>% filter(Category != "KEGG") %>%
+dis_slopes_biogrid_count = disease_overall_biogrid %>% filter(Internal != 0)  %>%
     group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
@@
 
-#Reactome
+#Full join KEGG coherence results together based on disease
+coherence_disease = full_join(coherence_biogrid, coherence_string, by = "Disease")
 
-coherence_string_filt_react = readRDS("./data/Coherence_Results/coherence_string_filt_react.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string_react = readRDS("./data/Coherence_Results/coherence_string_react.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid_react = readRDS("./data/Coherence_Results/coherence_biogrid_react.rds") %>% dplyr::select(Disease, Coherence)
-
-#CC
-
-coherence_string_filt_cc = readRDS("./data/Coherence_Results/coherence_string_filt_CC.rds") %>% dplyr::select(Disease, Coherence)
-coherence_string_cc = readRDS("./data/Coherence_Results/coherence_string_CC.rds") %>% dplyr::select(Disease, Coherence)
-coherence_biogrid_cc = readRDS("./data/Coherence_Results/coherence_biogrid_CC.rds") %>% dplyr::select(Disease, Coherence)
-
-#Full join KEGG coherence results together based on disease
-coherence_full = full_join(coherence_biogrid, coherence_string, by = "Disease")
-
-coherence_full = full_join(coherence_full, coherence_string_filt, by = "Disease")
+coherence_disease = full_join(coherence_disease, coherence_string_filt, by = "Disease")
 
@@
 
-colnames(coherence_full) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")
-
-#Save coherence full in this state for later manipulation
-coherence_disease = coherence_full
+colnames(coherence_disease) = c("Disease", "Biogrid_Coherence", "String_Coherence", "String_Filt_Coherence")
 
@@
 ```{r}
-#Finish all disease plot (From save coherence disease frame)
-
-#Joining REACTOME
-
-coherence_disease = left_join(coherence_disease, coherence_biogrid_react, by= c("Disease"= "Disease"))
-
-coherence_disease = left_join(coherence_disease, coherence_string_react, by= c("Disease"= "Disease"))
-
-coherence_disease = left_join(coherence_disease, coherence_string_filt_react, by= c("Disease"= "Disease"))
-
-#Joining CC
-
-coherence_disease = left_join(coherence_disease, coherence_biogrid_cc, by= c("Disease"= "Disease"))
-
-coherence_disease = left_join(coherence_disease, coherence_string_cc, by= c("Disease"= "Disease"))
-
-coherence_disease = left_join(coherence_disease, coherence_string_filt_cc, by= c("Disease"= "Disease"))
-
-coherence_disease = coherence_disease  %>% dplyr::rename(`Biogrid Coherence Reactome` = Coherence.x, `String Coherence Reactome` = Coherence.y, `String Filt Coherence Reactome` = Coherence.x.x, `Biogrid Coherence CC` = Coherence.y.y, `String Coherence CC` = Coherence.x.x.x, `String Filt Coherence CC` = Coherence.y.y.y)
-
 #Get categories from slope frame
@@
 
-coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Biogrid Coherence`:`String Filt Coherence CC`) %>% dplyr::distinct()
-
 #Adding gene counts
@@
 
-coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Number of SNPs`, `Number of genes`, `Biogrid Coherence`:`String Filt Coherence CC`)
+coherence_disease = coherence_disease %>% dplyr::select(Disease, Category, `Number of SNPs`, `Number of genes`, `Biogrid Coherence`:`String Filt Coherence`)
 ```
@@
 
-#Trait dataset 
-coherence_full = left_join(coherence_full,biogrid_p %>% dplyr::select(-Ave))
-
-coherence_full = left_join(coherence_full,string_p%>% dplyr::select(-Ave))
-
-coherence_full = left_join(coherence_full,string_filt_p%>% dplyr::select(-Ave))
-
 #Complete Disease dataset by adding p-values
@@
 
-#Rename KEGG pathways when necessary
-
-coherence_disease = coherence_disease %>% dplyr::rename(`String Coherence KEGG` = `String Coherence`, `String Filt Coherence KEGG` = `String Filt Coherence`, `Biogrid Coherence KEGG` = `Biogrid Coherence`)
-
-#Rename CC to GOCC to better reflect source
-colnames(coherence_disease) = gsub("CC", "GOCC", colnames(coherence_disease))
-
 #Writing to a tables

### 2019-09-16 Analysis/Coherence_Calculations_New.Rmd
fix doubling up of diseases

@@
 
-Cat = bind_rows(dis_slopes_string_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct(), dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(Disease, Category) %>% distinct())
+Cat = bind_rows(dis_slopes_string_count %>% ungroup() %>% dplyr::select(Disease, Category), dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(Disease, Category) ) %>% distinct()
 
@@
 
-coherence_disease =left_join(coherence_disease, Cat)
+coherence_disease =left_join(coherence_disease, Cat, "Disease")
 

### 2019-09-16 Analysis/Category_Tables.Rmd
Rename size_slope to Fig2

@@
 
-Disease_String_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Coherence.KEGG) %>% filter(!is.na(String.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
+Disease_String_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Coherence) %>% filter(!is.na(String.Coherence)) %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Coherence, na.rm = TRUE), Max_Coherence = max(String.Coherence, na.rm = TRUE), Mean_Coherence = mean(String.Coherence, na.rm = TRUE), Median_Coherence =median(String.Coherence, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
-Disease_String_Filt_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Filt.Coherence.KEGG) %>% filter(!is.na(String.Filt.Coherence.KEGG))   %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Filt.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(String.Filt.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(String.Filt.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(String.Filt.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
+Disease_String_Filt_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, String.Filt.Coherence) %>% filter(!is.na(String.Filt.Coherence))   %>% group_by(Category) %>% summarise(Min_Coherence = min(String.Filt.Coherence, na.rm = TRUE), Max_Coherence = max(String.Filt.Coherence, na.rm = TRUE), Mean_Coherence = mean(String.Filt.Coherence, na.rm = TRUE), Median_Coherence =median(String.Filt.Coherence, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
-Disease_Biogrid_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, Biogrid.Coherence.KEGG) %>% filter(!is.na(Biogrid.Coherence.KEGG)) %>% group_by(Category) %>% summarise(Min_Coherence = min(Biogrid.Coherence.KEGG, na.rm = TRUE), Max_Coherence = max(Biogrid.Coherence.KEGG, na.rm = TRUE), Mean_Coherence = mean(Biogrid.Coherence.KEGG, na.rm = TRUE), Median_Coherence =median(Biogrid.Coherence.KEGG, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
+Disease_Biogrid_Sum = Disease_Sum %>% dplyr::select(Disease, Category, Number.of.SNPs, Number.of.genes, Biogrid.Coherence) %>% filter(!is.na(Biogrid.Coherence)) %>% group_by(Category) %>% summarise(Min_Coherence = min(Biogrid.Coherence, na.rm = TRUE), Max_Coherence = max(Biogrid.Coherence, na.rm = TRUE), Mean_Coherence = mean(Biogrid.Coherence, na.rm = TRUE), Median_Coherence =median(Biogrid.Coherence, na.rm = TRUE), Min_Genes = min(Number.of.genes, na.rm = TRUE), Max_Genes = max(Number.of.genes, na.rm = TRUE), Mean_Genes = mean(Number.of.genes, na.rm = TRUE), Median_Genes = median(Number.of.genes, na.rm = TRUE) , Min_SNPs = min(Number.of.SNPs, na.rm = TRUE), Max_SNPs = max(Number.of.SNPs, na.rm = TRUE), Mean_SNPs = mean(Number.of.SNPs, na.rm = TRUE), Median_SNPs = median(Number.of.SNPs, na.rm = TRUE), Total_Networks = n()) %>% arrange(-Median_Coherence)
 
@@
 
-Cat_Within = Disease_Sum %>% group_by(Category) %>% summarise(`Number of SNPs`= sum(`Number.of.SNPs`), `Number of Genes` = sum(Number.of.genes), `Total Diseases` = n(),  `Mean Biogrid Coherence` =mean(Biogrid.Coherence.KEGG, na.rm = TRUE), `Mean String Coherence` = mean(String.Coherence.KEGG, na.rm = TRUE), `Mean String Filtered Coherence` = mean(String.Filt.Coherence.KEGG, na.rm = TRUE)) %>% arrange(Category)
+Cat_Within = Disease_Sum %>% group_by(Category) %>% summarise(`Number of SNPs`= sum(`Number.of.SNPs`), `Number of Genes` = sum(Number.of.genes), `Total Diseases` = n(),  `Mean Biogrid Coherence` =mean(Biogrid.Coherence, na.rm = TRUE), `Mean String Coherence` = mean(String.Coherence, na.rm = TRUE), `Mean String Filtered Coherence` = mean(String.Filt.Coherence, na.rm = TRUE)) %>% arrange(Category)
 
@@
 
-ggsave("../manuscript/Figures/Size_Slope_Plot.png", slope_plot, height=5, width=7, units="in", dpi = 600)
+ggsave("../manuscript/Figures/Fig2_Size_Slope_Plot.png", slope_plot, height=5, width=7, units="in", dpi = 600)
 ```
+
+```{}

### 2019-09-17 Analysis/Normalized_Coherence.Rmd
Code for generating random networks

@@
 ```
+
+# Generating random networks for comparison to msigdf (Category_Tables)
+
+```{r}
+#Function for generating random networks from a set of sizes
+#size_vec is vector of sizes
+
+generate_rand_slopes = function(size_vec, database) {
+  j=1
+  slope_list = c()
+  for (i in size_vec) {
+    # Empty data frame for disease networks
+      diseases = bind_rows()
+      # Generate random networks until there as many internal edges as current diseases
+      while(nrow(diseases)<i) {
+        # Generate random diseases network from PPI database[??? What this function does?]
+        diseases_curr = GeneSample(i, database, directed = FALSE) 
+        # Save internal and external edges
+        diseases_curr = data.frame(Gene = names(diseases_curr$internal_connectivity), Internal = diseases_curr$internal_connectivity, External = diseases_curr$external_connectivity)
+        # Filter out zeros
+        diseases_curr = diseases_curr %>% filter(Internal !=0) 
+        # Bind together non-zero internal/external edges (If all zero then diseases_curr is null and nothing happens)
+        diseases = bind_rows(diseases, diseases_curr)
+      }
+          # Save slope of current network
+          slope_list= c(slope_list,coef(lm(I(sqrt(External)) ~ I(sqrt(Internal))-1, data = diseases))[[1]])
+          # Iterate
+          j = j+1
+
+  }
+  data.frame(Count = size_vec, Slope = slope_list)
+}
+```
+
+```{r}
+#Read in data if you only want to run this section
+
+# msig_frame_string = readRDS("String_Msig_Slopes.rds")
+
+#Get msig counts (Comment out if using pre-calculated msig slopes)
+
+msig_counts = unique((msig_frame_string %>% filter(Category == "CP:KEGG"))$Num_Genes)
+
+#Generate the random slopes equal to msig lengths
+string_random = generate_rand_slopes(msig_counts, string)
+
+ saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
+```

### 2019-09-17 Analysis/Normalized_Coherence.Rmd

@@
 saveRDS(msig_frame_string_filt, "String_Filt_Msig_Slopes.rds")
+```
+
+# Getting all msig pathway slopes for string (Used in Supplementary Table 2)
+
+```{r}
+#Re-read msig
+msig = msigdbr()
+
+# Define all categories
+all_cats = unique(msig$gs_subcat)
+
+#Remove empty categories
+all_cats = subcats[which(subcats!="")]
+
+#Get slopes
+msig_frame_string_all = get_msig_slopes(sub_cats = all_cats, database = string)
+
+#Save results
+saveRDS(msig_frame_string_all, "./data/coherence_results/All_Msig_String.rds")
+
 ```

### 2019-09-17 manuscript/manuscript.Rmd
Previously uncommitted changes

@@
 High/low degree of interactions is suggestive of coherent/random molecular mechanisms, respectively.
-Comparing the degree of interactions may help to better predict and understand the coherence of the molecular mechanisms underlying complex phenotypes.
+Comparing the degree of interactions may help to better understand the coherence of the molecular mechanisms underlying complex phenotypes and the potential for therapeutic intervention.
 However, direct comparison of the degree of interactions is difficult due to differences in size and configuration of phenotype-associated gene networks.
@@
 
-**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (with exemplary high coherence) and random networks (the lower limit of coherence).
+**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (the reference for high coherence) and random networks (the lower limit of coherence).
 We observed a range of coherence estimates for each category of phenotypes.
@@
 
-This study compares the connectivity of networks formed by disease- or trait-associated genes and gene products.
+<!--This study compares the connectivity of networks formed by disease- or trait-associated genes and gene products.
 We hypothesized that genes associated with (categories of) diseases share common molecular mechanisms and, hence, form connected (i.e., coherent) networks.
@@
 Cancer, anthropometric traits, eye-, and cardiovascular diseases and traits formed networks of intermediate connectivity.
-Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.
-
-This study presents a novel measure of network coherence as a ratio of the normalized internal vs.
+Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.-->
+This study presents a novel measure of network coherence as a ratio of the internal vs.
 external degree distributions.
@@
 We used KEGG pathways [@Kanehisa:2000aa] as a reference for a paradigmatically high-coherent network [@Lee:2003aa], while sets of randomly sampled genes were used as the reference for the absence of coherence.
+Using KEGG and random networks matched in size to a phenotype-specific network, we derived normalized measure of coherence that can be compared across phenotypes.
 We hypothesized that the level of coherence may inform us about similarities and differences among phenotypes.
 Using two PPI databases (STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa]), we compared coherence of 133 phenotypes from GWAScatalog [@Hindorff:2009aa; @MacArthur:2017aa].
-Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have the most coherence.
+Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have high coherence.
 Our method enables direct comparison of coherence measures and highlight gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.
@@
 The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs are removed from the analysis (see Methods).
-Second, unfiltered STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING full”).
-The advantage of using unfiltered data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
-The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs.
+Second, complete STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING”).
+The advantage of using complete data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
+The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs having low confidence scores.
 Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, were used.
-The use of different data sources (STRING and BioGrid) and filtering (full or filtered STRING data) was intended to increase the generalizability of our conclusions.
+The use of different data sources (STRING and BioGrid) and filtering (complete or filtered STRING data) was intended to increase the generalizability of our conclusions.
 
 <!--Tables/table_1.csv-->
-**Table 1. Summary statistics of the analyzed categories.** Average coherence estimates are shown. "NA" values indicate that a phenotype was not considered due to the low number of genes with PPIs.
+**Table 1.
+Summary statistics of the analyzed categories.** Average coherence estimates are shown.
+"NA" values indicate that a category lacked phenotypes with sufficient number of PPIs.
+Sorted by "String Filtered" average coherence.
 
@@
 In genomics, nodes typically represent genes and edges correspond to some measure of interactions, e.g., gene co-expression or interactions between protein products [@Goncalves:2012aa; @Wang:2012aa].
-In terms of phenotypes, an intuitive expectation is that they will be represented by coherent modules of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
+In terms of phenotypes, an intuitive expectation is that they will be represented by coherent networks of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
 In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the network [@fortunato2010community; @Radicchi:2004aa].
@@
 A collection of degrees of all network nodes form a degree distribution.
-Comparing degree distributions among networks is an intuitive way to gain an understanding of network similarities and differences [@Barabasi:2004aa; @Feldman:2008aa].
-We hypothesized that comparing degree distributions of networks formed by phenotype-associated gene sets may inform us about similarities and differences among phenotypes in terms of coherence of their antecedent and underlying molecular mechanisms.
+Comparing degree distributions among networks is an intuitive way to gain an understanding of network similarities and differences in terms of connectivity (aka coherence) of the corresponding genes [@Barabasi:2004aa; @Feldman:2008aa].
+We hypothesized that comparing degree distributions of networks formed by phenotype-associated gene sets may inform us about similarities and differences among phenotypes in terms of coherence estimates, informing us about the underlying molecular mechanisms.
 
@@
 Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks of high coherence.
-This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathwaays (REACTOME genes) are expected to interact more frequently.
+This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathwaays (REACTOME genes) are presumed to interact more frequently.
 In our study, KEGG networks were used as the primary reference of high coherence, in parallel with GOCC and REACTOME collections.
@@
 The degree distributions of the ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).
-Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have degree 9, while all nodes in the 20-node network have degree 19, Figure 1B).
+Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have degree "9", while all nodes in the 20-node network have degree "19", Figure 1B).
 <!--Normalizing the degree distributions of the fully connected networks by dividing them by network size makes them identical as would be expected for networks with similar coherence.
@@
 To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
-We observed negative correlation of network size with the slopes of KEGG and random networks (-0.37/-0.85, respectively, Figure 2, Supplementary Table S2).
+We observed negative correlation of network size with the slopes of KEGG and random networks (-0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
 These observations confirm the notion that coherence estimation using internal vs.
@@
 external degree distributions created by these size-matched references to normalize the slope of the phenotype-specific network to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed removes the dependency between coherence and network size (Supplementary Figure S1).
+We found that the use of size-matched references indeed removes the dependency between coherence and network size (Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 To estimate the significance of coherence measures, we developed a permutation-based approach (see Methods).
-Notably, we observed coherence for all diseases to be highly significant (permutation p-value < 0.002, Supplementary Table S3).
-The distribution of coherence estimates was skewed towards 1.00 (highest coherence), with the median coherence being 0.86.
+Notably, we observed coherence for all diseases to be highly significantly different from random (permutation p-value < 0.002, Supplementary Table S3).
+Similarly, the distribution of coherence estimates was skewed towards 1.00 (highest coherence), with the median coherence being 0.86.
+Among traits, "Corneal structure", "Body mass index", "Weight" were the most coherent (1.02, 1.00, 0.99, respectively).
+[???]
+
 "Obesity-related traits" and "Height" were on the lowest range of coherence estimates (0.42 and 0.54, respectively).
@@
 
-We estimated coherence of 53 traits from five categories (Figure 3, Supplementary Figures S3-S5).
+We estimated coherence of 54 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S3-S5).
 Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
@@
 The STRING data for the tax ID "9606" (human) was used, and Ensembl protein IDs were converted to gene names using BioMart.
-Either the full dataset (11,761,040  entries) or data filtered by "combined_score" > 500 (range 0-1000) (1,373,946 entries, the main PPI dataset) was used.
+Either the complete dataset (11,761,040  entries) or data filtered by "combined_score" > 500 (range 0-1000) (1,373,946 entries, the main PPI dataset) was used.
 The BioGrid data was filtered to include "TaxID interactor" type equal to "9606", all "Interaction Types" were included, totaling 465,660 interactions.
@@
 
-**Supplementary Table S1. Summary statistics of the analyzed phenotypes.** Each worksheet corresponds to the PPI database used.
+**Supplementary Table S1.
+Summary statistics of the analyzed phenotypes.** Minimum, mean, median, maximum of coherence estimates, the number of genes, SNPs, the total number of phenotype networks.
+Each worksheet corresponds to the PPI database used.
 
 **Supplementary Table S2.
-Summary of networks from MSigDB categories.** Minimum, maximum, mean and median of the slopes and number of genes are shown for each collection, along with the total number of networks and the correlation between network size and slope.
+Summary of networks from MSigDB categories.** Minimum, mean, median, maximum of the slopes, the number of genes, and the total number of networks are shown for each collection, along with the correlation between network size and slope.
+"CP:/KEGG/Reactome/Biocarta" - canonical pathways, "GOBP/GOMF/GOCC" - gene ontology biological processes, molecular functions, cellular component, "TFT" - transcription factor targets, "CGN" - cancer gene neighborhoods, "CM" - cancer modules, "MIR" - microRNA targets, "CGP" - chemical and genetic perturbations.
 
 **Supplementary Table S3.
-Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) and the reference of high coherence (KEGG, GOCC, Reactome) are specified in the corresponding columns for coherence and permutation p-value estimations.
+Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) and the reference of high coherence (KEGG, GOCC, Reactome) are specified in the corresponding columns for coherence.
+Permutation p-values were estimated using STRING filtered database and KEGG reference of high coherence.
+"Trait" and "Disease" sheets correspond to subsets of the main table containing coherence estimates for traits/diseases, respectively, sorted by descending "String Filt Coherence KEGG" column.
 
-**Supplementary Figure S1. Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.
+**Supplementary Figure S1.
+Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases and KEGG pathways as a reference of high coherence.
 
 **Supplementary Figure S2.
-Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.
+Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the normalized coherence estimates using the corresponding PPI databases and KEGG/Reactome pathways and Gene Ontology Cellular Component networks as a reference of high coherence.
 
 **Supplementary Figure S3.
-Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
-Missing entries indicate that for a given PPI database a phenotype-associated network could not be built.
+Coherence estimates of traits using KEGG as a reference of high coherence.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+Plots are faceted by PPI database used to estimate coherence.
+Missing entries indicate that, for a given PPI database, a phenotype-associated network could not be built.
+
+**Supplementary Figure S4. Coherence estimates of traits using GOCC as a reference of high coherence.** See legend for Supplementary Figure S3.
+
+**Supplementary Figure S5. Coherence estimates of traits using Reactome as a reference of high coherence.** See legend for Supplementary Figure S3.
+
+**Supplementary Figure S6. Coherence estimates of diseases using KEGG as a reference of high coherence.** See legend for Supplementary Figure S3.
+
+**Supplementary Figure S7. Coherence estimates of diseases using GOCC as a reference of high coherence.** See legend for Supplementary Figure S3.
+
+**Supplementary Figure S8. Coherence estimates of diseases using Reactome as a reference of high coherence.** See legend for Supplementary Figure S3.
 

### 2019-09-17 Analysis/Normalized_Coherence.Rmd
??? BUG?

@@
     #Get internal/external connectivity
-    diseases_curr = diseases2frame("temp_rea_bio.txt", links = biogrid)
+    diseases_curr = diseases2frame("temp_rea_bio.txt", links = biogrid) # [??? Shouldn't it be database?]
     #Get disease name

### 2019-09-17 Analysis/Normalized_Coherence.Rmd
fix bug

--- 2019-09-17 Analysis/Normalized_Coherence.Rmd
@@
     #Get internal/external connectivity
-    diseases_curr = diseases2frame("temp_rea_bio.txt", links = biogrid)
+    diseases_curr = diseases2frame("temp_rea_bio.txt", links = database)
     #Get disease name

### 2019-09-17 manuscript/manuscript.Rmd
Add ref to size dependency

@@
 
-Direct comparison of degree distributions of the networks formed by phenotype-associated genes is hindered by the fact they depend on network size and configuration.
+Direct comparison of degree distributions of the networks formed by phenotype-associated genes is hindered by the fact they depend on network size and configuration [@Leskovec:2010:ECA:1772690.1772755].
 An intuitive example is to consider a ring and a fully connected network of size 10 and 20 nodes, respectively (Figure 1).
@@
 
-Our approach is similar in spirit to the well-researched community detection problem [@danon2005comparing; @fortunato2010community; @Radicchi:2004aa].
+Our approach is similar in spirit to the well-researched community detection problem [@danon2005comparing; @fortunato2010community; @Radicchi:2004aa; @Leskovec:2010:ECA:1772690.1772755].
 However, we approach this problem from a different angle by asking whether a given network as a whole can be considered a community within the global PPI network.
@@
 In addition to internal vs.
-external degree distributions, other network properties can be used to estimate coherence.
+external degree distributions, other network properties can be used to estimate coherence [@Leskovec:2010:ECA:1772690.1772755].
 For example, the centrality measure has been used to demonstrate that highly central hub genes are functionally essential and highly conserved [@Khurana:2013aa].

### 2019-09-17 manuscript/manuscript.Rmd
Discuss outliers

@@
 external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa].
-Our limited knowledge of protein-protein interactions limits the coherence estimation to phenotypes that have a sufficient number of genes with PPI annotations.
+Our limited knowledge of protein-protein interactions limits the coherence estimation to phenotypes that have a sufficient number of genes with PPI annotations, especially for phenotypes with low numbers of associated genes.
 Furthermore, networks formed by phenotype-associated genes should be sufficiently diverse to avoid cases in which all genes in a network have the same degree.
@@
 This problem is best illustrated by the BioGrid PPI database that has considerably fewer PPIs; hence, fewer phenotypes could be analyzed.
-Furthermore, using smaller PPI databases like BioGrid inevitably diminishes the size of phenotype-associated networks, limiting the internal degree distributions that can be derived from them.
+Differences in PPI databases may also lead to the presence of outliers that significantly alter coherence estimates.
+An example of such instability includes ESR2 and MOV10 genes in "Atrial Fibrilation" and "Blood Pressure", respectively, that have unusually high number of external edges in Biogrid- but not STRING-based networks.
+Consequently, the corresponding networks appear incoherent/highly coherent, when using Biogrid and STRING PPI databases, respectively.
 The differences in database-specific networks are alleviated by testing the significance of coherence using the permutation test; however, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].

### 2019-09-18 Analysis/Coherence_Calculations_New.Rmd
coherence tables

@@
 
+#Getting non-zero counts for diseases
+count_frame = full_join(dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(Disease, Count), dis_slopes_string_count %>% ungroup() %>% dplyr::select(Disease, Count), "Disease")
+
+count_frame = full_join(count_frame, dis_slopes_string_filt_count %>% ungroup() %>% dplyr::select(Disease, Count), "Disease") %>% arrange(Disease)
+```
+
+```{r}
+#Adding counts to full frame and reordering
+
+coherence_disease = coherence_disease %>% arrange(Disease) %>% mutate(`Size of Biogrid Network` = count_frame$Count.x, `Size of String Network` = count_frame$Count.y,`Size of String Filt Network`=count_frame$Count) %>% dplyr::select(Disease:`Number of genes`, `Size of Biogrid Network`:`Size of String Filt Network`, `Biogrid Coherence`:`String Filt P-value`)
+
 #Writing to a tables

### 2019-09-18 Analysis/Normalized_Coherence.Rmd
Make the min network size a setting

@@
 # Option for number of random networks to generate (Will generate num_rand-1 networks)
-num_rand = 11
+num_rand = 101
+# Minimum number of genes having non-zero internal edges for a network to be processed
+num_genes = 10
 # Construct file name to save the results
@@
     data.frame(Slope= coef(mod)[1])
-  }) %>% arrange(Slope) %>% filter(Count>4)
+  }) %>% arrange(Slope) %>% filter(Count > num_genes)
 }
@@
 external edges.
-Genes with zero internal edges are filtered and diseases with less than 5 non-zero internal edges are filtered.
+Genes with zero internal edges are filtered and diseases with less than `r num_genes` non-zero internal edges are filtered.
 

### 2019-09-18 Analysis/Analysis_Tables_New.Rmd
Major rewrite to use new data

@@
 max_genes <- 200 # Maximum number of genes for a network to be included
-
+min_genes <- 10  # Minimum number of genes for a network to be included
 fileNameIn0  <- "../manuscript/Tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
@@
 ```{r}
-mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% hist
-mtx0[, "String Filt Coherence KEGG"] %>% unlist %>% c %>% unname %>% summary
+mtx0[, "String Filt Coherence"] %>% unlist %>% c %>% unname %>% hist
+mtx0[, "String Filt Coherence"] %>% unlist %>% c %>% unname %>% summary
 ```
@@
 # Selected data for correlation
-# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value"
-mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG"))
-# Remove large networks
-mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
-# Go through each pairs of columns and calculate pairwise complete observations
-# pairwise_columns <- combn(ncol(mtx0_cor), 2)
-# for(j in 1:ncol(pairwise_columns)) {
-#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
-# }
+# Remove small/large networks
+mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min) > min_genes &
+                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max) < max_genes, ]
+# Remove size columns
+mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"))
 # Actual correlation
@@
 
-### FigS2_corr_references.png
-
-**Supplementary Figure S2.[FILE: Figures/FigS2_corr_references.png] Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients between corresponding entries.
-
-```{r}
-# Selected data for correlation
-# "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value"
-mtx0_cor <- mtx0 %>% dplyr::select(c("Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome", "Biogrid Coherence GOCC", "String Coherence GOCC", "String Filt Coherence GOCC"))
-# Remove large networks
-# mtx0_cor <- mtx0_cor[mtx0_cor$`Number of genes` < max_genes, ]
-# Go through each pairs of columns and calculate pairwise complete observations
-# pairwise_columns <- combn(ncol(mtx0_cor), 2)
-# for(j in 1:ncol(pairwise_columns)) {
-#   print(paste0(colnames(mtx0_cor)[pairwise_columns[1, j]], " - ", colnames(mtx0_cor)[pairwise_columns[2, j]], " : ", nrow(mtx0_cor[complete.cases(mtx0_cor[, c(pairwise_columns[1, j], pairwise_columns[2, j])]), c(pairwise_columns[1, j], pairwise_columns[2, j])])))
-# }
-# Actual correlation
-diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")
-
-paste0("Average Pearson: ", diff_cors[upper.tri(diff_cors)] %>% mean()) # Average correlation
-
-# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor)
-corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
-# Selected correlations
-png(filename = "../manuscript/Figures/FigS2_corr_references.png", height = 2000, width = 2000, res = 300)
-# corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
-corrplot.mixed(diff_cors, lower = "number", lower.col = col3(100), upper = "circle", upper.col = col3(100), tl.cex = 1, tl.pos = "lt", tl.col = "black")
-dev.off()
-```
-
 \pagebreak
@@
 ```{r}
-mtx_traits_selected_dotplot <- function(category = "Intelligence", mtx_traits = mtx_traits, databases = c("Biogrid", "STRING full", "STRING filtered"), reference = c("KEGG")) {
-  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(reference, colnames(mtx_traits))]
+mtx_traits_selected_dotplot <- function(category = "Intelligence", mtx_traits = mtx_traits, databases = c("Biogrid", "STRING full", "STRING filtered")) {
 
-  mtx_traits_selected_Biogrid <- data.frame(Disease = rownames(mtx_traits_selected), 
-                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
-                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("Biogrid", colnames(mtx_traits_selected))], 
-                                            PPI = paste0("Biogrid ", reference))
+  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(databases, colnames(mtx_traits))]
 
-  mtx_traits_selected_STRING_full <- data.frame(Disease = rownames(mtx_traits_selected), 
-                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))],
-                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String", colnames(mtx_traits_selected)) & !grepl("Filt", colnames(mtx_traits_selected))],
-                                            PPI = paste0("STRING full ", reference))
+  mtx_traits_selected_to_plot <- data.frame(Disease = rownames(mtx_traits_selected), 
+                                            Pvalue = mtx_traits_selected[, grepl(paste0(databases, " P-value"), colnames(mtx_traits_selected)) ], 
+                                            Coherence = mtx_traits_selected[, grepl(paste0(databases, " Coherence"), colnames(mtx_traits_selected)) ])
+  # Remove noninformative rows
+  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[!is.na(mtx_traits_selected_to_plot$Coherence), ]
+  # Sort by decreasing level of coherence
+  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = mtx_traits_selected_to_plot$Disease[order(mtx_traits_selected_to_plot$Coherence)])
 
-  mtx_traits_selected_STRING_filt <- data.frame(Disease = rownames(mtx_traits_selected), 
-                                            Pvalue = mtx_traits_selected[, grepl("P-value", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
-                                            Coherence = mtx_traits_selected[, grepl("Coherence", colnames(mtx_traits_selected)) & grepl("String Filt", colnames(mtx_traits_selected))], 
-                                            PPI = paste0("STRING filtered ", reference))
-  # Append plots from selected databases
-  mtx_traits_selected_to_plot <- rbind()
-  if( "Biogrid" %in% databases) {
-    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_Biogrid)
-  }
-  if( "STRING full" %in% databases) {
-    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_full)
-  }
-  if( "STRING filtered" %in% databases) {
-    mtx_traits_selected_to_plot <- rbind(mtx_traits_selected_to_plot, mtx_traits_selected_STRING_filt)
-  }
-
-  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[complete.cases(mtx_traits_selected_to_plot), ]
-
-  Coherence_levels <- mtx_traits_selected_to_plot[mtx_traits_selected_to_plot$PPI == paste0("STRING filtered ", reference), ]
-  if (length(setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) > 0) {
-    Coherence_levels <- rbind(Coherence_levels, mtx_traits_selected_to_plot[(mtx_traits_selected_to_plot$Disease %in% setdiff(mtx_traits_selected_to_plot$Disease, Coherence_levels$Disease)) & mtx_traits_selected_to_plot$PPI != paste0("STRING filtered ", reference), , drop = FALSE])
-  }
-
-  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = Coherence_levels$Disease[order(Coherence_levels$Coherence)])
-  PPI_levels <- unique(mtx_traits_selected_to_plot$PPI)
-  mtx_traits_selected_to_plot$PPI <- factor(mtx_traits_selected_to_plot$PPI, levels = c(PPI_levels[grepl("filtered", PPI_levels)], PPI_levels[grepl("full", PPI_levels)], PPI_levels[grepl("Biogrid", PPI_levels)]))
-  # mtx_traits_selected_to_plot$Pvalue <- -log10(mtx_traits_selected_to_plot$Pvalue)
-
-  p <- ggplot(data=mtx_traits_selected_to_plot, aes(x=Disease, y=Coherence, colour=PPI, size = Coherence)) + # label = Pvalue, 
+  p <- ggplot(data=mtx_traits_selected_to_plot, aes(x=Disease, y=Coherence, colour = "blue", size = Coherence)) + # label = Pvalue, 
     geom_point() +
-    facet_grid(. ~ PPI) + 
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
@@
 # STRING filtered plots only
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String Filt"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String Filt"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String Filt"))
+
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String Filt"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String Filt"))
 
@@
 
-### FigS3_summary_traits_KEGG.png
+### FigS2_summary_traits_Biogrid.png
 
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "KEGG"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "KEGG"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "KEGG"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "KEGG"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "KEGG"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "Biogrid"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "Biogrid"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "Biogrid"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "Biogrid"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "Biogrid"))
 
-png(filename = "../manuscript/Figures/FigS3_summary_traits_KEGG.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS2_summary_traits_Biogrid.png", height = 2500, width = 6000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 
-### FigS4_summary_traits_GOCC.png
+### FigS4_summary_traits_STRING.png
 
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "GOCC"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "GOCC"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "GOCC"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "GOCC"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "GOCC"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String"))
 
-png(filename = "../manuscript/Figures/FigS4_summary_traits_GOCC.png", height = 2500, width = 6000, res = 300)
-lay1 <- rbind(c(1, 2, 3))
-lay2 <- rbind(c(1, 2, 3))
-g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
-g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
-grid.arrange(g1, g2, ncol = 1)
-dev.off()
-```
-
-### FigS5_summary_traits_Reactome.png
-
-```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, reference = "Reactome"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, reference = "Reactome"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, reference = "Reactome"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, reference = "Reactome"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, reference = "Reactome"))
-
-png(filename = "../manuscript/Figures/FigS5_summary_traits_Reactome.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS4_summary_traits_STRING.png", height = 2500, width = 6000, res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 # STRING filtered plots only
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "STRING filtered", reference = "KEGG"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String Filt"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String Filt"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String Filt"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String Filt"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String Filt"))
+p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String Filt"))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String Filt"))
 
@@
 
-### FigS6_summary_diseases_KEGG.png
+### FigS4_summary_diseases_Biogrid.png
 
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "KEGG"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "KEGG"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "KEGG"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "KEGG"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "KEGG"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "KEGG"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "KEGG"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "Biogrid"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "Biogrid"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "Biogrid"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "Biogrid"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "Biogrid"))
+p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid"))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "Biogrid"))
 
-png(filename = "../manuscript/Figures/FigS6_summary_diseases_KEGG.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 3000, width = 8000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 
-### FigS7_summary_diseases_GOCC.png
+### FigS5_summary_diseases_STRING.png
 
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "GOCC"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "GOCC"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "GOCC"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "GOCC"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "GOCC"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "GOCC"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "GOCC"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String"))
+p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String"))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String"))
 
-png(filename = "../manuscript/Figures/FigS7_summary_diseases_GOCC.png", height = 3000, width = 8000, res = 300)
-lay1 <- rbind(c(1, 2, 3, 4))
-lay2 <- rbind(c(1, 2, 3, 4))
-g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
-g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
-grid.arrange(g1, g2, ncol = 1)
-dev.off()
-```
-
-### FigS8_summary_diseases_Reactome.png
-
-```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, reference = "Reactome"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, reference = "Reactome"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, reference = "Reactome"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, reference = "Reactome"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, reference = "Reactome"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, reference = "Reactome"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, reference = "Reactome"))
-
-png(filename = "../manuscript/Figures/FigS8_summary_diseases_Reactome.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS5_summary_diseases_STRING.png", height = 3000, width = 8000, res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 # "Disease", "Category", "Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Biogrid Coherence Reactome", "String Coherence Reactome", "String Filt Coherence Reactome""Biogrid Coherence CC", "String Coherence CC", "String Filt Coherence CC", "Biogrid P-value", "String P-value", "String Filt P-value"
-mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence KEGG", "String Coherence KEGG", "String Filt Coherence KEGG", "Modularity Biogrid", "Modularity String", "Modularity String Filt"))
+mtx0_cor <- mtx0 %>% dplyr::select(c("Number of SNPs", "Number of genes", "Biogrid Coherence", "String Coherence", "String Filt Coherence", "Modularity Biogrid", "Modularity String", "Modularity String Filt"))
 # Remove large networks

### 2019-09-19 Analysis/Analysis_Tables_New.Rmd
Publication quality figures

@@
 mtx0 <- read_csv(fileNameIn0) %>% as.data.frame()
+# Rename some diseases
+mtx0$Disease[mtx0$Disease == "Attention deficit hyperactivity disorder"] <- "ADHD"
+mtx0$Disease[mtx0$Disease == "Inflammatory bowel disease"] <- "IBD"
+mtx0$Disease[mtx0$Disease == "Systemic lupus erythematosus"] <- "SLE"
+mtx0$Disease[mtx0$Disease == "Age-related macular degeneration"] <- "Macular degeneration"
+mtx0$Disease[mtx0$Disease == "Lupus nephritis in systemic lupus erythematosus"] <- "Lupus nephritis in SLE"
+
 rownames(mtx0) <- mtx0$Disease
@@
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-    ylab("Normalized coherence") +
+    ylab("Coherence") +
     # geom_text_repel(size = 4, colour = "black") +
@@
 # STRING filtered plots only
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String Filt"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
 p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String Filt"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String Filt"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
 
 p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String Filt"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String Filt"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/Fig3_summary_traits.png", height = 2500, width = 5000, res = 300)
+png(filename = "../manuscript/Figures/Fig3_summary_traits.png", height = 8, width = 14, units = "in", res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "Biogrid"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
 p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "Biogrid"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "Biogrid"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "Biogrid"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
+# p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "Biogrid")) # Insufficient data
 p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "Biogrid"))
 
-png(filename = "../manuscript/Figures/FigS2_summary_traits_Biogrid.png", height = 2500, width = 6000, res = 300)
-lay1 <- rbind(c(1, 2, 3))
-lay2 <- rbind(c(1, 2, 3))
-g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular, p_metabolic), layout_matrix = lay1)
-g2 <- arrangeGrob(grobs = list(p_eye, p_intelligence), layout_matrix = lay2)
+png(filename = "../manuscript/Figures/FigS2_summary_traits_Biogrid.png", height = 6, width = 10, units = "in", res = 300)
+lay1 <- rbind(c(1, 2))
+lay2 <- rbind(c(1, 2))
+g1 <- arrangeGrob(grobs = list(p_anthropometric, p_cardiovascular), layout_matrix = lay1)
+g2 <- arrangeGrob(grobs = list(p_metabolic, p_intelligence), layout_matrix = lay2)
 grid.arrange(g1, g2, ncol = 1)
@@
 
-### FigS4_summary_traits_STRING.png
+### FigS3_summary_traits_STRING.png
 
 ```{r }
-p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String"))
+p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Intelligence", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Anthropometric trait", mtx_traits = mtx0, databases = "String"))
-p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String"))
+p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular trait", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye trait", mtx_traits = mtx0, databases = "String"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String"))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic trait", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/FigS4_summary_traits_STRING.png", height = 2500, width = 6000, res = 300)
+png(filename = "../manuscript/Figures/FigS3_summary_traits_STRING.png", height = 8, width = 14, units = "in", res = 300)
 lay1 <- rbind(c(1, 2, 3))
@@
 p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String Filt"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String Filt"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank())) 
 p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String Filt"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String Filt"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String Filt"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String Filt"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String Filt"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
+p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String Filt") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/Fig4_summary_diseases.png", height = 2500, width = 5000, res = 300)
+png(filename = "../manuscript/Figures/Fig4_summary_diseases.png", height = 8, width = 14, units = "in", res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "Biogrid"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "Biogrid"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
 p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "Biogrid"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "Biogrid"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "Biogrid"))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "Biogrid"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
+p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 3000, width = 8000, res = 300)
+png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 8, width = 14, units = "in", res = 300)
 lay1 <- rbind(c(1, 2, 3, 4))
@@
 p_intelligence <- plot(mtx_traits_selected_dotplot(category = "Autoimmune", mtx_traits = mtx0, databases = "String"))
-p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String"))
+p_anthropometric <- plot(mtx_traits_selected_dotplot(category = "Cancer", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 p_cardiovascular <- plot(mtx_traits_selected_dotplot(category = "Cardiovascular disease", mtx_traits = mtx0, databases = "String"))
-p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String"))
-p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String"))
+p_eye <- plot(mtx_traits_selected_dotplot(category = "Eye disease", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
+p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "String"))
-p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String"))
+p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "String") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/FigS5_summary_diseases_STRING.png", height = 3000, width = 8000, res = 300)
-lay1 <- rbind(c(1, 2, 3, 4))
-lay2 <- rbind(c(1, 2, 3, 4))
-g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
-g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)
-grid.arrange(g1, g2, ncol = 1)
+png(filename = "../manuscript/Figures/FigS5_summary_diseases_STRING.png", height = 12, width = 10, units = "in", res = 300)
+lay1 <- rbind(c(1, 2))
+lay2 <- rbind(c(1, 2))
+lay3 <- rbind(c(1, 2, 3))
+g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric), layout_matrix = lay1)
+g2 <- arrangeGrob(grobs = list(p_neurologic, p_psychiatric), layout_matrix = lay2)
+g3 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay3)
+
+grid.arrange(g1, g2, g3, ncol = 1)
 dev.off()

### 2019-09-19 manuscript/manuscript.Rmd
Search and replace trait to phenotype, mainly in Methods

@@
 
-Trait-associated gene networks were formed by mapping trait-associated genes to PPIs.
+Phenotype-associated gene networks were formed by mapping trait-associated genes to PPIs.
 Genes without PPI annotations (e.g., C5orf4, LOC100507462), antisense transcripts (e.g., CDKN2B-AS), readthrough (e.g., NPHP3-ACAD11), non-protein-coding transcripts (e.g., LINC00478) were excluded.
 Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
-Traits with less than five genes were omitted.
-Additionally, traits with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
+Phenotypes with less than five genes were omitted.
+Additionally, phenotypes with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
 external degree distributions (see below).
 These precautions were set to avoid situations where a small number of genes in a network can cause instability in the results.
-Importantly, the number of trait-associated genes having PPI annotations depends on the content of a PPI database; hence, some traits may be filtered when using one but not the other PPI database to build trait-associated networks.
+Importantly, the number of phenotype-associated genes having PPI annotations depends on the content of a PPI database; hence, some phenotypes may be filtered when using one but not the other PPI database to build phenotype-associated networks.
 
@@
 
-Trait-specific **internal degree distribution** was defined as the number of interactions between trait-associated genes, or, more precisely, their interacting protein products.
-As different traits have different numbers of associated genes forming networks of unknown configurations, the corresponding internal degree distribution was normalized to the expected interaction count for a network of that size (see Results).
+Phenotype-specific **internal degree distribution** was defined as the number of interactions between phenotype-associated genes, or, more precisely, their interacting protein products.
+As different phenotypes have different numbers of associated genes forming networks of unknown configurations, the corresponding internal degree distribution was normalized to the expected interaction count for a network of that size (see Results).
 
-Trait-specific **external degree distribution** was defined as the number of interactions between genes in a trait-associated network and all other genes in the PPI network.
+Phenotype-specific **external degree distribution** was defined as the number of interactions between genes in a phenotype-associated network and all other genes in the PPI network.
 Being measured in the context of the entire PPI network of fixed size, the external degree distribution did not require normalization.
 
-To compare trait-specific degree distributions, the normalized internal vs.
+To compare phenotype-specific degree distributions, the normalized internal vs.
 external degree distributions were plotted on a scatterplot.
 The square root of the degree was taken for more informative representation.
-For each trait, the internal vs.
+For each phenotype, the internal vs.
 external degree distribution plot was fit with a regression line through the origin to capture the slopes.
@@
 external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database.
-Thus, we normalized the slopes for the trait-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
-For each trait $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
+Thus, we normalized the slopes for the phenotype-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
+For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
 Higher values indicate a higher level of coherence similar to that of KEGG pathways.
-These values represent normalized coherence and can be compared across traits and PPI databases.
+These values represent normalized coherence and can be compared across phenotypes and PPI databases.
 
@@
 To test whether each slope is significantly different than random, we used a permutation test.
-For trait $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
-We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for trait $i$.
+For phenotype $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
+We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
 For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.

### 2019-09-20 manuscript/manuscript.Rmd
Describe new filtering in Methods
Current WIP

@@
 Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks of high coherence.
-This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathwaays (REACTOME genes) are presumed to interact more frequently.
-In our study, KEGG networks were used as the primary reference of high coherence, in parallel with GOCC and REACTOME collections.
+This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathways (REACTOME genes) are presumed to interact more frequently.
+In our study, KEGG networks were used in parallel with GOCC and REACTOME (referred hereafter as MSigDb networks) as a reference for high coherence.
 Conversely, as a reference for low coherence, we randomly sampled genes from a pool of genes in a given PPI database.
 These random gene lists represent expected network coherence arising by chance.
-The slopes of KEGG and random networks represent reference levels for high and low coherence, respectively.
+The slopes of MSigDb and random networks represent reference levels for high and low coherence, respectively.
 
@@
 Briefly, for each phenotype-specific network of a given size, we create size-matched references of high and low coherence.
-Specifically, we select the ten KEGG pathways closest in size and ten random networks matched in size as references for high/low coherence for the corresponding phenotype.
+Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections) and ten random networks matched in size for the corresponding network of phenotype-specific genes.
 Consequently, we use the slopes of internal vs.
@@
 Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
-Phenotypes with less than five genes were omitted.
+A non-zero interacting genes were defined as genes having at least one interaction with other phenotype-specific genes.
+Consequently, phenotypes with less than 10 non-zero interacting genes were omitted.
 Additionally, phenotypes with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.

### 2019-09-20 manuscript/manuscript.Rmd
Massage normalization methods

@@
 Briefly, for each phenotype-specific network of a given size, we create size-matched references of high and low coherence.
-Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections) and ten random networks matched in size for the corresponding network of phenotype-specific genes.
-Consequently, we use the slopes of internal vs.
+Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections) and 100 random networks matched in size for the corresponding network of phenotype-specific genes.
+Consequently, we use the median slopes of internal vs.
 external degree distributions created by these size-matched references to normalize the slope of the phenotype-specific network to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed removes the dependency between coherence and network size (Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
+We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 These precautions were set to avoid situations where a small number of genes in a network can cause instability in the results.
-Importantly, the number of phenotype-associated genes having PPI annotations depends on the content of a PPI database; hence, some phenotypes may be filtered when using one but not the other PPI database to build phenotype-associated networks.
 
@@
 Phenotype-specific **internal degree distribution** was defined as the number of interactions between phenotype-associated genes, or, more precisely, their interacting protein products.
-As different phenotypes have different numbers of associated genes forming networks of unknown configurations, the corresponding internal degree distribution was normalized to the expected interaction count for a network of that size (see Results).
-
 Phenotype-specific **external degree distribution** was defined as the number of interactions between genes in a phenotype-associated network and all other genes in the PPI network.
-Being measured in the context of the entire PPI network of fixed size, the external degree distribution did not require normalization.
-
-To compare phenotype-specific degree distributions, the normalized internal vs.
+The internal vs.
 external degree distributions were plotted on a scatterplot.
@@
 
-## Max-min normalization of coherence
+## Normalization of coherence
 
-The slopes of the internal vs.
-external degree distributions of the networks exhibiting high and random coherence (KEGG and random networks, respectively)  are dependent on the content of a PPI database.
-Thus, we normalized the slopes for the phenotype-associated networks using slopes formed by the degree distributions of similarly sized networks with high coherence ($\beta_{min}$ slope of KEGG networks) and random coherence ($\beta_{max}$ slope of random networks).
+To alleviate the dependency of degree distributions from the size of phenotype-associated gene sets and the content of a selected PPI database, we normalized slopes of phenotype-specific networks to the range of slopes formed by degree distributions of highly coherent and random networks.
+Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections, $\beta_{min}$ median slope) and 100 random networks ($\beta_{max}$ median slope) matched in size for the corresponding network of phenotype-specific genes.
 For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].

### 2019-09-20 Analysis/Category_Tables.Rmd
Fix new column names

@@
 
-msig_cor = msig %>% group_by(Category) %>% summarise(Slope_Cor = cor(Count, Slope, "complete")) %>% arrange(Slope_Cor)
+msig_cor = msig %>% group_by(Category) %>% summarise(Slope_Cor = cor(Num_Genes, Slope, "complete")) %>% arrange(Slope_Cor)
 
@@
 
-msig_sum = msig %>% group_by(Category) %>% summarise(Min_Slope = min(Slope, na.rm = TRUE), Max_Slope = max(Slope, na.rm = TRUE), Mean_Slope = mean(Slope, na.rm = TRUE), Median_Slope = median(Slope, na.rm = TRUE), Min_Genes = min(Count), Max_Genes = max(Count), Mean_Genes = mean(Count), Median_Genes = median(Count), Total_Sets = n()) %>% arrange(Mean_Slope)
+msig_sum = msig %>% group_by(Category) %>% summarise(Min_Slope = min(Slope, na.rm = TRUE), Max_Slope = max(Slope, na.rm = TRUE), Mean_Slope = mean(Slope, na.rm = TRUE), Median_Slope = median(Slope, na.rm = TRUE), Min_Genes = min(Num_Genes), Max_Genes = max(Num_Genes), Mean_Genes = mean(Num_Genes), Median_Genes = median(Num_Genes), Total_Sets = n()) %>% arrange(Mean_Slope)
 
@@
 
-kegg_string  = msig %>% filter(Category == "CP:KEGG") %>% dplyr::select(Count, Slope)
+kegg_string  = msig %>% filter(Category == "CP:KEGG") %>% dplyr::select(Num_Genes, Slope)
 
-Slope_Sum = bind_rows(rand_string %>% mutate(Method = "Random"), kegg_string %>% dplyr::select(Count, Slope) %>% mutate(Method = "KEGG"))
+Slope_Sum = bind_rows(rand_string %>% mutate(Method = "Random"), kegg_string %>% dplyr::rename(Count=Num_Genes) %>% dplyr::select(Count, Slope) %>% mutate(Method = "KEGG"))
 ```

### 2019-09-20 Analysis/Coherence_Calculations_New.Rmd
Add minimum num_genes

@@
 options(stringsAsFactors = FALSE)
+```
+
+```{r}
+# Minimum number of genes having non-zero internal edges for a network to be processed
+num_genes = 10
 ```
@@
       data.frame(Slope= coef(mod)[1])
-    }) %>% arrange(Slope) %>% filter(Count>4)
+    }) %>% arrange(Slope) %>% filter(Count > num_genes)
 

### 2019-09-21 Analysis/Analysis_Tables_New.Rmd
Correctly calculate average coherence

@@
 diff_cors <- cor(mtx0_cor, use = "pairwise.complete.obs")
+# Average correlation with network size
+print(paste0("Average correlation with network size: ", mean(diff_cors[grepl("Number", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))])))
+# Average correlation of coherences
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", rownames(diff_cors))]
+print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))]
+print(paste0("Average correlation of coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 

### 2019-09-21 manuscript/manuscript.Rmd
Describe Supplementary Figure S1

@@
 
-## Networks of KEGG pathways and randomly selected genes as reference for high and low coherence, respectively
+## Networks of KEGG/REACTOME pathways and GO Cellular Component collection vs. randomly selected genes serve as a reference for high vs. low coherence, respectively
 
@@
 To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
-We observed negative correlation of network size with the slopes of KEGG and random networks (-0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
+We observed negative correlation of network size with the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
 These observations confirm the notion that coherence estimation using internal vs.
@@
 Consequently, we use the median slopes of internal vs.
-external degree distributions created by these size-matched references to normalize the slope of the phenotype-specific network to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
+external degree distributions created by these size-matched references to normalize the slopes of the phenotype-specific networks to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
+We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (PCC = 0.29, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 
-The potential drawback of this reference size-matching approach is that phenotypes with large network sizes (e.g., "Obesity-related traits", 779 genes) may not have the size-matched reference of high coherence (e.g., the largest KEGG pathways is 374 genes).
-Consequently, coherence estimation for such networks may be affected by the smaller size of reference of high coherence.
-Therefore, we use networks from GOCC and REACTOME collections that have larger maximum number of genes (Supplementary Table S2) in parallel with KEGG networks as a reference of high coherence.
-We found that coherence estimations using KEGG, GOCC, and REACTOME networks as references of high coherence are highly concordant (Average Pearson correlation coefficient > 0.59, Supplementary Figure S2).
-The use of size-matched networks from KEGG, GOCC, and REACTOME collections enables robust estimation of coherence for phenotype-specific networks of different sizes.
-Given the similarity of the results obtained with different PPI databases and references of high coherence, we present results obtained with STRING filtered PPI database and KEGG networks as a reference of high coherence, with other observations presented in Supplementary Table S3 and Supplementary Figures S3-S8.
+Besides size dependency, coherence estimates depend on the choice of PPI database used to build networks and estimate the corresponding internal vs.
+external degree distributions.
+To evaluate consistency of coherence estimates, we evaluated pairwise correlation of normalized coherences obtained using different PPI databases.
+Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.77).
+Expectedly, coherence estimates using Biogrid database were less similar to that of "STRING" and "STRING filtered"; however, the overall correlation remained high (PCC = 0.56, Supplementary Figure S1).
+Notably, coherence estimates using "STRING filtered" database were most similar to that of "Biogrid" (PCC = 0.57), indicating that filtering step indeed removes "noisy" PPIs and improves quality of coherence estimates.
+Overall, these results demonstrate consistency in coherence estimates using different PPI databases.
 
@@
 For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
-Higher values indicate a higher level of coherence similar to that of KEGG pathways.
+Higher values indicate a higher level of coherence similar to that of MSigDb networks.
 These values represent normalized coherence and can be compared across phenotypes and PPI databases.
+Normalized coherence estimates for excellively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
 
@@
 **Supplementary Figure S1.
-Normalized coherence is unaffected by network size.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases and KEGG pathways as a reference of high coherence.
+Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.
 

### 2019-09-21 Analysis/Analysis_Tables_New.Rmd
Fix plots to use latest estimates

@@
 p_metabolic <- plot(mtx_traits_selected_dotplot(category = "Metabolic disease", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
-p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
+# p_neurologic <- plot(mtx_traits_selected_dotplot(category = "Neurologic", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
 p_psychiatric <- plot(mtx_traits_selected_dotplot(category = "Psychiatric", mtx_traits = mtx0, databases = "Biogrid") + theme(axis.title.y = element_blank()))
 
-png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 8, width = 14, units = "in", res = 300)
-lay1 <- rbind(c(1, 2, 3, 4))
-lay2 <- rbind(c(1, 2, 3, 4))
-g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_neurologic, p_psychiatric), layout_matrix = lay1)
+png(filename = "../manuscript/Figures/FigS4_summary_diseases_Biogrid.png", height = 8, width = 12, units = "in", res = 300)
+lay1 <- rbind(c(1, 2, 3))
+lay2 <- rbind(c(1, 2, 3))
+g1 <- arrangeGrob(grobs = list(p_intelligence, p_anthropometric, p_psychiatric), layout_matrix = lay1)
 g2 <- arrangeGrob(grobs = list(p_cardiovascular, p_metabolic, p_eye), layout_matrix = lay2)

### 2019-09-21 manuscript/manuscript.Rmd
Fix supplementary figure legends

@@
 
-To estimate the significance of coherence measures, we developed a permutation-based approach (see Methods).
-Notably, we observed coherence for all diseases to be highly significantly different from random (permutation p-value < 0.002, Supplementary Table S3).
-Similarly, the distribution of coherence estimates was skewed towards 1.00 (highest coherence), with the median coherence being 0.86.
-Among traits, "Corneal structure", "Body mass index", "Weight" were the most coherent (1.02, 1.00, 0.99, respectively).
-[???]
-
-"Obesity-related traits" and "Height" were on the lowest range of coherence estimates (0.42 and 0.54, respectively).
-"Weight", "Telomere length", and "Corneal structure" were on the highest range of coherence (1.00, Supplementary Table S3).
-These results indicate that coherence of phenotype-specific networks is generally high and significantly different from random.
-
-## Metabolic and intelligence-related traits as examples of high/low coherence
+## Metabolic and intelligence-related traits as examples of networks with overall highest/lowest coherence
 
@@
 Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
-Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose- and cholesterol-related traits (metabolic).
-Notably, metabolic and eye trait categories had the highest overall coherence (mean coherence 0.90 and 0.93, respectively).
-Several notable examples of low coherence include "Obesity-related traits", "Height", and "Longevity" (anthropometric traits), blood count-related traits (cardiovascular).
+Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic).
+Notably, metabolic trait category had the highest overall coherence (mean coherence 0.91 and 1.05, respectively, Supplementary Table S1).
+Several notable examples of low coherence include "Obesity-related traits" and "Height" (anthropometric traits), and blood count-related traits (cardiovascular).
 Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.76).
@@
 
-## Metabolic and neurologic diseases have high coherence
+## Metabolic diseases and cancer networks have high coherence
 
@@
 **Supplementary Figure S2.
-Coherence estimates are similar when using different references of high coherences and PPI databases.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the normalized coherence estimates using the corresponding PPI databases and KEGG/Reactome pathways and Gene Ontology Cellular Component networks as a reference of high coherence.
+Coherence estimates of traits using Biogrid as a reference PPI database.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+Plots are faceted by categories.
+Missing entries indicate that, for a given trait, a network could not be built and the coherence cannot be estimated.
 
-**Supplementary Figure S3.
-Coherence estimates of traits using KEGG as a reference of high coherence.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
-Plots are faceted by PPI database used to estimate coherence.
-Missing entries indicate that, for a given PPI database, a phenotype-associated network could not be built.
+**Supplementary Figure S3. Coherence estimates of traits using STRING as a reference of PPI database.** See legend for Supplementary Figure S2.
 
-**Supplementary Figure S4. Coherence estimates of traits using GOCC as a reference of high coherence.** See legend for Supplementary Figure S3.
+**Supplementary Figure S4. Coherence estimates of diseases using Biogrid as a reference PPI database.** See legend for Supplementary Figure S2.
 
-**Supplementary Figure S5. Coherence estimates of traits using Reactome as a reference of high coherence.** See legend for Supplementary Figure S3.
-
-**Supplementary Figure S6. Coherence estimates of diseases using KEGG as a reference of high coherence.** See legend for Supplementary Figure S3.
-
-**Supplementary Figure S7. Coherence estimates of diseases using GOCC as a reference of high coherence.** See legend for Supplementary Figure S3.
-
-**Supplementary Figure S8. Coherence estimates of diseases using Reactome as a reference of high coherence.** See legend for Supplementary Figure S3.
+**Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.
 

### 2019-09-21 manuscript/manuscript.Rmd
Describe latest results sans p-values

@@
 
-We estimated coherence of 54 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S3-S5).
+We estimated coherence of 54 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
 Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
@@
 
-We further estimated coherence of 80 diseases from seven categories (Figure 4, Supplementary Figures S6-S8, Supplementary Table S3).
-Metabolic diseases had high overall coherence (average coherence 0.93), with "Obesity" and "Type 2 diabetes" being the most coherent (0.99 and 0.94, respectively).
-Similarly, the average coherence for neurologic diseases was 0.86, with "Alzheimer's disease" and "Parkinson's disease" being highly coherent (0.94 and 0.92, respectively).
-The high coherence of metabolic and neurologic disease networks highlight the relatively good knowledge of their genetics.
+We further estimated coherence of 80 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3).
+Metabolic diseases had high overall coherence (average coherence 0.91), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
+Similarly, the average coherence for cancer diseases was 0.72, with "Testicular germ cell tumor" and "Pancreatic cancer" being highly coherent (0.92 and 0.74, respectively).
+The high coherence of metabolic and cancer disease networks highlight the relatively good knowledge about their genetics.
 
-## Intermediate coherence of autoimmune, cancer, and eye diseases
+## Intermediate coherence of autoimmune, neurologic, and eye diseases
 
-"Rheumatoid arthritis", "Ankylosing spongylitis", "Psoriasis" were the most coherent in the "Autoimmune" category (0.94, 0.91, 0.89, respectively).
-On the other end of the coherence spectrum, "Multiple sclerosis" and "Allergic rinitis" were the least coherent (0.76 for both).
-The average coherence of autoimmune diseases was high, 0.85, with all but two autoimmune diseases having high coherence (> 0.80).
-Diseases in the "Cancer" category were similarly highly coherent (average coherence 0.84).
-"Non-small lung cancer", "Testicular germ cell tumor", "Epithelial ovarian cancer" were the most coherent (0.94, 0.91, 0.88, respectively) while "Basal cell carcinoma", "Breast cancer" "Hodgkin's lymphoma" were the least coherent (0.64, 0.78, 0.78, respectively).
-Overall, networks of all but "Basal cell carcinoma" cancers had high coherence (> 0.78).
-Diseases in the "Eye disease" category were more heterogeneous, with coherence ranging from 0.71 for "Myopia (pathological)" disease to 0.97 for "Refractive error".
-The average coherence of diseases in the "Eye disease" category was relatively high (0.86).
-These results suggest good overall knowledge of the genetics of autoimmune, cancer, and eye-related diseases and highlight limited knowledge for some diseases in these categories.
+"Rheumatoid arthritis", "Systemic Lupus Erythematosus", and "Systemic sclerosis" were the most coherent in the "Autoimmune" category (0.86, 0.80, 0.79, respectively).
+On the other end of the coherence spectrum, "Allergic rinitis" and "Atopic dermatitis" were the least coherent (0.51, 0.60, respectively).
+The average coherence of autoimmune diseases was high, 0.70 (Table 1), with all autoimmune diseases, except "Allergic rinitis", having coherence estimates > 0.60 (Supplementary Table S3).
+Diseases in the "Neurologic" category were similarly highly coherent (average coherence 0.72).
+However, the average coherence of neurologic diseases using STRING database was 0.49, suggesting that, with larger number of networks being estimated the overall coherence of neurologic diseases is low.
+"Alzheimer's disease" and "Parkinson's disease" were the most coherent (0.88 and 0.84, respectively), while "Migraine" was the least coherent (0.54).
+Diseases in the "Eye disease" category were more heterogeneous, with coherence ranging from 0.35 for "Corneal astigmatism" to 1.01 for "Refractive error".
+The average coherence of diseases in the "Eye disease" category was relatively high (0.66).
+These results suggest good overall knowledge of the genetics of autoimmune, neurologic, and eye-related diseases and highlight limited knowledge for some diseases in these categories.
 
@@
 
-Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.82 and 0.83, respectively).
+Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.64 and 0.63, respectively).
 The corresponding coherence estimates were highly heterogeneous.
-Among cardiovascular diseases, "Atrial fibrillation", "Coronary artery disease" showed the highest level of coherence (0.98 and 0.89, respectively), while "Sudden cardiac arrest" and "Hypertension" had the lowest coherence (0.70 and 0.77, respectively).
-Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.93 and 0.89, respectively).
-Notably, "Schizophrenia" and "Major depressive disorder" showed the lowest coherence not only in the "Psychiatric disease" category (0.67 and 0.75, respectively) but among all phenotypes.
-Other psychiatric disorders, such as "Bipolar disorder", "Autism", "ADHD", showed intermediate coherence estimates (0.87, 0.83, 0.78, respectively).
-In summary, these low coherence estimates highlight the need for further research on genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms, and suggests "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies.
+Among cardiovascular diseases, "Coronary artery disease" showed the highest level of coherence (0.83), while "Hypertension" had the lowest coherence (0.53).
+Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.87 and 0.82, respectively).
+Notably, choerehce of "Schizophrenia" and "Major depressive disorder" networks (0.52 and 0.33, respectively) was at the lower end of coherence estimates among all phenotypes.
+Other psychiatric disorders, such as "Bipolar disorder", "ADHD", showed intermediate coherence estimates (0.76 and 0.55, respectively).
+In summary, these low coherence estimates highlight the need for further research on genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms.
+These results further suggest "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies in psychiatric genome-wide association studies.
 
@@
 **Supplementary Table S1.
-Summary statistics of the analyzed phenotypes.** Minimum, mean, median, maximum of coherence estimates, the number of genes, SNPs, the total number of phenotype networks.
+Summary statistics of the analyzed phenotype categories.** Minimum, mean, median, maximum of coherence estimates, the number of genes, SNPs, the total number of phenotype networks.
 Each worksheet corresponds to the PPI database used.

### 2019-09-21 manuscript/manuscript.Rmd
Polish discussion, methods

@@
 
-**Methods.** We introduce a measure of network coherence as a ratio of internal vs.
+**Methods.** We introduce a measure of network coherence as a slope of internal vs.
 external distributions of the degree of interactions.
@@
 We developed a method for normalizing the internal network connectivity that accounts for differences in network sizes formed by disease-associated genes.
-We compared the ratio of normalized internal vs.
+We compared the slope of normalized internal vs.
 external network connectivity between psychiatric, neurologic, autoimmune, cancer, cardiovascular, metabolic and other diseases and traits.
@@
 Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.-->
-This study presents a novel measure of network coherence as a ratio of the internal vs.
+This study presents a novel measure of network coherence as a slope of the internal vs.
 external degree distributions.
@@
 -->
-Given that phenotype-associated networks occur within the global network of PPIs, we developed a ratio-based metric of coherence.
-This metric is inspired by previous work showing that communities within a large network have high internal but low external levels of connectivity [@PhysRevE.81.046110; @Spirin:2003aa].
+Given that phenotype-associated networks occur within the global network of PPIs, we developed a metric of coherence based on the level of gene interactions.
+This metric is inspired by previous work showing that communities within a large network have high internal but low external levels of interactions (degrees) [@PhysRevE.81.046110; @Spirin:2003aa].
 Thus, for each phenotype-associated PPI network, we considered the relationship between its internal and external degree distributions.
@@
 We developed a general-purpose method to measure and compare coherence estimates of the molecular mechanisms in networks of genes.
-Intuitively, gene networks with high internal and low external interactions would be considered highly coherent; consequently, coherence is defined as the ratio between the internal and external degree distributions (the number of interactions per gene).
-We normalized coherence estimates to the range of high and random coherence, exemplified by KEGG pathways and random networks.
-To decouple coherence from its dependence on network size, we used KEGG and random networks matched in size to phenotype-specific networks.
-Applied to the analysis of 133 disease- and trait-associated gene sets, we found an unexpected inverse association of phenotype complexity with coherence.
-Consequently, complex traits like "Height", "Intelligence", "Longevity" had overall low coherence while metabolic traits and diseases had high coherence.
-Diseases in "Psychiatric" category had overall low coherence, with "Schizophrenia" being least coherent.
-Our method allows one to quantify the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.
+Intuitively, gene networks with high internal and low external interactions (degrees) would be considered highly coherent; consequently, coherence is defined as the slope through the internal and external degree distribution scatterplot.
+We normalized coherences (slopes) to the range of high and random coherence, exemplified by MSigDb networks and random networks.
+To decouple coherence from its dependence on network size, we used MSigDb and random networks matched in size to phenotype-specific networks.
+Applied to the analysis of 133 disease- and trait-associated gene sets, we found metabolic diseases and traits have overall high coherence, while psychiatric and cardiovascular diseases had generally low coherence.
+These results were consistent when using different PPI databases to build gene networks.
+Our method allows one to quantify and compare the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.
 
@@
 
-The definition of coherence as the ratio between internal vs.
+The definition of coherence as the slope between internal vs.
 external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa].
@@
 Furthermore, networks formed by phenotype-associated genes should be sufficiently diverse to avoid cases in which all genes in a network have the same degree.
-As the different PPI databases vary significantly in the number of PPI annotations, some phenotype-associated gene sets have sufficient PPI annotations when using one but not the other database.
+As the different PPI databases vary significantly in the number of PPI annotations, some phenotype-associated gene sets have sufficient PPI annotations only when using one but not the other database.
 This problem is best illustrated by the BioGrid PPI database that has considerably fewer PPIs; hence, fewer phenotypes could be analyzed.
-Differences in PPI databases may also lead to the presence of outliers that significantly alter coherence estimates.
+<!--Differences in PPI databases may also lead to the presence of outliers that significantly alter coherence estimates.
 An example of such instability includes ESR2 and MOV10 genes in "Atrial Fibrilation" and "Blood Pressure", respectively, that have unusually high number of external edges in Biogrid- but not STRING-based networks.
-Consequently, the corresponding networks appear incoherent/highly coherent, when using Biogrid and STRING PPI databases, respectively.
-The differences in database-specific networks are alleviated by testing the significance of coherence using the permutation test; however, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].
+Consequently, the corresponding networks appear incoherent/highly coherent, when using Biogrid and STRING PPI databases, respectively.--> Although our results show good correspondence between coherence estimates when using different PPI databases, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].
 
@@
 In addition to internal vs.
-external degree distributions, other network properties can be used to estimate coherence [@Leskovec:2010:ECA:1772690.1772755].
+external degree distributions, other gene-centric network metrics can be used to estimate coherence [@Leskovec:2010:ECA:1772690.1772755].
 For example, the centrality measure has been used to demonstrate that highly central hub genes are functionally essential and highly conserved [@Khurana:2013aa].
@@
 In summary, centrality measure parallels our observations of network coherence.
-Our future work will incorporate centrality and other network metrics into our framework to further refine our ability to compare coherence of phenotype-associated molecular mechanisms.
-
-[??? Potentially. discuss: We do not account for configuration]
+Our future work will incorporate centrality and other gene-centric network metrics into our framework to further refine our ability to compare coherence of phenotype-associated molecular mechanisms.
 
@@
 
-Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
-We developed a network coherence measure that estimates the degree of internal vs.
-external connectivity that is robust to the difference in network sizes.
+Our study investigated coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
+We developed a network coherence measure driven by the relationship between internal and external connectivity (degree distribution) that is robust to the difference in network sizes.
 Using well-curated lists of genomic variants and the associated genes, we built disease-specific networks using protein-protein interaction information from the STRING and BioGrid databases.
-We found a range of coherence estimates in each phenotype category, with metabolic diseases and traits being the most coherent.
-Psychiatric disorders were, on the other hand, the least coherent, with schizophrenia and major depressive disorder being the least coherent among all phenotypes.
+We found a range of coherence estimates in each phenotype category, with metabolic phenotypes being the most coherent.
+Psychiatric disorders had low coherence, with schizophrenia and major depressive disorder being the least coherent among all phenotypes.
 In summary, we provide a general-purpose method for quantifying and comparing coherence of molecular networks.
@@
 These values represent normalized coherence and can be compared across phenotypes and PPI databases.
-Normalized coherence estimates for excellively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
+
+Although using median slopes of MSigDb and random networks is expected to be a robust reference for high/random coherence, some phenotype-specific networks may have coherence larger than 1, that is, being more coherent than the median coherence of MSigDb networks.
+These situations may occur in smaller networks that have higher change to contain genes with unusually high internal degree distributions.
+Such outliers would inflate coherence estimations.
+Similar outliers may be observed in larger networks.
+For these reasons, coherence estimates for excellively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
 

### 2019-09-23 Analysis/Coherence_Calculations_New.Rmd
adapt to new style

@@
 #Biogrid
-biogrid_p = bind_rows(readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/biogrid_p_psych_1k.rds") %>% as.data.frame(.))
+biogrid_p = readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.) %>% dplyr::select(-new_coherence)
 
 #String
-string_p = bind_rows(readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_p_neuro_1k.rds") %>% as.data.frame(.))
+string_p = readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
 
 #String Filtered
-string_filt_p = bind_rows(readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_psych_1k.rds") %>% as.data.frame(.), readRDS("./data/Permutation_Results/string_filt_p_neuro_1k.rds") %>% as.data.frame(.))
+string_filt_p = readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
 

### 2019-09-24 Analysis/Coherence_Calculations_New.Rmd
new p-vals

@@
 
+#Filter out extra p-values
+
+coherence_disease = coherence_disease %>% mutate(`Biogrid P-value` = ifelse(is.na(`Biogrid Coherence`), NA, `Biogrid P-value` ))  %>% mutate(`String P-value` = ifelse(is.na(`String Coherence`), NA, `String P-value` ))  %>% mutate(`String Filt P-value` = ifelse(is.na(`String Filt Coherence`), NA, `String Filt P-value` ))
+
 #Writing to a tables

### 2019-09-24 Analysis/Analysis_Tables_New.Rmd
Add correlation for p-values

@@
 # Remove small/large networks
-mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min) > min_genes &
-                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max) < max_genes, ]
+mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min, na.rm = TRUE) > min_genes &
+                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max, na.rm = TRUE) < max_genes, ]
 # Remove size columns
-mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"))
+mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"), contains("P-value"))
+# Jitter STRING p-value a bit
+set.seed(1)
+mtx0_cor$`String P-value` <- jitter(mtx0_cor$`String P-value`, amount = 0.0001)
 # Actual correlation
@@
 # Average correlation of coherences
-diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", rownames(diff_cors))]
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", colnames(diff_cors))]
 print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
@@
 print(paste0("Average correlation of coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("P-value", colnames(diff_cors))]
+print(paste0("Average correlation of coherences and p-value: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 

### 2019-09-24 manuscript/manuscript.Rmd
Describe p-values

@@
 
-## Coherence of all phenotypes significantly differs from random
-
 ## Metabolic and intelligence-related traits as examples of networks with overall highest/lowest coherence
@@
 These results further suggest "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies in psychiatric genome-wide association studies.
+
+## Coherence of most phenotypes significantly differs from random
+
+Given the wide range of phenotype-specific coherence estimates, a natural question is whether they significantly different from random coherence.
+We assesses the significance of coherence estimates using permutation test (see Methods), individually for each PPI database.
+Permutation p-values obtained using Biogrid and STRING filtered databases were higly correlated (PCC = 0.71, Supplementary Figure S1), as would be expected for high-quality PPIs.
+Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to the large number of PPIs that enable phenotype-specific networks to be consistently different from random.
+Permutation p-values inversely correlated with coherence estimates (average PCC = -0.23), as would be expected for less coherent networks that resemble random networks.
+This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.73 (STRING filtered PCC = -0.51, Supplementary Figure S1).
+These observations confirmed consistent coherence and p-value estimates when using different PPI databases.
+
+The majority of phenotype-specific networks (~85%) were significantly coherent (STRING filtered p-value < 0.001, Supplementary Table S3).
+Among less significant phenotypes were "White blood cell count" (p-value = 0.102) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.05), "Major depressive disorder" (p-value = 0.03).
+When using Biogrid databases, "Educational attainment", "Myopia (pathological)", "Celiac disease" were among the non-significant phenotypes (p-values 0.29, 0.20,0.16, respectively).
+Interestingly, "Schizophrenia" network, despite being among the least coherent, was consistently significantly different from random (p-value < 0.001).
+Overall, significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence.
+These results further confirm our ability to identify phenotype-specific networks with low, non-significant coherence.
 

### 2019-09-24 manuscript/manuscript.Rmd
make more up to date

@@
 For phenotype $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
-We then generate $n=10000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
+We then generate $n=1000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
+Each random network is created by repeatedly generating networks of size $g_i$ and keeping non-zero genes.
+Once the number of non-zero genes is equal to $g_i$ we combine them into a single random network.
 For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.

### 2019-09-29 manuscript/manuscript.Rmd
minor discussion polish

--- 2019-09-24 manuscript/manuscript.Rmd
@@
 Typically, genes are linked to genetic variants by proximity (nearest gene), a strategy adopted by the NHGRI-EBI GWAS catalog [@MacArthur:2017aa].
-<!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, the majority of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Khurana:2013aa].
+<!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, >88% of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Hindorff:2009aa].
 <!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].-->The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy.

### 2019-09-29 manuscript/manuscript.Rmd
Grammarly edits

@@
 ---
-bibliography: [/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/networks.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/brain.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/enrichment.bib,/Users/mdozmorov/Documents/Work/VCU_grants/1_Dozmorov/3D_refs.bib]
+bibliography: [/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/networks.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/brain.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/enrichment.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/3D_refs.bib]
 csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
@@
 We observed a range of coherence estimates for each category of phenotypes.
-Metabolic traits and diseases were the most coherent, while psychiatric diseases and intelligence-related traits were the least coherent.
+Metabolic traits and diseases were the most coherent, while psychiatric disorders and intelligence-related traits were the least coherent.
 
 **Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences.
-Our results highlight the gaps in of our current knowledge of the genetics and molecular mechanisms of complext phenotypes and suggest priorities for future work.
+Our results highlight the gaps in of our current knowledge of genetics and molecular mechanisms of complex phenotypes and suggest priorities for future work.
 
@@
 external network connectivity between psychiatric, neurologic, autoimmune, cancer, cardiovascular, metabolic and other diseases and traits.
-We used KEGG pathways [30] as a reference for a coherent network with common molecular mechanisms, and we used sets of randomly sampled genes as a reference for the absense of coherence.
+We used KEGG pathways [30] as a reference for a coherent network with common molecular mechanisms, and we used sets of randomly sampled genes as a reference for the absence of coherence.
 We found that genes associated with metabolic diseases and traits and autoimmune diseases formed coherent networks similar to KEGG pathways.
@@
 We used KEGG pathways [@Kanehisa:2000aa] as a reference for a paradigmatically high-coherent network [@Lee:2003aa], while sets of randomly sampled genes were used as the reference for the absence of coherence.
-Using KEGG and random networks matched in size to a phenotype-specific network, we derived normalized measure of coherence that can be compared across phenotypes.
+Using KEGG and random networks matched in size to a phenotype-specific network, we derived a normalized measure of coherence that can be compared across phenotypes.
 We hypothesized that the level of coherence may inform us about similarities and differences among phenotypes.
@@
 Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have high coherence.
-Our method enables direct comparison of coherence measures and highlight gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.
+Our method enables direct comparison of coherence measures and highlights gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.
 
@@
 The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs are removed from the analysis (see Methods).
-Second, complete STRING data were used to maximize the use of PPI information at the expense of potential noise (referred hereafter as “STRING”).
+Second, complete STRING data were used to maximize the use of PPI information at the expense of potential noise (referred to hereafter as “STRING”).
 The advantage of using complete data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
@@
 Summary statistics of the analyzed categories.** Average coherence estimates are shown.
-"NA" values indicate that a category lacked phenotypes with sufficient number of PPIs.
+"NA" values indicate that a category lacked phenotypes with a sufficient number of PPIs.
 Sorted by "String Filtered" average coherence.
@@
 A degree of a node, or connectivity, is one of the most fundamental characteristics defining the number of other nodes connected to a given node.
-A collection of degrees of all network nodes form a degree distribution.
+A collection of degrees of all network nodes forms a degree distribution.
 Comparing degree distributions among networks is an intuitive way to gain an understanding of network similarities and differences in terms of connectivity (aka coherence) of the corresponding genes [@Barabasi:2004aa; @Feldman:2008aa].
@@
 The degree distributions of the ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).
-Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have degree "9", while all nodes in the 20-node network have degree "19", Figure 1B).
+Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have a degree "9", while all nodes in the 20-node network have degree "19", Figure 1B).
 <!--Normalizing the degree distributions of the fully connected networks by dividing them by network size makes them identical as would be expected for networks with similar coherence.
@@
 
-Besides size dependency, coherence estimates depend on the choice of PPI database used to build networks and estimate the corresponding internal vs.
+Besides size dependency, coherence estimates depend on the choice of the PPI database used to build networks and estimate the corresponding internal vs.
 external degree distributions.
-To evaluate consistency of coherence estimates, we evaluated pairwise correlation of normalized coherences obtained using different PPI databases.
+To evaluate the consistency of coherence estimates, we assessed the pairwise correlation of normalized coherences obtained using different PPI databases.
 Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.77).
@@
 These results highlight the relatively good knowledge of the genetics of metabolic traits which corresponds to highly coherent networks.
-Conversely, current knowledge about the genetics of intelligence-related traits appear insufficient as reflected in the low coherence of the corresponding networks.
+Conversely, current knowledge about the genetics of intelligence-related traits appears insufficient as reflected in the low coherence of the corresponding networks.
 
@@
 Diseases in the "Neurologic" category were similarly highly coherent (average coherence 0.72).
-However, the average coherence of neurologic diseases using STRING database was 0.49, suggesting that, with larger number of networks being estimated the overall coherence of neurologic diseases is low.
+However, the average coherence of neurologic disorders using STRING database was 0.49, suggesting that, with larger number of networks being estimated the overall coherence of neurologic diseases is low.
 "Alzheimer's disease" and "Parkinson's disease" were the most coherent (0.88 and 0.84, respectively), while "Migraine" was the least coherent (0.54).
@@
 Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.87 and 0.82, respectively).
-Notably, choerehce of "Schizophrenia" and "Major depressive disorder" networks (0.52 and 0.33, respectively) was at the lower end of coherence estimates among all phenotypes.
+Notably, coherence of "Schizophrenia" and "Major depressive disorder" networks (0.52 and 0.33, respectively) was at the lower end of coherence estimates among all phenotypes.
 Other psychiatric disorders, such as "Bipolar disorder", "ADHD", showed intermediate coherence estimates (0.76 and 0.55, respectively).
-In summary, these low coherence estimates highlight the need for further research on genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms.
+In summary, these low coherence estimates highlight the need for further research on the genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms.
 These results further suggest "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies in psychiatric genome-wide association studies.
@@
 
-Given the wide range of phenotype-specific coherence estimates, a natural question is whether they significantly different from random coherence.
-We assesses the significance of coherence estimates using permutation test (see Methods), individually for each PPI database.
-Permutation p-values obtained using Biogrid and STRING filtered databases were higly correlated (PCC = 0.71, Supplementary Figure S1), as would be expected for high-quality PPIs.
-Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to the large number of PPIs that enable phenotype-specific networks to be consistently different from random.
+Given the wide range of phenotype-specific coherence estimates, a natural question is whether they are significantly different from random coherence.
+We assess the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
+Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.71, Supplementary Figure S1), as would be expected for high-quality PPIs.
+Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to a large number of PPIs that enable phenotype-specific networks to be consistently different from random.
 Permutation p-values inversely correlated with coherence estimates (average PCC = -0.23), as would be expected for less coherent networks that resemble random networks.
@@
 Interestingly, "Schizophrenia" network, despite being among the least coherent, was consistently significantly different from random (p-value < 0.001).
-Overall, significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence.
+Overall, the significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence.
 These results further confirm our ability to identify phenotype-specific networks with low, non-significant coherence.
@@
 To decouple coherence from its dependence on network size, we used MSigDb and random networks matched in size to phenotype-specific networks.
-Applied to the analysis of 133 disease- and trait-associated gene sets, we found metabolic diseases and traits have overall high coherence, while psychiatric and cardiovascular diseases had generally low coherence.
+Applied to the analysis of 133 disease- and trait-associated gene sets, we found metabolic diseases and traits have overall high coherence, while psychiatric and cardiovascular diseases generally had low coherence.
 These results were consistent when using different PPI databases to build gene networks.
@@
 <!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, >88% of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Hindorff:2009aa].
-<!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].-->The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy.
+<!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].--> The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy.
 The development of chromatin conformation capture technologies now allows for understanding interactions among genomic regions on a genome-wide scale [@Lieberman-Aiden:2009aa; @Zhang:2013aa], helping to prioritize disease-associated gene-variant associations [@Pan:2018aa].
@@
 <!--Differences in PPI databases may also lead to the presence of outliers that significantly alter coherence estimates.
-An example of such instability includes ESR2 and MOV10 genes in "Atrial Fibrilation" and "Blood Pressure", respectively, that have unusually high number of external edges in Biogrid- but not STRING-based networks.
+An example of such instability includes ESR2 and MOV10 genes in "Atrial Fibrillation" and "Blood Pressure", respectively, that have an unusually high number of external edges in Biogrid- but not STRING-based networks.
 Consequently, the corresponding networks appear incoherent/highly coherent, when using Biogrid and STRING PPI databases, respectively.--> Although our results show good correspondence between coherence estimates when using different PPI databases, care should be exercised when selecting the PPI database for network analyses [@Bajpai:2019aa].
@@
 A large study of PPIs (BioPlex) compared centrality distributions among several disease categories.
-Cancers and immunological disease networks tended to have high centrality, while nervous system, congenital, neonatal and hereditary disease networks had low centrality [@Huttlin:2017aa].
+Cancers and immunological disease networks tended to have high centrality, while nervous system, congenital, neonatal, and hereditary disease networks had low centrality [@Huttlin:2017aa].
 Metabolic pathways were shown to have more duplicated copies of highly central genes making the networks tolerant to loss-of-function mutations [@Khurana:2013aa].
 In summary, centrality measure parallels our observations of network coherence.
-Our future work will incorporate centrality and other gene-centric network metrics into our framework to further refine our ability to compare coherence of phenotype-associated molecular mechanisms.
+Our future work will incorporate centrality and other gene-centric network metrics into our framework to further refine our ability to compare the coherence of phenotype-associated molecular mechanisms.
 
@@
 
-Our study investigated coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
+Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
 We developed a network coherence measure driven by the relationship between internal and external connectivity (degree distribution) that is robust to the difference in network sizes.
@@
 Psychiatric disorders had low coherence, with schizophrenia and major depressive disorder being the least coherent among all phenotypes.
-In summary, we provide a general-purpose method for quantifying and comparing coherence of molecular networks.
+In summary, we provide a general-purpose method for quantifying and comparing the coherence of molecular networks.
 
@@
 Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
-A non-zero interacting genes were defined as genes having at least one interaction with other phenotype-specific genes.
+Non-Zero interacting genes were defined as genes having at least one interaction with other phenotype-specific genes.
 Consequently, phenotypes with less than 10 non-zero interacting genes were omitted.
@@
 Although using median slopes of MSigDb and random networks is expected to be a robust reference for high/random coherence, some phenotype-specific networks may have coherence larger than 1, that is, being more coherent than the median coherence of MSigDb networks.
-These situations may occur in smaller networks that have higher change to contain genes with unusually high internal degree distributions.
+These situations may occur in smaller networks that have a higher chance to contain genes with unusually high internal degree distributions.
 Such outliers would inflate coherence estimations.
 Similar outliers may be observed in larger networks.
-For these reasons, coherence estimates for excellively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
+For these reasons, coherence estimates for excessively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating the correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
 

### 2019-09-29 manuscript/manuscript.Rmd
Proof-read abstract/intro

@@
 High/low degree of interactions is suggestive of coherent/random molecular mechanisms, respectively.
-Comparing the degree of interactions may help to better understand the coherence of the molecular mechanisms underlying complex phenotypes and the potential for therapeutic intervention.
-However, direct comparison of the degree of interactions is difficult due to differences in size and configuration of phenotype-associated gene networks.
+Comparing the degree of interactions may help to better understand the coherence of molecular mechanisms underlying complex phenotypes and the potential for therapeutic intervention.
+However, direct comparison of the degree of interactions is difficult due to different sizes and configurations of phenotype-associated gene networks.
 
@@
 The internal degree distribution is defined by interaction counts within a phenotype-specific gene network, while the external degree distribution counts interactions with other genes in the whole protein-protein interaction (PPI) network.
-We present a novel method for normalizing the internal degree distributions, making them directly comparable.
+We present a novel method for normalizing the coherence estimates, making them directly comparable.
 
@@
 **Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences.
-Our results highlight the gaps in of our current knowledge of genetics and molecular mechanisms of complex phenotypes and suggest priorities for future work.
+Our results highlight gaps in our current knowledge of genetics and molecular mechanisms of complex phenotypes and suggest priorities for future GWASs.
 
@@
 Furthermore, the corresponding genes are often spread across different chromosomes with no known connection to one another, hindering understanding of the molecular mechanisms.
-These limitations of current generation GWAS might stem from the yet incomplete knowledge of the genetic determinants of complex phenotypes, the potential heterogeneity of complex phenotypes and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].
+These limitations of current generation GWAS might stem from the yet incomplete knowledge of genetic determinants of complex phenotypes, the potential heterogeneity of complex phenotypes and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].
 
 Many studies have shown that genetically-driven complex phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa].
-This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to human phenotype, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
+This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to understanding human phenotype, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
 Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all molecular determinants of a certain phenotype.
@@
 
-Network properties, such as connectivity (aka degree distribution) [@Ghiassian:2015aa; @Lee:2008aa; @Barabasi:2004aa; @Feldman:2008aa], can be compared to better understand the relationships among the phenotypes [@Vidal:2011aa; @Barabasi:2011aa; @Wang:2011aa].
+Network properties, such as connectivity (aka degree distribution) [@Ghiassian:2015aa; @Lee:2008aa; @Barabasi:2004aa; @Feldman:2008aa], can be compared to better understand the relationships among phenotypes [@Vidal:2011aa; @Barabasi:2011aa; @Wang:2011aa].
 Networks formed by phenotype-associated genes, or, more precisely, their protein-protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa].
@@
 The internal degree distribution is defined as gene-specific interaction counts within a protein-protein interaction (PPI) network formed by phenotype-associated genes, while the external degree distribution counts interactions with other genes in the whole PPI network.
-We used KEGG pathways [@Kanehisa:2000aa] as a reference for a paradigmatically high-coherent network [@Lee:2003aa], while sets of randomly sampled genes were used as the reference for the absence of coherence.
-Using KEGG and random networks matched in size to a phenotype-specific network, we derived a normalized measure of coherence that can be compared across phenotypes.
+We used selected gene sets from the MSigDb database [@Subramanian:2005aa] as a reference for highly coherent networks, while sets of randomly sampled genes were used as the reference for the absence of coherence.
+Using MSigDb and random networks matched in size to phenotype-specific networks, we derived a normalized measure of coherence that can be compared across phenotypes.
 We hypothesized that the level of coherence may inform us about similarities and differences among phenotypes.
@@
 Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have high coherence.
-Our method enables direct comparison of coherence measures and highlights gaps in current understanding of the molecular mechanisms of many phenotypes, e.g., schizophrenia being the lowest coherence disorder in the already low-coherence "Psychiatric disease" category.
+Our method enables direct comparison of coherence measures and highlights gaps in the current understanding of molecular mechanisms of many phenotypes, e.g., Major Depressive Disorder having the lowest coherence in the already low-coherence "Psychiatric disease" category.
 

### 2019-09-30 manuscript/manuscript.Rmd
Only methods left to proof read

@@
 We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods).
-These were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).
+They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).
 
@@
 We opted for the use of three types of PPI data.
-First, STRING data filtered by the interaction confidence score above 500 (middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data.
-This filtering step ensured that only high-confidence PPIs were selected.
+First, STRING data filtered by interaction confidence score above 500 (middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data.
+This filtering step ensures that only high-confidence PPIs were selected.
 The advantage of using filtered data is the reliance on high-quality PPIs.
-The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs are removed from the analysis (see Methods).
-Second, complete STRING data were used to maximize the use of PPI information at the expense of potential noise (referred to hereafter as “STRING”).
+The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs were removed from the analysis (see Methods) leavind 116 phenotypes that could be analyzed.
+Second, full STRING data was used to maximize the use of PPI information at the expense of potential noise (referred to hereafter as “STRING”).
 The advantage of using complete data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
 The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs having low confidence scores.
-Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, were used.
-The use of different data sources (STRING and BioGrid) and filtering (complete or filtered STRING data) was intended to increase the generalizability of our conclusions.
+Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, was used.
+The use of different data sources (STRING and BioGrid) and filtering (full or filtered STRING data) was intended to increase the generalizability of our conclusions.
 
@@
 
-| Category               | Total Phenotypes | Number of SNPs | Number of Genes | Biogrid | String | String Filtered |
-|------------------------|:----------------:|:--------------:|:---------------:|:-------:|:------:|:---------------:|
+| Category               | Number of SNPs | Number of Genes | Total Diseases | Mean Biogrid Coherence | Mean String Coherence | Mean String Filtered Coherence |
+|------------------------|:--------------:|:---------------:|:--------------:|:----------------------:|:---------------------:|:------------------------------:|
+| Trait                  |                |                 |                |                        |                       |                                |
+| Eye trait              |       158      |       109       |        5       |           NA           |         0.724         |              1.050             |
+| Metabolic trait        |       790      |       646       |       10       |          0.738         |         0.779         |              0.863             |
+| Anthropometric trait   |      2090      |       1771      |       15       |          0.460         |         0.637         |              0.691             |
+| Cardiovascular trait   |       728      |       604       |       15       |          0.555         |         0.675         |              0.625             |
+| Intelligence           |       181      |       177       |        4       |          0.180         |         0.395         |              0.405             |
 | Disease                |                |                 |                |                        |                       |                                |
-| Metabolic disease      |         5        |       345      |       280       |  0.515  |  0.776 |      0.845      |
-| Autoimmune             |        20        |      1415      |       1212      |  0.470  |  0.801 |      0.702      |
-| Eye disease            |         8        |       270      |       242       |  0.300  |  0.509 |      0.654      |
-| Neurologic             |         8        |       382      |       343       |  0.710  |  0.498 |      0.637      |
-| Cancer                 |        18        |       713      |       610       |  0.548  |  0.658 |      0.604      |
-| Cardiovascular disease |         9        |       336      |       289       |  0.130  |  0.577 |      0.602      |
-| Psychiatric            |        11        |       678      |       611       |  0.560  |  0.528 |      0.583      |
-| Trait                  |                  |                |                 |         |        |                 |
-| Metabolic trait        |        10        |       790      |       646       |  0.766  |  0.779 |      0.842      |
-| Eye trait              |         5        |       158      |       109       |    NA   |  0.678 |      0.698      |
-| Anthropometric trait   |        19        |      2143      |       1822      |  0.600  |  0.624 |      0.678      |
-| Cardiovascular trait   |        15        |       728      |       604       |  0.463  |  0.655 |      0.537      |
-| Intelligence           |         5        |       236      |       198       |  0.310  |  0.334 |      0.373      |
+| Metabolic disease      |       331      |       269       |        4       |          0.470         |         0.825         |              0.907             |
+| Cancer                 |       670      |       570       |       15       |          0.490         |         0.689         |              0.723             |
+| Neurologic             |       372      |       333       |        7       |           NA           |         0.493         |              0.718             |
+| Autoimmune             |      1383      |       1184      |       18       |          0.425         |         0.820         |              0.701             |
+| Eye disease            |       229      |       202       |        5       |          0.330         |         0.630         |              0.663             |
+| Cardiovascular disease |       301      |       259       |        7       |          0.380         |         0.694         |              0.660             |
+| Psychiatric            |       678      |       611       |       11       |          0.540         |         0.552         |              0.642             |
 
@@
 In terms of phenotypes, an intuitive expectation is that they will be represented by coherent networks of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
-In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the network [@fortunato2010community; @Radicchi:2004aa].
+In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the global network [@fortunato2010community; @Radicchi:2004aa].
 
@@
 
-## Internal vs. external degree distributions as a measure of the coherence
+## Internal vs. external degree distributions as a measure of coherence
 
@@
 -->
-Given that phenotype-associated networks occur within the global network of PPIs, we developed a metric of coherence based on the level of gene interactions.
+Given that phenotype-specific networks occur within the global network of PPIs, we developed a metric of coherence based on the level of gene interactions.
 This metric is inspired by previous work showing that communities within a large network have high internal but low external levels of interactions (degrees) [@PhysRevE.81.046110; @Spirin:2003aa].
@@
 Intuitively, a highly coherent network is expected to have a large number of internal edges but a relatively small number of external edges.
-An extreme example of high coherence would be a fully internally connected network with zero interactions with external genes.
-Conversely, a low-coherence network is expected to have a relatively low number of internal edges compared to external edges, mirroring a random network.
+An extreme example of high coherence would be fully internally connected network with zero interactions with external genes.
+Conversely, a low-coherence network is expected to have a relatively low number of internal edges similar to the number of external edges.
 
@@
 external degree distributions.
-That is, we visualize internal (X-axis) vs.
-external (Y-axis) degrees for each phenotype-specific gene on a single plot and fit a regression line through it.
+That is, we visualized internal (X-axis) vs.
+external (Y-axis) degrees for each phenotype-specific gene on a single plot and fit a regression line through it, enforcing the fit through the origin.
 An example of a network with the highest coherence (full internal connectivity, zero external connections) would be represented by a horizontal line.
@@
 Consequently, slopes of internal vs.
-external degree distributions of phenotype-associated networks would represent the phenotype-specific levels of coherence.
+external degree distributions of phenotype-specific networks would represent the corresponding levels of coherence.
 In summary, comparing internal vs.
@@
 
-## Networks of KEGG/REACTOME pathways and GO Cellular Component collection vs. randomly selected genes serve as a reference for high vs. low coherence, respectively
+## Networks of KEGG/REACTOME pathways and GO Cellular Component collection vs. randomly selected genes serve as references for high vs. low coherence, respectively
 
-The coherence of phenotype-associated gene sets should be measured with respect to the reference level of networks with paradigmatically high and low coherence.
+The coherence of phenotype-associated gene sets should be measured with respect to the realistic references of high and low coherence.
 To establish a reference of high coherence, we considered networks from collections of the Molecular Signatures Database (MSigDB v6.2) [@Subramanian:2005aa].
@@
 We found that pathways from the Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa:2000aa] had the smallest average slope corresponding to the highest level of coherence (Supplementary Table S2).
-Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks of high coherence.
+Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks with high coherence.
 This is expected as genes expressed in the same cellular components (GOCC genes) or participating in the same metabolic pathways (REACTOME genes) are presumed to interact more frequently.
@@
 
-Direct comparison of degree distributions of the networks formed by phenotype-associated genes is hindered by the fact they depend on network size and configuration [@Leskovec:2010:ECA:1772690.1772755].
+Direct comparison of degree distributions of the networks formed by phenotype-specific genes is hindered by the fact they depend on network size and configuration [@Leskovec:2010:ECA:1772690.1772755].
 An intuitive example is to consider a ring and a fully connected network of size 10 and 20 nodes, respectively (Figure 1).
-The degree distributions of the ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).
+The degree distributions of ring networks can be directly compared regardless of network size (vectors of "2" of length 10 and 20, respectively, Figure 1A).
 Although the fully connected networks are similar in that they are fully connected (i.e., similarly coherent), their degree distributions differ (all nodes in the 10-node network have a degree "9", while all nodes in the 20-node network have degree "19", Figure 1B).
@@
 Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
-While degree distributions of the similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
+While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
 |
@@
 
-Given the dependence of internal and external degree distributions on network size, we designed a simple strategy to account for it.
-Briefly, for each phenotype-specific network of a given size, we create size-matched references of high and low coherence.
-Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections) and 100 random networks matched in size for the corresponding network of phenotype-specific genes.
-Consequently, we use the median slopes of internal vs.
-external degree distributions created by these size-matched references to normalize the slopes of the phenotype-specific networks to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (PCC = 0.29, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
+Given the dependence of degree distributions on network size, we designed a simple strategy to account for it.
+Briefly, for each phenotype-specific network of a given size, we created size-matched references of high and low coherence.
+Specifically, we selected 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections) and 100 random networks matched in size to the corresponding network of phenotype-specific genes.
+Consequently, we use median slopes of internal vs.
+external degree distributions created by these size-matched references to normalize the slopes of phenotype-specific networks to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
+We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (Average PCC = 0.01, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 To evaluate the consistency of coherence estimates, we assessed the pairwise correlation of normalized coherences obtained using different PPI databases.
-Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.77).
-Expectedly, coherence estimates using Biogrid database were less similar to that of "STRING" and "STRING filtered"; however, the overall correlation remained high (PCC = 0.56, Supplementary Figure S1).
-Notably, coherence estimates using "STRING filtered" database were most similar to that of "Biogrid" (PCC = 0.57), indicating that filtering step indeed removes "noisy" PPIs and improves quality of coherence estimates.
+Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.80, Supplementary Figure S1).
+Expectedly, coherence estimates using Biogrid database were less similar to that of "STRING" and "STRING filtered"; however, the overall correlation remained high.
+Notably, coherence estimates using "STRING filtered" database were most similar to that of "Biogrid" (PCC = 0.58), indicating that filtering step indeed removes "noisy" PPIs and improves quality of coherence estimates.
 Overall, these results demonstrate consistency in coherence estimates using different PPI databases.
@@
 
-We estimated coherence of 54 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
+We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
 Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
 Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic).
-Notably, metabolic trait category had the highest overall coherence (mean coherence 0.91 and 1.05, respectively, Supplementary Table S1).
+Notably, metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1).
 Several notable examples of low coherence include "Obesity-related traits" and "Height" (anthropometric traits), and blood count-related traits (cardiovascular).
-Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.76).
-These results highlight the relatively good knowledge of the genetics of metabolic traits which corresponds to highly coherent networks.
+Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.41).
+These results highlight the relatively good knowledge of genetics of metabolic traits which corresponds to highly coherent networks.
 Conversely, current knowledge about the genetics of intelligence-related traits appears insufficient as reflected in the low coherence of the corresponding networks.
@@
 
-We further estimated coherence of 80 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3).
-Metabolic diseases had high overall coherence (average coherence 0.91), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
+We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3).
+Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
 Similarly, the average coherence for cancer diseases was 0.72, with "Testicular germ cell tumor" and "Pancreatic cancer" being highly coherent (0.92 and 0.74, respectively).
-The high coherence of metabolic and cancer disease networks highlight the relatively good knowledge about their genetics.
+The high coherence of metabolic and cancer disease networks highlight the relatively good knowledge about their genetic component.
 
@@
 
-Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.64 and 0.63, respectively).
+Diseases in "Psychiatric" and "Cardiovascular" categories showed the lowest average level of coherence (0.64 and 0.66, respectively, Table 1).
 The corresponding coherence estimates were highly heterogeneous.
@@
 Among psychiatric disorders, "Smoking behavior" and "Alcohol dependence" had high coherence (0.87 and 0.82, respectively).
-Notably, coherence of "Schizophrenia" and "Major depressive disorder" networks (0.52 and 0.33, respectively) was at the lower end of coherence estimates among all phenotypes.
+Notably, coherences of "Schizophrenia" and "Major depressive disorder" networks (0.52 and 0.33, respectively) were at the lower end of coherence estimates among all phenotypes.
 Other psychiatric disorders, such as "Bipolar disorder", "ADHD", showed intermediate coherence estimates (0.76 and 0.55, respectively).
@@
 Given the wide range of phenotype-specific coherence estimates, a natural question is whether they are significantly different from random coherence.
-We assess the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
+We assessed the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
 Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.71, Supplementary Figure S1), as would be expected for high-quality PPIs.
-Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to a large number of PPIs that enable phenotype-specific networks to be consistently different from random.
+Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to a large number of PPIs that enabled phenotype-specific networks to be consistently different from random.
 Permutation p-values inversely correlated with coherence estimates (average PCC = -0.23), as would be expected for less coherent networks that resemble random networks.
@@
 Among less significant phenotypes were "White blood cell count" (p-value = 0.102) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.05), "Major depressive disorder" (p-value = 0.03).
-When using Biogrid databases, "Educational attainment", "Myopia (pathological)", "Celiac disease" were among the non-significant phenotypes (p-values 0.29, 0.20,0.16, respectively).
+When using Biogrid databases, "Educational attainment", "Myopia (pathological)", and "Celiac disease" were among the non-significant phenotypes (p-values 0.29, 0.20,0.16, respectively).
 Interestingly, "Schizophrenia" network, despite being among the least coherent, was consistently significantly different from random (p-value < 0.001).
-Overall, the significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence.
-These results further confirm our ability to identify phenotype-specific networks with low, non-significant coherence.
+Overall, the significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence that is frequently statistically insignificant.
+In summary, these results confirm our ability to identify coherence and of phenotype-specific networks with low, non-significant coherence.
 
@@
 We developed a general-purpose method to measure and compare coherence estimates of the molecular mechanisms in networks of genes.
-Intuitively, gene networks with high internal and low external interactions (degrees) would be considered highly coherent; consequently, coherence is defined as the slope through the internal and external degree distribution scatterplot.
-We normalized coherences (slopes) to the range of high and random coherence, exemplified by MSigDb networks and random networks.
+Intuitively, gene networks with high internal and low external number of interactions (degrees) would be considered highly coherent; consequently, coherence is defined as the slope through the internal and external degree distribution scatterplot.
+We normalized coherences (slopes) to the range of high and random coherence, exemplified by MSigDb and random networks.
 To decouple coherence from its dependence on network size, we used MSigDb and random networks matched in size to phenotype-specific networks.
-Applied to the analysis of 133 disease- and trait-associated gene sets, we found metabolic diseases and traits have overall high coherence, while psychiatric and cardiovascular diseases generally had low coherence.
-These results were consistent when using different PPI databases to build gene networks.
+Applied to the analysis of 116 disease- and trait-associated gene sets, we found that metabolic diseases and traits had overall high coherence, while psychiatric and cardiovascular diseases generally had low coherence.
+These results were consistent when using different PPI databases.
 Our method allows one to quantify and compare the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.
@@
 This may be considered a drawback as it has been shown that many networks consist of miniature communities known as motifs [@Milo:2002aa], or graphlets [@Przulj:2007aa].
-While the investigation of motif enrichment [@Milo:2002aa] is a viable approach to characterize a network, our goal was to develop a unified measure of coherence that can be compared across phenotypes.
+While the investigation of motif enrichment [@Milo:2002aa] is a viable approach to characterize a network, our goal was to develop a unified measure of coherence that can be compared across phenotype-specific networks.
 
-Our method relies on the ability to link the phenotype-associated genetic variants detected in genome-wide association studies to genes thought to be affected by them.
+Our method relies on the ability to link the phenotype-associated genetic variants to genes thought to be affected by them.
 Typically, genes are linked to genetic variants by proximity (nearest gene), a strategy adopted by the NHGRI-EBI GWAS catalog [@MacArthur:2017aa].
 <!--Numerous tools have been created to associate genetic variants with genes (MAGENTA [@Segre:2010aa], VEGAS [@Liu:2010aa], SKAT [@Ionita-Laza:2013aa], to name a few, reviewed in [@Wang:2011ab]).--> However, >88% of phenotype-associated genetic variants are located in regulatory elements outside of protein-coding regions [@Hindorff:2009aa].
-<!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].--> The pilot project by ENCODE showed that fewer than 10% of all interactions between TSSs and regulatory regions involved the closest TSS by linear distance [@Sanyal:2012aa], questioning the validity of the nearest gene mapping strategy.
-The development of chromatin conformation capture technologies now allows for understanding interactions among genomic regions on a genome-wide scale [@Lieberman-Aiden:2009aa; @Zhang:2013aa], helping to prioritize disease-associated gene-variant associations [@Pan:2018aa].
+<!--classical genetic studies of translocations and deletions have shown that distal regions, located at tens to hundreds of kilobases from a gene, affect transcription, indicating that regulatory elements can exert their effect over large genomic distances [@Kleinjan:2005aa].--> The pilot project by ENCODE showed that fewer than 10% of all interactions between promoters and regulatory regions involved the closest promoter by linear distance [@Sanyal:2012aa], questioning the validity of nearest gene mapping strategy.
+The development of chromatin conformation capture technologies now allows for understanding long-distance interactions among genomic regions on a genome-wide scale [@Lieberman-Aiden:2009aa; @Zhang:2013aa], helping to prioritize disease-associated gene-variant associations [@Pan:2018aa].
 Consequently, tools are being developed to consider long-distance interactions when mapping SNPs to genes (3DSNP [@Lu:2017aa], FUMA [@Watanabe:2017aa], PINES [@Bodea2016], HUGIn [@Martin:2017aa]), and studies redefining disease-associated genes using long-distance interactions started to emerge [@Fang:2019aa].
@@
 Metabolic pathways were shown to have more duplicated copies of highly central genes making the networks tolerant to loss-of-function mutations [@Khurana:2013aa].
-In summary, centrality measure parallels our observations of network coherence.
+These observations of centrality measure parallel our estimation of network coherence.
 Our future work will incorporate centrality and other gene-centric network metrics into our framework to further refine our ability to compare the coherence of phenotype-associated molecular mechanisms.
@@
 
-Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 133 phenotypes.
+Our study investigated the coherence of the molecular mechanisms of genes associated with genomic variants of 116 phenotypes.
 We developed a network coherence measure driven by the relationship between internal and external connectivity (degree distribution) that is robust to the difference in network sizes.
@@
 We found a range of coherence estimates in each phenotype category, with metabolic phenotypes being the most coherent.
-Psychiatric disorders had low coherence, with schizophrenia and major depressive disorder being the least coherent among all phenotypes.
+Psychiatric disorders had low coherence, with schizophrenia and major depressive disorder being among the least coherent.
 In summary, we provide a general-purpose method for quantifying and comparing the coherence of molecular networks.

### 2019-09-30 manuscript/manuscript.Rmd
Full methods proofread, permutation section remains TBD

@@
 
-Trait-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa].
+Phenotype-associated gene lists were obtained from the NHGRI-EBI catalog of genome-wide association studies (GWAS catalog, `gwascat` v.2.14.0 R package) [@Hindorff:2009aa; @MacArthur:2017aa].
 "MAPPED_GENE" column was used to extract trait-associated genes.
@@
 
-Human protein-protein interactions (PPIs) were obtained from STRING database v.11.0 [@Szklarczyk:2015aa] and BioGrid database v.3.5.174 [@Chatr-Aryamontri:2017aa] (accessed 06/25/2019).
+Human protein-protein interactions (PPIs) were obtained from STRING database v.11.0 [@Szklarczyk:2015aa] and BioGrid database v.3.5.174 [@Chatr-Aryamontri:2017aa].
 The STRING data for the tax ID "9606" (human) was used, and Ensembl protein IDs were converted to gene names using BioMart.
@@
 Genes without PPI annotations (e.g., C5orf4, LOC100507462), antisense transcripts (e.g., CDKN2B-AS), readthrough (e.g., NPHP3-ACAD11), non-protein-coding transcripts (e.g., LINC00478) were excluded.
-Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept.
+Genes that do not have PPIs with each other but interact with other genes in the global PPI network were kept for external degree calculation.
 Non-Zero interacting genes were defined as genes having at least one interaction with other phenotype-specific genes.
 Consequently, phenotypes with less than 10 non-zero interacting genes were omitted.
-Additionally, phenotypes with genes having identical internal degrees (hence, no degree distribution) were filtered due to the inability to derive a fit through the internal vs.
+Additionally, phenotypes with genes having identical internal degrees (hence, no internal degree distribution) were filtered due to the inability to derive a fit through the internal vs.
 external degree distributions (see below).
@@
 
-## Comparing internal vs. external degree (aka connectivity) distributions as a measure of coherence
+## Comparing internal vs. external degree distributions as a measure of coherence
 
@@
 external degree distribution plot was fit with a regression line through the origin to capture the slopes.
-Slopes were used as a measure of coherence with lower slopes being associated with higher coherence.
+Slopes were used as a measure of coherence.
+Lower slopes being associated with higher coherence, and vice versa.
 Degree distributions of the network nodes were obtained using the `igraph` R package v.
@@
 To alleviate the dependency of degree distributions from the size of phenotype-associated gene sets and the content of a selected PPI database, we normalized slopes of phenotype-specific networks to the range of slopes formed by degree distributions of highly coherent and random networks.
-Specifically, we select the 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections, $\beta_{min}$ median slope) and 100 random networks ($\beta_{max}$ median slope) matched in size for the corresponding network of phenotype-specific genes.
+Specifically, we select 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections, $\beta_{min}$ median slope) and 100 random networks ($\beta_{max}$ median slope) matched in size to the corresponding network of phenotype-specific genes.
 For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
@@
 - KEGG - Kyoto Encyclopedia of Genes and Genomes
+- MSigDb - Molecular Signatures Database 
 - PPI - Protein-Protein Interaction

### 2019-09-30 Analysis/01_GWAS_Prep.Rmd
Rename files

--- 2019-08-28 Analysis/GWAS_Prep.Rmd
@@
 library(gwascat) # BiocManager::install("gwascat")
+library(limma) # BiocManager::install("limma")
 library(dplyr)   # install.packages("dplyr")
@@
 library(readr)   # install.packages("readr")
+
 ```

### 2019-09-30 Analysis/03_Normalized_Coherence.Rmd
Disable saving to text files

--- 2019-09-18 Analysis/Normalized_Coherence.Rmd
@@
 #Remove empty categories
-all_cats = subcats[which(subcats!="")]
+all_cats = all_cats[which(all_cats != "")]
 
@@
 
+```{r eval=FALSE}
 # Write general function for saving runs to .csv
 
-```{r}
 save_norm_coh = function(coherence_tab, num_msig, num_rand, run, database) {
@@
 }
-```
 
 # Save the runs in .csv form
-
-```{r}
 #Saving biogrid results to csv

### 2019-10-06 Analysis/Coherence_Calculations_New.Rmd
latest tables

@@
 #Biogrid
-biogrid_p = readRDS("./data/Permutation_Results/biogrid_p_1k.rds") %>% as.data.frame(.) %>% dplyr::select(-new_coherence)
+biogrid_p = readRDS("./data/Permutation_Results/Complete_10k_Biogrid.rds") %>% as.data.frame(.) %>% dplyr::select(-new_coherence)
 
@@
 #String Filtered
-string_filt_p = readRDS("./data/Permutation_Results/string_filt_p_1k.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
+string_filt_p = readRDS("./data/Permutation_Results/Complete_10k_String_Filt.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
 

### 2019-10-06 Analysis/06_Analysis_Tables.Rmd

--- 2019-09-21 Analysis/Analysis_Tables_New.Rmd
@@
 # Average correlation with network size
-print(paste0("Average correlation with network size: ", mean(diff_cors[grepl("Number", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))])))
+diff_cors_selected <- diff_cors[grepl("Number", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))]
+print(paste0("Average correlation with network size: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+# Average correlation of STRING and STRING filtered coherences
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", colnames(diff_cors))]
+print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 # Average correlation of coherences
-diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", rownames(diff_cors))]
-print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))]
 print(paste0("Average correlation of coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+# Average correlation of coherences and p-value
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("P-value", colnames(diff_cors))]
+print(paste0("Average correlation of coherences and p-value: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 

### 2019-10-06 manuscript/manuscript.Rmd

@@
 To test whether each slope is significantly different than random, we used a permutation test.
-For phenotype $i$ with the gene network of size $g_i$, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
-We then generate $n=1000$ networks of randomly selected genes of the same size $g_i$ as the set for phenotype $i$.
-Each random network is created by repeatedly generating networks of size $g_i$ and keeping non-zero genes.
-Once the number of non-zero genes is equal to $g_i$ we combine them into a single random network.
+For phenotype $i$ with the gene network of size $g_i$ that contains $n_i$ number of non-zero interactions, we define $\beta_i$ as the slope derived by regressing the external degree distribution on the internal degree distribution.
+We then generate networks of randomly selected genes of size $g_i$ to ensure size-matched topology.
+Each random network will have a relatively few number of non-zero interactions; therefore, we collect genes non-zero interacting genes from the size-matched random networks until we have $\ge n_i$ of them, and combine them into a single random network.
+This procedure is repeated until we have $n=1000$ random networks of size $g_i$ having $n_i$ non-zero interactions.
 For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.

### 2019-10-06 Analysis/04_Coherence_Calculations.Rmd
make sure tabs are updated

--- 2019-10-06 Analysis/Coherence_Calculations_New.Rmd
@@
 
-#Give appropriate name
+#Give appropriate column names
 colnames(biogrid_p) = c("Disease", "Biogrid P-value", "Ave")

### 2019-10-06 manuscript/manuscript.Rmd
Adjust results for 10000 permutations

@@
 We assessed the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
-Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.71, Supplementary Figure S1), as would be expected for high-quality PPIs.
+Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.64, Supplementary Figure S1), as would be expected for high-quality PPIs.
 Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to a large number of PPIs that enabled phenotype-specific networks to be consistently different from random.
 Permutation p-values inversely correlated with coherence estimates (average PCC = -0.23), as would be expected for less coherent networks that resemble random networks.
-This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.73 (STRING filtered PCC = -0.51, Supplementary Figure S1).
+This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.69 (STRING filtered PCC = -0.53, Supplementary Figure S1).
 These observations confirmed consistent coherence and p-value estimates when using different PPI databases.
@@
 The majority of phenotype-specific networks (~85%) were significantly coherent (STRING filtered p-value < 0.001, Supplementary Table S3).
-Among less significant phenotypes were "White blood cell count" (p-value = 0.102) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.05), "Major depressive disorder" (p-value = 0.03).
-When using Biogrid databases, "Educational attainment", "Myopia (pathological)", and "Celiac disease" were among the non-significant phenotypes (p-values 0.29, 0.20,0.16, respectively).
-Interestingly, "Schizophrenia" network, despite being among the least coherent, was consistently significantly different from random (p-value < 0.001).
+Among less significant phenotypes were "White blood cell count" (p-value = 0.085) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.04), "Major depressive disorder" (p-value = 0.03).
+When using Biogrid databases, "Educational attainment", "Myopia (pathological)", and "Celiac disease" were among the non-significant phenotypes (p-values 0.30, 0.20, 0.16, respectively).
+Interestingly, "Schizophrenia" network, despite being among the least coherent, was consistently significantly different from random (p-value < 0.0001).
 Overall, the significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence that is frequently statistically insignificant.
@@
 To alleviate the dependency of degree distributions from the size of phenotype-associated gene sets and the content of a selected PPI database, we normalized slopes of phenotype-specific networks to the range of slopes formed by degree distributions of highly coherent and random networks.
-Specifically, we select 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections, $\beta_{min}$ median slope) and 100 random networks ($\beta_{max}$ median slope) matched in size to the corresponding network of phenotype-specific genes.
+Specifically, we select 30 MSigDb networks (10 from KEGG, GOCC, and REACTOME collections, $\beta_{min}$ median slope) and 1,000 random networks ($\beta_{max}$ median slope) matched in size to the corresponding network of phenotype-specific genes.
 For each phenotype $i$ with slope $\beta_i$, we calculate the normalized slope $\beta_{i\ norm} = \frac{\beta_{i} - \beta_{max}}{\beta_{min}- \beta_{max}}$, which gives a measure of coherence on the range of [0,1].
@@
 Each random network will have a relatively few number of non-zero interactions; therefore, we collect genes non-zero interacting genes from the size-matched random networks until we have $\ge n_i$ of them, and combine them into a single random network.
-This procedure is repeated until we have $n=1000$ random networks of size $g_i$ having $n_i$ non-zero interactions.
+This procedure is repeated until we have $n=10,000$ random networks of size $g_i$ having $n_i$ non-zero interactions.
 For each random network $j$, we identify its slope $\beta_{r}$ representing the internal vs.

### 2019-10-06 manuscript/manuscript.Rmd
Fix Supplemental Table S3 legend

@@
 **Supplementary Table S3.
-Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) and the reference of high coherence (KEGG, GOCC, Reactome) are specified in the corresponding columns for coherence.
-Permutation p-values were estimated using STRING filtered database and KEGG reference of high coherence.
-"Trait" and "Disease" sheets correspond to subsets of the main table containing coherence estimates for traits/diseases, respectively, sorted by descending "String Filt Coherence KEGG" column.
+Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) are specified in the corresponding column names.
+For each phenotype category, the results are sorted by the "String Filt Coherence" column in descending order.
+"NA" indicates the corresponding value cannot be estimated due to lack of sufficient number of genes annotated with PPIs.
 

### 2019-10-06 manuscript/manuscript.Rmd
Fix authors

@@
 
-# Coherence of Molecular Mechanisms in Genes Identified Through GWAS In Major Human Disease and Traits
+# Coherence of Molecular Mechanisms <!--in Genes Identified Through GWAS -->In Major Human Disease and Traits
 
@@
 Cresswell, Kellen G.^1+^ (cresswellkg@vcu.edu)  
-Reimers, Mark^2^ (reimersm@msu.edu)  
-Bacanu, Silviu-Alin^3^ (silviu-alin.bacanu@vcuhealth.org)  
+Bacanu, Silviu-Alin^2^ (silviu-alin.bacanu@vcuhealth.org)  
 Craver, Carl^4^ (ccraver@wustl.edu)  
-Kendler, Kenneth S.^3^ (kenneth.kendler@vcuhealth.org)
+Reimers, Mark^3^ (reimersm@msu.edu)  
+Kendler, Kenneth S.^2^ (kenneth.kendler@vcuhealth.org)
 
 ^1^ - Department of Biostatistics, Virginia Commonwealth University   
-^2^ - Neuroscience Program and Dept. Biomedical Engineering Michigan State  
-^3^ - Virginia Institute for Psychiatric and Behavior Genetics and the Department of Psychiatry, Virginia Commonwealth University  
+^2^ - Virginia Institute for Psychiatric and Behavior Genetics and the Department of Psychiatry, Virginia Commonwealth University  
+^3^ - Neuroscience Program and Dept. Biomedical Engineering Michigan State  
 ^4^ - Philosophy-Neuroscience-Psychology Program, Washington University in St. Louis  

### 2019-10-06 manuscript/Figures.Rmd
Make PPT of all figures

---
title: "Coherence of Molecular Mechanisms In Major Human Disease and Traits"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##

<center>
![**Figure 1. Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes. While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.](Figures/Fig1_network_distributions.png)
</center>

##

<center>
![**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.](Figures/Fig2_Size_Slope_Plot.png)
</center>

##

<center>
![**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).](Figures/Fig3_summary_traits.png)
</center>

##

<center>
![**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).](Figures/Fig4_summary_diseases.png)
</center>

##

<center>
![**Supplementary Figure S1. Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.](Figures/FigS1_corr_coherence.png)
</center>

##

<center>
![**Supplementary Figure S2. Coherence estimates of traits using Biogrid as a reference PPI database.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis). Plots are faceted by categories. Missing entries indicate that, for a given trait, a network could not be built and the coherence cannot be estimated.](Figures/FigS2_summary_traits_Biogrid.png)
</center>

##

<center>
![**Supplementary Figure S3. Coherence estimates of traits using STRING as a reference of PPI database.** See legend for Supplementary Figure S2.](Figures/FigS3_summary_traits_STRING.png)
</center>

##

<center>
![**Supplementary Figure S4. Coherence estimates of diseases using Biogrid as a reference PPI database.** See legend for Supplementary Figure S2.](Figures/FigS4_summary_diseases_Biogrid.png)
</center>

##

<center>
![**Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.](Figures/FigS5_summary_diseases_STRING.png)
</center>

### 2019-10-12 Analysis/data/Cluster_Scripts/bind_biogrid.R
add scripts from cluster

@@ -1,0 +1,13 @@
+require(dplyr)
+files = list()
+index = 1
+j = 10
+for(i in seq(1,21, by = 10)) {
+	
+		files[[index]] = readRDS(paste0("biogrid_p_", i, "_", j, ".rds"))
+j = j+10
+index = index+1
+
+}
+
+saveRDS(do.call(rbind, files), "Complete_10k_Biogrid.rds")

### 2019-10-12 Analysis/data/Cluster_Scripts/bind_string.R

@@ -1,0 +1,13 @@
+require(dplyr)
+files = list()
+index = 1
+j = 10
+for(i in seq(1,111, by = 10)) {
+	
+		files[[index]] = readRDS(paste0("string_p_10k_", i, "_", j, ".rds"))
+j = j+10
+index = index+1
+
+}
+
+saveRDS(do.call(rbind, files), "Complete_10k_String.rds")

### 2019-10-12 Analysis/data/Cluster_Scripts/bind_string_filt.R

@@ -1,0 +1,13 @@
+require(dplyr)
+files = list()
+index = 1
+j = 10
+for(i in seq(1,71, by = 10)) {
+	
+		files[[index]] = readRDS(paste0("string_filt_p_", i, "_", j, ".rds"))
+j = j+10
+index = index+1
+
+}
+
+saveRDS(do.call(rbind, files), "Complete_10k_String_Filt.rds")

### 2019-10-12 Analysis/data/Cluster_Scripts/biogrid_perm_11_20.R

@@ -1,0 +1,44 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_biogrid.csv")
+Count_Frame = read.csv("biogrid_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+
+links = read_tsv("BIOGRID-ALL-3.5.174.mitab.txt")
+
+links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+
+#Split based on Biogrid ID system then pull out gene names
+
+protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+
+protein1 = unlist(protein1)
+protein2 = unlist(protein2)
+
+#Place into matrix in form that disease2frame works on
+
+links = links %>% mutate(protein1 =protein1, 
+                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
+
+links = links %>% rename(hgnc_symbol_a = protein1) %>% rename(hgnc_symbol_b = protein2)
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[11:20]) {
+  print(i)
+  perm = slope_perm(Count_Frame, i, Norm_Reg =NULL, database = links, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "biogrid_p_11_20.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/biogrid_perm_1_10.R

@@ -1,0 +1,44 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_biogrid.csv")
+Count_Frame = read.csv("biogrid_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+
+links = read_tsv("BIOGRID-ALL-3.5.174.mitab.txt")
+
+links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+
+#Split based on Biogrid ID system then pull out gene names
+
+protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+
+protein1 = unlist(protein1)
+protein2 = unlist(protein2)
+
+#Place into matrix in form that disease2frame works on
+
+links = links %>% mutate(protein1 =protein1, 
+                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
+
+links = links %>% rename(hgnc_symbol_a = protein1) %>% rename(hgnc_symbol_b = protein2)
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[1:10]) {
+  print(i)
+  perm = slope_perm(Count_Frame, i, Norm_Reg =NULL, database = links, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "biogrid_p_1_10.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/biogrid_perm_21_29.R

@@ -1,0 +1,44 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_biogrid.csv")
+Count_Frame = read.csv("biogrid_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+
+links = read_tsv("BIOGRID-ALL-3.5.174.mitab.txt")
+
+links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+
+#Split based on Biogrid ID system then pull out gene names
+
+protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+
+protein1 = unlist(protein1)
+protein2 = unlist(protein2)
+
+#Place into matrix in form that disease2frame works on
+
+links = links %>% mutate(protein1 =protein1, 
+                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
+
+links = links %>% rename(hgnc_symbol_a = protein1) %>% rename(hgnc_symbol_b = protein2)
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[21:29]) {
+  print(i)
+  perm = slope_perm(Count_Frame, i, Norm_Reg =NULL, database = links, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "biogrid_p_21_29.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_101_110.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[101:110]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_101_110.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_111_120.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[110:117]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_110_117.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_11_20.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[11:20]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_11_20.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_1_10.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[1:10]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_1_10.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_21_30.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[21:30]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_21_30.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_31_40.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[31:40]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_31_40.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_41_50.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[41:50]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_41_50.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_51_60.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[51:60]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_51_10.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_61_70.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[61:70]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_61_70.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_71_80.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[71:80]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_71_80.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_81_90.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[81:90]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_81_90.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_cluster_91_100.R

@@ -1,0 +1,42 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[91:100]) {
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_p_10k_91_100.rds")
+}

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_11_20.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[11:20]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_11_20.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_1_10.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[1:10]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_1_10.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_21_30.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[21:30]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_21_30.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_31_40.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[31:40]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_31_40.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_41_50.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[41:50]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_41_50.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_51_60.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[51:60]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_51_60.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_61_70.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[61:70]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_61_70.rds")
+}
+

### 2019-10-12 Analysis/data/Cluster_Scripts/string_perm_filt_71_80.R

@@ -1,0 +1,46 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+
+network = read.csv("coherence_string_filt.csv")
+Count_Frame = read.csv("string_filt_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_p = rbind()
+for (i in unique(Count_Frame$Disease)[71:length(unique(Count_Frame$Disease))]) {
+  print(i)
+  perm = slope_perm(Count_Frame, Disease_Name = i, Norm_Reg =NULL, database = string, permutations = 10000, Disease = TRUE)
+  curr_p = cbind(i, perm)
+  over_p = rbind(over_p, curr_p)
+  saveRDS(over_p, "string_filt_p_71_80.rds")
+}
+

### 2019-10-13 Analysis/04_Coherence_Calculations.Rmd
permutation p values
Fix Attention deficit hyperactivity disorder that was run twice

@@
 #String
-string_p = readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
+string_p = readRDS("./data/Permutation_Results/Complete_10k_String.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
+string_p <- string_p[!duplicated(string_p$Disease), ] # Fix "Attention deficit hyperactivity disorder" that was run twice
 
@@
 ```
+
+```{r}
+```

### 2019-10-13 Analysis/04_Coherence_Calculations.Rmd
More fixes

--- 2019-10-06 Analysis/04_Coherence_Calculations.Rmd
@@
 #String
-string_p = readRDS("./data/Permutation_Results/string_p_1k.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
+string_p = readRDS("./data/Permutation_Results/Complete_10k_String.rds") %>% as.data.frame(.)  %>% dplyr::select(-new_coherence)
+string_p <- string_p[!duplicated(string_p[, 1]), ] # Fix "Attention deficit hyperactivity disorder" that was run twice
 

### 2019-10-13 manuscript/manuscript.Rmd
Update YAML to local paths

@@
 ---
-bibliography: [/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/networks.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/brain.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/enrichment.bib,/Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/3D_refs.bib]
-csl: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.ref/genomebiology.csl
+bibliography: [references/networks.bib,references/brain.bib,references/enrichment.bib,references/3D_refs.bib]
+csl: styles.ref/genomebiology.csl
 output:
   word_document:
-    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/styles.doc/Arial_11_single_space_normal_margins.docx
+    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
   pdf_document: default

### 2019-10-16 Analysis/07_Modularity_Analysis.Rmd
add modularity

---
title: "07_Modularity_Analysis"
author: "Kellen Cresswell"
date: "October 15, 2019"
output: pdf_document
---

```{r}
require(dplyr)
options(stringsAsFactors = FALSE)
```

```{r}
#Read in modularity

biogrid_mod = as.data.frame(readRDS("./data/Modularity_Results/biogrid_mod.rds"))
string_mod = as.data.frame(readRDS("./data/Modularity_Results/string_mod.rds"))
string_filt_mod = as.data.frame(readRDS("./data/Modularity_Results/string_filt_mod.rds"))
```

```{r}
#Read in coherence table

coherence_table = read.csv(".././manuscript/Tables/supplementary_table_S3.csv")

#Putting together frame for merging

mod_frame_b_s = full_join(biogrid_mod, string_mod, "i")

mod_frame = full_join(mod_frame_b_s, string_filt_mod, "i")

#Fixing column names

mod_frame = mod_frame %>% dplyr::rename(Disease = i, Biogrid.Modularity = mod.x, String.Modularity=mod.y, String.Filt.Modularity = mod)

#Join to coherence_table

coherence_table = left_join(coherence_table, mod_frame, "Disease")

#Fix mod cutoff being 5 and p-value being 10 for filtered

coherence_table = coherence_table %>% mutate(String.Filt.Modularity = ifelse(is.na(String.Filt.P.value), NA, String.Filt.Modularity   ))

#Remove . added by read.csv

colnames(coherence_table) = gsub("\\.", " ",colnames(coherence_table))
```

```{r}
#Write to table
write.csv(coherence_table, ".././manuscript/Tables/supplementary_table_S3.csv", 
          row.names = FALSE)
```

### 2019-10-16 Analysis/data/Cluster_Scripts/biogrid_modularity.R
add modularity data and functions

@@ -1,0 +1,45 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+source("./functions/diseases2mod.R")
+
+network = read.csv("coherence_biogrid.csv")
+Count_Frame = read.csv("biogrid_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+
+links = read_tsv("BIOGRID-ALL-3.5.174.mitab.txt")
+
+links = links %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
+
+#Split based on Biogrid ID system then pull out gene names
+
+protein1 = lapply(strsplit(links$`Alt IDs Interactor A`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+protein2 = lapply(strsplit(links$`Alt IDs Interactor B`, "locuslink:"), function(x) strsplit(x[2], "\\|")[[1]][1])
+
+protein1 = unlist(protein1)
+protein2 = unlist(protein2)
+
+#Place into matrix in form that disease2frame works on
+
+links = links %>% mutate(protein1 =protein1, 
+                         protein2 = protein2) %>% dplyr::select(protein1, protein2)
+
+links = links %>% rename(hgnc_symbol_a = protein1) %>% rename(hgnc_symbol_b = protein2)
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_mod = rbind()
+for (i in unique(Count_Frame$Disease)) {
+  Count_Curr = Count_Frame %>% filter(i %in% Disease)
+   Curr_Net = links %>% filter( (hgnc_symbol_a %in% Count_Curr$Genes) | (hgnc_symbol_b %in% Count_Curr$Genes))
+  mod = diseases2mod(Count_Curr, links)
+  curr_mod = cbind(i, mod)
+  over_mod = rbind(over_mod, curr_mod)
+  saveRDS(over_mod, "biogrid_mod.rds")
+}

### 2019-10-16 Analysis/data/Cluster_Scripts/string_filt_modularity.R

@@ -1,0 +1,47 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+source("./functions/diseases2mod.R")
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+string = string %>% filter(combined_score>=500)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_mod = rbind()
+for (i in unique(Count_Frame$Disease)) {
+  Count_Curr = Count_Frame %>% filter(i %in% Disease)
+   Curr_Net = string %>% filter( (hgnc_symbol_a %in% Count_Curr$Genes) | (hgnc_symbol_b %in% Count_Curr$Genes))
+  mod = diseases2mod(Count_Curr, string)
+  curr_mod = cbind(i, mod)
+  over_mod = rbind(over_mod, curr_mod)
+  saveRDS(over_mod, "string_filt_mod.rds")
+}

### 2019-10-16 Analysis/data/Cluster_Scripts/string_modularity.R

@@ -1,0 +1,45 @@
+source("perm_slopes.R")
+require(dplyr)
+require(readr)
+require(igraph)
+source("./functions/diseases2mod.R")
+
+network = read.csv("coherence_string.csv")
+Count_Frame = read.csv("string_edges_new.csv")
+
+Cat_Frame = Count_Frame %>% dplyr::select(Disease, Category) %>% distinct()
+
+network = left_join(network, Cat_Frame)
+
+string = read_table2("9606.protein.links.v11.0.txt.gz")
+
+string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
+                           protein2 = gsub("9606.", "", protein2))
+
+#Read in ensemble protein ID
+
+genes = readRDS("genes.rds")
+
+#Join gene names with string
+
+string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_a = hgnc_symbol)
+
+string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))
+
+string = string %>% rename(hgnc_symbol_b = hgnc_symbol)
+
+
+Count_Frame = Count_Frame %>% dplyr::filter(Internal!=0) %>%
+group_by(Disease) %>% mutate(Count = n()) %>% filter(Count>9)
+
+over_mod = rbind()
+for (i in unique(Count_Frame$Disease)) {
+  Count_Curr = Count_Frame %>% filter(i %in% Disease)
+   Curr_Net = string %>% filter( (hgnc_symbol_a %in% Count_Curr$Genes) | (hgnc_symbol_b %in% Count_Curr$Genes))
+  mod = diseases2mod(Count_Curr, string)
+  curr_mod = cbind(i, mod)
+  over_mod = rbind(over_mod, curr_mod)
+  saveRDS(over_mod, "string_mod.rds")
+}

### 2019-10-16 Analysis/functions/GeneSample.R

@@ -1,7 +1,7 @@
 GeneSample = function(num_genes, actions, directed = TRUE) {
   
   require(igraph)
-  gene_list = sample(unique(actions$hgnc_symbol_a), num_genes)
+  gene_list = base::sample(unique(actions$hgnc_symbol_a), num_genes)
   
   if (directed) {
   act_sub = actions %>% 

### 2019-10-16 Analysis/functions/diseases2mod.R

@@ -6,7 +6,7 @@ diseases2mod = function(disease_dis, links) {
   #Get internal vs. external connectivity
   
   disease_nets = lapply(disease_dis, function(x) {
-    gene_list = read.table(x)$V1
+    gene_list = disease_dis$Genes
     genes2graph(gene_list, links, directed = FALSE)
   })
   

### 2019-10-16 Analysis/06_Analysis_Tables.Rmd
Adjust to include modularity

--- 2019-09-21 Analysis/Analysis_Tables_New.Rmd
@@
 # Remove small/large networks
-mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min) > min_genes &
-                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max) < max_genes, ]
+mtx0_cor <- mtx0[ mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, min, na.rm = TRUE) > min_genes &
+                  mtx0 %>% dplyr::select(starts_with("Size")) %>% apply(., 1, max, na.rm = TRUE) < max_genes, ]
 # Remove size columns
-mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"))
+mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"), contains("P value"), contains("Modularity"))
+# Jitter STRING p-value a bit
+set.seed(1)
+mtx0_cor$`String P value` <- jitter(mtx0_cor$`String P value`, amount = 0.0001)
 # Actual correlation
@@
 # Average correlation with network size
-print(paste0("Average correlation with network size: ", mean(diff_cors[grepl("Number", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))])))
+diff_cors_selected <- diff_cors[grepl("Number", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))]
+print(paste0("Average correlation with network size: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+# Average correlation of STRING and STRING filtered coherences
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", colnames(diff_cors))]
+print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 # Average correlation of coherences
-diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)) & grepl("String", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors)) & grepl("String", rownames(diff_cors))]
-print(paste0("Average correlation of STRING coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("Coherence", colnames(diff_cors))]
 print(paste0("Average correlation of coherences: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+# Average correlation of coherences and p-value
+diff_cors_selected <- diff_cors[grepl("Coherence", rownames(diff_cors)), grepl("P value", colnames(diff_cors))]
+print(paste0("Average correlation of coherences and p-value: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
 
@@
   mtx_traits_selected_to_plot <- data.frame(Disease = rownames(mtx_traits_selected), 
-                                            Pvalue = mtx_traits_selected[, grepl(paste0(databases, " P-value"), colnames(mtx_traits_selected)) ], 
+                                            Pvalue = mtx_traits_selected[, grepl(paste0(databases, " P value"), colnames(mtx_traits_selected)) ], 
                                             Coherence = mtx_traits_selected[, grepl(paste0(databases, " Coherence"), colnames(mtx_traits_selected)) ])

### 2019-10-16 manuscript/manuscript.Rmd
Rewording results

@@
 In terms of graph theory, we are asking whether a phenotype-specific network is a community - a highly connected subgraph relatively well-separated from the rest of the network [@fortunato2010community].
-However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes forming networks of different sizes and configurations.
+However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes that form networks of different sizes and configurations.
 
@@
 Our study identifies significant differences between categories of diseases in terms of network properties of the disease-associated genes and highlights the limitations of the current understanding of the molecular determinants of psychiatric disorders.-->
-This study presents a novel measure of network coherence as a slope of the internal vs.
-external degree distributions.
+This study presents a novel measure of network coherence as a slope of the internal versus external degree distributions.
 The internal degree distribution is defined as gene-specific interaction counts within a protein-protein interaction (PPI) network formed by phenotype-associated genes, while the external degree distribution counts interactions with other genes in the whole PPI network.
@@
 
-To compare coherences, we measured coherence as the slope of a line through a scatterplot of internal vs.
+To compare coherences, we estimated coherence as the slope of a line through a scatterplot of internal vs.
 external degree distributions.
@@
 external (Y-axis) degrees for each phenotype-specific gene on a single plot and fit a regression line through it, enforcing the fit through the origin.
-An example of a network with the highest coherence (full internal connectivity, zero external connections) would be represented by a horizontal line.
-Conversely, an example of a network with the lowest coherence would be a set of non-interacting genes (zero internal connections and full external connectivity) would be represented by a vertical line.
+A network with the highest coherence (full internal connectivity, zero external connections) would be represented by a horizontal line.
+Conversely, a network with the lowest coherence (zero internal connectivity, full external connectivity) would be represented by a vertical line.
 Consequently, slopes of internal vs.
@@
 However, similar normalization of ring networks will make their degree distributions different (0.22 for the 10-node network and 0.11 for the 20-node network), leading to the illusory impression of differences in coherence between intuitively equally coherent networks.
--->These observations highlight the difficulty in comparing degree distributions of phenotype-associated gene networks due to two facts: 1) network sizes differ, and 2) network configuration (fully connected, ring, or intermediate connectivity) is unknown.
+-->These observations highlight the difficulty in comparing degree distributions of phenotype-associated gene networks due to two facts: 1) the network sizes differ, and 2) the network configuration (e.g., fully connected, ring, or intermediate connectivity) is unknown.
 
@@
 external degree distributions created by these size-matched references to normalize the slopes of phenotype-specific networks to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed alleviates the dependency between coherence and network size (Average PCC = 0.01, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
+We found that the use of size-matched references indeed alleviates the dependency of coherence on network size (Average PCC = 0.01, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 
-Besides size dependency, coherence estimates depend on the choice of the PPI database used to build networks and estimate the corresponding internal vs.
+In addition to size dependency, coherence estimates depend on the choice of the PPI database used to build networks and estimate the corresponding internal vs.
 external degree distributions.
@@
 We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
-Among anthropometric traits, "Weight" and "Body Mass Index" had the highest level of coherence.
+Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence.
 Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic).
-Notably, metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1).
-Several notable examples of low coherence include "Obesity-related traits" and "Height" (anthropometric traits), and blood count-related traits (cardiovascular).
-Traits in the "Intelligence" category had low overall coherence (mean coherence = 0.41).
-These results highlight the relatively good knowledge of genetics of metabolic traits which corresponds to highly coherent networks.
-Conversely, current knowledge about the genetics of intelligence-related traits appears insufficient as reflected in the low coherence of the corresponding networks.
+Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1).
+Several notable examples of low coherence include Obesity-related traits and Height (anthropometric traits), and blood count-related traits (cardiovascular).
+Traits in the Intelligence category had low overall coherence (mean coherence = 0.41).
 
@@
 Similarly, the average coherence for cancer diseases was 0.72, with "Testicular germ cell tumor" and "Pancreatic cancer" being highly coherent (0.92 and 0.74, respectively).
-The high coherence of metabolic and cancer disease networks highlight the relatively good knowledge about their genetic component.
 
@@
 The average coherence of diseases in the "Eye disease" category was relatively high (0.66).
-These results suggest good overall knowledge of the genetics of autoimmune, neurologic, and eye-related diseases and highlight limited knowledge for some diseases in these categories.
-
 +-------+
@@
 Other psychiatric disorders, such as "Bipolar disorder", "ADHD", showed intermediate coherence estimates (0.76 and 0.55, respectively).
-In summary, these low coherence estimates highlight the need for further research on the genetics of most cardiovascular and psychiatric diseases to better understand their molecular mechanisms.
-These results further suggest "Schizophrenia" and "Major depressive disorder" as the top candidates for further genetic studies in psychiatric genome-wide association studies.
 
@@
 Our method allows one to quantify and compare the estimated coherence of molecular mechanisms across phenotypes, and will only improve as more PPI information becomes available.
+
+The coherence measure allows us to gain insights into the genetic component of various phenotypes, as revealed by the current state of the corresponding GWASs.
+The high coherence of metabolic diseases and traits is suggestive of the strong genetic component driving molecular mechanisms of metabolism-specific gene networks.
+On the contrary, the low coherence of a phenotype suggests either insufficient knowledge of genetics or the true lack of genetic influence.
+The insufficient knowledge may be remedied by identifying novel genes through larger-scale GWASs and meta-analyses.
+However, the lack of genetic influence will continue to be reflected in low coherence measures irrespectively of the number of genes.
+Our results highlight the low coherence of the molecular mechanisms of intelligence-related traits, cardiovascular and psychiatric disorders, but cannot answer whether insufficient knowledge or the lack of genetic component drives it.
 

### 2019-10-19 manuscript/manuscript.Rmd
Add Carl's wording

@@
 The coherence measure allows us to gain insights into the genetic component of various phenotypes, as revealed by the current state of the corresponding GWASs.
-The high coherence of metabolic diseases and traits is suggestive of the strong genetic component driving molecular mechanisms of metabolism-specific gene networks.
-On the contrary, the low coherence of a phenotype suggests either insufficient knowledge of genetics or the true lack of genetic influence.
-The insufficient knowledge may be remedied by identifying novel genes through larger-scale GWASs and meta-analyses.
-However, the lack of genetic influence will continue to be reflected in low coherence measures irrespectively of the number of genes.
-Our results highlight the low coherence of the molecular mechanisms of intelligence-related traits, cardiovascular and psychiatric disorders, but cannot answer whether insufficient knowledge or the lack of genetic component drives it.
+The high coherence of metabolic diseases and traits arguably reflects the strong genetic component driving molecular mechanisms of metabolism-specific gene networks.
+On the contrary, the low coherence of a phenotype suggests either insufficient knowledge of genetics (as represented in current-generation databases) or the true lack of such a strong genetic influence, (and correspondingly, the heightened importance in these cases of non-genetic, and possibly environmental, factors.
+If the incoherence results from insufficient knowledge, this should be remedied as novel genes are identified through larger-scale GWASs and meta-analyses.
+Yet, the low coherence of a phenotype is expected to remain low with increasing gene discovery if, in contrast, the absence of coherence is due to the relative importance of non-genetic causes.
+Our results highlight the low coherence of the molecular mechanisms of intelligence-related traits and cardiovascular and psychiatric disorders, but our method alone cannot decide whether insufficient knowledge or the weakness of the genetic component ultimately explains the lack of internal connectivity characteristic of low-coherence traits.
 

### 2019-10-24 manuscript/manuscript.Rmd
Add comments from Silviu

@@
 Many studies have shown that genetically-driven complex phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa].
-This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to understanding human phenotype, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
+This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to understanding human phenotypes, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
 Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all molecular determinants of a certain phenotype.
@@
 Network properties, such as connectivity (aka degree distribution) [@Ghiassian:2015aa; @Lee:2008aa; @Barabasi:2004aa; @Feldman:2008aa], can be compared to better understand the relationships among phenotypes [@Vidal:2011aa; @Barabasi:2011aa; @Wang:2011aa].
-Networks formed by phenotype-associated genes, or, more precisely, their protein-protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa].
+Networks formed by interactions among phenotype-associated genes, or, more precisely, their protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa].
 This intuition reflects a well-known "guilt-by-association" principle that genes (or, more precisely, their products) that form an interaction network are more likely to share similar functions [@Farber:2013aa], or co-expression patterns [@Huang:2006aa; @Michalak:2008aa].
@@
 In terms of graph theory, we are asking whether a phenotype-specific network is a community - a highly connected subgraph relatively well-separated from the rest of the network [@fortunato2010community].
-However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes that form networks of different sizes and configurations.
+However, the direct comparison of connectivity across phenotype-specific networks is hindered by the fact that different phenotypes are associated with different numbers of genes forming networks of different sizes and configurations.
 
@@
 The advantage of using filtered data is the reliance on high-quality PPIs.
-The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs were removed from the analysis (see Methods) leavind 116 phenotypes that could be analyzed.
+The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs were removed from the analysis (see Methods) leaving 116 phenotypes that could be analyzed.
 Second, full STRING data was used to maximize the use of PPI information at the expense of potential noise (referred to hereafter as “STRING”).
@@
 In genomics, nodes typically represent genes and edges correspond to some measure of interactions, e.g., gene co-expression or interactions between protein products [@Goncalves:2012aa; @Wang:2012aa].
-In terms of phenotypes, an intuitive expectation is that they will be represented by coherent networks of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
+In terms of phenotypes, an intuitive expectation is that they will be represented by coherent networks of functionally related genes, similar gene expression profiles, shared genomic variants, higher PPI interactions, and higher co-morbidity [@Zhang:2010ab; @Goh:2007aa; @Driel:2006aa; @Wang:2012aa; @Hamaneh:2015aa; @Menche:2015aa; @Jia:2011aa], reviewed in [@Dozmorov:2018aa; @Wang:2011aa].
 In terms of graph theory, a coherent network is a community consisting of a group of nodes that are relatively highly connected to each other but sparsely connected to other nodes in the global network [@fortunato2010community; @Radicchi:2004aa].
@@
 To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
-We observed negative correlation of network size with the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
+We observed substantial negative correlation between network sizes and the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
 These observations confirm the notion that coherence estimation using internal vs.
@@
 Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1).
-Several notable examples of low coherence include Obesity-related traits and Height (anthropometric traits), and blood count-related traits (cardiovascular).
+Several notable examples of low coherence include Obesity-related traits, Height (anthropometric traits), and blood count-related traits (cardiovascular).
 Traits in the Intelligence category had low overall coherence (mean coherence = 0.41).
@@
 
-## Coherence of most phenotypes significantly differs from random
+## Coherence of most phenotypes is significantly larger than random networks
 
-Given the wide range of phenotype-specific coherence estimates, a natural question is whether they are significantly different from random coherence.
+Given the wide range of phenotype-specific coherence estimates, a natural question is whether they are significantly larger than the coherence of random networks.
 We assessed the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
@@
 This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.69 (STRING filtered PCC = -0.53, Supplementary Figure S1).
-These observations confirmed consistent coherence and p-value estimates when using different PPI databases.
+These observations confirmed that coherence and p-value estimates remained consistent when using different PPI databases.
 
-The majority of phenotype-specific networks (~85%) were significantly coherent (STRING filtered p-value < 0.001, Supplementary Table S3).
+The majority of phenotype-specific networks (~85%) were significantly more coherent than random networks (STRING filtered p-value < 0.001, Supplementary Table S3).
 Among less significant phenotypes were "White blood cell count" (p-value = 0.085) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.04), "Major depressive disorder" (p-value = 0.03).
@@
 Overall, the significance estimation of coherence supports our observation that phenotype-specific networks in "Intelligence" and "Psychiatric disease" categories generally have low coherence that is frequently statistically insignificant.
-In summary, these results confirm our ability to identify coherence and of phenotype-specific networks with low, non-significant coherence.
+In summary, these results confirm our ability to identify coherence of phenotype-specific networks and test their significance over random networks.
 
@@
 
-We developed a general-purpose method to measure and compare coherence estimates of the molecular mechanisms in networks of genes.
+We developed a general-purpose method to measure and compare the coherence estimates of the molecular mechanisms in networks of genes.
 Intuitively, gene networks with high internal and low external number of interactions (degrees) would be considered highly coherent; consequently, coherence is defined as the slope through the internal and external degree distribution scatterplot.
@@
 To decouple coherence from its dependence on network size, we used MSigDb and random networks matched in size to phenotype-specific networks.
-Applied to the analysis of 116 disease- and trait-associated gene sets, we found that metabolic diseases and traits had overall high coherence, while psychiatric and cardiovascular diseases generally had low coherence.
+Applied to the analysis of 116 disease- and trait-associated gene sets, we found that metabolic diseases and traits had overall high coherence, while psychiatric and cardiovascular diseases generally had lower coherence.
 These results were consistent when using different PPI databases.
@@
 
-The definition of coherence as the slope between internal vs.
-external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa].
+The definition of coherence as the slope between internal versus external degree distributions makes our measure dependent on the choice of PPI data [@Bajpai:2019aa].
 Our limited knowledge of protein-protein interactions limits the coherence estimation to phenotypes that have a sufficient number of genes with PPI annotations, especially for phenotypes with low numbers of associated genes.
@@
 
-In addition to internal vs.
-external degree distributions, other gene-centric network metrics can be used to estimate coherence [@Leskovec:2010:ECA:1772690.1772755].
+In addition to internal versus external degree distributions, other gene-centric network metrics can be used to estimate coherence [@Leskovec:2010:ECA:1772690.1772755].
 For example, the centrality measure has been used to demonstrate that highly central hub genes are functionally essential and highly conserved [@Khurana:2013aa].

### 2019-11-05 manuscript/manuscript.Rmd
Spacing fix
Make formatting to conform BMC Bioinformatics

@@
   word_document:
-    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
+    reference_docx: styles.doc/BMC_Bioinformatics.docx
   pdf_document: default
@@
 The average coherence of diseases in the "Eye disease" category was relatively high (0.66).
+
 +-------+

### 2019-11-05 manuscript/manuscript.Rmd
Reference style for BMC Bioinformatics

@@
 bibliography: [references/networks.bib,references/brain.bib,references/enrichment.bib,references/3D_refs.bib]
-csl: styles.ref/genomebiology.csl
+csl: styles.ref/bmc-bioinformatics.csl
 output:

### 2019-11-05 manuscript/manuscript.Rmd
Add Availability of data and materials section
Adjust abbreviations and contributions

@@
 
+## Availability of data and materials
+
+The code and the data supporting the conclusions of this article are available in the https://github.com/dozmorovlab/disease_coherence GitHub repository.
+
 # Abbreviations
 
-- GWAS - Genome-Wide Association Study
-- KEGG - Kyoto Encyclopedia of Genes and Genomes
-- MSigDb - Molecular Signatures Database 
-- PPI - Protein-Protein Interaction
-- SNP - Single Nucleotide Polymorphism
+GWAS - Genome-Wide Association Study; KEGG - Kyoto Encyclopedia of Genes and Genomes; MSigDb - Molecular Signatures Database; PPI - Protein-Protein Interaction; SNP - Single Nucleotide Polymorphism
 
@@
 The authors declare that they have no competing interests.
+
+## Authors’ contributions
+
+MD, KK, CC, and MR conceived the study.
+KC and MD performed the experiments.
+MD primarily wrote the manuscript.
+KK, MR, CC, and SB participated in manuscript writing.
+All authors read and approved the final manuscript.
 

### 2019-11-05 manuscript/manuscript.Rmd
Format for BMC Bioinformatics

@@
 
-**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 133 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (the reference for high coherence) and random networks (the lower limit of coherence).
+**Results.** Using STRING and BioGrid PPI databases, we compared the coherence of 116 phenotype-associated gene sets from GWAScatalog against size-matched KEGG pathways (the reference for high coherence) and random networks (the lower limit of coherence).
 We observed a range of coherence estimates for each category of phenotypes.
@@
 We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods).
-They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).
+They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Additional File 1).
 
@@
 MSigDB contains sets of genes having various types of functional relationships.
-We found that pathways from the Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa:2000aa] had the smallest average slope corresponding to the highest level of coherence (Supplementary Table S2).
+We found that pathways from the Kyoto Encyclopedia of Genes and Genomes (KEGG) [@Kanehisa:2000aa] had the smallest average slope corresponding to the highest level of coherence (Additional File 2).
 Networks assembled from Gene Ontology - Cellular Component (GOCC) collection and REACTOME pathways followed, representing alternative sources of networks with high coherence.
@@
 
+<!--
 +-------+
@@
 +-------+
+-->
+**Figure 1.
+Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
+While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
 
 To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
-We observed substantial negative correlation between network sizes and the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Supplementary Table S2).
+We observed substantial negative correlation between network sizes and the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Additional File 2).
 These observations confirm the notion that coherence estimation using internal vs.
@@
 +-------+
+-->
+**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
 
@@
 external degree distributions created by these size-matched references to normalize the slopes of phenotype-specific networks to the [0, 1] range (see Methods), where 0/1 correspond to random/high coherence, respectively.
-We found that the use of size-matched references indeed alleviates the dependency of coherence on network size (Average PCC = 0.01, Supplementary Figure S1) and allows for direct comparison of normalized coherences across phenotypes.
+We found that the use of size-matched references indeed alleviates the dependency of coherence on network size (Average PCC = 0.01, Additional File 4) and allows for direct comparison of normalized coherences across phenotypes.
 
@@
 To evaluate the consistency of coherence estimates, we assessed the pairwise correlation of normalized coherences obtained using different PPI databases.
-Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.80, Supplementary Figure S1).
+Correlation between coherence estimates obtained using "STRING" and "STRING filtered" was high (PCC = 0.80, Additional File 4).
 Expectedly, coherence estimates using Biogrid database were less similar to that of "STRING" and "STRING filtered"; however, the overall correlation remained high.
@@
 
-We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
+We estimated coherence of 49 traits from five categories (Figure 3, Additional File 3, Additional File 5, 6).
 Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence.
 Other examples of traits with high coherence include blood pressure-related traits (cardiovascular) and glucose-related traits (metabolic).
-Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Supplementary Table S1).
+Notably, the metabolic trait category had the highest overall coherence (mean coherence 0.86, Table 1), and these results were consistent when using other PPI databases (Additional File 1).
 Several notable examples of low coherence include Obesity-related traits, Height (anthropometric traits), and blood count-related traits (cardiovascular).
@@
 +-------+
+-->
+**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
 
@@
 
-We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3).
+We further estimated coherence of 67 diseases from seven categories (Figure 4, Additional File 7, 8, Additional File 3).
 Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
@@
 On the other end of the coherence spectrum, "Allergic rinitis" and "Atopic dermatitis" were the least coherent (0.51, 0.60, respectively).
-The average coherence of autoimmune diseases was high, 0.70 (Table 1), with all autoimmune diseases, except "Allergic rinitis", having coherence estimates > 0.60 (Supplementary Table S3).
+The average coherence of autoimmune diseases was high, 0.70 (Table 1), with all autoimmune diseases, except "Allergic rinitis", having coherence estimates > 0.60 (Additional File 3).
 Diseases in the "Neurologic" category were similarly highly coherent (average coherence 0.72).
@@
 +-------+
+-->
+**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
 
@@
 We assessed the significance of coherence estimates using a permutation test (see Methods), individually for each PPI database.
-Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.64, Supplementary Figure S1), as would be expected for high-quality PPIs.
+Permutation p-values obtained using Biogrid and STRING filtered databases were highly correlated (PCC = 0.64, Additional File 4), as would be expected for high-quality PPIs.
 Permutation p-values using STRING databases were predominantly significant (p-value < 0.001) due to a large number of PPIs that enabled phenotype-specific networks to be consistently different from random.
 Permutation p-values inversely correlated with coherence estimates (average PCC = -0.23), as would be expected for less coherent networks that resemble random networks.
-This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.69 (STRING filtered PCC = -0.53, Supplementary Figure S1).
+This dependency was especially pronounced for database-specific estimates, e.g., Biogrid coherence and p-value estimates correlated with PCC = -0.69 (STRING filtered PCC = -0.53, Additional File 4).
 These observations confirmed that coherence and p-value estimates remained consistent when using different PPI databases.
 
-The majority of phenotype-specific networks (~85%) were significantly more coherent than random networks (STRING filtered p-value < 0.001, Supplementary Table S3).
+The majority of phenotype-specific networks (~85%) were significantly more coherent than random networks (STRING filtered p-value < 0.001, Additional File 3).
 Among less significant phenotypes were "White blood cell count" (p-value = 0.085) and other blood count-related phenotypes, "Corneal astigmatism" (p-value = 0.04), "Major depressive disorder" (p-value = 0.03).
@@
 Similar outliers may be observed in larger networks.
-For these reasons, coherence estimates for excessively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating the correlation between coherences obtained using different PPI databases (Supplementary Figure S1).
+For these reasons, coherence estimates for excessively large networks ("Obesity-related traits", 779 genes, "Height", 418 genes) were excluded from calculating the correlation between coherences obtained using different PPI databases (Additional File 4).
 
@@
 
+# Additional files
+
+**Additional File 1.
+Summary statistics of the analyzed phenotype categories.** Minimum, mean, median, maximum of coherence estimates, the number of genes, SNPs, the total number of phenotype networks.
+Each worksheet corresponds to the PPI database used.
+
+**Additional File 2.
+Summary of networks from MSigDB categories.** Minimum, mean, median, maximum of the slopes, the number of genes, and the total number of networks are shown for each collection, along with the correlation between network size and slope.
+"CP:/KEGG/Reactome/Biocarta" - canonical pathways, "GOBP/GOMF/GOCC" - gene ontology biological processes, molecular functions, cellular component, "TFT" - transcription factor targets, "CGN" - cancer gene neighborhoods, "CM" - cancer modules, "MIR" - microRNA targets, "CGP" - chemical and genetic perturbations.
+
+**Additional File 3.
+Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) are specified in the corresponding column names.
+For each phenotype category, the results are sorted by the "String Filt Coherence" column in descending order.
+"NA" indicates the corresponding value cannot be estimated due to lack of sufficient number of genes annotated with PPIs.
+
+**Additional File 4.
+Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.
+
+**Additional File 5.
+Coherence estimates of traits using Biogrid as a reference PPI database.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+Plots are faceted by categories.
+Missing entries indicate that, for a given trait, a network could not be built and the coherence cannot be estimated.
+
+**Additional File 6. Coherence estimates of traits using STRING as a reference of PPI database.** See legend for Additional File 5.
+
+**Additional File 7. Coherence estimates of diseases using Biogrid as a reference PPI database.** See legend for Additional File 5.
+
+**Additional File 8. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Additional File 5.
+
+<!--
 # Supplementary material
@@
 **Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.
+-->
 

### 2019-12-17 manuscript/manuscript.Rmd
Add BMC Bioinformatics submission

--- 2019-11-05 manuscript/manuscript.Rmd
@@
   word_document:
-    reference_docx: styles.doc/BMC_Bioinformatics.docx
+    reference_docx: styles.doc/Arial_11_single_space_normal_margins.docx
   pdf_document: default
@@
 We collected 133 phenotype-associated gene lists from the NHGRI-EBI GWAS catalog [@MacArthur:2017aa] (see Methods).
-They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table 1).
+They were grouped into seven disease categories (“Autoimmune”, “Cancer”, “Cardiovascular disease”, “Eye disease”, “Metabolic disease”, “Neurologic”, “Psychiatric”) and five trait categories (“Anthropometric trait”, “Cardiovascular trait”, “Eye trait”, “Intelligence”, “Metabolic trait”, Table 1, Supplementary Table S1).
 
@@
 
+<!--
+**Figure 1.
+Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
+While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
+-->
+
 To observe whether network properties affect coherence estimation in experimental settings, we investigated the effect of network size on the slopes of KEGG and random networks.
@@
 +-------+
+
+<!--
+**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
+-->
 
@@
 
-We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figures S2-S3).
+We estimated coherence of 49 traits from five categories (Figure 3, Supplementary Table S3, Supplementary Figure S2, 6).
 Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence.
@@
 
+<!--
+**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+-->
+
 ## Metabolic diseases and cancer networks have high coherence
 
-We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figures S4-S5, Supplementary Table S3).
+We further estimated coherence of 67 diseases from seven categories (Figure 4, Supplementary Figure S4, 8, Supplementary Table S3).
 Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
@@
 +-------+
+
+<!--
+**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
+-->
 
@@
 
+<!--
+# Supplementary material
+
+**Supplementary Table S1.
+Summary statistics of the analyzed phenotype categories.** Minimum, mean, median, maximum of coherence estimates, the number of genes, SNPs, the total number of phenotype networks.
+Each worksheet corresponds to the PPI database used.
+
+**Supplementary Table S2.
+Summary of networks from MSigDB categories.** Minimum, mean, median, maximum of the slopes, the number of genes, and the total number of networks are shown for each collection, along with the correlation between network size and slope.
+"CP:/KEGG/Reactome/Biocarta" - canonical pathways, "GOBP/GOMF/GOCC" - gene ontology biological processes, molecular functions, cellular component, "TFT" - transcription factor targets, "CGN" - cancer gene neighborhoods, "CM" - cancer modules, "MIR" - microRNA targets, "CGP" - chemical and genetic perturbations.
+
+**Supplementary Table S3.
+Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) are specified in the corresponding column names.
+For each phenotype category, the results are sorted by the "String Filt Coherence" column in descending order.
+"NA" indicates the corresponding value cannot be estimated due to lack of sufficient number of genes annotated with PPIs.
+
+**Supplementary Figure S1.
+Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.
+
+**Supplementary Figure S2.
+Coherence estimates of traits using Biogrid as a reference PPI database.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+Plots are faceted by categories.
+Missing entries indicate that, for a given trait, a network could not be built and the coherence cannot be estimated.
+
+**Supplementary Figure S3. Coherence estimates of traits using STRING as a reference of PPI database.** See legend for Supplementary Figure S2.
+
+**Supplementary Figure S4. Coherence estimates of diseases using Biogrid as a reference PPI database.** See legend for Supplementary Figure S2.
+
+**Supplementary Figure S5. Coherence estimates of diseases using STRING as a reference PPI database.** See legend for Supplementary Figure S2.
+-->
+
 # References

### 2019-12-17 manuscript/submission.BMC_Bioinformatics/cover_letter_BMC_bioinformatics.Rmd
Add BMC Bioinformatics submission

---
output:
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NSF_grant_style.docx
  html_document: default
---

Dear Dr. Acosta,

Please, consider the amended version of our manuscript, formatted according to BMC Bioinformatics guidelines. We apologize for some oversights in formatting the submission.

Best regards,
Mikhail Dozmorov

Dear Editor,

We are submitting our manuscript "Coherence of Molecular Mechanisms In Major Human Disease and Traits" to be considered for publication in _BMC Bioinformatics_. This manuscript will be of interest to the network analysis community working with the analysis of molecular networks.

The main highlights of our work are:

- A novel method for comparing molecular networks. We introduce a metric for measuring the coherence of molecular mechanisms in the networks of phenotype-specific genes. This metric is independent of the size of molecular networks, allowing for comparing coherences across phenotype-specific networks of different sizes. The high coherence of a molecular network may inform about the therapeutically targetable molecular mechanisms.
- Coherence estimation and comparison across 116 phenotype-specific networks from seven disease- and five trait categories, including the estimation of statistical significance of coherence measures.
- Validation of our approach using different protein-protein interaction (PPI) databases.
- Identification of diseases and traits with high or low coherence. E.g., metabolic diseases and traits showed consistently high coherence, while psychiatric disorders and intelligence-related traits were the least coherent.

All data used in the manuscript was sourced from publicly available repositories (GWAScatalog, STRING, and Biogrid PPIs, MSigDb). All code and data needed to reproduce our results are made available on our GitHub repository https://github.com/dozmorovlab/disease_coherence. We also want to assure that this manuscript is not being considered for publication elsewhere.

With our best regards,  
On behalf of all authors,

Mikhail Dozmorov, Ph.D., Blick scholar  
Assistant Professor,  Department of Biostatistics 
Affiliate, Department of Pathology  
Virginia Commonwealth University  
One Capitol Square, room 738  
830 East Main Street  
Richmond,  Virginia  23298  
e-mail:  mikhail.dozmorov@vcuhealth.org  
Phone: 804-827-2055

### 2019-12-17 manuscript/submission.BMC_Bioinformatics/manuscript_BMC_Bioinformatics.Rmd
Add BMC Bioinformatics submission

--- 2019-11-05 manuscript/manuscript.Rmd
@@
 
-Dozmorov, Mikhail G.^1+^ (mikhail.dozmorov@vcuhealth.org)  
+Dozmorov, Mikhail G.^1+\*^ (mikhail.dozmorov@vcuhealth.org)  
 Cresswell, Kellen G.^1+^ (cresswellkg@vcu.edu)  
@@
 Reimers, Mark^3^ (reimersm@msu.edu)  
-Kendler, Kenneth S.^2^ (kenneth.kendler@vcuhealth.org)
+Kendler, Kenneth S.^2\*^ (kenneth.kendler@vcuhealth.org)
 
@@
 ^4^ - Philosophy-Neuroscience-Psychology Program, Washington University in St. Louis  
-^+^ - equal contribution
+^+^ - Equal contribution  
+^\*^ - Corresponding author
 
@@
 
-# Introduction
+# Background
 
@@
 The use of different data sources (STRING and BioGrid) and filtering (full or filtered STRING data) was intended to increase the generalizability of our conclusions.
-
-<!--Tables/table_1.csv-->
-**Table 1.
-Summary statistics of the analyzed categories.** Average coherence estimates are shown.
-"NA" values indicate that a category lacked phenotypes with a sufficient number of PPIs.
-Sorted by "String Filtered" average coherence.
-
-| Category               | Number of SNPs | Number of Genes | Total Diseases | Mean Biogrid Coherence | Mean String Coherence | Mean String Filtered Coherence |
-|------------------------|:--------------:|:---------------:|:--------------:|:----------------------:|:---------------------:|:------------------------------:|
-| Trait                  |                |                 |                |                        |                       |                                |
-| Eye trait              |       158      |       109       |        5       |           NA           |         0.724         |              1.050             |
-| Metabolic trait        |       790      |       646       |       10       |          0.738         |         0.779         |              0.863             |
-| Anthropometric trait   |      2090      |       1771      |       15       |          0.460         |         0.637         |              0.691             |
-| Cardiovascular trait   |       728      |       604       |       15       |          0.555         |         0.675         |              0.625             |
-| Intelligence           |       181      |       177       |        4       |          0.180         |         0.395         |              0.405             |
-| Disease                |                |                 |                |                        |                       |                                |
-| Metabolic disease      |       331      |       269       |        4       |          0.470         |         0.825         |              0.907             |
-| Cancer                 |       670      |       570       |       15       |          0.490         |         0.689         |              0.723             |
-| Neurologic             |       372      |       333       |        7       |           NA           |         0.493         |              0.718             |
-| Autoimmune             |      1383      |       1184      |       18       |          0.425         |         0.820         |              0.701             |
-| Eye disease            |       229      |       202       |        5       |          0.330         |         0.630         |              0.663             |
-| Cardiovascular disease |       301      |       259       |        7       |          0.380         |         0.694         |              0.660             |
-| Psychiatric            |       678      |       611       |       11       |          0.540         |         0.552         |              0.642             |
 
@@
 -->
-**Figure 1.
-Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
-While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
 
@@
 -->
-**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
 
@@
 -->
-**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
 
@@
 -->
-**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
 
@@
 
+# Abbreviations
+
+GWAS - Genome-Wide Association Study; KEGG - Kyoto Encyclopedia of Genes and Genomes; MSigDb - Molecular Signatures Database; PPI - Protein-Protein Interaction; SNP - Single Nucleotide Polymorphism
+
+# Declarations
+
+## Ethics approval and consent to participate
+
+Not applicable.
+
+## Consent for publication
+
+Not applicable.
+
+## Availability of data and materials
+
+The code and the data supporting the conclusions of this article are available in the https://github.com/dozmorovlab/disease_coherence GitHub repository.
+
+## Competing interests
+
+The authors declare that they have no competing interests.
+
+## Funding
+
+This work was supported by a Genetics and Human Agency Award from the Templeton Foundation to KSK, MR, CC, SB.
+The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation.
+
+## Authors’ contributions
+
+MD, KK, CC, and MR conceived the study.
+KC and MD performed the experiments.
+MD primarily wrote the manuscript.
+KK, MR, CC, and SB participated in manuscript writing.
+All authors read and approved the final manuscript.
+
 # Methods
@@
 
-## Availability of data and materials
-
-The code and the data supporting the conclusions of this article are available in the https://github.com/dozmorovlab/disease_coherence GitHub repository.
-
-# Abbreviations
-
-GWAS - Genome-Wide Association Study; KEGG - Kyoto Encyclopedia of Genes and Genomes; MSigDb - Molecular Signatures Database; PPI - Protein-Protein Interaction; SNP - Single Nucleotide Polymorphism
-
-# Acknowledgements
-
-This work was supported by a Genetics and Human Agency Award from the Templeton Foundation to KSK, MR, CC, SB.
-The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation.
-
-## Competing interests
-
-The authors declare that they have no competing interests.
-
-## Authors’ contributions
-
-MD, KK, CC, and MR conceived the study.
-KC and MD performed the experiments.
-MD primarily wrote the manuscript.
-KK, MR, CC, and SB participated in manuscript writing.
-All authors read and approved the final manuscript.
-
 # Additional files
@@
 
-# References
-
 <!--
@@
 -->
+
+# Figure legends
+
+**Figure 1.
+Degree distributions are affected by network size and configuration.** Degree distributions of A) ring networks and B) fully connected networks containing 10 and 20 nodes.
+While degree distributions of similarly coherent ring networks can be directly compared, degree distributions of fully connected networks depend on network size.
+
+**Figure 2. Network size is inversely associated with coherence.** Slopes of internal vs. external degree distributions (aka coherence) for KEGG and random networks are plotted against network size.
+
+**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+
+**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
+
+# Tables
+
+<!--Tables/table_1.csv-->
+**Table 1.
+Summary statistics of the analyzed categories.** Average coherence estimates are shown.
+"NA" values indicate that a category lacked phenotypes with a sufficient number of PPIs.
+Sorted by "String Filtered" average coherence.
+
+| Category               | Number of SNPs | Number of Genes | Total Diseases | Mean Biogrid Coherence | Mean String Coherence | Mean String Filtered Coherence |
+|------------------------|:--------------:|:---------------:|:--------------:|:----------------------:|:---------------------:|:------------------------------:|
+| Trait                  |                |                 |                |                        |                       |                                |
+| Eye trait              |       158      |       109       |        5       |           NA           |         0.724         |              1.050             |
+| Metabolic trait        |       790      |       646       |       10       |          0.738         |         0.779         |              0.863             |
+| Anthropometric trait   |      2090      |       1771      |       15       |          0.460         |         0.637         |              0.691             |
+| Cardiovascular trait   |       728      |       604       |       15       |          0.555         |         0.675         |              0.625             |
+| Intelligence           |       181      |       177       |        4       |          0.180         |         0.395         |              0.405             |
+| Disease                |                |                 |                |                        |                       |                                |
+| Metabolic disease      |       331      |       269       |        4       |          0.470         |         0.825         |              0.907             |
+| Cancer                 |       670      |       570       |       15       |          0.490         |         0.689         |              0.723             |
+| Neurologic             |       372      |       333       |        7       |           NA           |         0.493         |              0.718             |
+| Autoimmune             |      1383      |       1184      |       18       |          0.425         |         0.820         |              0.701             |
+| Eye disease            |       229      |       202       |        5       |          0.330         |         0.630         |              0.663             |
+| Cardiovascular disease |       301      |       259       |        7       |          0.380         |         0.694         |              0.660             |
+| Psychiatric            |       678      |       611       |       11       |          0.540         |         0.552         |              0.642             |
+
+# References

### 2020-02-10 manuscript/submission.BMC_Bioinformatics/manuscript_BMC_Bioinformatics_R1.Rmd
Start revision by adding modularity results/discussion/methods description

--- 2019-12-17 manuscript/submission.BMC_Bioinformatics/manuscript_BMC_Bioinformatics.Rmd
@@
 
+<!--New start-->
+## Coherence captures distinct network properties than modularity
+
+Our study considers a network of phenotype-associated genes in its entirety.
+However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa].
+Such subnetworks, while highly coherent by themselves, would have relatively low coherence overall due to the limited number of interactions among them (canonical definition of subnetwork).
+Consequently, a network with low coherence may consist of several modules and have high modularity.
+In turn, a network with high coherence may be represented by a single module and, consequently, have low modularity.
+Our assessment of network modularity using STRING and STRING filtered PPIs revealed negative association of modularity with network size, that is, the larger the network is, the less modular it becomes (Additional File 3, 4).
+We also did not observe a measurable association between modularity and coherence.
+Notably, network modularity using BioGrid PPIs showed a moderate positive association with network size and coherence, implying that, as networks get larger, they become more modular and yet highly coherent.
+Investigation of individual phenotype-specific networks built using BioGrid PPIs revealed they are expectedly smaller than networks build using STRING PPIs, making modularity estimates imprecise.
+In summary, our results did not reveal a consistent association between coherence and modularity, suggesting these metrics capture distinct network properties.
+
+Although our results did not show an association between coherence and modularity, our study is limited by using a single most common modularity measure [@Clauset_2004].
+Although numerous methods for module detection have been described [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section.
+Our future direction includes the development of a strategy for normalization of modularity by network size and comparing it with normalized coherence.
+
+## Modularity estimates
+
+Network modularity was estimated using the standard method developed by Clauset et al. [@Clauset_2004], implemented in the igraph::modularity function.
+<!--New end-->
+
 # Discussion
@@
 <!--
-## Single networks, or networks of modules?
-
-Our study considers a network of disease-associated genes in its entirety.
-However, networks may be comprised of interacting modules, i.e., relatively isolated subnetworks [@Hartwell:1999aa].
-Connectivity of such modules would be different from the overall network connectivity [@Ghiassian:2015aa].
-Although numerous methods exist for module detection in a network [@Chen:2008aa; @Dobrin:2009aa; @Jia:2011aa; @Agrawal:2018aa], comparing modules between diseases is plagued by normalization issues described in the Results section Intuitively, large networks will have a higher chance to have subnetworks (modules), and vice versa.
-It remains to be determined whether disease-associated genes form modular networks and whether the modular structure differs between (categories of) diseases.
-
 ## Tissue specificity

### 2020-02-25 manuscript/submission.BMC_Bioinformatics/Revision1/manuscript_BMC_Bioinformatics_R1.Rmd
Make repository for revised files

--- 2020-02-10 manuscript/submission.BMC_Bioinformatics/manuscript_BMC_Bioinformatics_R1.Rmd
@@
 
+<!-- New in Revision 1 -->
+## Networks with high/low coherence have distinct degree distributions
+
+Our measure of network coherence is expected to be proportional to the level of internal and/or external degree distributions.
+To investigate the relationship between coherence and degree distributions, we selected phenotype networks with low and high coherence.
+The thresholds for low and high normalized coherence were selected as <0.58 and >0.85, corresponding to the first and third quartiles of the coherence range, respectively.
+The Biogrid-generated networks were omitted because they had insufficient number of genes to form well-defined degree distributions.
+Expectedly, the internal degree distributions for high- and low-coherence networks were significantly different from random (Bonferroni-corrected two-tailed Wilcoxon p-values < 5.30E-12), and these results were consistent when using different PPI databases (Figure 3, Additional File 9).
+Similarly, the internal degree distribution for high-coherence networks was significantly larger than that of low-coherence networks (p-value = 2.01E-16/2.12E-4 for STRING/STRING filtered databases, respectively).
+Similar observations were observed for external degree distributions of high- and low-coherence networks (Figure 3, Additional File 9).
+In summary, these results support the notion that networks of high coherence have significantly more internal and external interactions as compared with low-coherence and random networks.
+<!--
++-------+
+| ![](Figures/Fig3_degree_distributions.png)      |
+| **Figure 3.
+Degree distributions differ for high/low-coherence networks.** Internal and external degree distributions for high- and low-coherence networks, and size-matched random networks.
+STRING and STRING Filtered indicate the corresponding PPI database.
+Bonferroni-corrected Wilcoxon p-values comparing the distributions are in Additional File 9.
+|
++-------+
+-->
+<!-- End New in Revision 1 -->
+
 ## Metabolic and intelligence-related traits as examples of networks with overall highest/lowest coherence
 
-We estimated coherence of 49 traits from five categories (Figure 3, Additional File 3, Additional File 5, 6).
+We estimated coherence of 49 traits from five categories (Figure 4, Additional File 3, Additional File 5, 6).
 Among anthropometric traits, Weight and Body Mass Index had the highest level of coherence.
@@
 | ![](Figures/Fig3_summary_traits.png)      |
-| **Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).   |
+| **Figure 4. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).   |
 +-------+
@@
 
-We further estimated coherence of 67 diseases from seven categories (Figure 4, Additional File 7, 8, Additional File 3).
+We further estimated coherence of 67 diseases from seven categories (Figure 5, Additional File 7, 8, Additional File 3).
 Metabolic diseases had high overall coherence (average coherence 0.91, Table 1), with "Obesity" and "Type 2 diabetes" being the most coherent (1.01 and 0.91, respectively).
@@
 | ![](Figures/Fig4_summary_diseases.png)      |
-| **Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).   |
+| **Figure 5. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).   |
 +-------+
@@
 
-<!--New start-->
+<!-- New in Revision 1 -->
 ## Coherence captures distinct network properties than modularity
@@
 Our future direction includes the development of a strategy for normalization of modularity by network size and comparing it with normalized coherence.
-
-## Modularity estimates
-
-Network modularity was estimated using the standard method developed by Clauset et al. [@Clauset_2004], implemented in the igraph::modularity function.
-<!--New end-->
+<!-- End New in Revision 1 -->
 
@@
 It provides additional protection against cases where coherence may appear high, yet, it may be a consequence of small network size providing unstable degree distributions.
-Testing the coherence estimates for significance helps against conclusions of the overall low coherence (e.g., anthropometric traits, Figure 3), when the coherence of individual traits is high (e.g., individual anthropometric traits, Figure 5B).
+Testing the coherence estimates for significance helps against conclusions of the overall low coherence (e.g., anthropometric traits, Figure 4), when the coherence of individual traits is high (e.g., individual anthropometric traits, Figure 5B).
 Coupled with the permutation test, our coherence estimation provides a robust framework for quantifying the coherence estimates of phenotype-associated genes, advancing our understanding of the corresponding molecular mechanisms.-->
@@
 -->
+<!-- New in Revision 1 -->
+## Modularity estimates
+
+Network modularity was estimated using the standard method developed by Clauset et al. [@Clauset_2004], implemented in the igraph::modularity function.
+<!--New end-->
 
@@
 
+<!-- New in Revision 1 -->
+**Additional File 9.
+Bonferroni-corrected Wilcoxon p-values comparing the internal and external degree distributions of phenotype-specific networks having high (>0.85) and low (<0.58) normalized coherence.** Each sheet corresponds to the type of degree distribution (internal and external), and the type of PPI database (STRING and STRINGfilt).
+<!-- End New in Revision 1 -->
 <!--
@@
 
-**Figure 3. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+<!-- New in Revision 1 -->
+**Figure 3.
+Degree distributions differ for high/low-coherence networks.** Internal and external degree distributions for high- and low-coherence networks and size-matched random networks.
+STRING and STRING Filtered indicate the corresponding PPI database.
+Bonferroni-corrected Wilcoxon p-values comparing the distributions are in Additional File 9.
+<!-- End New in Revision 1 -->
 
-**Figure 4. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
+**Figure 4. Coherence estimates of traits.** Size of dots represents the level of normalized coherence (X-axis) for individual traits (Y-axis).
+
+**Figure 5. Coherence estimates of diseases.** Size of dots represents the level of normalized coherence (X-axis) for individual diseases (Y-axis).
 

### 2020-02-25 Analysis/08_Generate_Internal_Distributions.Rmd
Add degree comparison analysis

---
title: "Density_Plots"
author: "Kellen Cresswell"
date: "February 22, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}
require(dplyr)
require(ggplot2)
library(readr)
source("functions/GeneSample.R")
library(cowplot)
library(gridExtra)
```

```{r}
#Repeating for string

data_dir = "./data/"

#Read in new string database

string = read_table2(paste0(data_dir,"9606.protein.links.v11.0.txt.gz"))

string = string %>% mutate(protein1 = gsub("9606.", "", protein1), 
                             protein2 = gsub("9606.", "", protein2))

#Read in ensemble protein ID

ensembl = biomaRt::useEnsembl(biomart="ensembl", dataset="hsapiens_gene_ensembl", mirror = "www")

#Get gene set
genes = biomaRt::getBM(attributes=c('ensembl_peptide_id','hgnc_symbol'), mart = ensembl)

#Join gene names with string

string = left_join(string, genes, by = c("protein1" = "ensembl_peptide_id", "protein2" = "ensembl_peptide_id"))

#Rename to work with our functions

string = string %>% dplyr::rename(hgnc_symbol_a = hgnc_symbol)

#Repeat for second set of proteins

string = left_join(string, genes, by = c("protein2" = "ensembl_peptide_id"))

string = string %>% dplyr::rename(hgnc_symbol_b = hgnc_symbol)

string_filt = string %>% filter(combined_score>=500)

dim(string)
```

```{r}
disease_overall = readRDS("biogrid_edges_new.rds")
disease_overall_string = readRDS("string_edges_new.rds")
disease_overall_string_filt = readRDS("string_filt_edges_new.rds")
```

```{r}
coherence = readRDS("data/Coherence_Results/coherence_biogrid.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
coherence_string = readRDS("data/Coherence_Results/coherence_string.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
coherence_string_filt =  readRDS("data/Coherence_Results/coherence_string_filt.rds") %>% dplyr::mutate(High = ifelse(Coherence>.85, "High", ifelse(Coherence<.58, "Low", "Neither"))) %>% dplyr::select(Disease, High)
```

```{r}
counts_biogrid = disease_overall %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())

counts_string = disease_overall_string %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())

counts_string_filt = disease_overall_string_filt %>% group_by(Disease) %>% filter(Internal !=0) %>% summarise(Count = n())
```

```{r}
disease_overall = left_join(disease_overall, coherence) %>% dplyr::filter(High %in% c("High", "Low"))
disease_overall_string = left_join(disease_overall_string, coherence_string)  %>% dplyr::filter(High %in% c("High", "Low"))
disease_overall_string_filt = left_join(disease_overall_string_filt, coherence_string_filt)  %>% dplyr::filter(High %in% c("High", "Low"))
```

```{r}
disease_overall = left_join(disease_overall, counts_biogrid)
disease_overall_string = left_join(disease_overall_string, counts_string)
disease_overall_string_filt = left_join(disease_overall_string_filt, counts_string_filt)

```

```{r}
#Generate random networks equivalent to high coherence (STRING)
random_string_high = bind_rows()

for (i in unique((disease_overall_string %>% filter(High == "High"))$Count)) {
 diseases_curr = GeneSample(i, string, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_high = bind_rows(diseases_curr, random_string_high)
}

#Generate random networks equivalent to low coherence (STRING)
random_string_low = bind_rows()

for (i in unique((disease_overall_string %>% filter(High == "Low"))$Count)) {
 diseases_curr = GeneSample(i, string, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_low = bind_rows(diseases_curr, random_string_low)
}

#Generate random networks equivalent to high coherence (STRING Filtered)
random_string_filt_high = bind_rows()

for (i in unique((disease_overall_string_filt %>% filter(High == "High"))$Count)) {
 diseases_curr = GeneSample(i, string_filt, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_filt_high = bind_rows(diseases_curr, random_string_filt_high)
}

#Generate random networks equivalent to low coherence (STRING Filtered)
random_string_filt_low = bind_rows()

for (i in unique((disease_overall_string_filt %>% filter(High == "Low"))$Count)) {
 diseases_curr = GeneSample(i, string_filt, directed = FALSE)
 diseases_curr = data.frame(Genes = names(diseases_curr$internal_connectivity), Internal = (diseases_curr$internal_connectivity), External = (diseases_curr$external_connectivity), Disease = "Random", Category = "Random", High = "Random", Count = i)
 diseases_curr  = diseases_curr %>% filter(Internal != 0)
  random_string_filt_low = bind_rows(diseases_curr, random_string_filt_low)
}

#Adding random to string and string filtered
disease_overall_string = bind_rows(disease_overall_string, random_string_high, random_string_low)

disease_overall_string_filt = bind_rows(disease_overall_string_filt, random_string_filt_high, random_string_filt_low)
```

```{r}
#Performing wilcox test of internal edges
string_wilcox = pairwise.wilcox.test(disease_overall_string$Internal, disease_overall_string$High, "bonferroni")
string_wilcox_filt = pairwise.wilcox.test(disease_overall_string_filt$Internal, disease_overall_string_filt$High, "bonferroni")

#Performing wilcox test of external edges
string_wilcox_ex = pairwise.wilcox.test(disease_overall_string$External, disease_overall_string$High, "bonferroni")
string_wilcox_filt_ex = pairwise.wilcox.test(disease_overall_string_filt$External, disease_overall_string_filt$High, "bonferroni")

#Performing KS test of internal edges
# source("https://raw.githubusercontent.com/happyrabbit/DataScienceR/master/R/pairwise_ks_test.R")
# string_ks = pairwise_ks_test(disease_overall_string$Internal, disease_overall_string$High)
# string_ks_filt = pairwise_ks_test(disease_overall_string_filt$Internal, disease_overall_string_filt$High)
# string_ks
# string_ks_filt

# Save in Supplementary Table S4 that will be Additional File 9
x <- list(data.frame(Coherence = c("Low", "Random"), string_wilcox$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_filt$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_ex$p.value), 
          data.frame(Coherence = c("Low", "Random"), string_wilcox_filt_ex$p.value))
names(x) <- c("STRING_internal", "STRINGfilt_internal", "STIRNG_external",  "STIRNGfilt_external")
writexl::write_xlsx(x, path = "../manuscript/Tables/supplementary_table_S4.xlsx")
```

```{r}
#Internal edges 
# gg_biogrid = ggplot(disease_overall %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.1) + labs(fill = "Coherence",x = "Internal Edges")
# plot(gg_biogrid)
# ggsave(gg_biogrid + theme_bw(20), file= './figures/Internal_Distribution_Biogrid.png', units = "in", width = 3.5, height = 2.5, scale = 2.8)

gg_string = ggplot(disease_overall_string %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "Internal Edges", y = "Density") + xlim(0, 75) + theme_cowplot() + theme(legend.position = c(0.6, 0.6)) + ggtitle("STRING") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
plot(gg_string)
# ggsave(gg_string, file= './figures/Internal_Distribution_String.png', units = "in", width = 4, height = 3)

gg_string_filt = ggplot(disease_overall_string_filt %>% dplyr::filter(Internal != 0), aes(x = Internal, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "Internal Edges", y = "Density") + xlim(0, 25) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING Filtered") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# plot(gg_string_filt)
# ggsave(gg_string_filt, file= './figures/Internal_Distribution_String_Filt.png', units = "in", width = 4, height = 3)

#External edges
gg_string_ex = ggplot(disease_overall_string, aes(x = External, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "External Edges", y = "Density") + xlim(0, 4000) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# print(gg_string_ex)
# ggsave(gg_string_ex, file= './figures/External_Distribution_String.png', units = "in", width = 4, height = 3)

gg_string_filt_ex = ggplot(disease_overall_string_filt, aes(x = External, fill = High)) + geom_density(alpha = 0.3) + labs(fill = "Coherence", x = "External Edges", y = "Density") + xlim(0, 500) + theme_cowplot() + theme(legend.position = "none") + ggtitle("STRING Filtered") + theme(plot.title = element_text(size=12, hjust = 0.5, margin = margin(t = 10, b = -10)))
# print(gg_string_filt_ex)
# ggsave(gg_string_filt_ex, file= './figures/External_Distribution_String_Filt.png', units = "in", width = 4, height = 3)

png(filename = "../manuscript/Figures/Fig3_degree_distributions.png", height = 5, width = 7, units = "in", res = 300)
lay1 <- rbind(c(1, 2))
lay2 <- rbind(c(1, 2))
g1 <- arrangeGrob(grobs = list(gg_string, gg_string_filt), layout_matrix = lay1)
g2 <- arrangeGrob(grobs = list(gg_string_ex, gg_string_filt_ex), layout_matrix = lay2)
grid.arrange(g1, g2, ncol = 1)
dev.off()
```

### 2020-02-25 manuscript/submission.BMC_Bioinformatics/Revision1/Response_R1.Rmd
Add Response

---
output:
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NSF_grant_style.docx
  html_document: default
---

Reviewer reports:

> This manuscript introduces a method, "network coherence", and applies it to human disease genes (from GWAS); in addition to introducing a new method, they make a claim that the coherence measure varies among specific disorders in a biologically significant manner. As such this work needs to be evaluated in these two aspects: First, whether the method is novel and sufficiently well-validated. Second, whether the data support the claims about human phenotypes.

> Taking up the second aspect first, there is a major problem with using PPINs from BioGRID and STRING: these networks are extremely biased by "well-studied genes" or other types of artifacts. This means using them to make conclusions about anything biological is horribly confounded with the effects of human decisions about which proteins to use as targets among other issues. What the authors have failed to do is provide any evidence that their results are not simply due to such biases. Note that making random networks does not solve this problem - it is indeed "significant" that genes like TGFB have a lot of interactions, but it's far from clear that's due to biology. I believe all of the papers the authors cite in support of the idea that coherence (et al.) is a documented phenomenon associated with human diseases (pg 3 line 20) all have this issue, and the point of ref 27 has not, to my knowledge, been rebutted.

We agree that the PPI databases are potentially biased towards well-studied genes. Yet, databases like STRING try to avoid it by aggregating PPI information from multiple sources, including the predicted interactions. We intentionally used the full STRING database to maximally avoid PPI bias towards the well-studied genes. Yet, at this state of the knowledge, PPI databases remain incomplete (c.f. Menche et al. 2015). This was the rationale to mention the incompleteness of the databases throughout the manuscript. E.g., we discussed that our method "... will only improve as more PPI information becomes available", and devoted the whole paragraph explaining the limitations of PPI databases.

While we acknowledge the incompleteness of the PPI databases and GWASs, the fundamental question is why some genes are studied much more than others? Historically, such genes emerge as a "low hanging fruit" in molecular biology experiments, consistently appear in GWAS studies, that is, representing the true biological signal. Consequently, they formed the backbone for forming the PPI databases, for building sets of GWAS loci. While we acknowledge the fact of bias, the source of it is likely biology-driven.

The bias and incompleteness of the PPI databases led us to use the permutation test. Both PPI databases and GWASs attempt to extract biological signal genome-wide. The fact that, e.g., TGFB got picked up by a GWAS study, and is highly represented in a PPI database is likely biology-driven. The fact that random networks have much less chance to pick up TGFB is due to the random, non-biological processes. The permutation test works in the same realm of PPIs that were used to build phenotype-specific networks. Being non-parametric, the permutation test equally treats genes with average degree distributions and those with extreme PPIs. If we had used parametric tests, chances of getting biased results would be high as outliers typically break the rules of parametric distributions leading to overinflated results.

> I have to question the limitation to GWAS loci. The authors concede that in effect their analysis cannot be used to conclude much of anything, due to gaps in knowledge about the genetics of the phenotypes (line 3 pg 14). Since the authors are aware of this problem, that they didn't use genes implicated across the spectrum of variant frequencies and effect sizes (including somatic variants in cancer) which are invisible to GWAS is surprising. So I wholeheartedly agree with their caveat, but the issue with the networks used is just as important. Overall, the results of the application of the method are very unconvincing to me.

We fully agree with the reviewer and intentionally chose our wording very carefully to highlight caveats in the current state of knowledge of both PPI databases and GWAS loci. Additionally, we carefully selected references highlighting the whole spectrum of current knowledge and views, e.g., intentionally including reference 27 as an alternative to the current state of knowledge supported by several other peer-reviewed references. Our goal was to present the method, the approach that can be used to compare molecular coherence of phenotypes while overcoming limitations of network analysis (differences in size, configuration). It was beyond the scope of our work to evaluate genes associated with rare variants, different effect sizes. The method itself can be applied to any GWAS results (see Discussion), including rare and somatic mutations, and reveal how coherent the associated gene networks are. We intentionally used the state-of-the-art list of GWAS results, and acknowledge the limitations of the current knowledge.

> None of this detracts, in principle, from the presentation of the method, since a method can exist and be novel in the absence of any useful or interesting application as these may emerge in the future. But as this is a bioinformatics paper, the standard format is to compare the results obtained on a gold standard, and to pit the method against other approaches either in reference to the gold standard (not that one could be defined here, except perhaps by simulations), or at least to the results obtained in actual use and describe the advantages, or at least differences, of the method to other approaches. The authors have done none of those things.

Indeed, the primary intention of our work is to present a principled method allowing for a comparison of the coherence of molecular mechanisms across phenotypes. While we are fully aware of the typical strategy for presenting novel bioinformatics methods (e.g., see https://www.biorxiv.org/content/10.1101/549170v3 for an example of such presentation), the problem here is that there are no other similar approaches that we are aware of. Nor is there a biological gold standard other than phenotype-specific gene networks used in our work. As described in the manuscript, random and KEGG networks are the best approximators of random and high coherence. The permutation framework is the most unbiased simulation strategy, in which we simulate random networks of the same size as phenotype-specific networks to get the significance of coherence estimates.

> In terms of novelty, as the authors discuss, there are other measures of coherence, and one that is particularly relevant is modularity - various measures of "cluster tightness" relative to a network context go back many decades. Since modularity is literally the measure of the ratio of the density of in-group and to-group edges, and the authors repeatedly described modularity in just those terms (e.g. on pg 13), it isn't clear that method of the authors is even needed.

We performed additional analyses of the association between modularity and coherence. Our results did not reveal a consistent association between coherence and modularity. We added the corresponding Results and Methods sections, discussed our observations, and updated the Additional Files 3 and 4 with modularity information. Again, our goal was to develop a general-purpose method for estimating and comparing coherence of molecular mechanisms across phenotypes, with the coherence normalization approach being one of the novel aspects. Normalization of modularity, while theoretically possible, would defy interpretability. Our results demonstrate that coherence is different from modularity. We hope that the simplicity, flexibility, and interpretability of our measure of coherence will be an important contribution to the field of network analysis in genomics.

> It is striking that the authors never provide an actual example of the node degree distributions for different phenotypes, since they form the basis for the regression analysis. As it stands, the novelty here seems minor, and directly comparing their method to other (including simpler) alternatives is essential.

We agree with the need to investigate and compare degree distributions. We added the results section and the new Figure 3 showing internal and external degree distributions for high- and low-coherence networks, and statistically compared them with size-matched random networks. Additional File 9 shows the results of the statistical comparison. In summary, the new results confirm the notion that networks of high coherence have a significantly larger number of internal and external connections than low-coherence networks and random networks.

Regarding the notion of comparison with other alternatives, please, see our response to the previous comment.

> Another major drawback of the approach as presented is that there is no estimate of variance in the measure, that is, coherence should be treated as a random variable. In order to say that phenotype X has a higher coherence than phenotype Y, one would like to know the distributional properties of the point estimates, as I'm sure the authors are aware. It is not the same as saying a coherence value is different from random. The authors should provide confidence intervals etc. or remove all claims that one set of phenotypes has a "lower coherence" than another. The closest the authors get is comparing STRING and BioGRID (the authors do not mention their lack of independence; these databases overlap extensively). And the comparison is very qualitative anyway, for example saying on p9 line 22 that the "overall correlation remained high" without any numbers given.

### 2020-03-04 Analysis/03_Normalized_Coherence.Rmd
Delete temp_rea_bio.txt after completion
integrate ci code

@@
 
+# Define function for calculating confidence intervals from edges
+
+```{r}
+ci_from_edges = function(edges) {
+  edges %>% filter(Internal != 0)  %>% 
+    group_by(Disease, Category) %>% mutate(Count = n()) %>% group_by(Disease, Category, Count) %>%
+    do({
+      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+      data.frame(Lower = confint(mod)[1], Upper = confint(mod)[2])
+    }) %>% arrange(Lower) %>% filter(Count > num_genes)
+}
+```
+
+The above function takes internal and external edges from each gene for each disease and calculates the 95% confidence interval for coherence.
+Genes with zero internal edges are filtered and diseases with less than `r num_genes` non-zero internal edges are filtered.
+
 # Get edges for diseases from Biogrid (Connectivity_Comparison.Rmd)
@@
 dis_slopes_biogrid_count = slopes_from_edges(disease_overall_biogrid)
-
+dis_slopes_biogrid_ci = ci_from_edges(disease_overall_biogrid)
 # String
@@
 dis_slopes_string_count = slopes_from_edges(disease_overall_string)
+dis_slopes_string_ci = ci_from_edges(disease_overall_string)
 
@@
 dis_slopes_string_filt_count = slopes_from_edges(disease_overall_string_filt)
+dis_slopes_string_filt_ci = ci_from_edges(disease_overall_string_filt)
 ```
@@
 # [??? What this function does? Write roxygen help]
-calc_norm_coherence = function(msig_frame, dis_frame, database, num_msig = 10, num_rand = 10) {
+calc_norm_coherence = function(msig_frame, dis_frame, database, num_msig = 10, num_rand = 10, ci = FALSE) {
   # Splitting into list by categories 
@@
 
+      if (!ci) {
       # Calculate coherence
@@
       coherence_tab = bind_rows(coherence, coherence_tab)
+
+  } else {
+  #Lower interval for coherence
+      coherence_low = (dis$Upper - rand_slope)/(mod_slope-rand_slope)
+      #Upper interval for coherence
+      coherence_up = (dis$Lower - rand_slope)/(mod_slope-rand_slope)
+
+      coherence = data.frame(Disease = dis$Disease, Coherence_Lower = coherence_low, Coherence_Upper = coherence_up,Msig_Norm = mod_slope, Rand_Norm = rand_slope, Count = dis$Count)
+
+      coherence_tab = bind_rows(coherence, coherence_tab)
+
+  }
 }
@@
 saveRDS(norm_coh_string_filt, "./data/Coherence_Results/coherence_string_filt.rds")
+```
+
+# Calculating CIs for normalized coherence
+
+```{r}
+# Calculating normalized coherence 
+norm_coh_biogrid_ci = calc_norm_coherence(msig_frame_biogrid, dis_slopes_biogrid_ci, database = biogrid, num_msig = num_msig, num_rand = num_rand, ci = TRUE)
+norm_coh_string_ci = calc_norm_coherence(msig_frame_string, dis_slopes_string_ci, database = string, num_msig = num_msig, num_rand = num_rand, ci = TRUE)
+norm_coh_string_filt = calc_norm_coherence(msig_frame_string_filt, dis_slopes_string_filt_ci, database = string_filt, num_msig = num_msig, num_rand = num_rand, ci = TRUE)
+
+#Saving to .rds files
+saveRDS(norm_coh_biogrid, "./data/Coherence_Results/coherence_biogrid_ci.rds")
+saveRDS(norm_coh_string, "./data/Coherence_Results/coherence_string_ci.rds")
+saveRDS(norm_coh_string_filt, "./data/Coherence_Results/coherence_string_filt_ci.rds")
 ```
@@
 saveRDS(string_random, "./data/Coherence_Results/string_random_edges.rds")
+
+# Small cleanup
+unlink("temp_rea_bio.txt")
 ```

### 2020-03-11 Analysis/03_Normalized_Coherence.Rmd
add code for msig slopes (still running)

@@
 # Reading in Biogrid data
-biogrid = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt.gz")
+biogrid = read_tsv("./data/BIOGRID-ALL-3.5.174.mitab.txt")
 biogrid = biogrid %>% filter(`Taxid Interactor A` == "taxid:9606" & `Taxid Interactor B` == "taxid:9606")
@@
 
+```{r}
+# [??? What this function does? Write roxygen help]
+
+get_msig_ci = function(sub_cats, database) {
+  require(msigdbr)
+  msig = msigdbr()
+
+  # Subset to include msig
+  msig = msig %>% filter(gs_subcat %in% sub_cats)
+
+  # Create empty list
+  mod_over = list()
+  # Set index for list entries
+  index = 1
+  # Loop through each pathway
+  for (j in unique(msig$gs_name)) {
+    #Subset to only include current pathway
+    curr_msig = msig %>% filter(gs_name == j)
+    #Write to table in format for disease2frame
+    write.table(curr_msig$gene_symbol, file = "temp_rea_bio.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
+    #Get internal/external connectivity
+    diseases_curr = diseases2frame("temp_rea_bio.txt", links = database)
+    #Get disease name
+    diseases_curr = diseases_curr %>% mutate(Disease = j)
+    mod_over[[index]] = diseases_curr
+    index =  index+1
+  }
+
+  # Bind together list
+  mod_over = bind_rows(mod_over)
+
+  # Applying categories
+  msig_cats  = msig %>% 
+    dplyr::select(Disease = gs_name, Category = gs_subcat) %>% 
+    distinct() #Get categories in order
+
+  # Joining categories
+  mod_over = left_join(mod_over, msig_cats)
+
+  #Getting gene counts
+  mod_over = mod_over %>% group_by(Disease) %>% filter(Internal != 0) %>% mutate(Num_Genes = n())
+
+  #Getting slopes from each diseases
+  mod_over = mod_over %>% group_by(Disease, Category, Num_Genes) %>%
+    do({
+      mod = lm(sqrt(External) ~ sqrt(Internal)-1, data = .)
+      data.frame(Lower = confint(mod)[1], Upper = confint(mod)[2])
+    }) 
+  #Rename x (Default do tidy result) to Slope
+  return(mod_over)
+}
+```
+
 # Getting slope between internal/external edges from GOCC, KEGG and REACTOME pathways
@@
 saveRDS(msig_frame_string_filt, "String_Filt_Msig_Slopes.rds")
+```
+
+# Getting 95% between internal/external edges from GOCC, KEGG and REACTOME pathways
+
+```{r}
+# Getting biogrid msig slopes
+msig_frame_biogrid_ci = get_msig_ci(sub_cats = subcats, database = biogrid)
+
+# Getting string msig slopes
+msig_frame_string_ci = get_msig_ci(sub_cats = subcats, database = string)
+
+# Getting string filtered msig slopes
+msig_frame_string_filt_ci = get_msig_ci(sub_cats = subcats, database = string_filt)
+
+# Saving the files
+saveRDS(msig_frame_biogrid_ci, "Biogrid_Msig_CI.rds")
+saveRDS(msig_frame_string_ci, "String_Msig_CI.rds")
+saveRDS(msig_frame_string_filt_ci, "String_Filt_Msig_CI.rds")
 ```

### 2020-03-15 Analysis/03_Normalized_Coherence.Rmd
confidene interval table

@@
 
-# Getting 95% between internal/external edges from GOCC, KEGG and REACTOME pathways
-
-```{r}
-# Getting biogrid msig slopes
-msig_frame_biogrid_ci = get_msig_ci(sub_cats = subcats, database = biogrid)
-
-# Getting string msig slopes
-msig_frame_string_ci = get_msig_ci(sub_cats = subcats, database = string)
-
-# Getting string filtered msig slopes
-msig_frame_string_filt_ci = get_msig_ci(sub_cats = subcats, database = string_filt)
-
-# Saving the files
-saveRDS(msig_frame_biogrid_ci, "Biogrid_Msig_CI.rds")
-saveRDS(msig_frame_string_ci, "String_Msig_CI.rds")
-saveRDS(msig_frame_string_filt_ci, "String_Filt_Msig_CI.rds")
-```
-
 # Getting all msig pathway slopes for string (Used in Supplementary Table 2)
@@
 dis_slopes_string_filt_ci = ci_from_edges(disease_overall_string_filt)
+```
+
+# Merging together the confidence intervals
+
+```{r}
+#Renaming names before joining and manipulating
+
+#BIOGRID
+dis_slopes_biogrid_ci = dis_slopes_biogrid_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`Biogrid Lower` = Lower, `Biogrid Upper` = Upper)
+
+#STRING
+dis_slopes_string_ci = dis_slopes_string_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`String Lower` = Lower, `String Upper` = Upper)
+
+#STRING Filtered
+dis_slopes_string_filt_ci = dis_slopes_string_filt_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`String Filt Lower` = Lower, `String Filt Upper` = Upper)
+
+#Biogrid and STRING
+ci = full_join(dis_slopes_biogrid_ci, dis_slopes_string_ci, by = c("Disease", "Category"))
+
+#Merging in STRING Filtered
+ci = full_join(ci, dis_slopes_string_filt_ci, by = c("Disease", "Category"))
+
+write.csv(ci, 'Confidence_Intervals.csv', quote = FALSE)
 ```

### 2020-03-17 Analysis/03_Normalized_Coherence.Rmd
Add raw slopes in addition to their CIs

--- 2020-03-11 Analysis/03_Normalized_Coherence.Rmd
@@
 dis_slopes_string_filt_ci = ci_from_edges(disease_overall_string_filt)
+```
+
+# Merging together the confidence intervals
+
+```{r}
+#Renaming names before joining and manipulating
+
+#BIOGRID
+dis_slopes_biogrid_count = dis_slopes_biogrid_count %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`Biogrid Slope` = Slope)
+dis_slopes_biogrid_ci = dis_slopes_biogrid_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`Biogrid Lower` = Lower, `Biogrid Upper` = Upper)
+dis_slopes_biogrid <- full_join(dis_slopes_biogrid_count, dis_slopes_biogrid_ci, by = c("Disease", "Category"))
+
+#STRING
+dis_slopes_string_count = dis_slopes_string_count %>% dplyr::rename(`String Slope` = Slope)
+dis_slopes_string_ci = dis_slopes_string_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`String Lower` = Lower, `String Upper` = Upper)
+dis_slopes_string <- full_join(dis_slopes_string_count, dis_slopes_string_ci, by = c("Disease", "Category"))
+
+#STRING Filtered
+dis_slopes_string_filt_count = dis_slopes_string_filt_count %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`String Filt Slope` = Slope)
+dis_slopes_string_filt_ci = dis_slopes_string_filt_ci %>% ungroup() %>% dplyr::select(-Count) %>% dplyr::rename(`String Filt Lower` = Lower, `String Filt Upper` = Upper)
+dis_slopes_string_filt <- full_join(dis_slopes_string_filt_count, dis_slopes_string_filt_ci, by = c("Disease", "Category"))
+
+#Biogrid and STRING
+ci = full_join(dis_slopes_biogrid, dis_slopes_string, by = c("Disease", "Category"))
+
+#Merging in STRING Filtered
+ci = full_join(ci, dis_slopes_string_filt, by = c("Disease", "Category"))
+
+write.csv(ci, '../manuscript/Tables/Confidence_Intervals.csv', quote = FALSE)
 ```

### 2020-03-18 Analysis/06_Analysis_Tables.Rmd
Add correlation of slopes and CIs with network sizes

@@
     toc: yes
+editor_options: 
+  chunk_output_type: console
 ---
@@
 fileNameIn0  <- "../manuscript/Tables/supplementary_table_S3.csv"   # diseases and traits, coherence new KEGG, REACTOME, CC. 2019-07-14, Kellen
+fileNameIn1 <- "../manuscript/Tables/Confidence_Intervals.csv" # Confidence intervals, generated in 03_Normalized_Coherence 2020-03-17
 # iconv -f UTF-8 -t ISO-8859-1 -c supplementary_table_S3.csv > tmp.csv
@@
 rownames(mtx0) <- mtx0$Disease
+```
+
+### Append confidence intervals
+
+```{r}
+# Should be done one time
+mtx <- read_csv(fileNameIn1)
+mtx0 <- left_join(mtx0, mtx, by = c("Disease", "Category"))
+write_csv(mtx0, fileNameIn0)
 ```
@@
 ```{r}
+# Add CI size
+mtx0 <- mtx0 %>% mutate(`CI Biogrid`=`Biogrid Upper`-`Biogrid Lower`, `CI String`=`String Upper`-`String Lower`, `CI String Filt`=`String Filt Upper`-`String Filt Lower`)
 # Selected data for correlation
@@
 # Remove size columns
-mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"), contains("P value"), contains("Modularity"))
+mtx0_cor <- mtx0_cor %>% dplyr::select(contains("Coherence"), contains("Number"), contains("P value"), contains("Modularity"), contains("Slope"), contains("CI"))
 # Jitter STRING p-value a bit
@@
 print(paste0("Average correlation of coherences and p-value: ", mean(diff_cors_selected[upper.tri(diff_cors_selected)])))
+# Average correlation of slopes and network sizes
+diff_cors_selected <- diff_cors[grepl("Slope", rownames(diff_cors)), grepl("Number of genes", colnames(diff_cors))]
+print(paste0("Average correlation of sloped and the number of genes: ", mean(diff_cors_selected)))
+# Average correlation of CIs and network sizes
+diff_cors_selected <- diff_cors[grepl("CI", rownames(diff_cors)), grepl("Number of genes", colnames(diff_cors))]
+print(paste0("Average correlation of CIs and the number of genes: ", mean(diff_cors_selected)))
 
@@
 # Selected correlations
-png(filename = "../manuscript/Figures/FigS1_corr_coherence.png", height = 2000, width = 2000, res = 300)
+png(filename = "../manuscript/Figures/FigS1_corr_coherence.png", height = 3000, width = 3000, res = 300)
 # corrgram::corrgram(diff_cors, order = FALSE, upper.panel=panel.cor, cex.labels = 4)
@@
 ```
+
+# Confidence Interval investigation
+
+```{r eval=FALSE}
+mtx_ci_dotplot <- function(category = "Intelligence", mtx_traits = mtx, databases = c("Biogrid", "String", "String.Filt")) {
+
+  mtx_traits_selected <- mtx_traits[mtx_traits$Category == category, grepl(databases, colnames(mtx_traits), fixed = TRUE)]
+
+  mtx_traits_selected_to_plot <- data.frame(Disease = rownames(mtx_traits_selected), 
+                                            Slope = mtx_traits_selected[, paste0(databases, ".Slope")], 
+                                            Lower = mtx_traits_selected[, paste0(databases, ".Lower")],
+                                            Upper = mtx_traits_selected[, paste0(databases, ".Upper")])
+  # Remove noninformative rows
+  mtx_traits_selected_to_plot <- mtx_traits_selected_to_plot[complete.cases(mtx_traits_selected_to_plot), ]
+  # Sort by decreasing level of slope
+  mtx_traits_selected_to_plot$Disease <- factor(mtx_traits_selected_to_plot$Disease, levels = mtx_traits_selected_to_plot$Disease[order(mtx_traits_selected_to_plot$Slope)])
+
+  p <- ggplot() +
+    geom_point(data=mtx_traits_selected_to_plot, mapping=aes(x=Disease, y=Slope), size=4, shape=21, fill="white") +
+    geom_errorbar(data=mtx_traits_selected_to_plot, mapping=aes(x=Disease, ymin=Lower, ymax=Upper), width=0.2, size=1, color="blue") +  
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+    ylab("Slope (unnormalized coherence") +
+    # geom_text_repel(size = 4, colour = "black") +
+    coord_flip() +
+    # theme_cowplot(12) +
+    theme(legend.position="none") +
+    theme(axis.text = element_text(size = 12)) + 
+    ggtitle(category)
+}
+
+```

### 2020-04-05 manuscript/submission.BMC_Bioinformatics/Revision1/cover_letter_BMC_bioinformatics_R1.Rmd
Latest changes for R1 submission

---
output:
  word_document:
    reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NSF_grant_style.docx
  html_document: default
---

Dear Dr. Talbot,

This submission is the full revision of the manuscript BINF-D-19-01133. The title of the revised manuscript is "A Method for Estimating Coherence of Molecular Mechanisms In Major Human Disease and Traits". This manuscript will be of interest to the network analysis community working with the analysis of molecular networks.

All comments raised during previous revisions have been addressed. The point-by-point response is attached.

The brief overview of our revised work are:

- Focus on a novel method for estimating and comparing the coherence of molecular-interaction networks. This method is independent of the size of molecular networks, allowing for comparing coherences across phenotype-specific networks of different sizes. The high coherence of a molecular network may inform about the therapeutically targetable molecular mechanisms.
- Coherence estimation and comparison across 116 phenotype-specific networks from seven disease- and five trait categories, including the estimation of statistical significance of coherence measures.
- Validation of our approach using different protein-protein interaction (PPI) databases.
- Identification of diseases and traits with high or low coherence. E.g., metabolic diseases and traits showed consistently high coherence, while psychiatric disorders and intelligence-related traits were the least coherent.
- Demonstration that coherence and the commonly used modularity measures capture different network properties.

All data used in the manuscript was sourced from publicly available repositories (GWAScatalog, STRING, and Biogrid PPIs, MSigDb). All code and data needed to reproduce our results are made available on our GitHub repository https://github.com/dozmorovlab/disease_coherence. We also want to assure that this manuscript is not being considered for publication elsewhere.

=== Start of Point-by-point response ===

This resubmission has been extensively reworked with new results added. Major changes include:

- New title to reflect the fact the manuscript presents the method
- New results section "Networks with high/low coherence have distinct degree distributions"
    - New Figure 3 showing the degree distributions. Renaming of the following figures
    - New Additional File 9 showing the statistics of differential analysis of the degree distributions
- New results section "Coherence captures distinct network properties than modularity"
    - Updated Additional File 3 table showing modularity estimation
    - Updated Additional File 4 showing the correlation of modularity with other measures
- New results describing the untransformed slopes and the corresponding confidence intervals
    - Updated Additional File 3 table showing slopes and the corresponding confidence intervals in numerical format
    - Updated Additional File 4 showing the correlation of slopes and the sized of their confidence intervals with other measures
- All results have been described in the corresponding Results sections, Methods sections and legends got updated

Reviewer reports:

> This manuscript introduces a method, "network coherence", and applies it to human disease genes (from GWAS); in addition to introducing a new method, they make a claim that the coherence measure varies among specific disorders in a biologically significant manner. As such this work needs to be evaluated in these two aspects: First, whether the method is novel and sufficiently well-validated. Second, whether the data support the claims about human phenotypes.

Thank you for clearly summarizing our work. We adjusted the title to reflect the fact that the paper presents the method as "A Method for Estimating Coherence of Molecular Mechanisms In Major Human Disease and Traits."

> Taking up the second aspect first, there is a major problem with using PPINs from BioGRID and STRING: these networks are extremely biased by "well-studied genes" or other types of artifacts. This means using them to make conclusions about anything biological is horribly confounded with the effects of human decisions about which proteins to use as targets among other issues. What the authors have failed to do is provide any evidence that their results are not simply due to such biases. Note that making random networks does not solve this problem - it is indeed "significant" that genes like TGFB have a lot of interactions, but it's far from clear that's due to biology. I believe all of the papers the authors cite in support of the idea that coherence (et al.) is a documented phenomenon associated with human diseases (pg 3 line 20) all have this issue, and the point of ref 27 has not, to my knowledge, been rebutted.

We agree that the PPI databases are potentially biased towards well-studied genes. Databases like STRING try to avoid it by aggregating PPI information from multiple sources, including the predicted interactions. We intentionally used the full STRING database to maximally avoid PPI bias towards the well-studied genes. Yet, at this state of the knowledge, PPI databases remain incomplete (c.f. Menche et al. 2015). This was the rationale to mention the incompleteness of the databases throughout the manuscript. E.g., we discussed that our method "... will only improve as more PPI information becomes available", and devoted the Discussion paragraph explaining the limitations of PPI databases.

While we acknowledge the human bias and incompleteness of the PPI databases and GWASs, the fundamental question is why some genes are studied much more than others? Historically, such genes emerge as a "low hanging fruit" in molecular biology experiments, consistently appear as significant in GWAS studies, that is, represent the true biological signal. Consequently, they formed the backbone for forming the PPI databases, for building sets of GWAS loci. While we acknowledge the fact of bias, the source of it is likely biology-driven.

The bias and incompleteness of the PPI databases led us to use the permutation test. Both PPI databases and GWASs attempt to extract biological signal genome-wide. The fact that, e.g., TGFB got picked up by a GWAS study, and is highly represented in a PPI database is likely biology-driven. The fact that random networks have much less chance to pick up TGFB is due to the random, non-biological processes. The permutation test works in the same realm of PPIs that were used to build phenotype-specific networks. Being non-parametric, the permutation test equally treats genes with average degree distributions and those with extreme PPIs. If we had used parametric tests, chances of getting biased results would be high as outliers typically break the rules of parametric distributions leading to overinflated results.

> I have to question the limitation to GWAS loci. The authors concede that in effect their analysis cannot be used to conclude much of anything, due to gaps in knowledge about the genetics of the phenotypes (line 3 pg 14). Since the authors are aware of this problem, that they didn't use genes implicated across the spectrum of variant frequencies and effect sizes (including somatic variants in cancer) which are invisible to GWAS is surprising. So I wholeheartedly agree with their caveat, but the issue with the networks used is just as important. Overall, the results of the application of the method are very unconvincing to me.

We agree with the reviewer and intentionally chose our wording very carefully to highlight caveats in the current state of knowledge of both PPI databases and GWAS loci and emphasize the suggestive nature of our results. Additionally, we carefully selected references highlighting the whole spectrum of current knowledge and views, e.g., intentionally including reference 27 as an alternative to the current state of knowledge supported by several other peer-reviewed references. We did use cancer GWAS hits provided by GWAScatalog. Our goal was to present the methodology that can be used to compare molecular coherence of phenotypes while overcoming limitations of network analysis (differences in size, configuration). It was beyond the scope of our work to evaluate genes associated with rare variants, different effect sizes. The method itself can be applied to any GWAS results (see Discussion), including rare and somatic mutations, and reveal how coherent the associated gene networks are. We intentionally used the state-of-the-art list of GWAS results from GWAScatalog, and acknowledge the limitations of the current knowledge.

> None of this detracts, in principle, from the presentation of the method, since a method can exist and be novel in the absence of any useful or interesting application as these may emerge in the future. But as this is a bioinformatics paper, the standard format is to compare the results obtained on a gold standard, and to pit the method against other approaches either in reference to the gold standard (not that one could be defined here, except perhaps by simulations), or at least to the results obtained in actual use and describe the advantages, or at least differences, of the method to other approaches. The authors have done none of those things.

Indeed, the primary intention of our work is to present a principled method allowing for a comparison of the coherence of molecular mechanisms across phenotypes. While we are fully aware of the typical strategy for presenting novel bioinformatics methods (e.g., see https://www.biorxiv.org/content/10.1101/549170v3 for an example of our approach to benchmarking), the problem here is that there are no other similar approaches that we are aware of. Nor is there a biological gold standard other than phenotype-specific gene networks used in our work. As described in the manuscript, random and KEGG networks are the best approximators of random and high coherence. The permutation framework is the most unbiased simulation strategy, in which we simulate random networks of the same size as phenotype-specific networks to get the significance of coherence estimates.

> In terms of novelty, as the authors discuss, there are other measures of coherence, and one that is particularly relevant is modularity - various measures of "cluster tightness" relative to a network context go back many decades. Since modularity is literally the measure of the ratio of the density of in-group and to-group edges, and the authors repeatedly described modularity in just those terms (e.g. on pg 13), it isn't clear that method of the authors is even needed.

We performed additional analyses of the association between modularity and coherence. We did not expect to find any strong association between coherence and modularity as the latter depends on network size. Our results confirmed this expectation and did not reveal a consistent association between coherence and modularity. We added the corresponding Results and Methods sections, discussed our observations, and updated the Additional Files 3 and 4 with modularity information. Again, our goal was to develop a general-purpose method for estimating and comparing coherence of molecular mechanisms across phenotypes, with the coherence normalization approach being one of the novel aspects. Normalization of modularity, while theoretically possible, would defy interpretability. Our results demonstrate that coherence is different from modularity. We hope that the simplicity, flexibility, and interpretability of our measure of coherence will be an important contribution to the field of network analysis in genomics.

> It is striking that the authors never provide an actual example of the node degree distributions for different phenotypes, since they form the basis for the regression analysis. As it stands, the novelty here seems minor, and directly comparing their method to other (including simpler) alternatives is essential.

We agree with the need to investigate and compare degree distributions. We added the results section and the new Figure 3 showing internal and external degree distributions for high- and low-coherence networks, and statistically compared them with size-matched random networks. Additional File 9 shows the results of the statistical comparison. In summary, the new results confirm the notion that networks of high coherence have a significantly larger number of internal and external connections than low-coherence networks and random networks. Regarding the notion of comparison with other alternatives, please, see our response to the previous comment.

> Another major drawback of the approach as presented is that there is no estimate of variance in the measure, that is, coherence should be treated as a random variable. In order to say that phenotype X has a higher coherence than phenotype Y, one would like to know the distributional properties of the point estimates, as I'm sure the authors are aware. It is not the same as saying a coherence value is different from random. The authors should provide confidence intervals etc. or remove all claims that one set of phenotypes has a "lower coherence" than another. The closest the authors get is comparing STRING and BioGRID (the authors do not mention their lack of independence; these databases overlap extensively). And the comparison is very qualitative anyway, for example saying on p9 line 22 that the "overall correlation remained high" without any numbers given.

We agree that confidence intervals are important information that should be presented. We now report the untransformed slopes and the corresponding confidence intervals in numerical form (Additional file 3) and in terms of their correlations with network size (Additional file 4). It should be noted, however, that the untransformed slopes and their confidence intervals cannot be directly compared as they are expectedly affected by network size. We described those results in the "Network size and configuration hinders direct comparison of degree distributions" section. Normalizing confidence intervals using our strategy for normalizing slopes to size-matched random-high coherence range defy interpretability of confidence intervals in that their overlap would not be indicative of (the presence or the lack of) significant difference. Overall, we tried to keep our narrative maximally accessible to the reader while being quantitatively accurate. We report all numerical data in the main text or the additional files, e.g., the "overall correlation remained high" results can be easily seen on the cited Additional File 4 figure.

=== End of Point-by-point response ===

With our best regards,  
On behalf of all authors,

Mikhail Dozmorov, Ph.D., Blick scholar  
Assistant Professor,  Department of Biostatistics 
Affiliate, Department of Pathology  
Virginia Commonwealth University  
One Capitol Square, room 738  
830 East Main Street  
Richmond,  Virginia  23298  
e-mail:  mikhail.dozmorov@vcuhealth.org  
Phone: 804-827-2055

### 2020-04-05 manuscript/submission.BMC_Bioinformatics/Revision1/manuscript_BMC_Bioinformatics_R1.Rmd
Tweak title
Latest changes for R1 submission

@@
 ---
-bibliography: [references/networks.bib,references/brain.bib,references/enrichment.bib,references/3D_refs.bib]
-csl: styles.ref/bmc-bioinformatics.csl
+bibliography: /Users/mdozmorov/Documents/Work/presentations/Templates/manuscript_template/references/references_MD.bib
+csl: ../../styles.ref/bmc-bioinformatics.csl
 output:
   word_document:
-    reference_docx: styles.doc/BMC_Bioinformatics.docx
+    reference_docx: ../../styles.doc/BMC_Bioinformatics.docx
   pdf_document: default
@@
 
-# Coherence of Molecular Mechanisms <!--in Genes Identified Through GWAS -->In Major Human Disease and Traits
+# <!-- New in Revision 1 -->A Method for Estimating <!-- End New in Revision 1 -->Coherence of Molecular Mechanisms <!--in Genes Identified Through GWAS -->In Major Human Disease and Traits
 
@@
 
-**Background.** Complex phenotypes such as height and intelligence, are thought to be a product of the collective effects of multiple phenotype-associated genes, or, more precisely, interactions among their protein products.
+**Background.** Phenotypes such as height and intelligence, are thought to be a product of the collective effects of multiple phenotype-associated genes and interactions among their protein products.
 High/low degree of interactions is suggestive of coherent/random molecular mechanisms, respectively.
-Comparing the degree of interactions may help to better understand the coherence of molecular mechanisms underlying complex phenotypes and the potential for therapeutic intervention.
+Comparing the degree of interactions may help to better understand the coherence of phenotype-specific molecular mechanisms and the potential for therapeutic intervention.
 However, direct comparison of the degree of interactions is difficult due to different sizes and configurations of phenotype-associated gene networks.
 
-**Methods.** We introduce a measure of network coherence as a slope of internal vs.
+**Methods.** <!-- New in Revision 1 -->We introduce a metric for measuring coherence of molecular-interaction networks<!-- End New in Revision 1 --> as a slope of internal vs.
 external distributions of the degree of interactions.
@@
 Metabolic traits and diseases were the most coherent, while psychiatric disorders and intelligence-related traits were the least coherent.
+<!-- New in Revision 1 -->We demonstrate that coherence and modularity measures capture distinct network properties.<!-- End New in Revision 1 -->
 
-**Conclusions.** We present a general-purpose method for estimating and comparing the coherence of phenotype-associated gene networks that accounts for the network size and shape differences.
+**Conclusions.** We present a general-purpose method for estimating and comparing the coherence of molecular-interaction gene networks that accounts for the network size and shape differences.
 Our results highlight gaps in our current knowledge of genetics and molecular mechanisms of complex phenotypes and suggest priorities for future GWASs.
@@
 Furthermore, the corresponding genes are often spread across different chromosomes with no known connection to one another, hindering understanding of the molecular mechanisms.
-These limitations of current generation GWAS might stem from the yet incomplete knowledge of genetic determinants of complex phenotypes, the potential heterogeneity of complex phenotypes and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].
+These limitations of current generation GWAS might stem from the yet incomplete knowledge of genetic determinants of complex phenotypes, the potential heterogeneity and/or the causes other than genetics [@Botstein:2003aa; @turkheimer2012genome; @Sullivan:2018aa].
 
-Many studies have shown that genetically-driven complex phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa].
+Many studies have shown that genetically-driven phenotypes are often associated with functionally related genes that are more likely to form networks of interacting protein products [@Hartwell:1999aa; @Krauthammer:2004aa; @Huang:2006aa; @Lage:2007aa; @Cordell:2009aa; @Emily:2009aa; @Feldman:2008aa].
 This observation has been verified systematically for a large number of diseases [@Menche:2015aa], thus confirming a fundamental hypothesis of the interactome-based approach to understanding human phenotypes, namely that disease- and trait-associated genes tend to form interaction modules [@Goh:2007aa].
-Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all molecular determinants of a certain phenotype.
-Consequently, methods to identify phenotype-associated networks of functionally-related genes have been developed [@Krauthammer:2004aa; @Cordell:2009aa; @Akula:2011aa].
+Such interaction modules may be viewed as connected subnetworks within the full interactome and may contain all the molecular determinants of a certain phenotype.
+Consequently, methods to quantitatively assess these phenotype-associated networks of functionally-related genes have been developed [@Krauthammer:2004aa; @Cordell:2009aa; @Akula:2011aa].
 
@@
 Networks formed by interactions among phenotype-associated genes, or, more precisely, their protein products, are thought to have high connectivity [@Huang:2006aa; @Krauthammer:2004aa; @Goh:2007aa].
-This intuition reflects a well-known "guilt-by-association" principle that genes (or, more precisely, their products) that form an interaction network are more likely to share similar functions [@Farber:2013aa], or co-expression patterns [@Huang:2006aa; @Michalak:2008aa].
+This intuition reflects a well-known "guilt-by-association" principle that genes (or, more precisely, their products) forming an interaction network are more likely to have similar functions [@Farber:2013aa] and co-expression patterns [@Huang:2006aa; @Michalak:2008aa].
 Although this view has been criticized [@Gillis:2012aa], several studies consistently observed phenotype-associated genes to be either connected as a single network or to participate in common phenotype-specific subnetworks [@Menche:2015aa; @Lee:2011aa; @Akula:2011aa; @Ideker:2008aa; @Iossifov:2008aa].
@@
 Using two PPI databases (STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa]), we compared coherence of 133 phenotypes from GWAScatalog [@Hindorff:2009aa; @MacArthur:2017aa].
-Our results show the tendency of complex phenotypes, such as intelligence-related traits, to have low coherence and metabolic traits and diseases to have high coherence.
+Our results show the tendency of intelligence-related traits to have low coherence and metabolic traits and diseases to have high coherence.
 Our method enables direct comparison of coherence measures and highlights gaps in the current understanding of molecular mechanisms of many phenotypes, e.g., Major Depressive Disorder having the lowest coherence in the already low-coherence "Psychiatric disease" category.
@@
 We created networks of phenotype-associated genes using PPI information from STRING [@Szklarczyk:2015aa] and BioGrid [@Chatr-Aryamontri:2017aa] databases (see Methods).
-We opted for the use of three types of PPI data.
-First, STRING data filtered by interaction confidence score above 500 (middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data.
-This filtering step ensures that only high-confidence PPIs were selected.
+We used three types of PPI data.
+First, STRING data filtered by interaction confidence score above 500 (the middle of the bimodal distribution of the score, 0-1000 range, referred hereafter as “STRING filtered”) was used as the primary source of curated PPI data.
+This filtering step selects for only high-confidence PPIs.
 The advantage of using filtered data is the reliance on high-quality PPIs.
-The disadvantage is that the number of PPIs may be insufficient for forming phenotype-associated networks; consequently, phenotypes with genes without PPIs were removed from the analysis (see Methods) leaving 116 phenotypes that could be analyzed.
-Second, full STRING data was used to maximize the use of PPI information at the expense of potential noise (referred to hereafter as “STRING”).
+The disadvantage is that, for some phenotypes, the number of PPIs may be insufficient for forming networks.
+Consequently, these phenotypes were removed from the analysis (see Methods) leaving 116 phenotype-associated gene sets that could be analyzed.
+Second, full STRING data (henceforth, "STRING") was used to maximize the use of PPI information at the expense of potential noise.
 The advantage of using complete data is that more trait-associated genes will have PPI information and, hence, can be analyzed.
-The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs having low confidence scores.
+The disadvantage, however, is that the results may be less reliable due to the presence of noisy PPIs with low confidence scores.
 Third, BioGrid PPI data, which has been curated to keep only high-confidence interactions, was used.
@@
 We observed substantial negative correlation between network sizes and the slopes of KEGG and random networks (Pearson Correlation Coefficient (PCC) -0.58/-0.86, respectively, Figure 2, Additional File 2).
-These observations confirm the notion that coherence estimation using internal vs.
+<!-- New in Revision 1 -->Similarly, slopes of the phenotype-specific networks and sizes of the corresponding confidence intervals negatively correlated with the number of genes in the networks (Average PCC -0.49/-0.70, respectively, Additional File 3, Additional File 4).
+This is expected as network size affects internal degree distributions and the corresponding slopes on the internal vs.
+external degree plots, with the larger networks allowing for a more precise estimation of slopes (smaller confidence intervals).
+<!-- End New in Revision 1 -->These observations confirm the notion that coherence estimation using internal vs.
 external degree distributions should be controlled for network size.
@@
 
-This work was supported by a Genetics and Human Agency Award from the Templeton Foundation to KSK, MR, CC, SB.
+This work was supported by a Genetics and Human Agency Award GF13361-152627 from the Templeton Foundation to KSK, MR, CC, SB.
 The content is solely the responsibility of the authors and does not necessarily represent the official views of the Templeton Foundation.
@@
 These values represent normalized coherence and can be compared across phenotypes and PPI databases.
+<!-- New in Revision 1 -->In addition, for each coefficient $\beta_i$, we extract confidence intervals and normalize them as the coefficients themselves.
+Note the normalization step alters common interpretation of confidence intervals; therefore, results of the permutation test of slopes should be used to compare coherences.<!-- End New in Revision 1 -->
 
@@
 Phenotype-specific coherence estimates.** The PPI database (STRING, STRING filtered, Biogrid) are specified in the corresponding column names.
-For each phenotype category, the results are sorted by the "String Filt Coherence" column in descending order.
-"NA" indicates the corresponding value cannot be estimated due to lack of sufficient number of genes annotated with PPIs.
+For each phenotype category, the results are sorted alphabetically.
+<!-- New in Revision 1 -->In addition to the normalized coherence and the permutation p-values indicating significant difference of phenotype-specific coherence from random, modularity, untransformed slopes, and the corresponding lower and upper confidence interval bounds are shown.
+<!-- New in Revision 1 -->"NA" indicates the corresponding value cannot be estimated due to lack of sufficient number of genes annotated with PPIs.
 
 **Additional File 4.
-Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, and the normalized coherence estimates using the corresponding PPI databases.
+Normalization of coherence estimates alleviates its network size dependence.** Red/blue gradient and numbers represent Pearson correlation coefficients for each pairwise comparison of the number of SNPs, genes, normalized coherence estimates, <!-- New in Revision 1 -->and the untransformed slopes and the sizes of their confidence intervals (CI)<!-- End New in Revision 1 --> using the corresponding PPI databases.
 

### 2020-04-06 manuscript/submission.BMC_Bioinformatics/Revision1/Response_R1.Rmd
Tweak title
Make requested formatting changes

@@
 output:
+  pdf_document: default
+  html_document: default
   word_document:
     reference_docx: /Users/mdozmorov/Documents/Work/presentations/Templates/grant_template/styles.doc/NSF_grant_style.docx
-  html_document: default
 ---
 
-Reviewer reports:
+# Revision changes
+
+This resubmission has been extensively reworked with new results added. Major changes include:
+
+- New title to reflect the fact the manuscript presents the method
+- New results section "Networks with high/low coherence have distinct degree distributions"
+    - New Figure 3 showing the degree distributions. Renaming of the following figures
+    - New Additional File 9 showing the statistics of differential analysis of the degree distributions
+- New results section "Coherence captures distinct network properties than modularity"
+    - Updated Additional File 3 table showing modularity estimation
+    - Updated Additional File 4 showing the correlation of modularity with other measures
+- New results describing the untransformed slopes and the corresponding confidence intervals
+    - Updated Additional File 3 table showing slopes and the corresponding confidence intervals in numerical format
+    - Updated Additional File 4 showing the correlation of slopes and the sized of their confidence intervals with other measures
+- All results have been described in the corresponding Results sections, Methods sections and legends got updated
+
+# Reviewer response
 
@@
 Second, whether the data support the claims about human phenotypes.
+
+Thank you for clearly summarizing our work.
+We adjusted the title to reflect the fact that the paper presents the method as "A Method for Estimating Coherence of Molecular Mechanisms In Major Human Disease and Traits."
 
@@
 We agree that the PPI databases are potentially biased towards well-studied genes.
-Yet, databases like STRING try to avoid it by aggregating PPI information from multiple sources, including the predicted interactions.
+Databases like STRING try to avoid it by aggregating PPI information from multiple sources, including the predicted interactions.
 We intentionally used the full STRING database to maximally avoid PPI bias towards the well-studied genes.
@@
 E.g., we discussed that our method "...
-will only improve as more PPI information becomes available", and devoted the whole paragraph explaining the limitations of PPI databases.
+will only improve as more PPI information becomes available", and devoted the Discussion paragraph explaining the limitations of PPI databases.
 
-While we acknowledge the incompleteness of the PPI databases and GWASs, the fundamental question is why some genes are studied much more than others?
-Historically, such genes emerge as a "low hanging fruit" in molecular biology experiments, consistently appear in GWAS studies, that is, representing the true biological signal.
+While we acknowledge the human bias and incompleteness of the PPI databases and GWASs, the fundamental question is why some genes are studied much more than others?
+Historically, such genes emerge as a "low hanging fruit" in molecular biology experiments, consistently appear as significant in GWAS studies, that is, represent the true biological signal.
 Consequently, they formed the backbone for forming the PPI databases, for building sets of GWAS loci.
@@
 
-We fully agree with the reviewer and intentionally chose our wording very carefully to highlight caveats in the current state of knowledge of both PPI databases and GWAS loci.
+We agree with the reviewer and intentionally chose our wording very carefully to highlight caveats in the current state of knowledge of both PPI databases and GWAS loci and emphasize the suggestive nature of our results.
 Additionally, we carefully selected references highlighting the whole spectrum of current knowledge and views, e.g., intentionally including reference 27 as an alternative to the current state of knowledge supported by several other peer-reviewed references.
-Our goal was to present the method, the approach that can be used to compare molecular coherence of phenotypes while overcoming limitations of network analysis (differences in size, configuration).
+We did use cancer GWAS hits provided by GWAScatalog.
+Our goal was to present the methodology that can be used to compare molecular coherence of phenotypes while overcoming limitations of network analysis (differences in size, configuration).
 It was beyond the scope of our work to evaluate genes associated with rare variants, different effect sizes.
 The method itself can be applied to any GWAS results (see Discussion), including rare and somatic mutations, and reveal how coherent the associated gene networks are.
-We intentionally used the state-of-the-art list of GWAS results, and acknowledge the limitations of the current knowledge.
+We intentionally used the state-of-the-art list of GWAS results from GWAScatalog, and acknowledge the limitations of the current knowledge.
 
@@
 Indeed, the primary intention of our work is to present a principled method allowing for a comparison of the coherence of molecular mechanisms across phenotypes.
-While we are fully aware of the typical strategy for presenting novel bioinformatics methods (e.g., see https://www.biorxiv.org/content/10.1101/549170v3 for an example of such presentation), the problem here is that there are no other similar approaches that we are aware of.
+While we are fully aware of the typical strategy for presenting novel bioinformatics methods (e.g., see https://www.biorxiv.org/content/10.1101/549170v3 for an example of our approach to benchmarking), the problem here is that there are no other similar approaches that we are aware of.
 Nor is there a biological gold standard other than phenotype-specific gene networks used in our work.
@@
 We performed additional analyses of the association between modularity and coherence.
-Our results did not reveal a consistent association between coherence and modularity.
+We did not expect to find any strong association between coherence and modularity as the latter depends on network size.
+Our results confirmed this expectation and did not reveal a consistent association between coherence and modularity.
 We added the corresponding Results and Methods sections, discussed our observations, and updated the Additional Files 3 and 4 with modularity information.
@@
 In summary, the new results confirm the notion that networks of high coherence have a significantly larger number of internal and external connections than low-coherence networks and random networks.
-
 Regarding the notion of comparison with other alternatives, please, see our response to the previous comment.
@@
 And the comparison is very qualitative anyway, for example saying on p9 line 22 that the "overall correlation remained high" without any numbers given.
+
+We agree that confidence intervals are important information that should be presented.
+We now report the untransformed slopes and the corresponding confidence intervals in numerical form (Additional file 3) and in terms of their correlations with network size (Additional file 4).
+It should be noted, however, that the untransformed slopes and their confidence intervals cannot be directly compared as they are expectedly affected by network size.
+We described those results in the "Network size and configuration hinders direct comparison of degree distributions" section.
+Normalizing confidence intervals using our strategy for normalizing slopes to size-matched random-high coherence range defy interpretability of confidence intervals in that their overlap would not be indicative of (the presence or the lack of) significant difference.
+Overall, we tried to keep our narrative maximally accessible to the reader while being quantitatively accurate.
+We report all numerical data in the main text or the additional files, e.g., the "overall correlation remained high" results can be easily seen on the cited Additional File 4 figure.

### 2020-04-06 manuscript/submission.BMC_Bioinformatics/Revision1/cover_letter_BMC_bioinformatics_R1.Rmd
Make requested formatting changes

@@
 
-All comments raised during previous revisions have been addressed. The point-by-point response is attached.
+All comments raised during previous revisions have been addressed. The point-by-point response is submitted as a supplementary file.
 
@@
 
-=== Start of Point-by-point response ===
-
-This resubmission has been extensively reworked with new results added. Major changes include:
-
-- New title to reflect the fact the manuscript presents the method
-- New results section "Networks with high/low coherence have distinct degree distributions"
-    - New Figure 3 showing the degree distributions. Renaming of the following figures
-    - New Additional File 9 showing the statistics of differential analysis of the degree distributions
-- New results section "Coherence captures distinct network properties than modularity"
-    - Updated Additional File 3 table showing modularity estimation
-    - Updated Additional File 4 showing the correlation of modularity with other measures
-- New results describing the untransformed slopes and the corresponding confidence intervals
-    - Updated Additional File 3 table showing slopes and the corresponding confidence intervals in numerical format
-    - Updated Additional File 4 showing the correlation of slopes and the sized of their confidence intervals with other measures
-- All results have been described in the corresponding Results sections, Methods sections and legends got updated
-
-Reviewer reports:
-
-> This manuscript introduces a method, "network coherence", and applies it to human disease genes (from GWAS); in addition to introducing a new method, they make a claim that the coherence measure varies among specific disorders in a biologically significant manner.
-As such this work needs to be evaluated in these two aspects: First, whether the method is novel and sufficiently well-validated.
-Second, whether the data support the claims about human phenotypes.
-
-Thank you for clearly summarizing our work.
-We adjusted the title to reflect the fact that the paper presents the method as "A Method for Estimating Coherence of Molecular Mechanisms In Major Human Disease and Traits."
-
-> Taking up the second aspect first, there is a major problem with using PPINs from BioGRID and STRING: these networks are extremely biased by "well-studied genes" or other types of artifacts.
-This means using them to make conclusions about anything biological is horribly confounded with the effects of human decisions about which proteins to use as targets among other issues.
-What the authors have failed to do is provide any evidence that their results are not simply due to such biases.
-Note that making random networks does not solve this problem - it is indeed "significant" that genes like TGFB have a lot of interactions, but it's far from clear that's due to biology.
-I believe all of the papers the authors cite in support of the idea that coherence (et al.) is a documented phenomenon associated with human diseases (pg 3 line 20) all have this issue, and the point of ref 27 has not, to my knowledge, been rebutted.
-
-We agree that the PPI databases are potentially biased towards well-studied genes.
-Databases like STRING try to avoid it by aggregating PPI information from multiple sources, including the predicted interactions.
-We intentionally used the full STRING database to maximally avoid PPI bias towards the well-studied genes.
-Yet, at this state of the knowledge, PPI databases remain incomplete (c.f.
-Menche et al.
-2015).
-This was the rationale to mention the incompleteness of the databases throughout the manuscript.
-E.g., we discussed that our method "...
-will only improve as more PPI information becomes available", and devoted the Discussion paragraph explaining the limitations of PPI databases.
-
-While we acknowledge the human bias and incompleteness of the PPI databases and GWASs, the fundamental question is why some genes are studied much more than others?
-Historically, such genes emerge as a "low hanging fruit" in molecular biology experiments, consistently appear as significant in GWAS studies, that is, represent the true biological signal.
-Consequently, they formed the backbone for forming the PPI databases, for building sets of GWAS loci.
-While we acknowledge the fact of bias, the source of it is likely biology-driven.
-
-The bias and incompleteness of the PPI databases led us to use the permutation test.
-Both PPI databases and GWASs attempt to extract biological signal genome-wide.
-The fact that, e.g., TGFB got picked up by a GWAS study, and is highly represented in a PPI database is likely biology-driven.
-The fact that random networks have much less chance to pick up TGFB is due to the random, non-biological processes.
-The permutation test works in the same realm of PPIs that were used to build phenotype-specific networks.
-Being non-parametric, the permutation test equally treats genes with average degree distributions and those with extreme PPIs.
-If we had used parametric tests, chances of getting biased results would be high as outliers typically break the rules of parametric distributions leading to overinflated results.
-
-> I have to question the limitation to GWAS loci.
-The authors concede that in effect their analysis cannot be used to conclude much of anything, due to gaps in knowledge about the genetics of the phenotypes (line 3 pg 14).
-Since the authors are aware of this problem, that they didn't use genes implicated across the spectrum of variant frequencies and effect sizes (including somatic variants in cancer) which are invisible to GWAS is surprising.
-So I wholeheartedly agree with their caveat, but the issue with the networks used is just as important.
-Overall, the results of the application of the method are very unconvincing to me.
-
-We agree with the reviewer and intentionally chose our wording very carefully to highlight caveats in the current state of knowledge of both PPI databases and GWAS loci and emphasize the suggestive nature of our results.
-Additionally, we carefully selected references highlighting the whole spectrum of current knowledge and views, e.g., intentionally including reference 27 as an alternative to the current state of knowledge supported by several other peer-reviewed references.
-We did use cancer GWAS hits provided by GWAScatalog.
-Our goal was to present the methodology that can be used to compare molecular coherence of phenotypes while overcoming limitations of network analysis (differences in size, configuration).
-It was beyond the scope of our work to evaluate genes associated with rare variants, different effect sizes.
-The method itself can be applied to any GWAS results (see Discussion), including rare and somatic mutations, and reveal how coherent the associated gene networks are.
-We intentionally used the state-of-the-art list of GWAS results from GWAScatalog, and acknowledge the limitations of the current knowledge.
-
-> None of this detracts, in principle, from the presentation of the method, since a method can exist and be novel in the absence of any useful or interesting application as these may emerge in the future.
-But as this is a bioinformatics paper, the standard format is to compare the results obtained on a gold standard, and to pit the method against other approaches either in reference to the gold standard (not that one could be defined here, except perhaps by simulations), or at least to the results obtained in actual use and describe the advantages, or at least differences, of the method to other approaches.
-The authors have done none of those things.
-
-Indeed, the primary intention of our work is to present a principled method allowing for a comparison of the coherence of molecular mechanisms across phenotypes.
-While we are fully aware of the typical strategy for presenting novel bioinformatics methods (e.g., see https://www.biorxiv.org/content/10.1101/549170v3 for an example of our approach to benchmarking), the problem here is that there are no other similar approaches that we are aware of.
-Nor is there a biological gold standard other than phenotype-specific gene networks used in our work.
-As described in the manuscript, random and KEGG networks are the best approximators of random and high coherence.
-The permutation framework is the most unbiased simulation strategy, in which we simulate random networks of the same size as phenotype-specific networks to get the significance of coherence estimates.
-
-> In terms of novelty, as the authors discuss, there are other measures of coherence, and one that is particularly relevant is modularity - various measures of "cluster tightness" relative to a network context go back many decades.
-Since modularity is literally the measure of the ratio of the density of in-group and to-group edges, and the authors repeatedly described modularity in just those terms (e.g.
-on pg 13), it isn't clear that method of the authors is even needed.
-
-We performed additional analyses of the association between modularity and coherence.
-We did not expect to find any strong association between coherence and modularity as the latter depends on network size.
-Our results confirmed this expectation and did not reveal a consistent association between coherence and modularity.
-We added the corresponding Results and Methods sections, discussed our observations, and updated the Additional Files 3 and 4 with modularity information.
-Again, our goal was to develop a general-purpose method for estimating and comparing coherence of molecular mechanisms across phenotypes, with the coherence normalization approach being one of the novel aspects.
-Normalization of modularity, while theoretically possible, would defy interpretability.
-Our results demonstrate that coherence is different from modularity.
-We hope that the simplicity, flexibility, and interpretability of our measure of coherence will be an important contribution to the field of network analysis in genomics.
-
-> It is striking that the authors never provide an actual example of the node degree distributions for different phenotypes, since they form the basis for the regression analysis.
-As it stands, the novelty here seems minor, and directly comparing their method to other (including simpler) alternatives is essential.
-
-We agree with the need to investigate and compare degree distributions.
-We added the results section and the new Figure 3 showing internal and external degree distributions for high- and low-coherence networks, and statistically compared them with size-matched random networks.
-Additional File 9 shows the results of the statistical comparison.
-In summary, the new results confirm the notion that networks of high coherence have a significantly larger number of internal and external connections than low-coherence networks and random networks.
-Regarding the notion of comparison with other alternatives, please, see our response to the previous comment.
-
-> Another major drawback of the approach as presented is that there is no estimate of variance in the measure, that is, coherence should be treated as a random variable.
-In order to say that phenotype X has a higher coherence than phenotype Y, one would like to know the distributional properties of the point estimates, as I'm sure the authors are aware.
-It is not the same as saying a coherence value is different from random.
-The authors should provide confidence intervals etc.
-or remove all claims that one set of phenotypes has a "lower coherence" than another.
-The closest the authors get is comparing STRING and BioGRID (the authors do not mention their lack of independence; these databases overlap extensively).
-And the comparison is very qualitative anyway, for example saying on p9 line 22 that the "overall correlation remained high" without any numbers given.
-
-We agree that confidence intervals are important information that should be presented.
-We now report the untransformed slopes and the corresponding confidence intervals in numerical form (Additional file 3) and in terms of their correlations with network size (Additional file 4).
-It should be noted, however, that the untransformed slopes and their confidence intervals cannot be directly compared as they are expectedly affected by network size.
-We described those results in the "Network size and configuration hinders direct comparison of degree distributions" section.
-Normalizing confidence intervals using our strategy for normalizing slopes to size-matched random-high coherence range defy interpretability of confidence intervals in that their overlap would not be indicative of (the presence or the lack of) significant difference.
-Overall, we tried to keep our narrative maximally accessible to the reader while being quantitatively accurate.
-We report all numerical data in the main text or the additional files, e.g., the "overall correlation remained high" results can be easily seen on the cited Additional File 4 figure.
-
-=== End of Point-by-point response ===
-
 With our best regards,  
